### 2016-03-29 msmfs.tex
First commit

%rubber: module pdflatex

\documentclass[11pt,a4paper]{article}

\usepackage[sans]{tcc_doc} % SDP style file
\usepackage[english]{babel}
\usepackage{amssymb,amsmath} % for the math symbols
\usepackage{geometry}
\usepackage{marginnote}
\usepackage{datetime}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{mathtools}
\graphicspath{ {images/} }

% TODO:
% 1. Check if all new symbols have been explained when used for the first
% time.
% 2. Check if all symbols are actually being used.
% 3. Check the symbol definitions: rates, numbers, etc. to use the same main letter
% 4. Is there a way to use proper referencing of applicable and reference
% documents?

%%%%%%%%% START OF USER SETTINGS %%%%%%%%%%%%%%%%%%

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Implementation of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-SDP-151123-2\xspace}
% Context
\newcommand{\context}{(SDP.PIP)}
% Revision
\newcommand{\revision}{2\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwel\\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Draft\xspace}

% The definitions below are used for the signature tables in the preamble of
% the document
\newcommand{\firstsigname}{Name 1 - to be filled in}
\newcommand{\firstsigdesignation}{Designation 1 - to be filled in}
\newcommand{\firstsigaffiliation}{Affiliation 1 - to be filled in}

\newcommand{\secondsigname}{Name 2 - to be filled in}
\newcommand{\secondsigdesignation}{Designation 2 - to be filled in}
\newcommand{\secondsigaffiliation}{Affiliation 2 - to be filled in}

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
  \end{tabularx}
}

%%%%%%%%%%%%% END OF USER SETTINGS %%%%%%%%%%%%%%%%%%%

% Table with signatures
\newcommand{\signaturetable}[3]{
  \begin{tabularx}{\textwidth}{|X|X|X|}
    \hline
    Name & Designation & Affilitation\\
    \hline
    {#1} & {#2} & {#3}\\
    \hline
    Signature \& Date: & & \\
    & & \\
    & & \\
    \hline
  \end{tabularx}
}

% Table with affiliations
\newcommand{\organisationtable}{
  \begin{center}
    \sffamily{\bf ORGANISATION DETAILS}\end{center}
  \begin{table}[htb]
    \centering
    \begin{tabular}[htb]{|l|l|}
      \hline
      Name & Science Data Processor Consortium\\
      \hline
    \end{tabular}
  \end{table}
}

%%%%%%%%%%%%%%%%%%%% START YOUR DOCUMENT HERE %%%%%%%%%%%%%%%%%

\newcommand{\subsubsubsection}[1]{\noindent{\bf{#1}}}

%%%% Please DO USE the symbol definitions listed below to preserve
%%%% consistency.

%%%% If you need a new symbol, please add it here. If really want to modify
%%%% one, do so, just take care to do a search-and-replace for the entire document
% Symbol definitions
\newcommand{\bytes}{b} % Length of a complex number in bytes
\newcommand{\cbytes}{b} % Length of a real number in bytes
\newcommand{\rbytes}{b} % Length of a complex number in bytes
\newcommand{\bl}{B} % Baseline length
\newcommand{\maxbl}{B_{\mathrm{max}}} % Maximum baseline length
\newcommand{\membw}{B_\mathrm{mem}} % Memory bandwidth
\newcommand{\pciebw}{B_\mathrm{PCIe}} % PCI-e bandwidth
\newcommand{\viscompr}{C} % Compression factor of visibilities for gridding
                          % for facets
\newcommand{\vism}{\mathbb{D}} % Matrix of observed visibilities
\newcommand{\da}{D_{\mathrm A}} % Extent of aperture
\newcommand{\ds}{D_{\mathrm s}} % Diameter of antennas/stations
\newcommand{\reff}{f_\mathrm{ref}} % Reference frequency
\newcommand{\fcc}{F_{\mathrm{cc}}} % Convolution kernel computational cost
\newcommand{\fci}{F_{\mathrm{ci}}} % Factor between bandwidth and FLOPS (bytes/ops)
\newcommand{\fgrid}{F_{\mathrm{grid}}} % Gridding cost
\newcommand{\fpatch}{f_\mathrm{patch}} % Fraction of the maximum baseline below which the uv coverage is almost filled
\newcommand{\fpr}{F_\mathrm{pr}} % Cost for phase rotation
\newcommand{\gainm}{\mathbb{G}} % 2x2 block diagonal matrix of complex gains
% To change to N_something
\newcommand{\modskyvism}{\mathbb{M}} % Matrix of model sky visibilities
\newcommand{\skymodbufsize}{M_\mathrm{buf,skymod}} % Size of the sky model buffer
\newcommand{\visbufsize}{M_\mathrm{buf,vis}} % Size of the visibility buffer
\newcommand{\visbufsizespec}{M^\mathrm{spec}_\mathrm{buf,vis}} % Size of the visibility buffer needed for spectral line processing
\newcommand{\visbufsizecont}{M^\mathrm{cont}_\mathrm{buf,vis}} % Size of the visibility buffer needed for continuum imaging
\newcommand{\bufsize}{M_\mathrm{cu,buf}} % Size of buffer
\newcommand{\slowmemsize}{M_\mathrm{cu,pool}} %	Size of slow working memory
\newcommand{\cumemsize}{M_\mathrm{cu,work}} % Size of working memory of compute unit
\newcommand{\imgsize}{M_\mathrm{image}} % Size of image (bytes)
\newcommand{\imgsizeoneax}{M_\mathrm{image,ax}}	% Image size along one axis
\newcommand{\nbyteperpix}{N_\mathrm{byte,pix}} % Number of bytes per pixel
%\newcommand{\nbytepervis}{N_\mathrm{byte,vis}} % Number of bytes per pixel
\newcommand{\targgridsize}{M_\mathrm{target}} %	Target grid size (per facet)
\newcommand{\uvgridsize}{M_{uv,\mathrm{grid}}} % Size of the uv-grid (bytes)
\newcommand{\vissize}{M_\mathrm{vis}} % Size of visibility data (bytes)
\newcommand{\wgridsize}{M_\mathrm{weight\_grid}} % Size of weight grid (bytes)
\newcommand{\gridcachesize}{M_{w,\mathrm{cache}}} % Cache size of
                                % gridding kernel
\newcommand{\ntel}{N_\mathrm{a}} % Number of antennas or stations in the array
\newcommand{\nateam}{N_\mathrm{A}} % Number of A-team sources
\newcommand{\ngridpakern}{N_\mathrm{AA}} % Number of grid points for A-kernel support (along one axis)
\newcommand{\naccvis}{N_\mathrm{apv}} % Number of accesses per visibility
\newcommand{\navg}{N_\mathrm{avg}} % Number of averages
\newcommand{\nbins}{N_\mathrm{bin}} % Number of bins
\newcommand{\nbeam}{N_\mathrm{beam}} % Number of beams simultaneously observed by the array
\newcommand{\rma}{N_{rma}} % Number of flops in real multiply and add
\newcommand{\cma}{N_{cma}} % Number of flops in complex multiply and add
\newcommand{\nbits}{N_\mathrm{bit}} %	Number of bits
\newcommand{\nbl}{N_\mathrm{bl}} % Number of baselines
\newcommand{\nbyteacc}{N_\mathrm{bpa}} % Number of bytes per access
\newcommand{\nbyte}{N_\mathrm{byte}} % Number of bytes
\newcommand{\nch}{N_\mathrm{ch}XXXWANT TO USE NF HERE} % Number of channels
\newcommand{\ncompu}{N_\mathrm{cu}} % Number of compute units
\newcommand{\ndatait}{N_\mathrm{datait}} % Number of data items required per source (e.g. postion)
\newcommand{\ngridp}{N_\mathrm{cvff}} % Number of grid points along one axis for the gridding convolution function
\newcommand{\nfch}{N_\mathrm{f}} % Number of frequency channels
\newcommand{\nfchcorr}{N_\mathrm{f,corr}} % Number of frequency channels output by the correlator
\newcommand{\nfnosmear}{N_\mathrm{f,smear}} % Number of frequency channels set by smearing limit
\newcommand{\nfchgrid}{N_\mathrm{f,grid}} % Number of frequency channels to be gridded
\newcommand{\nfchdegrid}{N_\mathrm{f,de-grid}} % Number of frequency channels to be de-gridded
\newcommand{\nfchout}{N_\mathrm{f,out}} % Number of output frequency channels
                                % (to be FFT-ed and reprojected or output from
                                % the CSP pulsar timing backend
\newcommand{\nfchdist}{N_\mathrm{f,distribute}} % Minimum number of frequencies needed to maintain distributability
\newcommand{\nfkernel}{N_\mathrm{f,kernel}} %Number of convolution kernels needed to cover frequency axis for a particular baseline
\newcommand{\nfacet}{N_\mathrm{facet}} % Number of facets along one axis
\newcommand{\facetoverlap}{r_\mathrm{facet}} %Facet overlap overhead
\newcommand{\nflop}{N_\mathrm{FLOP}} % Number of floating point operations
\newcommand{\nflopsamp}{N_\mathrm{FLOPsamp}} % Number of floating point
                                % operations per sample
\newcommand{\nflopvis}{N_\mathrm{FLOPvis}} % Number of floating point
                                % operations per visibility 
\newcommand{\npixgrid}{N_\mathrm{grid}} % Number of pixels in the grid (i.e. the square of the linear dimension)
\newcommand{\ngridwkern}{N_\mathrm{GW}} % Number of grid points for w-kernel
                                % support (along one axis)
\newcommand{\nit}{N_\mathrm{it}} % Number of iterations
\newcommand{\nawkern}{N_\mathrm{kernel}} % Linear size of combined A and w-kernels
\newcommand{\nmaj}{N_\mathrm{major}} % Number of major CLEAN cycles
\newcommand{\nmin}{N_\mathrm{minor}} % Number of minor CLEAN cycles
\newcommand{\nselfcal}{N_\mathrm{SelfCal}} %Number of self cal loops
% TODO: Factors should have another letter than N
\newcommand{\facgridoffdiagmuller}{N_\mathrm{mm}} % Factor to account for
                                % gridding of off-diagonal Mueller terms
% TODO: Factors should have another letter than N
\newcommand{\nops}{N_\mathrm{ops}} % Number of operations
\newcommand{\npatchpix}{N_\mathrm{patch,pix}} % Number of pixels of PSF to be used in minor cycle decorrelation
\newcommand{\npix}{N_\mathrm{pix}} % Number of pixels along one axis of a grid
\newcommand{\npixfacet}{N_\mathrm{pix,facet}} % Number of pixels along one axis of a single image facet
\newcommand{\npol}{N_\mathrm{pol}} % Number of polarisation products
\newcommand{\nsubint}{N_\mathrm{subint}} % Number of sub-integrations
\newcommand{\ntslot}{N_t} % Number of time slots
\newcommand{\ntaylor}{N_\mathrm{Tt}} % Number of Taylor terms
\newcommand{\ntslotave}{N_{t,\mathrm{av}}} % Number of time slots that are averaged (in de-mixing)
\newcommand{\nvis}{N_\mathrm{vis}} % Number of visibilities
\newcommand{\nvisrate}{R_{vis}} % Number of visibilities per second\\
% TODO: what kind of qfac?
\newcommand{\qfacbw}{Q_\mathrm{bw}} % Quality factor
\newcommand{\qfacreuse}{Q_\mathrm{fcv}} % Factor to allow for the reuse of functions between neighbouring frequency channels
\newcommand{\qfacfov}{Q_\mathrm{FoV}} % Quality factor defining how much bigger the diameter of the field of view is than the first zero of the Airy function
\newcommand{\facgridovers}{Q_\mathrm{GCF}} % Gridding oversampling factor
\newcommand{\facmsmf}{Q_\mathrm{MSMF}} % Factor to account for multiple subtractions (multi-frequency multi-scale CLEAN)
\newcommand{\qfacbeamovers}{Q_\mathrm{pix}} % Quality factor defining the
                                % oversampling of the synthesised beam
% TODO: what kind of qfac?
\newcommand{\qfac}{Q_t} % Quality factor
\newcommand{\corrfacw}{Q_w} % Correction factor for w values (0 to 1)
\newcommand{\earthrad}{R_\mathrm{Earth}} % Radius of the Earth ($6400\times 10^3$\,m)
\newcommand{\comprate}{R} % Compute rate (FLOPS)
\newcommand{\compratespec}{R^\mathrm{spec}} % Compute rate for spectral line processing
\newcommand{\compratecont}{R^\mathrm{cont}} % Compute rate for continuum imaging
\newcommand{\compratefast}{R^\mathrm{fast}} % Compute rate for fast imaging
\newcommand{\covm}{\mathbb{R}} % Covariance matrix
\newcommand{\covmfilt}{\mathbb{R}_\mathrm{filt}} % Filtered covariance matrix
\newcommand{\bwtoworkmem}{R_\mathrm{bw,mem}} % Bandwidth to working memory
\newcommand{\maxmembw}{R_\mathrm{bw,max}} % Maximum memory bandwidth of
                                % compute unit
\newcommand{\convkernflops}{R_\mathrm{CCF}} % Convolution kernel compute rate (FLOPS)
\newcommand{\fftflops}{R_\mathrm{FFT}} % FFT compute rate (FLOPS)
\newcommand{\flagflops}{R_\mathrm{flag}} % FLOP rate to do flagging on input visibilities
\newcommand{\griddingflops}{R_\mathrm{grid}} % Gridding compute rate (FLOPS)
\newcommand{\phaserotflops}{R_\mathrm{PR}} % Phase rotation compute rate (FLOPS)
\newcommand{\reprojflops}{R_\mathrm{RP}} % Re-projection compute rate (FLOPS)
\newcommand{\rcorr}{R_\mathrm{s}} % Correlator output rate (samples/sec)
\newcommand{\peakflop}{R_\mathrm{peak}} % Peak FLOPS of compute unit
% TODO - verschil met max mem bw of compute unit?
% TODO - verschil memory bw and bw to buffer?
\newcommand{\IO}{{R_\mathrm{IO}}} %IO rate
\newcommand{\maxiobwbuff}{R_\mathrm{bw,I/O,max}} % Maximum I/O bandwidth of compute unit to buffer
\newcommand{\bwbuff}{R_\mathrm{bw,I/O}} % Bandwidth to buffer
\newcommand{\bwbuffcont}{R^\mathrm{cont}_\mathrm{bw,I/O}} % Bandwidth to
                                % buffer for cotinuum imaging
\newcommand{\bwbuffspec}{R^\mathrm{spec}_\mathrm{bw,I/O}} % Bandwidth to
                                % buffer for spectral line processing
\newcommand{\nsource}{N_\mathrm{source}} % Number of sources
\newcommand{\telsensit}{S} % Telescope sensitivity
\newcommand{\corrdumpt}{t_\mathrm{dump}} % Correlator dump time (sec)
\newcommand{\compt}{t_{\mathrm{comp}}} % Computation time
\newcommand{\gridkernupdate}{t_\mathrm{kernel,update}} % Update timescale for gridding kernel
\newcommand{\lsmupdate}{t_\mathrm{LSM}} % Update timescale for the LSM
\newcommand{\obst}{t_{\mathrm{obs}}} % Observation time
\newcommand{\snapshot}{t_\mathrm{snap}} % Snapshot duration
\newcommand{\minsnapshot}{t_\mathrm{snap,min}} % Minimum duration of a snapshot in which w-correction is not dominated by Earth's curvature
\newcommand{\optsnapshot}{t_\mathrm{snap,opt}} % Optimum snapshot duration (sec)
\newcommand{\subint}{t_\mathrm{subint}} % Subintegration time
\newcommand{\uv}{uv} % Visibility coordinates
\newcommand{\uvw}{u,v,w} % Visibility coordinates
\newcommand{\w}{w} % Visibility w coordinate
\newcommand{\nbytepervis}{N_\mathrm{byte,vis}} %Memory per individual visibility 
\newcommand{\maxwdev}{\Delta w_\mathrm{max}} % Maximum deviation of the w
                                % coordinate due to the Earth's curvature
\newcommand{\minwdev}{\Delta w_\mathrm{min}} % Minimum deviation of the w
                                % coordinate due to the Earth's curvature
\newcommand{\spfiltm}{\mathbb{W}_\mathrm{filt}} % Spatial filter matrix
\newcommand{\fres}{\Delta f} % Frequency resolution
\newcommand{\tressmear}{t_\mathrm{smear}} % Smearing Time resolution (sec)
\newcommand{\tres}{t_{res}} % Time resolution (sec)
\newcommand{\fracconvfuncamp}{\epsilon_\omega} % Fraction of maximal amplitude of envelope of convolution function
\newcommand{\compeff}{\eta_\mathrm{comp}} % Computational efficiency
\newcommand{\fracfband}{\eta_f} % Fraction of frequency band
\newcommand{\beamsq}{\theta_\mathrm{bs}} % Beam squint
\newcommand{\fovdiam}{\theta_\mathrm{FoV}} % Angular diameter of the field of view (rad)
\newcommand{\pixres}{\theta_\mathrm{pix}} % Angular resolution of pixel size (rad)
\newcommand{\beamdiam}{\theta_\mathrm{PSF}} % Angular diameter of the synthesised beam (rad)
\newcommand{\wavel}{\lambda} % Wavelength
\newcommand{\maxwavel}{\lambda_\mathrm{max}} % Maximum wavelength
\newcommand{\minwavel}{\lambda_\mathrm{min}} % Minimum wavelength
\newcommand{\maxwavelsubb}{\lambda_{\mathrm{sub,max}}} % Maximum wavelength of an imaging sub-band (channels over which image and pixel size are matched)
\newcommand{\minwavelsubb}{\lambda_\mathrm{sub,min}} % Minimum wavelength of an imaging sub-band (channels over which image and pixel size are matched)
\newcommand{\earthrot}{\Omega_\mathrm{E}} % Angular velocity of the Earth (7.27 10-5 rad/sec)
\newcommand{\epsilonf}{\epsilon_f}  %Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging
% Other command definitions
\newcommand{\Qkernel}{Q_\mathrm{kernel}} %Convolution kernel quality factor, for kernel re-use in frequency
\newcommand{\bs}{\texttt{\char`\\}}

\begin{document}

% load automatic pages
\tccfrontpage

\sdptableofcontents

% Add here the abbreviations used in your document
\sdplistofabbreviations
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
    \item[CDR] Critical Design Review \vspace{-0.2cm}
    \item[CPU] Central Processing Unit\vspace{-0.2cm}
    \item[CSP] Central Signal Processor\vspace{-0.2cm}
    \item[CUFFT] \textcolor{red}{Define it.}\vspace{-0.2cm}
    \item[DFT] \textcolor{red}{Only used once, in a title. Remove/expand?}\vspace{-0.2cm}
    \item[DM] \textcolor{red}{Define it.}\vspace{-0.2cm}
    \item[EoR] Epoch of Reionisation\vspace{-0.2cm}
    \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
    \item[FLOP] Floating Point Operations \vspace{-0.2cm}
    \item[FLOPS] Floating Point Operations per Second \vspace{-0.2cm}
    \item[FoV] Field of View\vspace{-0.2cm}
    \item[GPU] Graphics Processing Unit \vspace{-0.2cm}
    \item[ICD] Interface Control Document\vspace{-0.2cm}
    \item[LMA] Levenberg-Marquardt Algorithm \vspace{-0.2cm}
    \item[LOFAR] LOw Frequency ARray\vspace{-0.2cm}
    \item[LSM] Local Sky Model\vspace{-0.2cm}
    \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[NVRAM] \textcolor{red}{Define it.} \vspace{-0.2cm}
    \item[OCLD] Optimised Candidate List and Data \vspace{-0.2cm}
    \item[PDR] Preliminary Design Review\vspace{-0.2cm}
    \item[RFI] Radio Frequency Interference\vspace{-0.2cm}
    \item[PSF] \textcolor{red}{Define it.} \vspace{-0.2cm}
    \item[PSS] Pulsar Search\vspace{-0.2cm}
    \item[PST] Pulsar Timing\vspace{-0.2cm}
    \item[SDP] Scientific Data Processor\vspace{-0.2cm}
    \item[SKA] Square Kilometre Array\vspace{-0.2cm}
    \item[SKAO] SKA Organisation\vspace{-0.2cm}
    \item[WCS] \textcolor{red}{Define it.} \vspace{-0.2cm}
%    \item[] \vspace{-0.2cm}
\end{basedescript}

\noindent \emph{Reminder:}\\
Computational Cost = Number of FLOP that need to be executed to turn x into y (a.k.a.\ Computational Intensity)\\
Computational Rate = Number of FLOP executed per unit of time -- FLOPS\\
Computational Power = Computational Rate / Efficiency -- FLOPS\\
Multiply and Add = 2 Flops i.e. no Fused multiply and add assumed \\
Complex Multiply and Add = 6 Flops i.e. not o Fused multiply and add assumed\\

% Add here the symbols used in your document
\clearpage \sdplistofsymbols
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
\item[$\bytes$] Length of a complex number in bytes \vspace{-0.2cm}
\item[$\cbytes$] Length of a complex number in bytes \vspace{-0.2cm}
\item[$\rbytes$] Length of a real number in bytes \vspace{-0.2cm}
\item[$\bl$] Baseline length / GPU memory bandwidth \vspace{-0.2cm}
\item[$\maxbl$] Maximum baseline length \vspace{-0.2cm}
\item[$\membw$] Memory bandwidth \vspace{-0.2cm}
\item [$\pciebw$] PCI-e bandwidth \vspace{-0.2cm}
\item[$\viscompr$] Compression factor of visibilities for grdding for facets
  \textcolor{red}{This symbol is not used in the document, consider removing.}
  \vspace{-0.2cm}
\item[$\vism$] Matrix of observed visibilities \vspace{-0.2cm}
\item[$\da$] Extent of aperture \vspace{-0.2cm}
\item[$\ds$] Diameter of antennas/stations \vspace{-0.2cm}
\item[$\reff$] Reference frequency\vspace{-0.2cm}
\item[$\fcc$] Convolution kernel computational cost\vspace{-0.2cm}
\item[$\fci$] Factor between bandwidth and FLOPS (bytes/operations)\vspace{-0.2cm}
\item[$\fgrid$] Gridding cost\vspace{-0.2cm}
\item[$\fpatch$] Fraction of the maximum baseline below which the $\uv$
  coverage is almost filled\vspace{-0.2cm}
\item[$\fpr$] Cost for phase rotation\vspace{-0.2cm}
\item[$\gainm$] 2$\times$2 block diagonal matrix of complex
  gains\vspace{-0.2cm}
\item[$\modskyvism$] Matrix of model sky visibilities\vspace{-0.2cm}
\item[$\skymodbufsize$] Size of the sky model buffer\vspace{-0.2cm}
\item[$\visbufsize$] Size of the visibility buffer \vspace{-0.2cm}
\item[$\visbufsizespec$] Size of the visibility buffer needed for spectral line processing\vspace{-0.2cm}
\item[$\visbufsizecont$] Size of the visibility buffer needed for continuum imaging\vspace{-0.2cm}
\item[$\bufsize$] Size of buffer \vspace{-0.2cm}
\item[$\slowmemsize$] Size of slow working memory \vspace{-0.2cm}
\item[$\cumemsize$] Size of working memory of compute unit \vspace{-0.2cm}
\item[$\imgsize$] Size of image (bytes) \vspace{-0.2cm}
\item[$\imgsizeoneax$] Image size along one axis \vspace{-0.2cm}
\item[$\targgridsize$] Target grid size (per facet) \vspace{-0.2cm}
\item[$\uvgridsize$] Size of the $\uv$ grid (bytes) \vspace{-0.2cm}
\item[$\vissize$] Size of visibility data (bytes) \vspace{-0.2cm}
\item[$\wgridsize$] Size of weight grid (bytes)\vspace{-0.2cm}
\item[$\gridcachesize$] Cache size of gridding kernel \vspace{-0.2cm}
\item[$\nbins$] Number of bins \vspace{-0.2cm}
\item[$\ntel$] Number of antennas or stations in the array
  \vspace{-0.2cm}
\item[$\nateam$] Number of A-team sources \vspace{-0.2cm}
\item[$\ngridpakern$] Number of grid points for A-kernel support (along one
  axis) \vspace{-0.2cm}
\item[$\naccvis$] Number of accesses per visibility \vspace{-0.2cm}
\item[$\navg$] Number of averages\vspace{-0.2cm}
\item[$\nbeam$] Number of beams simultaneously observed by the array
  \vspace{-0.2cm}
\item[$\nbits$] Number of bits\vspace{-0.2cm}
\item[$\rma$] Number of flops in real multiply and add\vspace{-0.2cm}
\item[$\cma$] Number of flops in complex multiply and add\vspace{-0.2cm}
\item[$\nbl$] Number of baselines \vspace{-0.2cm}
\item[$\nbyteacc$] Number of bytes per access \vspace{-0.2cm}
\item[$\nbyte$] Number of bytes \vspace{-0.2cm}
\item[$\nbyteperpix$] Number of bytes per pixel \vspace{-0.2cm}
\item[$\nbytepervis$] Number of bytes per individual visibility

\item[$\nch$] Number of channels \vspace{-0.2cm}
\item[$\ncompu$] Number of compute units \vspace{-0.2cm}
\item[$\ndatait$] Number of data items required per source (e.g. position)\vspace{-0.2cm}
\item[$\ngridp$] Number of grid points along one axis for the gridding
  convolution function \vspace{-0.2cm}
\item[$\nfch$] Number of frequency channels/sub-bands \vspace{-0.2cm}
\item[$\nfchcorr$] Number of frequency channels output by the correlator
  \vspace{-0.2cm}
\item[$\nfchgrid$] Number of frequency channels to be gridded
  \vspace{-0.2cm}
\item[$\nfchdegrid$] Number of frequency channels to be de-gridded
  \vspace{-0.2cm}
\item[$\nfchout$] Number of frequency channels to be FFT-ed and reprojected
  (output) or output from the CSP pulsar timing backend\vspace{-0.2cm}
\item[$\nfchdist$] Minimum number of frequency channels to maintain distributability in the SDP
\item[$\nfkernel$] Number of convolution kernels needed to cover frequency axis for a given baseline
\item[$\nfacet$] Number of facets along one axis \vspace{-0.2cm}
\item[$\nflop$] Number of floating point operations
  \vspace{-0.2cm}
\item[$\nflopsamp$] Number of floating point operations per sample
  \vspace{-0.2cm}
\item[$\nflopvis$] Number of floating point operations per visibility
\item[$\npixgrid$] Number of pixels in the grid (i.e. the square of the linear
  dimension) \vspace{-0.2cm}
\item[$\ngridwkern$] Number of grid points for w-kernel support (along one
  axis) \vspace{-0.2cm}
\item[$\nit$] Number of iterations \vspace{-0.2cm}
\item[$\nawkern$] Linear size of combined A and $\w$-kernels \vspace{-0.2cm}
\item[$\nmaj$] Number of major CLEAN cycles (within each self-cal loop if appropriate) \vspace{-0.2cm}
\item[$\nselfcal$] Number of self-calibration cycles \vspace{-0.2cm}
\item[$\nmin$] Number of minor CLEAN cycles (within each major cycle loop if appropriate)\vspace{-0.2cm}
\item[$\facgridoffdiagmuller$] Factor to account for gridding of off-diagonal
  M\"uller terms \vspace{-0.2cm}
\item[$\nops$] Number of operations \vspace{-0.2cm}
\item[$\npatchpix$] Number of pixels of PSF to be used in minor cycle
  decorrelation \vspace{-0.2cm}
\item[$\npix$] Number of pixels along one axis of a grid\vspace{-0.2cm}
\item[$\npixfacet$] Number of pixels on one side of a facet \vspace{-0.2cm}
\item[$\npol$] Number of polarisation products \vspace{-0.2cm}
\item[$\nsource$] Number of sources \vspace{-0.2cm}
\item[$\ntslot$] Number of time slots \vspace{-0.2cm}
\item[$\ntaylor$] Number of Taylor terms (in MS-MFS)\vspace{-0.2cm}
\item[$\ntslotave$] Number of time slots that are averaged (in de-mixing)
  \vspace{-0.2cm}
\item[$\nvis$] Number of visibilities in an observation \vspace{-0.2cm}
\item[$\qfacbw$] Quality factor \vspace{-0.2cm}
\item[$\qfacreuse$] Factor to allow for the reuse of functions between
  neighbouring frequency channels \vspace{-0.2cm}
\item[$\qfacfov$] Quality factor defining how much bigger the diameter of the
  field of view is than the first zero of the Airy function \vspace{-0.2cm}
\item[$\facgridovers$] Gridding oversampling factor \vspace{-0.2cm}
\item[$\facmsmf$] Factor to account for multiple subtractions (multi-frequency
  multi-scale CLEAN\vspace{-0.2cm}
\item[$\qfacbeamovers$] Quality factor defining the oversampling of the
  synthesised beam \vspace{-0.2cm}
\item[$\qfac$] Quality factor \vspace{-0.2cm}
\item[$\Qkernel$] Convolution kernel quality factor, for kernel re-use in frequency
\item[$\corrfacw$] Correction factor for w values (0 to 1) \vspace{-0.2cm}
\item[$\earthrad$] Radius of the Earth ($6400\times 10^3$\,m) \vspace{-0.2cm}
\item[$\covm$] Covariance matrix \vspace{-0.2cm}
\item[$\covmfilt$] Filtered covariance matrix \vspace{-0.2cm}
\item[$\comprate$] Compute rate (FLOPS) \vspace{-0.2cm}
\item[$\compratespec$] Compute rate (FLOPS) for spectral line processing \vspace{-0.2cm}
\item[$\compratecont$] Compute rate (FLOPS) for continuum imaging \vspace{-0.2cm}
\item[$\compratefast$] Compute rate (FLOPS) for fast imaging \vspace{-0.2cm}
\item[$\bwtoworkmem$] Bandwidth to working memory \vspace{-0.2cm}
\item[$\maxmembw$] Maximum memory bandwidth of compute unit \vspace{-0.2cm}
\item[$\peakflop$] Peak FLOPS of compute unit \vspace{-0.2cm}
\item[$\maxiobwbuff$] Maximum I/O bandwidth of compute unit to buffer
  \vspace{-0.2cm}
\item[$\bwbuff$] Bandwidth to buffer \vspace{-0.2cm}
\item[$\bwbuffcont$] Bandwidth to buffer for continuum imaging\vspace{-0.2cm}
\item[$\bwbuffspec$] Bandwidth to buffer for spectral line processing\vspace{-0.2cm}
\item[$\convkernflops$] Convolution kernel compute rate (FLOPS)\vspace{-0.2cm}
\item[$\fftflops$] FFT compute rate (FLOPS)\vspace{-0.2cm}
\item[$\griddingflops$] Gridding compute rate (FLOPS)\vspace{-0.2cm}
\item[$\phaserotflops$] Phase rotation compute rate (FLOPS)\vspace{-0.2cm}
\item[$\reprojflops$] Re-projection compute rate (FLOPS)\vspace{-0.2cm}
\item[$\rcorr$] Correlator output rate\vspace{-0.2cm} TRY TO REMOVE
\item[$\facetoverlap$]  ratio for linear size increase of facets to account for overlap, typically 1.2-1.5 \vspace{-0.2cm}
\item[$\telsensit$] Telescope sensitivity \vspace{-0.2cm} TRY TO REMOVE
\item[$\corrdumpt$] Correlator dump time (sec) \vspace{-0.2cm}
\item[$\compt$] Computation time \vspace{-0.2cm}
\item[$\gridkernupdate$] Update timescale for gridding kernel \vspace{-0.2cm}
\item[$\lsmupdate$] Update timescale for the LSM \vspace{-0.2cm}
\item[$\obst$] Observation time \vspace{-0.2cm}
\item[$\snapshot$] Snapshot duration \vspace{-0.2cm}
\item[$\minsnapshot$] Minimum duration of a snapshot in which w-correction is
  not dominated by the Earth's curvature \vspace{-0.2cm}
\item[$\optsnapshot$] Optimum snapshot duration (sec) \vspace{-0.2cm}
\item[$\subint$] Sub-integration time\vspace{-0.2cm}
\item[$\uvw$] Visibility coordinates \vspace{-0.2cm}
\item [$\w$] Visibility $w$ coordinate \vspace{-0.2cm}
\item[$\maxwdev$] Maximum deviation of the $w$ coordinate due to the Earth's
  curvature \vspace{-0.2cm}
\item[$\minwdev$] Minimum deviation of the $w$ coordinate due to the Earth's
  curvature \vspace{-0.2cm}
\item[$\spfiltm$] Spatial filter matrix \vspace{-0.2cm}
\item[$\fres$] Frequency resolution \vspace{-0.2cm}
\item[$\tres$] Time resolution (sec) \vspace{-0.2cm}
\item[$\tressmear$]  Smearing Time resolution (sec)  \vspace{-0.2cm}
\item[$\fracconvfuncamp$] Fraction of maximal amplitude of envelope of
  convolution function \vspace{-0.2cm}
\item[$\compeff$] Computational efficiency \vspace{-0.2cm}
\item[$\fracfband$] Fraction of frequency band \vspace{-0.2cm}
\item[$\beamsq$] Beam squint \vspace{-0.2cm}
\item[$\fovdiam$] Angular diameter of the field of view (rad) \vspace{-0.2cm}
\item[$\pixres$] Angular resolution of pixel size (rad) \vspace{-0.2cm}
\item[$\beamdiam$] Angular diameter of the synthesised beam (rad)
  \vspace{-0.2cm}
\item[$\wavel$] Wavelength \vspace{-0.2cm}
\item[$\maxwavel$] Maximum wavelength \vspace{-0.2cm}
\item[$\minwavel$] Minimum wavelength \vspace{-0.2cm}
\item[$\maxwavelsubb$] Maximum wavelength of an imaging sub-band (channels
  over which image and pixel size are matched) \vspace{-0.2cm}
\item[$\minwavelsubb$] Minimum wavelength of an imaging sub-band (channels
  over which image and pixel size are matched) \vspace{-0.2cm}
\item[$\earthrot$] Angular velocity of the Earth
  ($7.27\times10^{-5}$\,rad/sec) \vspace{-0.2cm}
\item[$\epsilonf$] Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging \vspace{-0.2cm}
\end{basedescript}

\sdplistoffigures

\sdplistoftables

% Add here the executive summary
\sdpsummary

\newpage

% Add to the table the list of applicable and reference documents (this is NOT
% meant for your usual bibiliography, only for SKA docs)
\sdpreferencedocs

\subsection*{Applicable Documents}

The following documents are applicable to the extent stated herein. In the
event of conflict between the contents of the applicable documents and this
document, \emph{the applicable documents} shall take precedence.

 \begin{center}{
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
     {\bf [AD01]} & Dewdney, P. E. (2013). SKA1 System Baseline Design. SKA Office\\
     {\bf [AD02]} & McCool, R., Cornwell, T. (2013). Miscellaneous Corrections
     to the Baseline Design\\
     {\bf [AD03]} & SKA Phase 1 System (Level 1) Requirements Specification, T.Cornwell, SKA-OFF.SE.ARC-SKO-SRS-001\_2\\
     {\bf [AD04]} & PDR.01 SDP.ARCH document\\
     {\bf [AD05]} & SDP Costing spreadsheet\\
     {\bf [AD06]} & Cornwell, T.J. (2015). SKA1 Telescope Calibration
     Framework. SKA Office (draft version)\\
     {\bf [AD07]} & CSP--SDP ICD\\
     \hline
   \end{tabularx}}
\end{center}

\subsection*{Reference Documents}

The following documents are referenced in this document. In the event of
conflict between the contents of the referenced documents and this document,
\emph{this document} shall take precedence.

 \begin{center}{
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
     {\bf [RD01]} & SKA-TEL-SDP-0000041, F. Malan: iPython Performance Model Description\\
    {\bf [RD02]} & SKA-TEL-SDP-0000027, R. Nijboer: Pipelines Element Subsystem Design\\
    {\bf [RD03]} & SKA-TEL-SDP-0000038, R. Bolton: High Priority Science Objectives: Performance Analysis\\
    {\bf [RD04]} & SKA-TEL-SDP-0000028, A.-J. Boonstra: Ingest Pipeline\\
    {\bf [RD05]} & SKA-TEL-SDP-0000029, S. Salvini: Pipelines: Calibration\\
    {\bf [RD06]} & SKA-TEL-SDP-0000030, A. Scaife: Imaging Pipeline\\
    {\bf [RD07]} & SKA-TEL-SDP-0000031, M. Johnston-Hollitt: Science Analysis Pipeline\\
    {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent
    Averaging\\
    {\bf [RD09]} & SKA-TEL-SDP-0000058, S. Salvini: Fast Fourier Transforms\\
    {\bf [RD10]} & SKA-TEL-SDP-0000057, C. Skipper: Time and Channel Averaging\\
    {\bf [RD11]} & Alexander et al., Software and Computing CoDR, Analysis of requirements derived from the DRM, WP2-050.020.010-RR-001\\
    {\bf [RD12]} & SKA-TEL-SDP-0000016, R. Bolton: SDP Archive Size
    Estimates\\
    {\bf [RD13]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
    {\bf [RD14]} & \url{https://confluence.ska-sdp.org/pages/viewpage.action?title=MS+in+UML&spaceKey=ARCH}\\
    {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821,
    \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
    {\bf [RD16]} & \url{http://www.lofar.org/operations/lib/exe/fetch.php?media=:public:documents:casa_image_for_lofar_0.03.00.pdf}\\
     \hline
   \end{tabularx}}
\end{center}

 \begin{center}{
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
    {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis
    imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
    {\bf [RD18]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
    {\bf [RD19]} & Cornwell, T. J., Voronkov, M. A., Humphreys, B. (2012). Wide field imaging for the Square Kilometre Array.\\
    {\bf [RD20]} & An Efficient Work-Distribution Strategy for Gridding Radio-Telescope Data on GPUs, John W. Romein, ICS’12, June 25–29, 2012, San Servolo Island, Venice, Italy. Source code available at: \url{ftp://ftp.astron.nl/outgoing/romein/Gridding-0.2.tar.bz}\\
   {\bf [RD21]} & Mitchell, D. a. (2014). Analysis of w-projection kernel size. SKA SDP Consortium\\
   {\bf [RD22]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
   {\bf [RD23]} & A parametric model for the calibration and imaging costs of SKA1, T.J. Cornwell, 12 Feb 2013, SKA Draft Memo\\
     {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
   {\bf [RD25]} & Tasse, C. a. (2013). Applying full polarization A-Projection to very wide field of view instruments: An imager for LOFAR. A\&A.\\
   {\bf [RD26]} & Faceted imaging in AIPS, Kogan and Greisen, AIPS Memo 113, 2009\\
   {\bf [RD27]} & Radio-interferometric imaging of very large fields - The
   problem of non-coplanar arrays, Cornwell, T. J.; Perley, R. A., Astronomy
   and Astrophysics (ISSN 0004-6361), vol. 261, no. 1, p. 353-364\\
    {\bf [RD28]} & Wide Field Imaging at Low Frequencies, Sault, R. J.,  The Universe at Low Radio Frequencies, Proceedings of IAU Symposium 199, held 30 Nov - 4 Dec 1999, Pune, India. Edited by A. Pramesh Rao, G. Swarup, and Gopal-Krishna, 2002., p.508\\
    {\bf [RD29]} & Roofline: An Insightful Visual Performance Model for
    Multicore Architectures, Williams, Waterman and Patterson,
    DOI:10.1145/1498765.1498785, April 2009, Vol. 52, No. 4, Communications of
    the ACM\\
    {\bf [RD30]} & Whiting, M., 2012, Duchamp: a 3D source finder for spectral-line data, Mon. Not. R. Astron. Soc., Volume 421, Issue 4,  pages 3242–3256, April 2012 \\
    {\bf [RD31]} & Hales, C. A., Murphy, T., Curran, J. R., Middelberg, E., Gaensler, B. M., Norris, R. P., blobcat: software to catalogue flood-filled blobs in radio images of total intensity and linear polarization, Mon. Not. R. Astron. Soc., Volume 425,  Issue 2, pages 1365-2966, September 2012\\
    {\bf [RD32]} & Hancock, P.J., Murphy, T., Gaensler, B.M., Hopkins, A.,
    Curran, J.R., (2012), Compact continnum source finding for next generation
    radio surveys, Mon. Not. R. Astron. Soc. 422, 1812–1824\\
   {\bf [RD33]} & Whiting, M.,  Humphreys, B., (2012) Source-Finding for the
   Australian Square Kilometre Array Pathfinder, Publications of the
   Astronomical Society of Australia, Volume 29, Special Issue 03, 2012,
   pp 371-381.\\
   {\bf [RD34]} Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
   {\bf [RD35]} Conway, J. E., Cornwell, T. J., & Wilkinson, P. N. (1990). Multi-Frequency Synthesis - a New Technique in Radio Interferometric Imaging. Monthly Notices of the Royal Astronomical Society, 246, 490.\\

   \hline
 \end{tabularx}}
\end{center}

% The actual content goes here
\newpage
\section{Introduction}

\subsection{Purpose of the document}

The purpose of this document is to describe the Multi-Scale MultiFrequency Synthesis algorithm, the variant implementations in CASA and ASKAPsoft, and the scaling for insertion in the SDP Performance Model.

\subsection{Scope of the document}

\clearpage
\section{Overview of MSMFS}
\label{sec:overview}

MS-MFS was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g. [RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data. The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law.

\clearpage
\section{Major cycle}
\label{sec:major}
\subsection{Computational task}

\subsection{ASKAP implementation}

\subsection{CASA implementation}

\clearpage
\section{Minor cycle}
\label{sec:minor}

\subsection{Computational task}

\subsection{ASKAP implementation}

\subsection{CASA implementation}

\clearpage
\section{A projection}
\label{sec:minor}

\clearpage
section{Wide-band behaviour}
\label{sec:wideband}

\begin{itemize}
\item Does naive broadband work?
\item Bhatnagar et al WB algorithm	
\end{itemize}

\clearpage
\section{Implementation in SDP}
\label{sec:implementation}

\clearpage
\section{Resource usage}
\label{sec:resource}

% \label{section:main}

% Add the bibliography
\clearpage \addcontentsline{toc}{section}{References}
\bibliography{your_bibfile}%

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-04-19 msmfs.tex
Merging in Urvashi's equations

@@
 \usepackage{amssymb,amsmath} % for the math symbols
+\usepackage{mdwlist}
 \usepackage{geometry}
@@
 \usepackage{mathtools}
+\usepackage{hhtensor}
 \graphicspath{ {images/} }
+
+\usepackage[boxruled,linesnumbered]{algorithm2e}
+\usepackage{fmtcount}
 
@@
 % Lead author (goes in the footer)
-\newcommand{\leadauthor}{T.J.\ Cornwel\\xspace}
+\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
 % Release date (YYYY-MM-DD)
@@
 
-\noindent \emph{Reminder:}\\
-Computational Cost = Number of FLOP that need to be executed to turn x into y (a.k.a.\ Computational Intensity)\\
-Computational Rate = Number of FLOP executed per unit of time -- FLOPS\\
-Computational Power = Computational Rate / Efficiency -- FLOPS\\
-Multiply and Add = 2 Flops i.e. no Fused multiply and add assumed \\
-Complex Multiply and Add = 6 Flops i.e. not o Fused multiply and add assumed\\
+\newcommand{\nuno}{{\left(\frac{\nu}{\nu_0}\right)}}
+\newcommand{\dnuno}{{\left(\frac{\nu-\nu_0}{\nu_0}\right)}}
+
+\newcommand{\dg}{^\dag}
+\newcommand{\X}{\vec{x}}
+\newcommand{\Xd}{\vec{{x}^\dag}}
+\newcommand{\B}{\vec{b}}
+\newcommand{\Bd}{\vec{b^\dag}}
+\newcommand{\A}{{\tens{A}}}
+\newcommand{\Ad}{{\tens{A^\dag}}}
+\newcommand{\F}{{\tens{F}}}
+\newcommand{\Fd}{{\tens{F^\dag}}}
+\newcommand{\He}{{\tens{H}}}
+\newcommand{\Sa}{{\tens{S}}}
+\newcommand{\Sd}{{\tens{S^\dag}}}
+\newcommand{\Sna}{\tens{{S_{\nu}}}}
+\newcommand{\Snd}{\tens{{S_{\nu}^\dag}}}
+\newcommand{\T}{{\tens{T}}}
+\newcommand{\W}{{\tens{W}}}
+\newcommand{\Wd}{{\tens{W^\dag}}}
+\newcommand{\Pb}{{P_b}}
+
+\newcommand{\Wim}{{\tens{W^{im}}}}
+\newcommand{\Wimd}{{\tens{{W^{im}}^\dag}}}
+\newcommand{\Wnt}{{\tens{W^{mfs}_t}}}
+\newcommand{\Wntd}{{\tens{{W^{mfs}_t}^\dag}}}
+\newcommand{\Wnp}{{\tens{W^{mfs}_p}}}
+\newcommand{\Wnpd}{{\tens{{W^{mfs}_p}^\dag}}}
+\newcommand{\Wnq}{{\tens{W^{mfs}_q}}}
+\newcommand{\Wnqd}{{\tens{{W^{mfs}_q}^\dag}}}
+\newcommand{\Wimn}{{\tens{W^{im}_{\nu}}}}
+\newcommand{\Wimnd}{{\tens{{W^{im}_{\nu}}^\dag}}}
+
+\newcommand{\wnt}{{w_{\nu}^t}}
+\newcommand{\wnq}{{w_{\nu}^q}}
+\newcommand{\wntq}{{w_{\nu}^{t+q}}}
+\newcommand{\Wntn}{{\tens{w^{mfs}_{t,\nu}}}}
+\newcommand{\Wntnd}{{\tens{{w^{mfs}_{t,\nu}}^\dag}}}
+\newcommand{\Wnpn}{{\tens{W^{mfs}_{p,\nu}}}}
+\newcommand{\Wnpnd}{{\tens{{W^{mfs}_{p,\nu}}^\dag}}}
+
+\newcommand{\pd}{{\partial}}
+\newcommand{\mi}{{m_{I}}}
+\newcommand{\R}{{R}}
+\newcommand{\Rd}{{R^\dag}}
+\newcommand{\I}{{\vec{I}}}
 
@@
     {\bf [RD16]} & \url{http://www.lofar.org/operations/lib/exe/fetch.php?media=:public:documents:casa_image_for_lofar_0.03.00.pdf}\\
-     \hline
-   \end{tabularx}}
-\end{center}
-
- \begin{center}{
- \begin{tabularx}{\textwidth}{|l|X|}
-     \hline
-     \bf{Reference} & \bf{Reference}\\
-     \bf{Number} & \\
-     \hline
     {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis
@@
      {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
+     \hline
+   \end{tabularx}}
+\end{center}
+
+ \begin{center}{
+ \begin{tabularx}{\textwidth}{|l|X|}
+     \hline
+     \bf{Reference} & \bf{Reference}\\
+     \bf{Number} & \\
+     \hline
    {\bf [RD25]} & Tasse, C. a. (2013). Applying full polarization A-Projection to very wide field of view instruments: An imager for LOFAR. A\&A.\\
@@
 [RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data.
-The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law.
+The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law expanded into a Taylor series.
+
+The algorithm is an elaboration of the Multiscale CLEAN. It consists of two parts:
+\begin{description}
+\item[Major cycle] The image residuals for the current model are calculated by Fourier transform of the images constituting the power law expansion and degridding to obtain the predicted visibility, subtraction from the observed visibility, then gridding, and Fourier transformation to the image plane.
+There are two ways of performing this cycle depending on whether the Taylor series is calculated in image or uv space.
+Typically 5 - 10 major cycles are required.
+\item[Minor cycle] The minor cycle is a greedy algorithm which identifies the dominant candidate scale and removes that using the appropriate PSF centered on that component.
+The minor cycle repeats this until convergence.
+Typically hundreds or thousands of iterations are required.
+\end{description}
+
+A canonical major/minor cycle algorithm is shown in Figure \ref{fig:major/minor}.
+The two functions Predict and MakeImage are sub-pipelines.
+ASKAPSoft and CASA both follow this form but with different ways of performing Predict (\ref{fig:predictuv} and \ref{fig:predictimage}) and MakeImage (\ref{fig:makeimageuv} and \ref{fig:makeimageimage})
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_MajorMinor.pdf}
+  \caption{Structure of Major/Minor cycle algorithm}
+  \label{fig:majorminor}
+\end{figure}
+
+The major cycle returns to the visibility data each iteration, whereas the minor cycle is entirely image-based.
+
+MSMFS is relatively high in complexity because the usual CLEAN process is coupled over both scale (MSClean) and frequency or Taylor terms (MFS).
+The minor cycle must conduct a search in scale and Taylor term for each peak search.
+The major cycle requires calculation of the residual Taylor terms, which implies gridding and degridding all relevant frequency information.
+Memory use is high since a large number of images must be kept and updated as the iteration proceeds.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_Predict_UV.pdf}
+  \caption{Structure of UV-weighting Predict, as used in ASKAPsoft. }
+  \label{fig:predictuv}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_UV.pdf}
+  \caption{Structure of UV-weighting MakeImage, as used in ASKAPsoft. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
+  \label{fig:predictimage}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_Predict_Image.pdf}
+  \caption{Structure of Image plane-weighting Predict, as used in CASA. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
+  \label{fig:makeimageuv}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_Image.pdf}
+  \caption{Structure of Image plane-weighting MakeImage, as used in CASA. Note that the three loops (denoted by the rectangle with curved ends) can be permuted at will.}
+  \label{fig:makeimageimage}
+\end{figure}
 
 \clearpage
-\section{Major cycle}
-\label{sec:major}
-\subsection{Computational task}
+\section{Serial}
+\label{sec:serial}
 
-\subsection{ASKAP implementation}
+The serial algorithm is as shown in Algorithm \ref{alg:MSMFS_1}.
 
-\subsection{CASA implementation}
+\begin{equation}
+\vec{I}^{model} = \sum_{s=0}^{N_s-1}  \vec{I}^{shp}_{s} \star \vec{I}^{sky,\delta}_s
+\label{Eq:ms_model}
+\end{equation}
+
+where $N_s$ is the number of spatial scales used to construct the image, and
+$\vec{I}^{sky,\delta}_{s}$ represents a collection of $\delta$-functions that describe the locations
+and integrated amplitudes of flux components of scale $s$ in the image. $\vec{I}^{shp}_s$ is a tapered truncated parabola of width proportional to $s$.
+The symbol $\star$ denotes convolution.
+
+\begin{eqnarray}
+\label{eqn:mf_model}
+\vec{I}^{model}_{\nu} = \sum_{t=0}^{N_t-1} \wnt \vec{I}^{sky}_{t} ~~~\mathrm{where}~~~ \wnt&=&\dnuno^t 
+\end{eqnarray}
+
+where $N_t$ is the order of the Taylor series expansion, and 
+the $I^m_t$ represent multi-scale Taylor coefficient images.
+
+The image flux model at each frequency can be written as a linear sum of  
+coefficient images at different spatial scales. 
+\begin{equation}
+\vec{I}^{model}_{\nu} = \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} \wnt \left[ \vec{I}^{shp}_s \star \vec{I}^{sky}_\frac{s}{t}\right] ~~~~\mathrm{where}~~~\wnt = \dnuno^t 
+\label{eqn:msmf_model}
+\end{equation}
+
+Here, $N_s$ is the number of discrete spatial scales used to represent the image and  
+$N_t$ is the order of the series expansion of the spectrum. $\vec{I}^{sky}_\frac{s}{t}$ represents 
+a collection of $\delta$-functions that describe the locations
+and integrated amplitudes of flux components of scale $s$ in the image of the $t^{th}$ series 
+coefficient.
+
+In words, the MSMFS algorithm decomposes the image into a set of blobs, using a greedy algorithm in which the peak (in scale space) is found by searching for the peak in the reference frequency residual images each convolved with scale.
+For bright pixels, the principal solution is calculated by applying the inverse Hessian of the PSF convolved with scale.
+For faint pixels, just the peak in the reference frequency image is used.
 
 \clearpage
-\section{Minor cycle}
-\label{sec:minor}
 
-\subsection{Computational task}
+\begin{algorithm}[t!]
+\scriptsize
+\label{alg:MSMFS_1}
+%  \SetKwRepeat{Repeat}{Repeat}{Until}
+%  \SetKwFor{ForEach}{ForEach}{Do}{End}
+%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{shp}_s~\forall~s\in\{0,N_s$-$1\}, ~ [\Sna]~ \forall \nu$ }
+%\KwResult{ $\vec{I}^{psf}_{{sp}\atop{tq}},[{H^{peak}_s}], \vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ }
+  \KwData{calibrated visibilities : $\vec{V}^{corr}_{\nu}~~\forall \nu$}
+  \KwData{$uv$-sampling function : $\Sna$}
+  \KwData{image noise threshold and loop gain $\sigma_{thr}, g_s$}
+  \KwData{scale basis functions : $\I^{shp}_s ~ \forall s\in\{0,N_s-1\}$}
+  \KwResult{model coefficient images : $\I^{m}_{q}~ \forall q\in\{0,N_t-1\}$}
+  \KwResult{spectral index and curvature : $\I^{m}_{\alpha},\I^{m}_{\beta}$}
 
-\subsection{ASKAP implementation}
+  \vspace{0.5cm} 
+%{Compute the dirty image $\vec{I}^{dirty}$ and psf $\vec{I}^{psf}$}\;
+  \For{ $t \in \{0,N_t-1\}, q \in \{t,N_t-1\}$}
+  {
+        { Compute the spectral PSF } $\vec{I}^{psf}_{tq}$\;
+        \For{ $s \in \{0,N_s-1\}, p \in \{s,N_s-1\}$}
+	{
+		{Compute the scale-spectral PSF} $\vec{I}^{psf}_{{sp}\atop{tq}} = \vec{I}^{shp}_s \star \vec{I}^{shp}_p \star \vec{I}^{psf}_{tq} $\;
+	}
+  }
+  \For { $s \in \{0,N_s-1\}$}
+  {
+     Construct $[{H^{peak}_s}]$ from $mid(I^{psf}_{{s,s}\atop{t,q}})$ and compute $[{H^{peak}_s}^{-1}]$\;
+  }
+%%%%%%%%%%%%%%%%%%%%%%
+%%%\vspace{0.5cm}
+%%  \caption[MS-MFS CLEAN : Pre-Deconvolution Setup]
+%%          {MS-MFS CLEAN : Pre-Deconvolution Setup}
+%%\end{algorithm}
+%%%\newpage
+%%
+%%\begin{algorithm}[H]
+%%  \SetLine
+%%  \linesnumbered
+%%  \dontprintsemicolon
+%%%%%%%%%%%%%%%%%%%%%%%  
+%%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{shp}_s, \vec{I}^{psf}_{{sp}\atop{tq}},[{H^{peak}_s}]~~~\forall~~s \in \{0,N_s-1\}, p \in \{s,N_s-1\}, \sigma_{thr}$ }
+%%\KwResult{$I^{m}_{q}~ \forall q \in \{0,N_t-1\}$}
+%%%%%%%%%%%%%%%%%%%%%  
+  Initialize the model $\vec{I}^{m}_t$ for all $t \in \{0,N_t-1\}$ and compute $f_{sidelobe}$ \;
+  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{res}_0 < \sigma_{thr}$ }
+  {
+    \For{$t \in \{0,N_t$-$1\}$}
+    {
+      Compute the residual image $\vec{I}^{res}_t$ \tcc*[r]{Makeimage}
+      \For{$s \in \{0,N_s$-$1\}$}
+      {
+	      Compute $\vec{I}^{res}_{{s},{t}} = \vec{I}^{shp}_s \star \vec{I}^{res}_t$
+      }
+    }
+    Calculate $f_{limit}$ from $\vec{I}^{res}_{0,0}$\;
+    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{res}_{0,0} < f_{limit} $ } 
+    {
+%     Compute $I^{m}_q~\forall q\in \{0.N_t-1\}$ and update $\vec{I}^{res}_{{s},{t}}~\forall s,t$ (Algorithm \vref{ALGO:MSMFS_2})\;
+     \For{$s \in \{0,N_s$-$1\}$} 
+     {
+      \uIf{Peak of $\vec{I}^{res}_{s,0} > 10~\sigma_{thr} $}
+      {
+       \ForEach{pixel}
+       {
+          Construct $I^{rhs}_s$, an $N_t\times 1$ vector from $I^{res}_{s,t} ~~\forall ~ t \in \{0,N_t$-$1\}$\;
+          Compute principal solution $I^{sol}_s = [{H^{peak}_s}^{-1}] I^{rhs}_s$\;
+       }
+       Choose $I^{sol} = max\{I^{sol}_{t=0},~\forall~s\in\{0,N_s$-$1\}\}$ \;
+      }
+      \Else
+      {
+       Find the location of the peak in $\vec{I}^{res}_{s,0},~\forall~s\in\{0,N_s$-$1\}$\;
+       Construct $I^{rhs}_s$, from $I^{res}_{s,t}$ for the chosen $s$, at this location\;
+       Compute $I^{sol} = [{H^{peak}_s}^{-1}] I^{rhs}_s$ at this location\;
+      }
+     }
+       \For{$t \in \{0,N_t-1\}$}
+       {
+        Update the model image : $I^{m}_t = I^{m}_t + g_s ~ I^{shp}_{s_i} \star I^{sol}_t$ \;
+        \For{$s \in \{0,N_s$-$1\}$}
+	{
+          Update the residual image : $I^{res}_{s,t} = I^{res}_{s,t} - g ~\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1}[I^{psf}_{{sp}\atop{tq}} \star I^{sol}_q]$\;
+	}
+       }
+    }
+   Compute model visibilities $V^{m}_{\nu}$ from  $I^{m}_t~\forall t\in \{0.N_t-1\}$\tcc*[r]{Predict} 
+   Compute a new residual image $I^{res}$ from residual visibilities $V^{corr}_{\nu}-V^{m}_{\nu}$ \tcc*[r]{Makeimage}
+}
+Calculate $\vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ from $I^{m}_t~\forall t\in \{0.N_t-1\}$ and restore the results \;
+If required, remove average primary beam : $\vec{I}^{new}_{\nu_0}=\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}, \vec{I}^{new}_{\alpha}=\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}, \vec{I}^{new}_{\beta}=\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}$\;
+\vspace{0.5cm}
 
-\subsection{CASA implementation}
+  \caption[MS-MFS Algorithm]
+          {MS-MFS Algorithm : }
+\end{algorithm}
+
+Notes:
+\begin{description}
+\item[Line 13]	
+\end{description}
+
+\clearpage
+\section{Partitioning}
+\label{sec:partitioning}
+
+The SDP pipeline framework [RD:??], allows partitioning across multiple axes. In this case, the natural partitions are:
+
+\begin{description}
+\item[Frequency]
+\item[Time]
+\item[Facet]
+\item[Scales]	
+\end{description}
 
@@
 \clearpage
-section{Wide-band behaviour}
+\section{Wide-band behaviour}
 \label{sec:wideband}
@@
 \clearpage
-\section{Implementation in SDP}
-\label{sec:implementation}
+\section{Modelling MSMFS in the SDP Performance Manager}
+\label{sec:modelling}
 

### 2016-04-19 aa/msmfs_aa.tex
More files needed for urvashi's equations

%                                                                 aa.dem
% AA vers. 6.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                 (c) Springer-Verlag HD
%                                                revised by EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%
\documentclass[structabstract]{stylefiles/aa}  
%\documentclass[traditabstract]{aa} % for the abstract without structuration 
                                   % (traditional abstract)

\usepackage[boxruled]{stylefiles/algorithm2e}
\usepackage[dvips]{graphicx, color}
\usepackage{epsfig}
\usepackage{mdwlist}
%\usepackage{amsmath}
%\usepackage{mathptmx}
\usepackage{txfonts}
\setcounter{secnumdepth}{3}

\usepackage{stylefiles/fmtcount}
\usepackage{natbib}
\bibpunct{[}{]}{;}{a}{}{;}
\renewcommand{\cite}{\citep}

\usepackage{varioref}

\newcommand{\nuno}{{\left(\frac{\nu}{\nu_0}\right)}}
\newcommand{\dnuno}{{\left(\frac{\nu-\nu_0}{\nu_0}\right)}}

\newcommand{\dg}{^\dag}
\newcommand{\X}{\vec{x}}
\newcommand{\Xd}{\vec{{x}^\dag}}
\newcommand{\B}{\vec{b}}
\newcommand{\Bd}{\vec{b^\dag}}
\newcommand{\A}{{\tens{A}}}
\newcommand{\Ad}{{\tens{A^\dag}}}
\newcommand{\F}{{\tens{F}}}
\newcommand{\Fd}{{\tens{F^\dag}}}
\newcommand{\He}{{\tens{H}}}
\newcommand{\Sa}{{\tens{S}}}
\newcommand{\Sd}{{\tens{S^\dag}}}
\newcommand{\Sna}{\tens{{S_{\nu}}}}
\newcommand{\Snd}{\tens{{S_{\nu}^\dag}}}
\newcommand{\T}{{\tens{T}}}
\newcommand{\W}{{\tens{W}}}
\newcommand{\Wd}{{\tens{W^\dag}}}
\newcommand{\Pb}{{P_b}}

\newcommand{\Wim}{{\tens{W^{im}}}}
\newcommand{\Wimd}{{\tens{{W^{im}}^\dag}}}
\newcommand{\Wnt}{{\tens{W^{mfs}_t}}}
\newcommand{\Wntd}{{\tens{{W^{mfs}_t}^\dag}}}
\newcommand{\Wnp}{{\tens{W^{mfs}_p}}}
\newcommand{\Wnpd}{{\tens{{W^{mfs}_p}^\dag}}}
\newcommand{\Wnq}{{\tens{W^{mfs}_q}}}
\newcommand{\Wnqd}{{\tens{{W^{mfs}_q}^\dag}}}
\newcommand{\Wimn}{{\tens{W^{im}_{\nu}}}}
\newcommand{\Wimnd}{{\tens{{W^{im}_{\nu}}^\dag}}}

\newcommand{\wnt}{{w_{\nu}^t}}
\newcommand{\wnq}{{w_{\nu}^q}}
\newcommand{\wntq}{{w_{\nu}^{t+q}}}
\newcommand{\Wntn}{{\tens{w^{mfs}_{t,\nu}}}}
\newcommand{\Wntnd}{{\tens{{w^{mfs}_{t,\nu}}^\dag}}}
\newcommand{\Wnpn}{{\tens{W^{mfs}_{p,\nu}}}}
\newcommand{\Wnpnd}{{\tens{{W^{mfs}_{p,\nu}}^\dag}}}

\newcommand{\pd}{{\partial}}
\newcommand{\mi}{{m_{I}}}
\newcommand{\R}{{R}}
\newcommand{\Rd}{{R^\dag}}
\newcommand{\I}{{\vec{I}}}

%\usepackage{amsmath}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{document}
%\tableofcontents
%\newpage
%
   \title{A multi-scale multi-frequency deconvolution algorithm
for synthesis imaging in radio interferometry}

%   \subtitle{I. Overviewing the $\kappa$-mechanism}

   \author{U. Rau
          \inst{1}
          \and
          T.J. Cornwell \inst{2} %\fnmsep \thanks{Just to show the usage of the elements in the author field}
          }

   \institute{ National Radio Astronomy Observatory, Socorro, NM, USA\\
              \email{rurvashi@aoc.nrao.edu}
         \and
             Australia Telescope National Facility, CSIRO, Sydney, AU \\
             \email{tim.cornwell@atnf.csiro.au}
%             \thanks{blah.}
             }

   \date{Received XXX XX, 2009; accepted XXX XX, 2009}
   %%\date{Received December 15, 2008; accepted XXX XX, 2009}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {The use of broad-band receivers in radio interferometery impacts standard continuum
    synthesis imaging in three ways. 
    First, correlations measured at different frequencies sample the visibility function of the
   sky brightness at different spatial frequencies.
    Second, the sky brightness can change significantly across the large frequency 
    range that these receivers are sensitive to.
    Third, direction-dependent instrumental effects such as the element response pattern
     also vary with frequency, and this 
    affects the spectral characteristics of the measured sky brightness during wide-field
    continuum imaging.
    Existing algorithms designed to account for some of the above, can be shown to be
    insufficient to achieve target image fidelity and dynamic range over the wide bands that 
    most new instruments are being built with.}
  % aims heading (mandatory)
   {This paper describes a multi-scale multi-frequency deconvolution algorithm for
    the minor-cycle of 
synthesis imaging that combines the information measured about the source spectrum
    with the additional spatial-frequency-coverage offered by multi-frequency-synthesis to 
    reconstruct source structure.} %, while accounting for frequency-dependent antenna primary-beams.}
  % methods heading (mandatory)
   {
The MS-MFS (multi-scale multi-frequency synthesis) algorithm discussed in this
paper extends the MF-Clean algorithm to reconstruct spectral curvature in 
addition to spectral index and combines it with a multi-scale deconvolution approach
similar to MS-Clean.
%The MS-MFS algorithm discussed in this paper models the wide-band sky-brightness
%     distribution as a linear combination of spatial and spectral basis functions, 
%      and combines a linear-least squares approach along with iterative chi-sq minimization.
%    This method can be used in conjunction with existing wide-field imaging algorithms.
   We also discuss a simpler hybrid of spectral-line and continuum imaging methods
   that may suffice in certain situations.
%    Frequency-dependent fields-of-view are corrected via existing wide-field imaging
%    algorithms.
   }
  % results heading (mandatory)
   {We show via simulations and application to wideband (E)VLA data (1 GHz to 2 GHz),
    that it is possible to reconstruct both spatial and spectral structure of 
    compact and extended emission at the continuum sensitivity level and
    angular resolution allowed by the highest sampled frequency.
   We show imaging examples using simulations, multi-frequency VLA data
    and wideband EVLA data,  to illustrate the capabilities of the MS-MFS algorithm
    and conditions under which these techniques are feasible and give accurate results.
   }
  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{multi-frequency-synthesis --
                multi-scale-deconvolution --
                spectral index --
		spectral curvature --
               }

   \authorrunning{U.Rau and T.J.Cornwell}
   \titlerunning{Multi-Scale Multi-Frequency Synthesis Imaging in Radio Interferometry}
   \maketitle
%

\section{Introduction}

%Instruments like the EVLA and e-MERLIN, are among a new 
%generation of broad band radio interferometers designed to provide continuum sensitivities and
%high dynamic range imaging capabilities superior to that of existing radio interferometers.
%The wide bandwidths will allow detailed measurements of the spectral structure of astrophysical
%sources, but to achieve the desired continuum sensitivities, imaging algorithms need to be 
%sensitive to the effects of combining measurements from a large range of frequencies 
% (namely varying ranges of sampled spatial scales, varying antenna fields-of-view, and
% the spectral structure of the sky brightness distribution).

A new generation of broad-band radio interferometers is currently being
designed and built to provide 
high-dynamic-range imaging capabilities superior to that of existing instruments. 
The large instantaneous bandwidths offered by new front-end systems 
increases the raw continuum sensitivity of the instrument,
%The introduction of broad-band receivers into radio interferometry has
%opens up new opportunities for the study of wide-band continuum
%emission from a vast range of astrophysical objects.  In addition to
%increasing the imaging sensitivity of the instrument, such systems
as well as 
greatly enhances our ability to measure the detailed spectral structure 
of the incident radiation across large continuous bandwidths. 
%
One prominent reason for using broad-band receivers on an imaging
interferometer has been to obtain a continuum image that makes 
use of the increased sensitivity and spatial frequency 
coverage offered by combining measurements from multiple frequencies. 
So far, the bandwidths used have been relatively
narrow ($<25\%$), and effects due to spectral structure of the 
sky-brightness distribution have been considered 
only in the context of reducing errors in the continuum image.% \citep{MFCLEAN_CCW,MFCLEAN}.
Any spectral information is obtained only as a by-product, and attention has not 
been paid to their accuracy.
But now, the new bandwidths ($~100\%$) are large enough to allow the spectral 
structure of the sky brightness distribution 
to also be reconstructed to produce a meaningful astrophysical measurement.   
To do so, 
%while taking full advantage of these instruments, we need imaging 
we need imaging 
algorithms that model and reconstruct both spatial and spectral structure simultaneously,
and that are also 
%sensitive to the frequency dependence of the
%instrument as well as the sky brightness distribution. 
sensitive to various effects of combining measurements from a large range
 of frequencies
 (namely varying ranges of sampled spatial scales and 
varying array-element response functions)

%The use of broad-band receivers in radio interferometery impacts standard continuum
%    synthesis imaging in three ways. 
%\begin{enumerate}
%\item   Correlations measured at different frequencies sample the visibility function of the
%   sky brightness at different spatial frequencies. 
%\item  The sky brightness can change significantly across the large frequency 
%    range that these new receivers are sensitive to.
%\item  Instrumental effects such as the element response pattern vary with frequency.
%\end{enumerate}
%%The following discussion describes existing methods designed to account for some of
%the above, and motivates the work that is the focus of this paper.

%    Existing algorithms designed to account for some of the above, can be shown to be
%    insufficient to achieve target image fidelity and dynamic range over the wide bands that 
%    most new instruments are being built with.

% and
% the spectral structure of the sky brightness distribution).

%Continuum emission from most astrophysical radio sources shows 
%significant spectral structure over the frequency ranges that these
%new receivers will instantaneously measure.
%To make a continuum image at the desired sensitivity, 
%it is essential to measure or reconstruct the spectral structure 
%the sky-brightness distribution 
%before constructing an image of the 
%integrated flux, and to do this while accounting for the frequency 
%dependence of the instrument (namely varying ranges of sampled 
%spatial scales, and varying antenna fields-of-view).
%
%While the main goal of wide-band imaging is to obtain a high dynamic-range
%continuum image, the reconstructed spectral structure can also be
%a useful astrophysical measurement.

The simplest method of wide-band image reconstruction is to image 
each frequency channel separately and combine the results at the end. 
However, single-channel imaging is restricted to the narrow-band
$uv$-coverage and sensitivity of the instrument, and source 
%Also, the angular resolution of the telescope 
%changes with frequency, 
spectra can be studied only at the
angular resolution allowed by the lowest frequency in the sampled range.
For complicated extended emission, the single-frequency $uv$-coverage
may not be sufficient to produce a consistent solution across frequency.
While such imaging may suffice for many science goals, it does not
take full advantage of what an instantaneously wide-band instrument provides,
namely the sensitivity and spatial-frequency coverage obtained by 
combining measurements from multiple receiver frequencies.
%

Multi-Frequency-Synthesis (MFS) is the technique of combining measurements at multiple
discrete receiver frequencies during synthesis imaging.
MFS was initially done (REF) to increase the aperture-plane coverage
of sparse arrays by using narrow-band receivers and switching frequencies during the observations.
Wide bandwidth systems ( ~10\% ) later presented the problem of bandwidth smearing, which was
eliminated (REF) by splitting the wide band into narrow-band channels and mapping them onto
their correct spatial frequencies during imaging. It was assumed that
at the receiver sensitivities of the time, 
the measured sky brightness was constant across the observed bandwidth.
The next step was to consider a frequency-dependent sky brightness distribution.
\citep{MFCLEAN_CCW} describe a double-deconvolution algorithm based on
%and \citep{MFCLEAN} describe multi-frequency
%deconvolution algorithms based on 
the instrument's responses to a series of spectral basis functions, 
in particular, the first two terms of a Taylor series.
A map of the average spectral index is derived from the coefficient maps.
%
\citep{MFCLEAN} describe the MF-Clean algorithm which uses a formulation
similar to double-deconvolution but calculate Taylor-coefficients via a
least-squares solution. MF-Clean is implemented in the miriad software package.
The MF-Clean formulation ignores the process of resampling 
multi-frequency visibilities onto a single grid of spatial frequencies, and for 
arrays with many overlapping spatial-frequency tracks, errors are incurred.
More recently, \citet{LIKHACHEV} re-derive the least-squares method used
in MF-Clean to include more than two series coefficients.
\citep{SPATIOSPECTRALMEM} describe spatio-spectral 
MEM, an entropy based method in which single-channel imaging is done
along with a smoothness constraint applied across frequency.

%
%Both methods model an image as a collection of point 
%sources with linear spectra, solve for their intensities and slopes,
%and derive an average spectral-index for each source from the
%fitted slope. 
%The MF-Clean methods were developed for relatively narrow bandwidths ($<25\%$), 
%and can be shown \citep{EVLAMEMO101} to be insufficient to model 
%typical spectral structure across the large frequency ranges that 
%new wide-band receivers are sensitive to.
%

So far, these CLEAN-based multi-frequency deconvolution algorithms used 
point-source (zero-scale) flux components to model the sky emission.
% choice not well suited for extended emission.
We show in this paper that with the MF-Clean approach, 
deconvolution errors that occur with a point-source
model are enhanced in the spectral index image because of error propagation effects, and
that the use of a multi-scale technique can minimize this.
%
%In (REF1,REF1), XXX show that for high dynamic range imaging, point-source models
%are not well suited for extended emission, and describe several multi-scale
%techniques that are more capable at reconstructing complex extended emission accurately.
%Multiscale deconvolution techniques model images using flux
%components of varying scale size, and are more accurate at deconvolving large scale emission
%features. 
\citep{MSCLEAN} describe the MS-CLEAN algorithm which is a 
matched filtering technique using templates constructed from the instrument 
response to various large scale flux components. 
\citep{ERIC_MSCLEAN} describe a method simular to MS-Clean, and 
\citep{Asp_Clean} describe the ASP-CLEAN algorithm that explicitly fits 
for the parameters of Gaussian flux components and uses scale size to aid the separation of
signal from noise.

For high dynamic range imaging across wide fields-of-view,
direction-dependent instrumental effects that need to be accounted for are 
the w-term and the antenna power pattern, both of which are frequency-dependent. 
\citep{AWProjection}describe an algorithm for the correction of time-variable
wide-field instrumental effects for narrow-band interferometric imaging.

In this paper, we describe MS-MFS (multi-scale multi-frequency synthesis) as
an algorithm that combines variants of the MF-Clean and MS-Clean
approaches to simultaneously reconstruct both spatial and spectral structure of the 
sky-brightness distribution. 
Frequency-dependent primary-beam correction is considered as a 
post-deconvolution correction step\footnote{The integration of 
direction-dependent correction algorithms such as AW-Projection
with MS-MFS will be discussed in a subsequent paper.}.
In section ZZ, we show imaging examples using simulations, multi-frequency VLA data
 and wideband EVLA data,  to illustrate the capabilities and limits of the
MS-MFS algorithm.

%__________________________________________________________________

\subsection{Wide-Band Imaging}\label{Sec:WIDEBANDIMAGING}
%__________________________________________________________________

%
%The use of broad-band receivers in radio interferometery impacts standard continuum
%    synthesis imaging in three ways. 
%First, correlations measured at different frequencies sample the visibility function of the
%   sky brightness at different spatial frequencies. 
%Second, the sky brightness can change significantly across the large frequency 
%    range that these new receivers are sensitive to.
% Third, direction-dependent instrumental effects such as the element response pattern
%     also vary with frequency.
%This section describes existing methods designed to account for some of
%the above, and motivates the work that is the focus of this paper.
%

We begin with a discussion of how well we can reconstruct both spatial and 
spectral information from an incomplete set of visibility samples at
multiple observing frequencies.
% and describe how our choice of a flux model influences the 
%image reconstruction process when each observing frequency measures a different 
%set of spatial frequencies.
%Section \ref{Sec:MSMFS-model} then defines a multi-scale multi-frequency image model  
%and section \ref{Sec:MSMFS-eqns} describes a deconvolution algorithm that uses
%this flux model.

The spatial frequencies sampled at each observing frequency $\nu$ are between  
${u}_{min} = \frac{\nu}{c}{b}_{min} $ and ${u}_{max} = \frac{\nu}{c}{b}_{max}$, 
where ${u}$ is used here as a generic label for the $uv$-distance\footnote
{
The $uv$-distance is defined as $\sqrt{u^2+v^2}$ and is the 
radial distance of the spatial frequency measured by the baseline from the
 origin of the $uv$-plane, in units of wavelength $\lambda$.
} and 
$b$ represents the length of the baseline vector (in units of meters) projected onto the plane
perpendicular to the direction of the source.
%The maximum spatial frequency measured at each frequency defines the angular 
%resolution of the instrument ($\theta_{\nu} = 1/u_{max}(\nu)$).
%
The range of spatial frequencies between ${u}_{min}$ at $\nu_{max}$ and
${u}_{max}$ at $\nu_{min}$ represents the region where the visibility function
is sampled at all frequencies in the band, and there is sufficient information
to reconstruct both spatial and spectral structure.
The spatial frequencies outside this region are sampled only by a fraction of the 
band and the accuracy of a broad-band reconstruction
depends on how well the spectral and spatial structure are
constrained by an appropriate choice of a flux model. % and the rest of the measurements.

%\begin{enumerate}
%\item
For a flat-spectrum source, measurements at multiple frequencies sample the 
same sky brightnes distribution at different ranges of spatial scales, increasing 
the signal-to-noise of the measurements in regions of overlap, and providing 
better uv-coverage and angular resolution.
Therefore standard deconvolution algorithms applied to measurements combined 
via MFS will suffice to reconstruct source structure across the full range of 
spatial scales measured across the band.

%For sources with spectral structure, different frequencies measure different sky
% brightness distributions, and a direct combination of frequencies via MFS would 
%generate spurious spatial structure. The traditional method of imaging has been to
% make separate images at each frequency and and derive spectral structure after
% tapering the measurements from higher frequencies to match the angular 
%resolution of the lowest frequency. The imaging sensitivity and fidelity is however
%restricted to be that of a single channel and its uv-coverage, and in most cases,
%it is possible to do better.
%

%\noindent A few examples are used to illustrate the importance of an appropriate flux model.
%\item 
A compact, unresolved source with spectral structure is measured as a point
source at all frequencies, and $u_{max}$ at $\nu_{max}$ gives the maximum angular resolution at
which this source can be imaged. Since the visibility function of a point source is flat across
the entire spatial frequency plane, its spectrum is adequately sampled by the
multi-frequency measurements.
Using a flux model in which each source is a $\delta$-function with a smooth 
polynomial spectrum, 
it is possible to reconstruct the spectral structure of the source
at the maximum possible angular resolution.

%\item 
For resolved sources with spectral structure, the accuracy of the reconstruction
across all spatial scales between $u_{min}$ at $\nu_{min}$ and  $u_{max}$ at $\nu_{max}$
depends on an appropriate choice of flux model, and the constraints that it provides.
For example, a source emitting broad-band synchrotron radiation can be described by
a fixed brightness 
distribution at one frequency with a power-law spectrum associated with each location. 
Images can be made at the maximum angular resolution (given by $u_{max}$ at $\nu_{max}$)
with the 
assumption that different observing frequencies probe the same spatial 
structure but measure different amplitudes (usually a valid assumption). 
%Measurements at the highest frequencies constrain the angular size of the flux 
%components, and the combined spectra are constrained by the rest of the measurements.
This constraint is strong enough to correctly reconstruct even moderately resolved sources
that are completely unresolved at the low end of the band but resolved at the higher end.
On the other hand, a source whose structure itself changes by 100\% in amplitude 
across the band would break the above assumption (band-limited signals).
%that structure seen at the highest resolution is what is present all across the band.
One example is with multi-frequency observations of solar magnetic loops
where the different frequencies probe different layers in the upper chromosphere
and can have very different structures. 
In this case, a complete reconstruction would be possible only in the region of
overlapping spatial frequencies (between $u_{min}$ at $\nu_{min}$ and  $u_{max}$ at $\nu_{max}$),
unless the flux model includes constraints that bias the solution towards one
appropriate for such sources.
%The imaging would still however operate at the continuum sensitivity level, and is therefore
%better than single-channel imaging.

%\item 
The lower end of the spatial frequency range presents a different problem. 
The size of the central hole in the $uv$-coverage
increases with frequency. Spectra are not measured adequately for
%extended 
emission whose visibility function is non-zero only %in
%the range 
below $u_{min}$ at $\nu_{max}$ and a flat-spectrum large-scale source can
be indistinguishable from a relatively smaller source with a steep spectrum.
%For scale sizes in this range,
Additional constraints in the form 
of total-flux values for each frequency may be required for an accurate reconstruction.
% or spectrum is simple enough that half the band is enough to get it.
%\end{enumerate}

Finally, for wide-field imaging, the frequency-dependence of the primary-beam
can introduce artificial spectral effects that result in a 100\% variation of the
flux across the band. To recover both spatial and spectral structure of
the sky brightness across a large field of view, the frequency dependence of
the primary beam must be modeled and removed before or 
during multi-frequency synthesis imaging.  
%In addition to the effects of $uv$-coverage and source structure, various 
%frequency depdendent instrumental effects also need to be accounted for. 
%The angular size of the primary beam of the antenna decreases with
%an increase in observing frequency (see Fig.\ref{WBPB1}, section \ref{Sec:FreqDep}).
%Sources away from the pointing center of the beam
%are attenuated by different amounts across the 
%frequency band and this introduces artificial spectral structure into the
%measurements. To recover both spatial and spectral structure of
%the sky brightness across a large field of view, the frequency dependence of
%the primary beam must be modeled and removed during multi-frequency synthesis
%imaging.

%In addition to the sky flux being frequency dependent, the instrumental effects are also 
%frequency dependent.
%The HPBW of the primary beam of each array element scales as $HPBW = \frac{c}{\nu D}$%. 
%The reconstructed image will therefore have artificial spatial and spectral structure that can% be
%measured and removed to undo these effects.

To summarize, just as standard interferometric image reconstruction uses 
{\it a priori} information about
the spatial structure of the sky to estimate the
visibility function in unmeasured regions of the $uv$-plane,
multi-frequency image reconstruction
algorithms need to use {\it a priori} information about the spectral
structure of the sky brightness. 
By combining such models 
with the known frequency-dependence of the 
spatial-frequency coverage and element response function, 
it is possible to reconstruct the broad-band sky brightness distribution % across wide fields of view
from incomplete spectral and spatial-frequency sampling.

%\subsection{Outline of this paper}

%Section \ref{Sec:WIDEBANDIMAGING} discusses the information contained in
%multi-frequency measurements from the point of view of a combined spatial and
%spectral reconstruction. 
%Section \ref{Sec:OVERVIEW} summarizes various existing algorithms that can be
%used for wide-band imaging, and points out what improvements are required.
%Section \ref{Sec:ALGORITHM} describes the formulation and implementation of the
%MS-MFS algorithm (multi-scale multi-frequency synthesis) as 
%a linear least-squares technique combined with conventional deconvolution. 
%This is an extension and combination of two existing methods with a few 
%differences that affect the accuracy and stability of the results. 
%Section \ref{Sec:IMAGINGRESULTS} shows some imaging results that demonstrate the capabilities
% and limits of the MS-MFS algorithm, and discusses various sources of error
%inherent in the reconstruction process. 
%Section \ref{Sec:CONCLUSIONS} summarizes these results.
%

\section{Multi-scale Multi-frequency deconvolution}\label{Sec:ALGORITHM}

%The design of an image reconstruction algorithm begins with an appropriate choice of
%parameterization of the sky brightness distribution and various instrumental effects.
%This parameterization is then folded into the transfer function of the instrument
%whose measurement process we are trying to invert. An optimization process is then
%followed to obtain best-fit estimates for the parameters which can then be 
%interpreted physically. 
%
The MS-MFS algorithm described here is based on the iterative image-reconstruction
framework\footnote{A 
steepest-descent chi-square minimization is done by iterating between two steps.
A major cycle computes the RHS of the normal equations, and the
minor cycle performs an approximate inverse of the Hessian matrix to generate
an update direction (image model components). }
described in \citep{RAU_IEEE_2009}. 
Sections xxx to yyy formulate the algorithm used in the minor-cycle of iterative
deconvolution for MS-MFS. 
Differences between the multi-scale and multi-frequency parts of MS-MFS with
the original MS-Clean and MF-Clean approaches are highlighted in sections XX and YY.
The implementation of this algorithm in the CASA package is summarized in
section ZZ.

\subsection{Parameterization of spatial structure}
For a multi-scale model, an image can be written as a linear combination of images
at different angular resolutions (Ref MS-CLEAN).
A multiscale representation of an image is given by 
\begin{equation}
\vec{I}^{model} = \sum_{s=0}^{N_s-1}  \vec{I}^{shp}_{s} \star \vec{I}^{sky,\delta}_s
\label{Eq:ms_model}
\end{equation}
where $N_s$ is the number of discrete spatial scales used to represent the image, and
$\vec{I}^{sky,\delta}_{s}$ represents a collection of $\delta$-functions that describe the locations
and integrated amplitudes of flux components of scale $s$ in the image.
$\vec{I}^{shp}_s$ is a tapered truncated parabola of width proportional to $s$.
The symbol $\star$ denotes convolution. 
%$\T_s=[\F]\vec{I}^{shp}_{s}$ is a 
%spatial-frequency taper equivalent to the convolution kernel $\vec{I}^{shp}_{s}$.
In order to always allow for the modeling of unresolved sources, 
we choose the first scale function $\vec{I}^{shp}_{s=0}$ to be a $\delta$-function.
%Successive basis functions then correspond to inverted parabolas of 
%larger widths (as $s$ increases).

\subsection{Parameterization of spectral structure}\label{Sec:freqmodel}
The spectrum of each flux component is modeled by a polynomial in frequency
( a Taylor series expansion about $\nu_0$ ).
\begin{eqnarray}
%\label{Eq:mf_model}
\label{Eq:mf_model}
\vec{I}^{model}_{\nu} = \sum_{t=0}^{N_t-1} \wnt \vec{I}^{sky}_{t} ~~~\mathrm{where}~~~ \wnt&=&\dnuno^t 
%\label{Eq:tfunc_2}
%~~~~~\mathrm{or}~~~~ \wnt&=&\left[\log\nuno\right]^t
\end{eqnarray}
where $N_t$ is the order of the Taylor series expansion, and 
the $I^m_t$ represent multi-scale Taylor coefficient images (moment maps ??? ).
This decomposition is linear in the coefficients as well as in the basis functions.

These Taylor coefficients can be interpreted by choosing an astro-physically appropriate
spectral model and performing a Taylor expansion to derive expressions that each coefficient
maps to.
One choice for a spectral model is a power law with a varying index, represented by a 
second-order polynomial in $log(I)~vs~log\nuno$ space.
The variation of the spectral index with frequency is described by an average spectral
index $\alpha$ and a curvature term $\beta$.
\begin{equation}
I_{\nu}^{sky} = I_{\nu_0}^{sky} \nuno^{\alpha + \beta log \nuno}
\label{EQN_POWERLAW1}
\end{equation}
The main reason behind this choice of interpretation is the fact that continuum synchrotron emission is 
usually modeled (and observed) as a power law distribution with frequency. Across the wide
frequency ranges that new receivers are now sensitive to, 
spectral breaks, steepening and turnovers need to be factored into
models, and the simplest way to include them and ensure smoothness, is spectral curvature.
(FIGURE).
%% Figure Caption
%% (1) Synchrotron spectrum with steepening, turnover, break.
(Wideband imaging algorithms describes in CCW/SW use a fixed spectral index across the band,
and handle slight curvature by performing multiple rounds of imaging after removing the 
dominant/average $\alpha$ at each stage. 
They also suggest using higher order polynomials to handle spectral curvature.)

A Taylor expansion of Eqn \ref{EQN_POWERLAW1} yields the following expressions for the first
three coefficients ($t=0,1,2$ in Eqn \ref{EQN_TS}) from which the spectral index $\alpha$ and curvature $\beta$ images 
can be computed algebraically.
\begin{equation}
I^m_0 = I^{sky}_{\nu_0} ~~;~~~ I^m_1 = \alpha I^{sky}_{\nu_0} ~~;~~~ I^m_2 = \left(\frac{\alpha(\alpha-1)}{2} + \beta\right) I^{sky}_{\nu_0}
\label{EQN_COEFFS}
\end{equation}
Note that with this choice of parameterization, 
we are using a polynomial to model an exponential, and this has caveats, etc which
are discussed in the section on errors.
Also, note that there can be other interpretations of the Taylor coefficients %poln ?
and other expansions. 
A power-series expansion about $\alpha$ and $\beta$ will yield a logarithmic
expansion i.e. I vs log nu. (CHECK if/when this is better)
It is however impractical to work directly in Log I and Log nu space because that
involves taking logs and exp of image pixel amplitudes and this is highly unstable numerically.

Across wide fields of view, an approximate correction of the primary-beam and its
frequency dependence can also be folded into this formulation. 
We can write 
\begin{equation}
I_{\nu}^{sky} = P_{\nu}  I_{\nu}^{true} =  P_{\nu_0} I_{\nu_0}^{true} \nuno^{[\alpha_{true}+\alpha_{PB}] + [\beta_{true} + \beta_{PB} ] log \nuno}
\label{EQN_POWERLAW2}
\end{equation}
where $P_{\nu_0}$ is the primary beam at the reference frequency, and
$\alpha_{PB}$ and $\beta_{PB}$ are spectral index and curvature of the frequency
dependence of the primary beam.   An image reconstructed using such a 
formulation can be corrected in a post-deconvolution step, if the primary-beam and
its frequency dependence is known {\it a priori}.  This however corrects only for
an average primary-beam and its frequency dependence, and a more accurate
solution requires the AW-Projection algorithm. 
%%%%% Implementation

\subsection{Multi-scale multi-frequency model}
For multi-scale and multi-frequency deconvolution, the sky-brightness distribution
can be parameterized in a multi-scale basis (\ref{Eq:ms_model}), with the amplitudes
of each component described by a polynomial across frequency (\ref{Eq:mf_model}).  
A region of emission in which the spectrum varies with position
will then be modeled as a sum of wide-band flux components.
The image flux model at each frequency can be written as a linear sum of  
coefficient images at different spatial scales. 
\begin{equation}
\vec{I}^{model}_{\nu} = \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} \wnt \left[ \vec{I}^{shp}_s \star \vec{I}^{sky}_{s\atop t}\right] ~~~~\mathrm{where}~~~\wnt = \dnuno^t 
\label{Eq:msmf_model}
\end{equation}
Here, $N_s$ is the number of discrete spatial scales used to represent the image and  
$N_t$ is the order of the series expansion of the spectrum. 
$\vec{I}^{sky}_{s\atop t}$ represents a collection of $\delta$-functions that describe the locations
and integrated amplitudes of flux components of scale $s$ in the image of the $t^{th}$ series 
coefficient.

\subsection{Measurement equations}\label{Sec:meqn}
The {Measurement Equations\footnote
{
The measurement equation of an imaging instrument describes its transfer function (the effect of
the measurement process on the input signal).
For an ideal interferometer (a perfect spatial frequency filter, with no instrumental gains),
it can be written in matrix notation as follows. 
Let $I^{sky}_{m\times 1}$ be a pixelated image of the sky and let $V^{obs}_{n\times 1}$
be a vector of $n$ visibilities. 
Let $S_{n\times m}$ be a projection operator that describes
the uv-coverage as a mapping of $m$ discrete spatial 
frequencies (pixels on a grid) to $n$ visibility samples (usually $n>m$).
Let $F_{m\times m}$ be the Fourier transform operator.
Then, $[{\Sa}_{n\times m}] [F_{m\times m}] \vec{I}^{sky}_{m\times 1} = \vec{V}^{obs}_{n\times 1} $
}
}for a sky brightness distribution parameterized by Eqn.\ref{Eq:msmf_model} are

\begin{eqnarray}
\vec{V}^{obs}_{\nu} &=& [\Sna][\F]\vec{I}^{model}_{\nu}
= \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} \wnt [\Sna][T_s][\F] \vec{I}^{sky}_{s\atop t}
\label{Eq:msmfs_meqn}
\end{eqnarray}
where $w_{\nu}$ are Taylor-weights, $[\Sna]$ represents the 
spatial-frequency sampling function for frequency $\nu$, and 
the image-domain convolution with $\vec{I}^{shp}_s$ is written as a spatial-frequency
taper function $[T_s]_{m\times m} = diag([\F] \vec{I}^{shp}_s)$.

Combining measurements from all frequencies together, we can write
\begin{eqnarray}
\vec{V}^{obs} &=& \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} [\Wnt][\Sa][T_s] [\F] \vec{I}^{sky}_{s\atop t}
\label{Eq:msmfs_meqn}
\end{eqnarray}
where $[\Wnt]$ is a diagonal $nN_c\times nN_c$ matrix of weights, comprised of $N_c$ blocks each
of size $n\times n$ for each frequency channel ($\nu$), containing $\wnt$. 
The multi-frequency $uv$-coverage of the synthesis array is represented by $[\Sa_{nN_c\times m}]$.

The full measurement matrix therefore has the shape
$nN_c\times mN_sN_t$, which when multiplied by the set of $N_sN_t$ model sky vectors each
of shape $m\times 1$, produces $nN_c$ visibilities.

\noindent For $N_t=3, N_s=2$ the measurement equations can be written as follows, in
block matrix form. The subscript $p$ denotes the $p^{th}$ spatial scale
and the subscript $q$ denotes the $q^{th}$ Taylor coefficient of the spectrum
polynomial.

{
\begin{equation}
\begin{array}{l}
\left[\begin{array}{llllll} 
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
%   \left[\W^{mfs}_0 \Sa T_0 \F\right] & \left[\W^{mfs}_1 \Sa T_0 \F\right] & \left[\W^{mfs}_2 \Sa T_0 \F\right] & 
%   \left[\W^{mfs}_0 \Sa T_1 \F\right] & \left[\W^{mfs}_1 \Sa T_1 \F\right] & \left[\W^{mfs}_2 \Sa T_1 \F\right] \\  
   \left[A_{{0}\atop{0}}\right] & \left[ A_{{0}\atop{1}} \right] & \left[A_{{0}\atop{2}}\right] & 
   \left[A_{{1}\atop{0}}\right] & \left[A_{{1}\atop{1}}\right] & \left[A_{{1}\atop{2}}\right] \\  
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
   \end{array} \right] \\
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
\noalign{\medskip}
{~\mathrm{where}~~\left[ A_{{p}\atop{q}}\right] = [\W^{mfs}_q][\Sa][T_p][\F]} \\
\noalign{\medskip}
{~\mathrm{for}~p\in\{0,N_s-1\}~\mathrm{and}~q\in\{0,N_t-1\}    }  
\end{array}
\left[\begin{array}{l} 
\noalign{\medskip}
                       \vec{I}^{sky}_{{0}\atop{0}} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{{0}\atop{1}} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{{0}\atop{2}} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{{1}\atop{0}} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{{1}\atop{1}} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{{1}\atop{2}}\\
\noalign{\medskip}
		       \end{array}\right] =\vec{V}^{obs}%_{nN_c\times 1}
\label{meqn_msmfs_math}
\end{equation}
}

\subsection{Normal equations}\label{Sec:neqn}

The {Normal Equations\footnote
{\label{FN:Neqn}
The Normal Equations are the linear system of equations whose solution 
gives a weighted least-squares estimate of a set of parameters in a model  
($\chi^2$ minimization). 
For an ideal interferometer, it is given by 
$[\Fd \Sd W \Sa \F ] I^{sky}_{m\times 1} = [\Fd \Sd W] V^{obs}_{n\times 1} = I^{dirty}_{m\times 1}$
where $W_{n\times n}$ is a diagonal matrix of signal-to-noise based measurement weights
and $\Sd$ denotes the mapping of measured visibilities onto a spatial frequency grid.
The Hessian (matrix on the LHS) is by construction a circulant { convolution} operator with 
a shifted version of $I^{psf}_{m\times 1} = diag[\Fd \Sd W \Sa]$ in each row. 
The {dirty} image on the RHS (produced by direct Fourier inversion of weighted visibilities)
is therefore the convolution of $I^{sky}_{m\times 1}$ with $I^{psf}$,
and these equations can be solved by a {deconvolution}. 
}
}
can be written in block matrix form, with each block-row 
(for scale size $s$, and Taylor term $t$) is given by
\begin{eqnarray}
\label{Eq:msmfs_neqn_1}
\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1} \left[\He_{{s,p}\atop{t,q}}\right] \vec{I}^{sky}_{{p}\atop{q}} &=& \vec{I}^{dirty}_{{s}\atop{t}}~~~  \forall~ s \in [0,N_s-1], t\in[0,N_t-1]
\end{eqnarray}
where each $\left[\He_{{s,p}\atop{t,q}} \right]$ is an $m\times m$ block of the 
Hessian matrix, and $\vec{I}^{dirty}_{{s}\atop{t}}$ is one of $N_s N_t$ dirty images.
\begin{eqnarray}
\left[\He_{{s,p}\atop{t,q}} \right] &=& \left[A_{{s}\atop{t}}^{\dag}\right][\Wim] \left[A_{{p}\atop{q}}\right] \\
 &=&    [\Fd T_s  \Sd \W^{mfs}_t]  [\Wim^{mfs}]  [\W^{mfs}_q \Sa T_p \F] \\
 &=&    [\Fd T_s \F] [\Fd  \Sd \W^{mfs}_t  \Wim^{mfs}  \W^{mfs}_q \Sa  \F] [\Fd T_p \F]\\
% &=&    [\Fd T_s \F] [\He^{mfs}_{t,q}] [\Fd T_p \F]
 &=& [\Fd T_s \F] \left\{  \sum_{\nu} \wntq [\Fd\Snd\Wimn\Sna\F] \right\} [\Fd T_p \F] \\
 &=& [\Fd T_s \F] \left\{  \sum_{\nu} \wntq [\He_{\nu}\} \right] [\Fd T_p \F]
\end{eqnarray}
%\begin{eqnarray}
% [\He^{mfs}_{t,q}] &=& [\Fd  \Sd \W^{mfs}_t  \Wim^{mfs}  \W^{mfs}_q \Sa  \F]\\
%   &=& \sum_{\nu} \wntq [\Fd\Snd\Wimn\Sna\F] \\
% &=& \sum_{\nu} \wntq [\He_{\nu}]
%\end{eqnarray}
%\begin{eqnarray}
%\label{Eq:Hessian}
% [\He_{\nu}] &=&  [\Fd\Snd\Wimn\Sna\F]
%\end{eqnarray}
$[\He_{\nu}] =  [\Fd\Snd\Wimn\Sna\F]$ is the Hessian matrix formed using only one
 frequency channel, and
is a convolution operator containing a shifted version of the single-frequency 
PSF $\I^{psf}_{\nu} = diag[\Fd \Sd W \Sa]$ in each row (see footnote \vref{FN:Neqn}).
$[\Fd \T_s\F]$ is also a convolution operators with $I^{shp}_s$ as 
the kernel. 
The process of convolution is associative and commutative, and 
therefore, $\left[\He_{{s,p}\atop{t,q}}\right]$  is also a convolution operator 
whose kernel is given by 
\begin{equation}
\label{Eq:msmfs_neqn_2.5}
\vec{I}^{psf}_{{s,p}\atop{t,q}} = I^{shp}_s  \star \left\{ \sum_{\nu} \wntq \vec{I}^{psf}_{\nu} \right\} \star I^{shp}_p 
%&=& I^{shp}_s \star I^{psf}_{t,q} \star I^{shp}_p
\end{equation}
Therefore, to compute the Hessian matrix, it suffices to compute one such kernel
per Hessian block.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The dirty images on the RHS of Eqn.\ref{Eq:msmfs_neqn_1} can be written as follows.
\begin{eqnarray}
\label{Eq:msmfs_neqn_3}
\vec{I}^{dirty}_{{s}\atop{t}}  &=& [\Fd T_s \F][\Fd\Sd\Wntd\Wim] \vec{V}^{obs} \\
&=& [\Fd T_s \F] \left\{  \sum_{\nu} \wnt [\Fd\Snd\Wimn] \vec{V}_{\nu}^{obs}  \right\} \\
&=& I^{shp}_s \star \left\{\sum_{\nu} \wnt \vec{I}^{dirty}_{\nu}  \right\} 
%&=& [\Fd T_s \F] \vec{I}^{dirty}_t = I^{shp}_s \star \vec{I}^{dirty}_t\\
\label{Eq:mf_neqn_3a}
%\vec{I}^{dirty}_{t}  &=& [\Fd\Sd\Wntd\Wim] \vec{V}^{corr}\\
% &=& \sum_{\nu} \wnt [\Fd\Snd\Wimn] \vec{V}_{\nu}^{corr} = \sum_{\nu} \wnt \vec{I}^{dirty}_{\nu}\\
%\label{Eq:specPSF}
%\vec{I}^{psf}_{t,q}&=& \sum_{\nu} \wntq [\Fd\Snd\Wimn] \vec{1} = \sum_{\nu} \wntq \vec{I}^{psf}_{\nu}
\end{eqnarray}
where ${\I}_{\nu}^{dirty} = [\Fd \Snd \W_{\nu}] \vec{V}^{obs}_{\nu} $ is the dirty
image formed by direct Fourier inversion of weighted visibilities from one frequency channel.

When all scales and Taylor terms are combined, 
the full Hessian matrix contains
$N_t N_s \times N_t N_s$ blocks each of size $m\times m$, 
and $N_t$ Taylor coefficient images each of size $m\times 1$, 
for all $N_s$ spatial scales.

The normal equations in block matrix form for the example 
in Eqn.\ref{meqn_msmfs_math} for $N_t=3, N_s=2$ is shown
in Eqn.\ref{Eq:msmfs_neqn_matrix}.
The Hessian matrix consists of $N_s\times N_s=2\times 2$ blocks 
(the four quandrants of the matrix), each for one pair of spatial scale $s,p$.
Within each quadrant, the $N_t\times N_t = 3\times 3$
matrices correspond to various pairs of $t,q$ (Taylor coefficient indices).
This layout shows how the multi-scale and multi-frequency aspects of
this imaging problem are combined and illustrates the 
dependencies between the spatial and spectral basis functions.
%Note that the $3\times3$ block in the top left quadrant corresponds to the
%entire Hessian matrix in Eqn.~\ref{Eq:mfs_neqn3} (since $\I^{shp}_0$ is a
%$\delta$-function).

\begin{equation}\small
\left[\begin{array}{llllll} 
\noalign{\medskip}
   \left[H_{{ 0, 0}\atop{ 0, 0}}\right] & \left[H_{{ 0, 0}\atop{ 0, 1}}\right] & \left[H_{{ 0, 0}\atop{ 0, 2}}\right] & \left[H_{{ 0, 1}\atop{ 0, 0}}\right] & \left[H_{{ 0, 1}\atop{ 0, 1}}\right] & \left[H_{{ 0, 1}\atop{ 0, 2}}\right] \\  
\noalign{\medskip}
   \left[H_{{ 0, 0}\atop{ 1, 0}} \right] & \left[H_{{ 0, 0}\atop{ 1, 1}}\right] & \left[H_{{ 0, 0}\atop{ 1, 2}}\right] & \left[H_{{ 0, 1}\atop{ 1, 0}}\right] & \left[H_{{ 0, 1}\atop{ 1, 1}}\right] & \left[H_{{ 0, 1}\atop{ 1, 2}}\right] \\  
\noalign{\medskip}
   \left[H_{{ 0, 0}\atop{ 2, 0}} \right] & \left[H_{{ 0, 0}\atop{ 2, 1}}\right] & \left[H_{{ 0, 0}\atop{ 2, 2}}\right] & \left[H_{{ 0, 1}\atop{ 2, 0}}\right] & \left[H_{{ 0, 1}\atop{ 2, 1}}\right] & \left[H_{{ 0, 1}\atop{ 2, 2}}\right] \\  
\noalign{\medskip}
%   \hline
\noalign{\medskip}
   \left[H_{{ 1, 0}\atop{ 0, 0}} \right] & \left[H_{{ 1, 0}\atop{ 0, 1}}\right] & \left[H_{{ 1, 0}\atop{ 0, 2}}\right] & \left[H_{{ 1, 1}\atop{ 0, 0}}\right] & \left[H_{{ 1, 1}\atop{ 0, 1}}\right] & \left[H_{{ 1, 1}\atop{ 0, 2}}\right] \\  
\noalign{\medskip}
   \left[H_{{ 1, 0}\atop{ 1, 0}} \right] & \left[H_{{ 1, 0}\atop{ 1, 1}}\right] & \left[H_{{ 1, 0}\atop{ 1, 2}}\right] & \left[H_{{ 1, 1}\atop{ 1, 0}}\right] & \left[H_{{ 1, 1}\atop{ 1, 1}}\right] & \left[H_{{ 1, 1}\atop{ 1, 2}}\right] \\  
\noalign{\medskip}
   \left[H_{{ 1, 0}\atop{ 2, 0}} \right] & \left[H_{{ 1, 0}\atop{ 2, 1}}\right] & \left[H_{{ 1, 0}\atop{ 2, 2}}\right] & \left[H_{{ 1, 1}\atop{ 2, 0}}\right] & \left[H_{{ 1, 1}\atop{ 2, 1}}\right] & \left[H_{{ 1, 1}\atop{ 2, 2}}\right] \\  
\noalign{\medskip}
   \end{array} \right]
\left[\begin{array}{l} 
\noalign{\medskip}
		       \vec{I}^{sky}_{{ 0}\atop{ 0}} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{{ 0}\atop{ 1}}\\ 
\noalign{\medskip}
		       \vec{I}^{sky}_{{ 0}\atop{ 2}}\\ 
\noalign{\medskip}
%\hline
\noalign{\medskip}
		       \vec{I}^{sky}_{{ 1}\atop{ 0}} \\ 
\noalign{\medskip}
		       \vec{I}^{sky}_{{ 1}\atop{ 1}}\\ 
\noalign{\medskip}
		       \vec{I}^{sky}_{{ 1}\atop{ 2}}\\
\noalign{\medskip}
			       \end{array}\right] =
\left[\begin{array}{l} 
\noalign{\medskip}
                       \vec{I}^{dirty}_{{ 0}\atop{ 0}}  \\ 
\noalign{\medskip}
                       \vec{I}^{dirty}_{{ 0}\atop{ 1}} \\ 
\noalign{\medskip}
		       \vec{I}^{dirty}_{{ 0}\atop{ 2}}\\ 
\noalign{\medskip}
%\hline
\noalign{\medskip}
		       \vec{I}^{dirty}_{{ 1}\atop{ 0}} \\ 
\noalign{\medskip}
		       \vec{I}^{dirty}_{{ 1}\atop{ 1}} \\ 
\noalign{\medskip}
		       \vec{I}^{dirty}_{{ 1}\atop{ 2}} \\
\noalign{\medskip}
			       \end{array}\right] 
\label{Eq:msmfs_neqn_matrix}
\end{equation}

This is the system of equations to be solved to obtain estimates of the model parameters $I^{sky}_{p\atop q}$.
The spatial-frequency sampling  of a real interfetometer is always incomplete 
($[\Sa]$ is rank-deficient).
Therefore, each Hessian block, and the entire Hessian matrix is singular, an exact inverse
does not exist. An accurate reconstruction can be obtained only via successive approximation
(iterative numerical optimization).
%An approximate solution of the normal equations is performed in each minor cycle
%to choose flux components,

\subsection{Principal Solution}
An approximate solution of the normal equations can be computed via diagonal
approximations of all Hessian blocks.
%Even in the ideal case of complete sampling where it may not be singular, 
%its dimensions make a brute-force inversion and application intractable.
Each Hessian block is a convolution operator with a shifted version of 
$I^{psf}_{{s,p}\atop{t,q}}$ in each row, and 
elements on the diagonal within each Hessian block are the same. 
The Hessian matrix with a diagonal approximation for each block $\left[\He_{{s,p}\atop{t,q}}\right]$ 
can be written as one $N_tN_s \times N_tN_s$ element matrix ($[H^{peak}]$), 
that applied to all pixels independently.  An approximate solution of the normal equations
can be obtained by inverting $[H^{peak}]$ once, and applying it to all pixels of the dirty
images, one pixel at a time.

This process of doing a pixel-by-pixel inversion of the diagonal-approximate 
Hessian matrix gives the
principal solution\footnote{The principal solution (as defined in \citet{BRACEWELL_ROBERTS} and used in 
\citet{DECONV_LECTURE}) is a term specific to radio interferometry
and represents the dirty image normalized by the sum of weights.  
It is the image formed purely from the measured data, with no contribution from the
invisible distribution of images (unmeasured spatial frequencies).
It is also an approximate solution of the normal equations, calculated using a diagonal
approximation of the Hessian (elements on the diagonal are the sum of weights).
For isolated sources, the values measured at the peaks of the
principal solution images are the true sky values as represented in the 
image model.
} of the system. 
%For the simplest case of $N_s=1, N_r=1$, the principal solution is obtained
%by normalizing the Fourier-inverted weighted visibilities by the sum of weights (values on the
%diagonal of the Hessian matrix).
%
%
%
%The principal solution is found by using a diagonal approximation of each
Such a solution will be correct only at the locations of the centers
of flux components (source peaks), and must be augmented with an iterative optimization 
approach to ensure accuracy.  In the case of perfect sampling (where the Hessian blocks
are truly diagonal), the principal solution will directly give images
of series-coefficients.

%When the principal solution is to be used within an iterative joint deconvolution,
%a few simplifying assumptions may be needed to trim computational costs.
%For a source with complicated spatial structure the number of distinct spatial
%scale basis functions is typically $N_s \approx 10$, and for power-law spectra with
%indices around -1.0, $N_t =4$ or $N_t=5$ terms in the series are required to
%accurately model the power law with a polynomial (across a 2:1 bandwidth). 
%Therefore typically, $N_sN_t \approx 50$. Although the inversion of $[H^{peak}]$ 
%may be tractable, the computational cost of a $50\times 50$ matrix multiplication
%applied per pixel to a set of $10^6$ pixels over  a large number of
%iterations may be prohibitive in comparison to the numerical accuracy that this
%exact inversion provides. 
%

\subsubsection{Properties of $[H^{peak}]$}

\begin{enumerate}
\item The elements on the diagonal of $[H^{peak}]$ are a measure of the instrument's sensitivity 
to a flux component of unit total flux whose shape and spectrum is given by each of the $N_sN_t$
possible pairs of spatial and spectral basis functions.
\begin{eqnarray}
H^{peak}_{{s,p}\atop{t,q}} &=& mid\left\{\vec{I}^{psf}_{{s,p}\atop{t,q}}\right\} = tr\left[\sum_{\nu} \wntq [T_s\Snd\Wimn\Sna T_p] \right] \\
	& &\forall~ s,p ~\in ~\{0 ... N_s-1\}~~,~~t,q ~\in ~\{0 ... N_t-1\}\nonumber
\label{Eq:hpeak_msmfs}
\end{eqnarray}
\item The off-diagonal elements measure the orthogonality\footnote
{\label{FN:ortho}The following definition of orthogonality is used here. Two vectors are orthogonal
if their inner product is zero. The orthogonality of a pair of scale functions 
is measured by the integral of the product of their $uv$-taper functions.
To account for $uv$-coverage, this integral is weighted by the sampling function
(see Eqn.~\ref{Eq:hpeak}).
}
 between the various
basis functions, for the given $uv$-coverage and weighting scheme.
They measure the amount of overlap between basis
functions in the measurement domain. 
Smaller values indicate a more orthogonal set of basis functions,
and the instrument is better able to distinguish between the chosen spatial scales.

\item The condition number of this matrix (or of blocks within this matrix)
will indicate if the chosen set of basis functions and spatial-frequency coverage 
provide enough constraints to provide a stable solution, and 
can be used as a metric to choose a suitable basis set.
For a simple example, if a 3-term solution is attempted with data from only two distinct
frequencies, $[H^{peak}]$ will be singular.
Or, for some choice of multi-frequency $uv$-coverage, the visibilities
measured by the instrument for two different spatial scales may become hard to distinguish. 
Then, the cross-term element of $[H^{peak}]$ corresponding to
this combination could have a higher value, indicating that the two parameters are highly coupled,
and there is insufficient information in the data and sampling pattern to distinguish 
between the scales. A similar situation can arise to create ambiguity between 
spatial or spectral structure (an extreme example is multi-frequency measurements from only
one baseline.
%Figure XXX shows a 1D example of
%such a situation, where it can be difficult to 
%distinguish between an extended source with a flat spectrum and a compact source 
%whose flux falls with increasing frequency (steep spectrum).
%\item The convolution kernels from the first row of Hessian blocks
% ($s=0,t=0, p\in\{0,N_s-1\}, q\in\{0,N_t-1\}$) represent the
%instrument's response functions to a flux component of unit total flux whose shape is
%given by the $p^{th}$ scale basis function and whose spectrum is given
%by the $q^{th}$ Taylor function. 
\item In general, $[H^{peak}]$ will be a positive-definite 
symmetric matrix whose inverse can be easily computed {\it via} a 
Cholesky decomposition\footnote
{A Cholesky decomposition is a decomposition of a symmetric positive-definite matrix
into the product of a lower triangular matrix and its conjugate transpose. 
It is used in the solution of system of equations $[A]\vec{x}=\vec{b}$ where
$[A]$ is symmetric positive-definite. 
The normal equations of a linear least-squares problem are usually in this form. 
In our case, this linear least-squares problem corresponds to the 
representation of the sky brightness as a linear combination of basis functions
\citep{NR}.
%[$http://en.wikipedia.org/wiki/Cholesky\_decomposition$].
}.  
Also, the value of $N_s$ is usually $< 10 $, making the inversion 
of $[H^{peak}]$ tractable.

\item 
Several approximations can be made about the structure
of $[H^{peak}]$ to simplify its inversion, and it is 
important to understand the numerical implications of these trade-offs.
%
One is a block-diagonal approximation of the full 
Hessian ({\it i.e.} using only those blocks of the Hessian in
Eqn.~\ref{Eq:msmfs_neqn_matrix} for which $s=p$; top-left
and bottom-right quadrants).
This approximation ignores the cross-terms between spatial scales and assumes that
the scale basis functions are orthogonal. 
This is never true for a set of tapered truncated paraboloids, but
this approximation works because of the iterative $\chi^2$-minimization
process.
Now, a multi-frequency principal solution 
can be done separately on each remaining $N_t\times N_t$ block,  
one spatial scale at a time.
%The Hessian blocks that form each $[H^{peak}_s]$ are outlined
% in Eqn.~\ref{Eq:msmfs_neqn_matrix}.
%Note that solving the multi-frequency $[H^{peak}]$ for each scale
This automatically does a normalization across scales that corresponds to a diagonal
approximation of the multi-scale Hessian 
(see section \ref{Sec:diff_msclean} for 
alternate ways of computing the multi-scale solution).
%This diagonal approximation across scales  is never accurate because 
%a set of tapered truncated paraboloids cannot form an orthogonal basis set.
%However, this approximation works because an iterative $\chi^2$-minimization
%process tolerates inaccurate steps during each iteration.   
The update  step of the iterative deconvolution still needs to evaluate
the full LHS of the normal equations while subtracting out a flux component

\end{enumerate}

\subsection{MS-MFS algorithm}\label{Sec:MSMFS}

This section describes an iterative joint deconvolution process that 
solves the normal equations (Eqn.\ref{Eq:msmfs_neqn_1}) and produces a
set of $N_t$ series-coefficient images at $N_s$ different spatial scales.
The algorithm presented in this section is listed with more details
in Algorithm \vref{ALGO:MSMFS_1} and Algorithm \vref{ALGO:MSMFS_2})

%\subsubsection{Pre-compute Hessian}
\noindent\paragraph{\bf Pre-compute Hessian : }
Convolution kernels for all distinct blocks in the $N_s N_t \times N_s N_t$ Hessian 
are evaluated {\it via} Eqn.~\ref{Eq:msmfs_neqn_2.5}. 
All kernels are normalized by $w_{sum}$ such that the peak of $\vec{I}^{psf}_{{0,0}\atop{0,0}}$
is unity, and the relative weights between Hessian blocks is preserved. 
%This is equivalent to defining the weight image $\vec{I}^{wt}$ as the diagonal of
%the $[H_{{00}\atop{00}}]$ Hessian block, and normalizing all the RHS vectors by it. 
A set of $N_s$ matrices each of shape $N_t\times N_t$ and denoted as $[H^{peak}_s]$
are constructed from the diagonal blocks of the full Hessian (blocks for which $s=p$ 
in Eqn.~\ref{Eq:msmfs_neqn_matrix}).
Their inverses are computed and stored in $[{H^{peak}_s}^{-1}]$.

%\subsubsection{Iterative deconvolution}

\noindent\paragraph{\bf Initialization :} 
All $N_sN_t$ model images are initialized to zero (or an {\it a priori} model).

\noindent\paragraph{\bf Major and minor cycles : }
The normal equations are solved iteratively by repeating steps 
\ref{stepRHS_msmfs} to \ref{stepPredict_msmfs} until some termination criterion is reached.
Steps \ref{stepRHS_msmfs} and \ref{stepPredict_msmfs} form one major cycle, 
and repetitions of Steps \ref{stepFind_msmfs} to \ref{stepUpdate_2_msmfs} form the minor cycle.
\begin{enumerate}
%\item {\bf Initialization :} 
%All $N_sN_t$ model images are initialized to zero (or an {\it a priori} model).
\item\label{stepRHS_msmfs} {\bf Compute residual images :} The RHS vectors 
(residual or dirty images) $\vec{I}^{dirty}_{{s}\atop{t}} ~\forall~t\in\{0,N_t$-$1\}$
of the normal equations are computed {\it via} Eqn.~\ref{Eq:msmfs_neqn_3} by first
computing the multi-frequency dirty images and then convolving them by the
scale basis functions.

\item\label{stepFind_msmfs} {\bf Find a Flux Component :}
The principal solution is computed for all pixels, one scale at a time.
\begin{equation}
I^{pix,psol}_s = [{H^{peak}_s}^{-1}] I^{pix,dirty}_s ~~~~~~~\mathrm{for~each~pixel,~and~scale} ~s
\label{Eq:msmfs_psol}
\end{equation}
Here,  $[H^{peak}_s]$ is the $s^{th}$ block (of size $N_t\times N_t$)
in the list of diagonal-blocks of $[H^{peak}]$, 
and $I^{pix,dirty}_s$ is the $N_t\times 1$ vector constructed 
from $\vec{I}^{dirty}_{{s}\atop{t}} ~\forall~t\in\{0,N_t$-$1\}$.

The principal solution consists of $N_s$ sets of 
$N_t$ Taylor-coefficient images.
For iteration $i$, the $N_t$ element solution set 
with the dominant $q=0$ component across all scales and pixel locations
is chosen the current flux component. Other heuristics may also be employed
to make this choice, for example, pick the set of components that makes the
largest impact on the value of $\chi^2$. 
Let the scale size of this chosen subset be $p$.

The result of this step is a set of $N_t$ model images, each containing one 
$\delta$-function that marks the location of the center of a flux component
of shape $\I^{shp}_{p,(i)}$. The amplitudes of these $N_t$ $\delta$-functions
are the Taylor coefficients that model the spectrum of the total flux of 
this component. Let these model images be denoted as 
$\left\{\vec{I}^{model}_{{{p}\atop{q}},(i)}\right\};q\in[0,N_t]$.

\item\label{stepUpdate_msmfs} {\bf Update model images :}
A single multi-scale model image is accumulated for each Taylor coefficient.
\begin{equation}
\I^{model}_{q} = \I^{model}_{q} + g \left( \I^{model}_{{{p}\atop{q}},(i)} \star \I^{shp}_{p} \right) ~~~~~~~~ \forall q\in[0,N_t]
\label{Eq:msmfs_updatemodel}
\end{equation}
where $g$ is a loop-gain that takes on values between 0 and 1 and controls the 
step size for each iteration in the $\chi^2$-minimization process.

\item\label{stepUpdate_2_msmfs} {\bf Update RHS :}
The RHS residual images are updated by evaluating and subtracting out the entire
LHS of the normal equations. Since the chosen flux component corresponds
to just one scale, the evaluation of the LHS is a summation over only Taylor terms.
\begin{equation}
\I^{res}_{{s}\atop{t}} = \I^{res}_{{s}\atop{t}} - g\left( \sum_{q_i=0}^{N_t-1} \left[  \I^{psf}_{{s,p}\atop{t,q}}  \star \I^{model}_{{{p}\atop{q}},(i)} \right] \right)
\label{Eq:msmfs_updaterhs}
\end{equation}
{\bf Repeat from Step \ref{stepFind_msmfs}} until the minor-cycle flux limit is reached.

\item\label{stepPredict_msmfs} {\bf Predict }:
Model visibilities are computed from each Taylor-coefficient image, 
in the same way as in Eqn.~\ref{Eq:mfs_predict} for multi-frequency imaging.
Residual visibilities are computed as $\vec{V}^{res}_{\nu}= \vec{V}^{corr}_{\nu} - \vec{V}^{model}_{\nu}$.

{\bf Repeat from Step \ref{stepRHS_msmfs}} until a global convergence criterion is satisfied.
\end{enumerate}
\enlargethispage{\baselineskip}

\noindent\paragraph{\bf Restoration :}
After convergence, the model spectral coefficient images can be interpreted in
different ways. If required, the final image products can be smoothed
with the restoring beam and the residuals are added back in. 
%Some forms applicable for radio astronomy are described below as additional
%operations that need to be performed on the model images. 
\begin{enumerate}
\item The most obvious data products are the spectral-coefficient images themselves, 
which can be directly smoothed by the restoring beam. The residual images that are added
back in should be the principal solution computed from the final residuals, to
ensure that any undeconvolved flux has the correct flux values.
\item  For the study of broad-band radio emission, the spectral coefficients can be
interpreted in terms of a power law in frequency with varying index
(as described in Section \ref{Sec:freqmodel}).
The data products are images of the reference-frequency flux $\vec{I}^{sky}_{\nu_0}$,
the spectral-index $\vec{I}^{\alpha}$ and the spectral curvature $\vec{I}^{\beta}$.
%Section \ref{Sec:ErrPowerLaw} describes the errors involved in this calculation.
\begin{eqnarray}
\label{Eq:calcab_1}
\vec{I}^{sky}_{\nu_0} &=& \vec{I}^{model}_0 \\
\label{Eq:calcab_2}
\vec{I}^{\alpha} &=& {\vec{I}^{model}_1}/{\vec{I}^{model}_0}  \\
\label{Eq:calcab_3}
\vec{I}^{\beta} &=& \left[{\vec{I}^{model}_2}/{\vec{I}^{model}_0}\right] - \left[{{\vec{I}^{\alpha}(\vec{I}^{\alpha}-1)}}{/2}\right]
\end{eqnarray}
Spectral index and curvature images can be calculated only in regions where
the values in $\vec{I}^{model}_0$ are above a chosen threshold.
%In this case, it is appropriate to smooth the final $\vec{I}^{sky}_{\nu_0}$ image
%with a restoring beam, but not the spectral index or curvature images. 
\item An image cube can be constructed by evaluating the spectral polynomial
{\it via} Eqn.~\ref{Eq:tfunc_1} for each frequency. This form of data product is
useful for sources whose emission is not well modeled by a power law,
but is a smooth polynomial in frequency. Band-limited signals that taper off
smoothly in frequency are one example.
\item An image of the continuum flux can be constructed by evaluating and adding up
the flux at all frequencies. This continuum image is different from the 
reference-frequency image which represents the flux measured at only one frequency.
\end{enumerate}

\noindent \paragraph{\bf Primary-beam correction : } 
A correction for the average primary-beam and its frequency dependence can be
done as a post-deconvolution step. The primary-beams are first evaluated or measured 
as a function of frequency, and the frequency-dependence per pixel modeled by a
power-law or a polynomial (perferably the same spectral polynomial used for the image
reconstruction). Primary-beam correction can be done as follows.
%\begin{equation}
%\vec{I}^{new}_{\nu_0}=\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}, \vec{I}^{new}_{\alpha}=\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}, \vec{I}^{new}_{\beta}=\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}
%\end{equation}
\begin{eqnarray}
\vec{I}^{new}_{\nu_0}&=&\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}\\
 \vec{I}^{new}_{\alpha}&=&\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}\\
 \vec{I}^{new}_{\beta}&=&\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}
\end{eqnarray}
Note that if a polynomial is fit for the frequency-dependence of the primary beam, 
and $\vec{\Pb}_{\nu_0}, \vec{\Pb}_{\alpha}, \vec{\Pb}_{\beta}$ computed from it, the above 
operation is numerically identical to doing a polynomial division in terms of two sets of
coefficients (for $N_t<=3$). A brute-force polynomial-division using more series coefficients
will yield a more accurate solution.   Note however, that such a correction will not be
accurate if there are time-dependent variations in the primary-beam, and will require
integration with the AW-Projection algorithm discussed in \citep{AW-Projection}.

\subsubsection{Software Implementation : }
The MS-MFS algorithm described in section \ref{Sec:MSMFS} has been implemented and 
released {\it via} the CASA\footnote{\href{http://casa.nrao.edu}
{Common Astronomy Software Applications} is being developed at the
\href{http://www.nrao.edu}{National Radio Astronomy Observatory}} 
software package (version 2.4 onwards). The data products are $N_t$
spectral-coefficient images, a spectral index image, and a curvature images (if $N_t>2$).
The parameters that control the algorithm are (a) $\nu_0$ : a reference frequency chosen near the middle of the sampled
frequency range, about which the Taylor expansion is performed, 
(b) $N_t$ : the number of coefficients of the Taylor polynomial to solve for, and 
(c) $N_s$ and $\I^{shp}_s$ : a set of scale sizes in units of image pixels to use
for the multi-scale representation of the image.
The resulting wide-band image model can then be used within a standard 
self-calibration loop.

%\noindent{\bf Data Products : }
%The basic products of the MS-MFS algorithm are a set of 
%$N+1$ multi-scale coefficient images that describe the spectrum of the
%sky brightness at each pixel (coefficients of an $N^{th}$-order polynomial).
%%(see section \ref{Sec:mfs_model}). 
%The $0^{th}$-order coefficient image is the Stokes I intensity image at the reference
%frequency (not the continuum image defined as the
%integrated flux across the full sampled bandwidth). To create the continuum
%image, the polynomial has to be evaluated and summed over all frequency channels.
%Derived quantities such as the spectral index and spectral curvature are 
%computed from the coefficient images (see Eqns.~\ref{Eq:calcab_1} 
%to \ref{Eq:calcab_3}).
%
%\noindent{\bf Control parameters :} 
%The three main parameters that control the operation of the MS-MFS algorithm
%are (a) $\nu_0$ : a reference frequency chosen near the middle of the sampled
%frequency range, about which the Taylor expansion is performed, 
% (b) $N_t$ : the number of coefficients of the Taylor polynomial to solve for, and 
% (c) $N_s$ and $\I^{shp}_s$ : a set of scale sizes in units of image pixels to use
%for the multi-scale representation of the image.
%
%
%\noindent{\bf Wide-band self-calibration : }
%The broad-band flux model generated by the MS-MFS algorithm can be used within 
%a self-calibration loop in exactly the same manner as standard self-calibration.
%The purpose of such a self-calibration would be to 
%improve the accuracy of the calibration.
%

%\newpage
\begin{algorithm}[Ht!]
\label{ALGO:MSMFS_1}
  \SetLine
  \linesnumbered
  \dontprintsemicolon
%  \SetKwRepeat{Repeat}{Repeat}{Until}
%  \SetKwFor{ForEach}{ForEach}{Do}{End}
%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{shp}_s~\forall~s\in\{0,N_s$-$1\}, ~ [\Sna]~ \forall \nu$ }
%\KwResult{ $\vec{I}^{psf}_{{sp}\atop{tq}},[{H^{peak}_s}], \vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ }
  \KwData{calibrated visibilities : $\vec{V}^{corr}_{\nu}~~\forall \nu$}
  \KwData{$uv$-sampling function : $[\Sa_{\nu}]$}
  \KwData{image noise threshold and loop gain $\sigma_{thr}, g_s$}
  \KwData{scale basis functions : $\I^{shp}_s ~ \forall s\in\{0,N_s-1\}$}
  \KwResult{model coefficient images : $\I^{m}_{q}~ \forall q\in\{0,N_t-1\}$}
  \KwResult{spectral index and curvature : $\I^{m}_{\alpha},\I^{m}_{\beta}$}

  \vspace{0.5cm} 
%{Compute the dirty image $\vec{I}^{dirty}$ and psf $\vec{I}^{psf}$}\;
  \For{ $t \in \{0,N_t-1\}, q \in \{t,N_t-1\}$}
  {
        { Compute the spectral PSF } $\vec{I}^{psf}_{tq}$\;
        \For{ $s \in \{0,N_s-1\}, p \in \{s,N_s-1\}$}
	{
		{Compute the scale-spectral PSF} $\vec{I}^{psf}_{{sp}\atop{tq}} = \vec{I}^{shp}_s \star \vec{I}^{shp}_p \star \vec{I}^{psf}_{tq} $\;
	}
  }
  \For { $s \in \{0,N_s-1\}$}
  {
     Construct $[{H^{peak}_s}]$ from $mid(I^{psf}_{{s,s}\atop{t,q}})$ and compute $[{H^{peak}_s}^{-1}]$\;
  }
%%%%%%%%%%%%%%%%%%%%%%
%%%\vspace{0.5cm}
%%  \caption[MS-MFS CLEAN : Pre-Deconvolution Setup]
%%          {MS-MFS CLEAN : Pre-Deconvolution Setup}
%%\end{algorithm}
%%%\newpage
%%
%%\begin{algorithm}[H]
%%  \SetLine
%%  \linesnumbered
%%  \dontprintsemicolon
%%%%%%%%%%%%%%%%%%%%%%%  
%%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{shp}_s, \vec{I}^{psf}_{{sp}\atop{tq}},[{H^{peak}_s}]~~~\forall~~s \in \{0,N_s-1\}, p \in \{s,N_s-1\}, \sigma_{thr}$ }
%%\KwResult{$I^{m}_{q}~ \forall q \in \{0,N_t-1\}$}
%%%%%%%%%%%%%%%%%%%%%  
  Initialize the model $\vec{I}^{m}_t$ for all $t \in \{0,N_t-1\}$ and compute $f_{sidelobe}$ \;
  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{res}_0 < \sigma_{thr}$ }
  {
    \For{$t \in \{0,N_t$-$1\}$}
    {
      Compute the residual image $\vec{I}^{res}_t$ \;
      \For{$s \in \{0,N_s$-$1\}$}
      {
	      Compute $\vec{I}^{res}_{{s},{t}} = \vec{I}^{shp}_s \star \vec{I}^{res}_t$
      }
    }
    Calculate $f_{limit}$ from $\vec{I}^{res}_{0,0}$\;
    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{res}_{0,0} < f_{limit} $ } 
    {
%     Compute $I^{m}_q~\forall q\in \{0.N_t-1\}$ and update $\vec{I}^{res}_{{s},{t}}~\forall s,t$ (Algorithm \vref{ALGO:MSMFS_2})\;
     \For{$s \in \{0,N_s$-$1\}$}
     {
      \uIf{Peak of $\vec{I}^{res}_{s,0} > 10~\sigma_{thr} $}
      {
       \ForEach{pixel}
       {
          Construct $I^{rhs}_s$, an $N_t\times 1$ vector from $I^{res}_{s,t} ~~\forall ~ t \in \{0,N_t$-$1\}$\;
          Compute principal solution $I^{sol}_s = [{H^{peak}_s}^{-1}] I^{rhs}_s$\;
       }
       Choose $I^{sol} = max\{I^{sol}_{t=0},~\forall~s\in\{0,N_s$-$1\}\}$ \;
      }
      \Else
      {
       Find the location of the peak in $\vec{I}^{res}_{s,0},~\forall~s\in\{0,N_s$-$1\}$\;
       Construct $I^{rhs}_s$, from $I^{res}_{s,t}$ for the chosen $s$, at this location\;
       Compute $I^{sol} = [{H^{peak}_s}^{-1}] I^{rhs}_s$ at this location\;
      }
     }
       \For{$t \in \{0,N_t-1\}$}
       {
        Update the model image : $I^{m}_t = I^{m}_t + g_s ~ I^{shp}_{s_i} \star I^{sol}_t$ \;
        \For{$s \in \{0,N_s$-$1\}$}
	{
          Update the residual image : $I^{res}_{s,t} = I^{res}_{s,t} - g ~\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1}[I^{psf}_{{sp}\atop{tq}} \star I^{sol}_q]$\;
	}
       }
    }
   Compute model visibilities $V^{m}_{\nu}$ from  $I^{m}_t~\forall t\in \{0.N_t-1\}$\;
   Compute a new residual image $I^{res}$ from residual visibilities $V^{corr}_{\nu}-V^{m}_{\nu}$\;
  }
Calculate $\vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ from $I^{m}_t~\forall t\in \{0.N_t-1\}$ and restore the results \;
If required, remove average primary beam : $\vec{I}^{new}_{\nu_0}=\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}, \vec{I}^{new}_{\alpha}=\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}, \vec{I}^{new}_{\beta}=\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}$\;
\vspace{0.5cm}

  \caption[MS-MFS Algorithm]
          {MS-MFS Algorithm : }
\end{algorithm}

%\newpage
%\begin{algorithm}[t!]
%\label{ALGO:MSMFS_2}
%  \SetLine
%  \linesnumbered
%  \dontprintsemicolon
%%%%%%%%%%%%%%%%%%%%%%  
%\KwData{ residual images : $\vec{I}^{res}_{s,t}$}
%\KwData{ scale basis functions : $\vec{I}^{shp}_s$}
%\KwData{ scale-Spectral PSFs : $\vec{I}^{psf}_{{sp}\atop{tq}}~~~\forall~~s \in \{0,N_s-1\}, p \in \{s,N_s-1\}$}
%\KwData{ Hessian for each scale : $[{H^{peak}_s}]~~~\forall~~s \in \{0,N_s-1\}$ }
%\KwResult{model coefficient images : $I^{m}_{q}~ \forall q \in \{0,N_t-1\}$}
%\KwResult{updated residual images : $I^{res}_{s,t}~ \forall s \in \{0,N_s-1\},~ t \in \{0,N_t-1\}$}
%\vspace{0.5cm}
%%%%%%%%%%%%%%%%%%%%%%  
%
%%    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{res}_{0,0} < f_{limit} $ } 
%%    {
%     \For{$s \in \{0,N_s$-$1\}$}
%     {
%      \uIf{Peak of $\vec{I}^{res}_{s,0} > 10~\sigma_{thr} $}
%      {
%       \ForEach{pixel}
%       {
%          Construct $I^{rhs}_s$, an $N_t\times 1$ vector from $I^{res}_{s,t} ~~\forall ~ t \in \{0,N_t$-$1\}$\;
%          Compute principal solution $I^{sol}_s = [{H^{peak}_s}^{-1}] I^{rhs}_s$\;
%       }
%       Choose $I^{sol} = max\{I^{sol}_{t=0},~\forall~s\in\{0,N_s$-$1\}\}$ \;
%      }
%      \Else
%      {
%       Find the location of the peak in $\vec{I}^{res}_{s,0},~\forall~s\in\{0,N_s$-$1\}$\;
%       Construct $I^{rhs}_s$, from $I^{res}_{s,t}$ for the chosen $s$, at this location\;
%       Compute $I^{sol} = [{H^{peak}_s}^{-1}] I^{rhs}_s$ at this location\;
%      }
%     }
%       \For{$t \in \{0,N_t-1\}$}
%       {
%        Update the model image : $I^{m}_t = I^{m}_t + g_s ~ I^{shp}_{s_i} \star I^{sol}_t$ \;
%        \For{$s \in \{0,N_s$-$1\}$}
%	{
%          Update the residual image : $I^{res}_{s,t} = I^{res}_{s,t} - g ~\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1}[I^{psf}_{{sp}\atop{tq}} \star I^{sol}_q]$\;
%	}
%       }
%%    }
%
%\vspace{0.5cm}
%  \caption[MF-MFS CLEAN : minor cycle steps]
%          {MF-MFS CLEAN : minor cycle steps}
%\end{algorithm}

\subsection{Relation to MF-Clean}
A point-source multi-frequency deconvolution algorithm can be derived by setting  $N_s=1$ and
using $I^{shp}_0 = \delta$-function in the derivations in sections \ref{Sec:meqn} and \ref{Sec:neqn}.
This section discusses the difference between the resulting multi-frequency Hessian and RHS vectors
and those described in the MF-Clean algorithm \citep{MFCLEAN}, and shows that the MF-Clean
approach will incur errors for arrays with dense and irregular spatial-frequency coverage where
tracks from different baselines intersect.

\noindent The normal equations can be written in block matrix form (for example, for $N_t=3$).
\begin{equation}
\left[\begin{array}{lll} 
   [H_{0,0}] & [H_{0,1}] & [H_{0,2}]\\  
\noalign{\medskip}
   [H_{1,0}] & [H_{1,1}] & [H_{1,2}] \\  
\noalign{\medskip}
   [H_{2,0}] & [H_{2,1}] & [H_{2,2}] \\  
   \end{array} \right]
\left[\begin{array}{l} \vec{I}^{sky}_{0} \\ 
\noalign{\medskip}
                       \vec{I}^{sky}_{1} \\ 
\noalign{\medskip}
		       \vec{I}^{sky}_{2}\end{array}\right] =
\left[\begin{array}{l} \vec{I}^{dirty}_{0}\\
\noalign{\medskip}
		       \vec{I}^{dirty}_{1} \\
\noalign{\medskip}
		       \vec{I}^{dirty}_{2}\end{array}\right] 
\label{Eq:mfs_neqn3}
\end{equation}
where each block $[\He_{t,q}]$ is a convolution operator with $\vec{I}^{psf}_{t,q}$ as its kernel.
\begin{eqnarray}
%\label{Eq:mfs_neqn_1}
%\sum_{q=0}^{N_t-1} [H_{t,q}] \vec{I}^{sky}_{q} &=& \vec{I}^{dirty}_{t} ~~~~~~~~~~~~~ \forall~ t ~\in ~\{0 ... N_t-1\} \\
%\label{Eq:mfs_neqn_2a}
% [H_{t,q}] &=& \sum_{\nu} \wntq [\He_{\nu}]\\
\label{Eq:mf_neqn_2b}
\vec{I}^{psf}_{t,q}  &=& \sum_{\nu} \wntq \vec{I}^{psf}_{\nu}\\
\label{Eq:mf_neqn_3a}
\vec{I}^{dirty}_{t}  &=& \sum_{\nu} \wnt \vec{I}^{dirty}_{\nu}
%\label{Eq:specPSF}
%\vec{I}^{psf}_{t,q}&=& \sum_{\nu} \wntq [\Fd\Snd\Wimn] \vec{1} = \sum_{\nu} \wntq \vec{I}^{psf}_{\nu}
\end{eqnarray}
%
The MF-Clean algorithm described in \citep{MFCLEAN} follows a
matched-filtering approach using functions called spectral-psfs, 
%\begin{equation}
%\vec{I}^{psf}_t = \sum_{\nu}\wnt \vec{I}^{psf}_{\nu}
%\label{Eq:mfclean_1}
%\end{equation}
which are equivalent to the convolution kernels from the first row
of Hessian blocks ($q=0$) in Eqn.\ref{Eq:mf_neqn_2b}.
%\item The convolution kernels from the first row of Hessian blocks
% ($s=0,t=0, p\in\{0,N_s-1\}, q\in\{0,N_t-1\}$) represent the
%instrument's response functions to a flux component of unit total flux whose shape is
%given by the $p^{th}$ scale basis function and whose spectrum is given
%by the $q^{th}$ Taylor function. 
Hessian elements and RHS vectors are calculated by convolving spectral-psfs with 
themselves and the residual images.
\begin{eqnarray}
\label{Eq:mfclean_2}
\vec{I}^{psf}_{t,q}  &=& \left\{ \sum_{\nu}\wnt \vec{I}^{psf}_{\nu}  \right\} \star \left\{ \sum_{\nu}\wnq \vec{I}^{psf}_{\nu}  \right\}\\
\label{Eq:mfclean_3}
\vec{I}^{dirty}_{t}  &=& \left\{ \sum_{\nu}\wnt \vec{I}^{psf}_{\nu}  \right\} \star \left\{\sum_{\nu} \vec{I}^{dirty}_{\nu}  \right\}
\end{eqnarray}
%Note the difference between Eqns \ref{Eq:mfclean_2} and \ref{Eq:mfclean_3}
% with Eqns.\ref{Eq:mf_neqn_2b} and \ref{Eq:mf_neqn_3a}.
Formally, this matched filtering approach is exactly equal to the
calculations shown in 
Eqns.\ref{Eq:mf_neqn_2b} and \ref{Eq:mf_neqn_3a} only under the conditions that there
is no overlap on the spatial frequency plane between measurements from different
observing frequencies, and all measurements are weighted equally across the spatial-frequency
plane (uniform weighting).
Appendix \ref{App:A} contains a derivation that demonstrates this.

In practice this difference manifests itself as follows.
Consider a spatial-frequency grid cell onto which measurements from
two different baselines and frequencies map. Let $V_1,V_2$ be the measured visibilities
at two frequencies $1,2$ and let $w_1,w_2$ be their Taylor-weights.
A matched-filtering approach 
calculates $(w_1+w_2)(V_1+V_2)$, whereas Eqns.\ref{Eq:mf_neqn_2b} and \ref{Eq:mf_neqn_3a}
require the computation of $(w_1 V_1) + (w_2 V_2)$.  The two are equivalent only for
flat spectrum sources where $V_1=V_2$ or when there is no overlap between the spatial
frequency grid cells measured 
from different observing frequencies ($V_1$ and $V_2$ map to different spatial frequencies).

The MF-Clean algorithm was initially developed for the ATCA telescope,
an East-West array of antennas with circular $uv$-coverage patterns and minimal
spatial-frequency overlap across channels. This matched filtering approach therefore
worked well.  However, when applied to data from the VLA (where $uv$-tracks intersect
each other and there is considerable spatial-frequency overlap), numerical
instabilities limited the fidelity of the final image, especially with
extended emission.  Changing the computations
to those in Eqns.\ref{Eq:mf_neqn_2b} and \ref{Eq:mf_neqn_3a}
eliminated this instability (determined using simulated VLA data).

%%%%%%%%%%%%%%%%%%%%%%

\subsection{Relation to MS-Clean}
A narrow-band (or flat-spectrum) multi-scale deconvolution algorithm can be derived by
setting $N_t=1$ in the derivations in sections \ref{Sec:meqn} and \ref{Sec:neqn}.
%The main difference between this method and those described in  \citep{MSCLEAN} and \citep{ERIC_MSCLEAN}
%is in the calculation of the multi-scale principal solution before searching for the dominant flux component. 
%

\noindent The normal equations can be written in block matrix form (for example, for $N_s=2$).
\begin{equation}
\left[\begin{array}{ll} 
\noalign{\medskip}
   [H_{0,0}] & [H_{0,1}] \\  
\noalign{\medskip}
   [H_{1,0}] & [H_{1,1}]  \\  
\noalign{\medskip}
   \end{array} \right]
\left[\begin{array}{l} \vec{I}^{sky,\delta}_{0} \\ 
\noalign{\medskip}
                       \vec{I}^{sky,\delta}_{1} \end{array}\right] =
\left[\begin{array}{l} \vec{I}^{dirty}_{0}\\
\noalign{\medskip}
		       \vec{I}^{dirty}_{1} \end{array}\right] 
\label{neqn_ms_math}
\end{equation}
%where each block $[\He_{s,p}]$ is a convolution operator with $\vec{I}^{psf}_{s,p}$ as its kernel.
%\begin{eqnarray}
%%\mathrm{where}~~~ [H_{s,p}] &=& [\Fd T_{s}\F][B][\Fd T_{p}\F] \\%~~~~~~~~\mathrm{where}~~~~~~ [B]=[\Fd\Sd\Wim\Sa\F]\\
%\vec{I}^{psf}_{s,p} &=& \vec{I}^{shp}_s \star \vec{I}^{psf} \star \vec{I}^{shp}_p \\
%		\vec{I}^{dirty}_{s}  &=& \vec{I}^{shp} \star \vec{I}^{dirty} %~~~~~~~~\mathrm{where}~~~~~~ \vec{I}^{dirty} = [\Fd\Sd\Wim] V^{corr}
%\label{Eq:ms_neqn_3}
%\end{eqnarray}
%$[B]$ is the Beam matrix ($[H]$ in Eqn.~\ref{Eq:Hessian}) 
%and $\I^{dirty}$ is the standard dirty image (Eqn.~\ref{Eq:Dirty}).
%
The elements of $[\He^{peak}]$, the $N_s\times N_s$ matrix formed from the multi-scale Hessian blocks
are given by 
\begin{equation}
H^{peak}_{s,p} = mid\left\{\vec{I}^{psf}_{s,p}\right\} = tr([T_s][\Sd\W\Sa][T_p])~~~~\forall~ s,p ~\in ~\{0 ... N_s-1\}
\label{Eq:hpeak}
\end{equation}
The elements on the diagonal of $[H^{peak}]$ correspond to $s=p$ and 
are a measure of the sensitivity of the instrument to a particular spatial scale. 
With uniform weighting, the kernels $\vec{I}^{psf}_{s,s}$ on the diagonal blocks are 
the autocorrelations of the different scale functions $\vec{I}^{shp}_s$, as measured by the
interferometer, and this gives the area under the main beam of the PSF for each spatial scale.
%The off-diagonal elements given by $s\ne p$
%are a measure of the orthogonality  of the basis set,
%for the given $uv$-coverage and weighting scheme.
The principal solution can be computed by inverting $[\He^{peak}]$ and applying it to the
RHS vectors.

There are two differences between this approach, and those described in \citet{MSCLEAN} and \citet{ERIC_MSCLEAN}.
\begin{enumerate}
\item {\bf Finding a flux component : }
In both forms of MS-Clean, 
the amplitude and scale of a
flux component are chosen by searching for the peak in the list of dirty
images after having applied a scale bias, 
an empirical term that de-emphasises large spatial scales.
The scale bias $b_s = 1-0.6~s/s_{max}$  used in \citet{MSCLEAN}
(where $s_{max}$ is the width of the largest scale basis function)
is a linear approximation of how the inverse of the area under each
scale function changes with scale size\footnote
{g
When $s/s_{max} = 1.0$ the bias term is $1.0 - 0.6 = 0.4$ which is approximately equal
 to the inverse of the area under a Gaussian of unit peak and 
 width, given by ${1.0}/{\sqrt{2\pi}} = 0.398$.}.
%The smoothed residual images at each scale are normalized by this estimate
%before peak values are chosen.
It is meant to be used to normalize residual images that have been smoothed with 
scale functions that have unit peak, before flux components are chosen.
The algorithm described in \citet{ERIC_MSCLEAN} uses 
$b_s \approx 1.0/s^{2x}$ where $x\in\{0.2,0.7\}$, to approximate
a normalization by the area under a Gaussian,
for the case when images are smoothed by applying a $uv$-taper 
that tends to unity for the zero spatial frequency.
%
Both these normalization schemes are 
approximations of using a diagonal approximation of $[H^{peak}]$ 
%and discarding all cross-terms 
when computing the principal solution before picking out flux components. 
%A diagonal approximation will however not give correct values
%for sources composed of overlapping flux components of different spatial scales.

Once we have this understanding, we can see that the full Hessian $[H^{peak}]$ 
(and not just a diagonal approximation)
can be inverted to get the normalization exactly right, especially for sources
that contain overlapping flux components of different spatial scales. 
%(i.e. taking into account
%the non-orthogonality between the different spatial scales).
It can be shown 
that by applying the inverse of the full $[H^{peak}]$ to the RHS vectors,
 we are able to 
get a more accurate estimate of the total-flux of the component at each scale 
than by just reading off a peak from a series of dirty images biased by the MS-Clean $b_s$. 
%This difference has been demonstrated on simulations (Section \ref{Sec:MS-pc})
%where the inverse of
%$[H^{peak}]$ was applied to all pixels of a series of smoothed dirty-images,
%but the relative performance of this approach (compared to the existing methods)
%is yet to be analysed within the complete iterative deconvolution
%framework. It is likely that the technique described
%in Section \ref{MS-deconv} would get more accurate minor cycle estimates 
%and therefore converge in fewer iterations.
However, this solution gives correct values only at the locations of the
centers of the flux-components, and introduces large errors in the psf sidelobes.
Therefore, for reasons of stability, a diagonal approximation is a more 
appropriate choice (demonstrated on simulated VLA data).

\item{\bf Minor cycle updates :}
The update steps in \citet{MSCLEAN} and Section \ref{MS-MFS}
evaluate the full LHS of the normal equations
to update the smoothed residual images and subtract out flux components 
within the image domain.
This allows each minor cycle iteration to search for the 
optimal flux component across all scales  
without having to recompute smoothed residual images by going to the visibility domain
in each iteration.
\citet{ERIC_MSCLEAN} ignores the cross-terms, performs
a full set of minor cycle iterations on one scale at a time, and recomputes
smoothed residual images {\it via} the visibility domain after every full set of
minor cycle iterations.
\end{enumerate}

\noindent A choice among these three methods
will depend on trade-offs between the accuracy within each minor cycle iteration, 
the computational cost per step, and optimized global convergence
patterns to control the total number of iterations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

\section{Hybrids of Narrow-Band and Continuum Techniques}

The preceeding sections discussed multi-frequency solutions that used data from all measured
frequencies together, so as to take advantage of the combined spatial-frequency coverage.
However, there are some situations where single-channel methods used in combination
with multi-frequency-synthesis (and no built-in spectral model) will be able to deliver
scientifically useful wide-band reconstructions at the continuum sensitivity.

%There are many hybrid approaches where the advantage of arbitrary spectral
%reconstructions allowed by single-channel imaging is combined with MFS
%below a certain flux level to make use of continuum sensitivity. 
%These approaches can be very effective for simple fields of point sources
% (ref TechRep/thesis.)

%\noindent{\bf Single-channel imaging : }

%

The basic idea of a hybrid wide-band method is to combine 
%narrow-band channel imaging
%with one or more of the continuum methods in a multi-stage approach. 
%The goal is to combine 
the advantages of single-channel imaging (simplicity and 
insensitivity to source spectra) with those of continuum imaging 
(deconvolution with full continuum sensitivity).
%
\begin{enumerate}
\item Deconvolve each channel separately upto the single-channel sensitivity limit.
At this stage, only sources brighter than $\sigma_{chan}$ will be detected and deconvolved.
\item Remove the contribution of bright (spectrally varying) sources by subtracting out
visibilities predicted from the model image cube.
At this stage, the peak residual brightness is at the level of the single-channel
noise limit $\sigma_{chan}$. 
\item Perform regular multi-frequency-synthesis imaging (flat-spectrum assumption)
on the continuum residuals to extract flux that lies between $\sigma_{chan}$ and $\sigma_{cont}$.
Accordiing to \citep{CCW_MFCLEAN}, and as shown in section XX, 
errors due to this flat-spectrum assumption become visible only above a dynamic
range of 1000 (for $\alpha=-1.0$ and a 2:1 bandwidth ratio). 
Therefore, as long as the sensitivity improvement between a single-channel
the full band is less than 1000, this second step will incur no errors even if the
remaining flux has spectral structure. 
This requirement translates to $N_{chan}<10^6$, which is usually satisfied\footnote
{Even if visibilities are measured at a very high frequency resolution, they can be
averaged across frequency ranges upto the bandwidth-smearing limit for the
desired image field-of-view. 
%For the EVLA in A-configuration between 1 and 2 GHz, this 
%corresponds to few-hundred channels for a single-primary-beam field-of-view.
}
\item Add model images from both steps, and restore the results. 
For unresolved sources, it may be appropriate to use a clean-beam fitted for
the highest frequency, but in general, to not bias spectral information, all channels
should be restored using a clean beam fitted to the lowest frequency in the range.
\end{enumerate}
The advantages of this approach are its simplicity, and that it can handle wide-band
reconstructions with band-limited signals and spectral-lines.
The disadvantages are that the angular resolution of the images and spectral information 
will be restricted to that given by the lowest frequency (a factor of two worse than what is
possible), and high-fidelity deconvolution of resolved structure
requires sufficient single-frequency spatial-frequency
coverage to unambiguously reconstruct all the spatial structure of interest, at all frequencies.

%
%However, all hybrid methods are 
%inherently single-channel methods that depend on having
%sufficient spatial-frequency coverage per channel to  reconstruct all
%the structure of interest. Also, any spectral information can be obtained
%only at the angular resolution of the lowest frequency.

%__________________________________________________________________

\section{Imaging results}\label{Sec:IMAGINGRESULTS}
\subsection{Simulation}

\paragraph{Objective : }
The goals of this test are to assess the ability of the MS-MFS algorithm 
to reconstruct both spatial and spectral 
information about a source in terms of a linear combination of compact and 
extended flux components with 
polynomial spectra (flux model described in section \ref{Sec:MSMFS-model})
as well as to test how appropriate this flux model is when the 
true sky brightness is a complex extended source whose spectral
characteristics vary smoothly across its surface.

\paragraph{Sky brightness : }
Wide-band EVLA observations were simulated for a sky brightness distribution 
consisting of one point source with spectral index of $-$2.0 and 
two overlapping Gaussians with spectral indices of $-$1.0 and $+$1.0. 
Fig.\ref{Fig:msmfs_sim_1} shows the reference
frequency image of this simulated source, 
plots of the spectrum at different locations on the source, and the resulting
spectral index and curvature maps. 
The spectral
index across the resulting extended source varies smoothly between $-$1.0 and $+$1.0,
with a spectral turnover in the central region corresponding to a spectral 
curvature of approximately 0.5. 
Fig.\ref{Fig:msmfs_sim_11} shows the first
three Taylor coefficient maps that describe this source.

\paragraph{MS-MFS Imaging : }
Two wide-band imaging runs were done using the MS-MFS algorithm and the results
compared. The first used a multi-scale flux model (section \ref{Sec:MSMFS-model}) 
in which $N_t=3$ and $N_s=4$ with scale sizes
defined by widths of $0,6,18,24$ pixels and the second used a point-source 
flux model in which $N_t=3$ and $N_s=1$ with one scale
function given by the $\delta$-function (to emulate the MF-CLEAN algorithm described 
in section \ref{Sec:mfs_model}).
A $5\sigma$ flux threshold of about $20\muup$Jy was used as the
termination criterion.

\paragraph{Results : }
The results from these imaging runs are shown in %Fig.\ref{Fig:msmfs_sim_2} 
Fig.~\ref{Fig:msmfs_sim_12} (three Taylor coefficients), 
Figure \ref{Fig:msmfs_sim_3} shows residual images 
over a larger region of the sky, and Fig.~\ref{Fig:msmfs_sim_2} shows the
intensity at the reference frequency, spectral index and spectral curvature.
All figures show the results with both MS-MFS and MF-CLEAN. 
%as images of 
%the intensity at the reference frequency (top), the 
%spectral index (middle) and spectral curvature (bottom).
%The left column shows the results with a multi-scale flux model 
%and the right column shows the results with a point-source 
%flux model.

%
\begin{figure}[t!]
%%\epsfig{figure=algomsmfs/msmfs_model.eps,scale=0.78}
\epsfig{figure=pics/fig_msmfs_1.eps,scale=0.3}
\caption[Example : Simulated wide-band sky brightness distribution]
{\small Simulated wide-band sky brightness distribution : 
These images represent the wide-band sky brightness distribution that was
used to simulate EVLA data to test the MS-MFS algorithm. 
%This true sky brightness is
%shown in terms of the final data products that the MS-MFS algorithm produces (total intensity
%spectral index and spectral curvature). in order to be able to evalua imaging
%performance. 
The image on the top left shows the
total intensity image of the source at the reference frequency $\I_{\nu_0}$. The plots
on the bottom left show spectra (and their power law parameters) at 4 different locations.
%different parts of the field extended source and for one isolated point source.
The spectral index varies smoothly between about $+$1 and $-$1 across the extended source 
and is $-$2.5 for the point source. 
The spectral curvature has significant values only in the central region of the extended source
where the spectrum turns over within the sampled range.
The images on the right show these trends in the form of spectral index (top) and
spectral curvature (bottom) maps. 
}
\label{Fig:msmfs_sim_1}
%\end{figure}
\vspace*{0.5cm}
%\begin{figure}
\epsfig{figure= pics/fig_msmfs_2.eps,scale=0.3}
\caption[Example : True Taylor coefficient images]
{\small True Taylor coefficient images :
These images show the first three Taylor coefficients for the polynomial expansion
of the wide-band flux distribution shown in Fig.~\ref{Fig:msmfs_sim_1}.
These images are the (left) intensity at the reference frequency $I_{0} = I_{\nu_0}$,
(middle) first-order Taylor-coefficient $I_1 = \alpha I_{\nu_0}$
and (right) second-order Taylor-coefficient 
$I_2 = \left({\alpha(\alpha-1)}/{2} + \beta \right) I_{\nu_0}$ 
(see Eqn.~\ref{Eq:taylorexpansion}). All images are displayed at the same flux scale.
}
\label{Fig:msmfs_sim_11}
\end{figure}

\begin{figure}
\epsfig{figure=pics/fig_msmfs_12.eps,scale=0.3}
\caption[Example : Reconstructed Taylor coefficient images]
{\small Reconstructed Taylor coefficient images :
These images show the first three Taylor coefficients (similar to Fig.~\ref{Fig:msmfs_sim_11})
obtained using two different wide-band flux models.
The top row shows the results of using a multi-scale wide-band 
flux model (MS-MFS) and the bottom row shows the results of using a point-source wide-band flux model
(MF-CLEAN, or MS-MFS with only one spatial scale given by a $\delta$-function).
All images are displayed at the same flux scale.
}
\label{Fig:msmfs_sim_12}
\end{figure}

\begin{figure}[t!]
\begin{center}
\epsfig{figure=pics/bigpic_multiscale_residual.eps,scale=0.15}
\epsfig{figure=pics/bigpic_pointsource_residual.eps,scale=0.15}
\end{center}
\caption[Example : Residual images]
{\small Residual images : This figure shows the residual images obtained after
applying MS-MFS to wide-band EVLA data simulated for the
sky brightness distribution shown in Fig.\ref{Fig:msmfs_sim_2}.
The residual image on the left is obtained when a multi-scale flux model was used (MS-MFS). 
The RMS noise on source is about 20 $\muup$Jy and off source is 5 $\muup$Jy.  
Compare this with the residual image on the right from a point-source deconvolution (MF-CLEAN) where
the on source RMS is about 0.2 mJy and off source is 50 $\muup$Jy. 
(Note that the displayed data ranges are different for these two images.
 The flux scale for the image on the left is $\pm0.3 \times 10^{-4}$ and 
 for the right is $\pm0.3\times 10^{-3}$.)
This clearly demonstrates the advantage of using a
multi-scale flux model.
}
\label{Fig:msmfs_sim_3}
\end{figure}

\begin{figure}[t!]
\begin{center}
\epsfig{figure=pics/fig_msmfs_3.eps,scale=0.3}
\end{center}
\caption[Example : MS-MFS final imaging data products]
{\small MS-MFS final imaging data products : 
These images show the results of applying MS-MFS to wide-band 
EVLA data simulated for the
sky brightness distribution described in Fig.\ref{Fig:msmfs_sim_1}.
The left column shows the results of using a multi-scale wide-band 
flux model (MS-MFS) and the right
column shows the results of using a point-source wide-band flux model
(MF-CLEAN, or MS-MFS with only one spatial scale given by a $\delta$-function). 
The top, middle and
bottom rows correspond to the intensity image at the reference frequency
$I_{\nu_0}$, the spectral index $\alpha$ and spectral curvature $\beta$ maps respectively.
The flux scale for each left/right pair of images is the same, and
the sharp source boundaries in the spectral index and curvature maps are because 
of a flux threshold used to compute them.
With a multi-scale flux model (MS-MFS, left), the reconstructions of $\alpha$
and $\beta$ are accurate to within 0.1 in high signal-to-noise regions.
With a point-source flux model (MF-CLEAN, right), deconvolution errors break
extended emission into flux components of the size of the resolution element
and these errors transfer non-linearly to the spectral index and curvature maps.
Table \ref{Tab:evlasim_err} compares the true and reconstructed values of $I_{\nu_0},\alpha,\beta$
for three regions of this sky brightness distribution.
}
\label{Fig:msmfs_sim_2}
\end{figure}

\clearpage
\noindent The main points to note from these images are listed below.
\begin{enumerate}
%\item Multi-scale vs point source : residuals and image dynamic range, fidelity, error bars on a,b.
\item With a multi-scale multi-frequency flux model (MS-MFS) 
the spectral index across the extended source was reconstructed 
to an accuracy of $\delta\alpha < 0.05$ with the maximum error being in the
central region where the spectral index goes to zero and $N_t=3$ is too high 
for an accurate fit (section \ref{Sec:Errs} describes how the choice of
$N_t$ affects the solution process). 
The spectral curvature across the extended source was estimated to an accuracy of
$\delta\beta<0.1$ in the central region with the maximum error of 
$\delta\beta \approx 0.2$ in the regions 
where the curvature signal goes to zero and the source surface brightness
is also minimum (the outer edges of the source).

\item With a multi-frequency point-source model (MF-CLEAN)
the accuracy of the spectral index and curvature maps was limited to
$\delta\alpha\approx 0.1,\delta\beta\approx 0.5$.
This is because the use of a point source model will break any 
extended emission into components the size of the resolution element
and this leads to deconvolution errors well above the
off-source noise level (note the difference between the
intensity images $I(\nu_0)$ produced with MS-MFS {\it vs} MF-CLEAN).
%a multiscale and point-source
%(right) model \cite{CORNWELL_CLEAN_1983,DANS_THESIS}). 
Error propagation during the computation of spectral index and 
curvature as ratios of these noisy reconstructed images leads to
high error levels in the result.

%\item For the isolated point source, both multi-frequency flux models were able to
%reconstruct the spectral index and curvature 
%to accuracies of $< 0.002$ {\tt CHECK!!!!!!!!!!!!!!!!}. 
%These numbers coincide with
%$\delta\alpha$ calculated for an image-domain error of 0.001. 
%
%Figure \ref{Fig:msmfs_sim_3} shows the residual images 
%over a larger region of the sky for the two test runs described above.
%(multi-scale image model (left) and a point-source model (right)).  
\item The imaging run that used a multi-scale image model was terminated at a $5\sigma$
noise threshold. The peak residual is about 20 $\muup$Jy and the off-source RMS is $5 \muup$Jy
(close to the theoretical RMS of 3 $\muup$Jy as listed in Table \ref{Tab:evlasim}).
The imaging run that used a point-source model was terminated after at least
four successive major cycles failed to reduce the peak residual below 200$\muup$Jy despite
an apparant decrease in the residuals during the minor cycle iterations.
The off source RMS in the result is about 50 $\muup$Jy.

\end{enumerate}

\subsection{Multi-frequency VLA observations of Cygnus-A}

\paragraph{Objective : }
Wide-band VLA observations of the bright radio galaxy Cygnus~A were used to
test the MS-MFS algorithm on real data as well as to test standard
calibration methods on wide-band data.
Most of the images so far made of Cygnus~A and
its spectral structure have been from large amounts of multi-configuration
narrow-band VLA data \cite{CYGA_1991} designed so as to measure the spatial structure as
completely as possible at two widely separated frequencies.
The goal of this test was to use multi-frequency
snapshot observations of Cygnus~A to evaluate how well the MS-MFS algorithm is 
able to simultaneously reconstruct its spatial and spectral structure
from measurements in which the single-frequency $uv$-coverage was insufficient
to accurately reconstruct all the spatial structure at that frequency.

\begin{figure}[t!]
\epsfig{figure=pics/cyga_uvcoverage.eps,scale=0.4}
\caption[VLA multi-frequency $uv$-coverage]
{\small VLA multi-frequency $uv$-coverage : 
	This figure shows the multi-frequency $uv$-coverage of VLA observations of Cygnus~A,
taken as a series of narrow-band snapshot observations. The plots on the left show the
$uv$-coverage from one frequency channel (20 snapshots at 1.7 GHz).
By zooming into the central region (bottom left) and comparing the spacing between the
measurements to the size of the $uv$ grid cells being used for imaging we can show that
the single-frequency measurements are incomplete. 
The plot on the right shows the multi-frequency $uv$-coverage using nine frequency tunings.
A zoom-in of the same central region (bottom right) shows that for the chosen $uv$ grid cell
size (or image field of view over which the image is to be reconstructed) the combined
sampling leaves no unmeasured grid cells. The imaging results from these observations will
test our ability to reconstruct both spatial and spectral information from
incomplete spatial frequency samples at a discrete set of frequencies.}
\label{Fig:cyga_uvcoverage}
\end{figure}

\paragraph{Cygnus~A } 
%Wide-band VLA observations of the bright radio galaxy Cygnus~A were used to
%test the MS-MFS algorithm on real data as well as to test standard
%calibration methods on wide-band data.
Cygnus~A an extremely bright (1000 Jy) radio galaxy with 
a pair of bright compact hotspots about 1 arcmin away from each other on either
side of a very compact core, 
and extended radio lobes associated with the hotspots that have  
broad-band synchrotron emission at multiple spatial scales. 
From many existing measurements \cite{CYGA_1996}, this radio source is
known to have a spatially varying spectral
index ranging from near zero at the core, -0.5 at the bright hotspots and up to
-1.0 or more in the radio lobes.

\paragraph{Observations : } Wide-band data were taken as described in 
Table \ref{Tab:vladata_cyg} using the
VLA 4-IF mode which allowed four simultaneous data streams containing
RR and LL correlations at two independent frequency tunings.
A set of 18 frequencies were chosen such that they spanned
the entire frequency range allowed by the new EVLA receivers (1--2 GHz).
Visibilities that used antennas with the older receivers were flagged for regions of 
the band not covered by the receivers (below 1.2 GHz and above 1.8 GHz). 
The $uv$-coverage for this dataset for the RR correlations is shown
in Fig.\ref{Fig:cyga_uvcoverage}.
The data were inspected visually and visibilities that were affected by 
strong radio frequency interference were flagged (masked).

\paragraph{Calibration :} 
Standard techniques were used to calibrate these data.
Flux calibration at each frequency was done {\it via} observations of 3C286.
Phase calibration was done using 
an existing narrow-band image of Cygnus~A at 1.4 GHz 
\cite{CYGA_1991} as a model.

At the time of these observations, the VLA correlator was getting inputs from
a combination of VLA and EVLA antennas. A gain control system that was temporarily
put in place to accomodate the use of new EVLA antennas with the VLA correlator
treated the two independent frequency tunings
in the 4-IF mode differently\footnote
{
To allow the use to new EVLA antennas with
the old VLA correlator, an automatic gain control had to be used at each EVLA antenna
to mimic the old VLA antennas and ensure that the input power levels to the VLA correlator 
were within the range over which it has a linear response. The type of gain control 
was being done differently for the two simultaneous
frequency tunings in the VLA 4-IF mode. The A/C IF stream used an automatic
gain controller based on power levels measured in 1 second and the B/D IF stream
used a static look-up table to decide attenuation levels.
This resulted in a difference in power levels
for the A/C and B/D data streams for all baselines that involved EVLA antennas
when the source being observed was bright enough to contribute to increasing the
overall system temperature.
}. This caused errors in the correlator input for very strong sources (Cygnus~A)
that increased the input power level beyond the linear power range of the VLA correlator.
%and as a result of this effect, 
Observations of the calibrator source 3C286 were not affected by this problem.
We were therefore able to calibrate all the frequency tunings for Cygnus~A and use 
the resulting wide-band spectrum along with the known integrated flux and spectral index of
Cygnus~A to idenfity which of the frequency tunings of Cygnus~A 
were affected. It was found that every alternate frequency (the second of each pair 
of simultaneous frequency tunings (B/D) in the VLA 4-IF mode) was affected.
Therefore to safely eliminate the effect of this problem for our tests, 
one of the two simultaneous frequency tunings were flagged from the recorded 
visibilities reducing the number of spectral windows from 18 to 9.
The final dataset used for imaging consisted of nine spectral windows
each of a width of about 4 MHz and separated by about 100 MHz.

\paragraph{Imaging : }
These data were imaged using two methods, the  
MS-MFS algorithm and a hybrid method consisting of STACK + MFS on residuals
(see section \ref{Sec:hybrids} for a description of this method).
Their results were compared
to evaluate the merits of the MS-MFS algorithm over the much simpler hybrid
method that used a combination of existing standard methods.
The data products evaluated were the total-intensity image, the continuum residual image
and the spectral index map.
The effect of the primary beam was ignored in these imaging runs because 
the angular size of Cygnus~A is about 2 arcmin, which at L-band is within
a few percent of the HPBW of the primary beam, 
a region where the antenna primary beam and its spectral effects can be ignored.

\begin{enumerate}
\item {\bf MS-MFS : }
The MS-MFS algorithm was run with a $2^{nd}$-order polynomial to model the source
spectrum and a set of 10 scale basis functions of different spatial scales
to model the spatial structure ($N_t=3, N_s=10$).
Iterations were terminated
using a 30 mJy stopping threshold. A theoretical continuum point-source 
sensitivity of 0.38 mJy was 
calculated for this dataset using an increased system temperature of $T_{sys}=250$
(due to the high total power of Cygnus~A).

\item {\bf Hybrid : }
The second approach was a hybrid algorithm
in which the MS-CLEAN algorithm was run separately on the data from each 
spectral window and then a single MS-CLEAN run was performed on the continuum
residuals (the STACK + MFS on residuals hybrid algorithm described in 
\ref{Sec:hybrids}). The total intensity image was constructed as an average of the
single channel image plus the result of the second stage on the continuum residuals.
This method is the same as that used in section \ref{Sec:hybrid} to test
the hybrid algorithm for the case of dense single-frequency $uv$-coverage. Note however
that the observations being described in this section do not have dense single-frequency
$uv$-coverage, and the purpose of applying this hybrid method is to emphasize the
errors that can occur if this method is used inappropriately.
\end{enumerate}

\paragraph{Results : }

Figure \ref{Fig:cyga_intres} shows the reconstructed total-intensity images (top row)
and  the residual images (bottom row) obtained from these two methods. 
Figure \ref{Fig:cyga_spx} shows the spectral maps constructed 
{\it via} the two methods described above as well as from existing images at 
1.4 and 4.8 GHz.

\begin{enumerate}
\item {\bf Intensity and Residuals :}  
Both methods gave a peak brightness of 77 Jy/beam at the hotspots 
and a peak brightness of about 400 mJy/beam for the
fainter extended parts of the halo. 
The residual images for both methods  
showed correlated residuals due to the use of a multi-scale
flux model composed of a discrete set of scales (small-scale correlated structure within 
the area covered by the source, but no visible large-scale deconvolution errors due to
missing large-scale flux).

The off-source noise level 
achieved in the continuum image with MS-MFS was about 25 mJy,
giving a maximum dynamic range of
about 3000. The peak on-source residuals were at the level of 30 mJy.
Further iterations did not reduce these
residuals, and the use of a higher-order polynomial $N_t>3$ introduced more errors in
the spectral index map (see section \ref{Sec:ErrPowerLaw} for a discussion about errors
on the spectral index as a function
of $N_t$ and the SNR of the measurements). 
%
The off source RMS reached by the hybrid method was about 30 mJy, with the peak residuals
in the region of the source of 50 mJy. Deeper imaging in either stage did not
reduce these residuals.

Note also that both methods were almost two orders of magnitude above the theoretical point-source
sensitivity shown in Table~\ref{Tab:vladata_cyg} (calculated for
an equivalent wide-band observation).
However, the achieved RMS levels were consistent with the best RMS levels previously achieved
with the VLA at 1.4 GHz for this particular source at 
L-band ($\sim$20 mJy, [Perley, R. (private communication)]).

%Therefore, for these images, the limiting error looks to be due to the
%multi-scale representation, and not residual spectral effects...

\item {\bf Spectral Index : }
The image on the top left is the result of the MS-MFS algorithm and shows 
spectral structure at multiple scales across the source.  
For comparison, the image at the bottom is a spectral-index map constructed
from existing narrow-band images at 1.4 and 4.8 GHz, each constructed from 
a combination of VLA A, B, C and D configuration data \cite{CYGA_1991}.
These two images (top-left and bottom) show a very similar spatial distribution
of spectral structure. This shows that despite having a comparatively small amount
of data (20 VLA snapshots at 9 frequencies) the use of an algorithm that models
the sky brightness distribution appropriately is able to extract the same information
from the data as standard methods applied to large amounts of data.
The estimated errors on the spectral index map are $<0.1$ for the brighter regions
of the source (near the hotspots) and $\ge 0.2$ for the fainter parts of the lobes
and the core.

The image on the top right shows the spectral index map constructed from a
spectral cube (a set of 9 single-channel images) containing the results of
running the MS-CLEAN algorithm separately on each frequency and then smoothing
the results down to the angular resolution at the lowest frequency in the range.
Note that the single-frequency observations consisted of 20 snapshots of Cygnus~A.
This $uv$-coverage is too sparse to have measured all the spatial structure
present in the source, and the non-uniqueness of the single-frequency reconstructions
caused the images at the different frequencies to differ from each
other enough to adversely affect the spectra derived from these images.

\item {\bf Spectral Curvature : }
Note that although Cygnus~A itself has more than sufficient signal-to-noise to measure
any spectral curvature, very low level deconvolution errors (3 orders of magnitude below
the bright 77 Jy/beam hotspot) dominate the region around the 
very bright hotspots and this is sufficient to destroy the spectral curvature images.
That is, the signal-to-error ratio of the higher-order coefficient images
is too low to measure a physically plausible curvature
term (corresponding to a change in $\alpha$ of $<0.2$ across 700 MHz at 1.4 GHz).
\end{enumerate}

\paragraph{Wide-band Self Calibration : }

A few tests were done to test whether a self-calibration process that used 
wide-band flux models would yield any improvement on the gain solutions or
imaging results.

Two sets of calibration solutions were computed and compared.
For the first set of solutions, several rounds of amplitude and phase self-calibration were run,
beginning with a point-source model 
and using the MS-MFS algorithm to iteratively build up a wide-band flux model.
Self-calibration was terminated after new gain solutions were indistinguishable
from that of the previous run.
The second set of solutions was found by using a single 
1.4 GHz model for amplitude and phase self-calibration (with gain amplitudes
normalized to unity to preserve the source spectrum). 
No significant difference was found
and the second set of solutions were chosen for imaging.

As an additional test, the final wide-band flux model generated {\it via} the
MS-MFS algorithm was used to predict model visibilities for a 
wide-band self-calibration step (amplitude and phase) to test
if this process yielded any different gain solutions.
Again, on these data, there was no noticeable improvement in the continuum residuals 
or on the stability of the spectral-index solution in low signal-to-noise regions.

This suggests that either the use of a common 1.4 GHz model image for 
all individual frequencies did not introduce much error, or that the residual
errors are dominated by the effects of multi-scale wide-band deconvolution
and the flux model assumed by the MS-MFS algorithm.  Further tests are required with
much simpler sky brightness distributions and real wide-band data, in order to clearly
ascertain when wide-band self-calibration will be required for high-dynamic range
imaging.

%In an unrelated test (done by SB), these data were used to test wide-band primary-beam
%correction {\it via} looking at Stokes V images with and without primary beam correction.
%A Stokes V signal of XXX was measured before correction - reduced to YYY afterwards.
%Cygnus~A hotspots are bright enough that a 1\% instrumental polarization so close
%to the center of the beam (1 arcmin out of a FWHM of 28 arcmin) is a 100mJy signal !!

\begin{figure}[t!]
\begin{center}
\epsfig{figure=pics/cyga_msmfs_stokesi.eps,scale=0.2}
\epsfig{figure=pics/cyga_hybrid_stokesi.eps,scale=0.2}
\epsfig{figure=pics/cyga_msmfs_residual.eps,scale=0.2}
\epsfig{figure=pics/cyga_hybrid_residual.eps,scale=0.2}
\end{center}
\caption[Cygnus~A : Intensity and residual images]
{\small Cygnus~A : Intensity and residual images : 
These images show the total intensity (top row) and residual images (bottom row)
obtained by applying two wide-band imaging methods to Cygnus~A data taken
as described in Table \ref{Tab:vladata_cyg}.
The images on the left are the result of the MS-MFS algorithm and those on 
the right are with the STACK + MFS hybrid in which MS-CLEAN was used for all the
deconvolutions (single-channel deconvolutions followed by 
second deconvolution on the continuum residuals. 
The total intensity images show no significant differences. 
Both residual images show correlated residuals of the type expected for 
the MS-CLEAN algorithm that uses a discrete set of scale sizes
(the error pattern obtained by choosing a nearby but not exact spatial scale
 for a flux component will be a ridge running along the edge of each flux component).
The peak and off source residuals for the MS-MFS algorithm are
30 mJy and 25 mJy and with the hybrid algorithm are 50 mJy and 30 mJy 
respectively, showing a very mild improvement in continuum sensitivity
with the MS-MFS algorithm.
}
\label{Fig:cyga_intres}
\end{figure}

\begin{figure}[t!]
\begin{center}
\epsfig{figure=pics/cyga_msmfs_spx.eps,scale=0.2}
\epsfig{figure=pics/cyga_hybrid_spx.eps,scale=0.2}
\end{center}
\begin{center}
\epsfig{figure=pics/cyga_ccarrili_spx.eps,scale=0.2}
\end{center}
\caption[Cygnus~A : Spectral Index image]
{\small Cygnus~A : Spectral Index image : 
These images show spectral index maps of Cygnus~A constructed {\it via}
the MS-MFS algorithm (top left) and the hybrid algorithm (top right)
applied to the data described in Table \ref{Tab:vladata_cyg}.
The image at the bottom is a spectral index map constructed from 
two narrow-band images at 1.4 and 4.8 GHz obtained
from VLA A,B,C and D configuration data at these two frequencies \cite{CYGA_1991}. 
The spatial structure seen in the MS-MFS spectral index image is very similar
to that seen in the bottom image. For comparison, the spectral index
map on the top-right clearly shows errors arising due to non-unique 
solutions at each separate frequency as well as smoothing to the
angular resolution at the lowest frequency.
}
\label{Fig:cyga_spx}
\end{figure}

\subsection{M87}

M87 is a bright (200 Jy) radio galaxy located at the center of the Virgo cluster.
The spatial distribution of broad-band synchrotron emission from this source 
consists of a bright central region (spanning a few arcmin) containing a 
flat-spectrum core, a jet (with known spectral index of $-$0.55) and two radio lobes
with steeper spectra ($-$0.5 $> \alpha > -$0.8) 
\citep{Rottmann_1996_AA,Owen_2000}.
This central region is 
surrounded by a large diffuse radio
halo (7 to 14 arcmin) with many bright narrow filaments ($\approx 10''\times 3'$).

Multi-frequency VLA data were taken in a similar way as described for
Cygnus-A, with a series of 10 snapshots at 16 different frequencies 
within the sensitivity range of the EVLA L-band receivers.

Fig.\ref{Fig:m87_lobes} shows the intensity, spectral
index and spectral curvature maps of the bright central region at an
angular resolution of 3 arcsec (C+B-configuration).
Fig.\ref{Fig:m87_spectrum} shows a plot of the spectrum formed from the 
integrated flux in the central bright region.

The peak brightness at the center of the final restored intensity image
was 15 Jy with an off-source RMS of 1.8 mJy and an
on-source RMS of about between 3 and 10 mJy.
The residual images show low-level correlated residuals at the location of 
the source but deconvolution errors are almost absent from the rest of the 
image, indicating that the best off-source RMS noise level for these data has
almost been reached.
The maximum dynamic range (ratio of peak brightness to off-source RMS)
is about 8000, with the on-source dynamic range (ratio of peak brightness to
on-source RMS) of about 1000.

The spectral index map\footnote
{The spectral index between two frequency bands $A$ and $B$ will be denoted 
as $\alpha_{AB}$. For example, the symbol $\alpha_{PL}$ corresponds to the
frequency range between P-band (327 MHz) and L-band (1.4 GHz), and
$\alpha_{LL}$ corresponds to two frequencies within L-band (here, 1.1 and 1.8 GHz).
A similar convention will be used for spectral curvature $\beta$.
} of the bright central region (at 3 arcsec resolution)
shows a near flat-spectrum core with $\alpha_{LL} = -0.25$,
a jet with $\alpha_{LL} = -0.5$ and lobes with $-0.6 > \alpha_{LL} > -0.7$.

These numbers show that in the bright central region and in the halo 
there is sufficient signal-to-noise to measure the spectral index but any 
realistic spectral curvature (for broad-band synchrotron emission)
is detectable only within the central bright region.

This bright central region had sufficient
($>$100) signal-to-noise to be able to detect spectral curvature. 
The third panel in Fig.\ref{Fig:m87_lobes} shows the spectral curvature
measured within this region.
Note that the error bars on the spectral curvature are at the same level as 
the measurement itself. Therefore, a reliable estimate can only be obtained as
an average over this entire bright region. The average curvature is
measured to be $\beta_{LL} = -0.5$ which corresponds to a change in $\alpha$ 
across L-band by
$\triangle \alpha = \beta \frac{\triangle\nu}{\nu_0} \approx -0.2$.

These numbers were compared with two-point spectral indices computed between
327 MHz (P-band), 1.4 GHz (L-band), and 4.8 GHz (C-band) from existing images
\cite{Owen_2000},[Owen, F. (private communication)]. 
Across the bright central region, $-0.36 > \alpha_{PL} > -0.45$ and 
$-0.5 > \alpha_{LC} > -0.7$. The measured values ($-0.5 > \alpha_{LL} > -0.7$
and $\triangle\alpha \approx 0.2$) are consistent with these
independent calculations.
%($-0.5 > \alpha_{LL} > -0.7$ and $\triangle\alpha \approx 0.2$.

The points in Fig.\ref{Fig:m87_spectrum} shows the integrated flux over the central
bright region of M87
(shown in $\log(I)$ vs $\log(\nu/\nu_0)$ space)
from the 16 single-spectral-window images. The curved line passing through these points
is the average spectrum that the MS-MFS algorithm automatically fit for this region.
It corresponds to $\alpha \approx -0.52$ and $\triangle \alpha \approx 0.2$ across
the source. The straight dashed lines correspond to constant spectral indices of $-$0.42 and
$-$0.62 and show that the change in $\alpha$ across the band is approximately 0.2 (as also
calculated from $\beta_{LL} = -0.5$ that the MS-MFS algorithm produced).
Note that the scatter seen on the points in the plot is at the 1\% level of the
values of the points (signal-to-noise of 100). Also evident from the plot is the
fact that the curvature signal is at a signal-to-noise ratio of 1.
These results show that a signal-to-noise of $>100$ is required to measure 
a change in spectral index of 0.2 across 700 MHz at 1.4 GHz.

\begin{figure}[t!]
\begin{center}
%\epsfig{figure=algomsmfs/m87_centre_stokesi.eps,scale=0.255}
%\epsfig{figure=algomsmfs/m87_centre_alpha.eps,scale=0.255}
%\epsfig{figure=algomsmfs/m87_centre_beta.eps,scale=0.255}
\epsfig{figure=pics/picM87_center_restored.eps,scale=0.15}
%\epsfig{figure=pics/picM87_center_residual.eps,scale=0.15}
\epsfig{figure=pics/picM87_center_alpha.eps,scale=0.15}
\epsfig{figure=pics/picM87_center_beta.eps,scale=0.15}
\end{center}
\caption[M87 core/jet/lobe : Intensity, Spectral index, Curvature]
{\small M87 core/jet/lobe : Intensity, Spectral index, Curvature : 
These images show 3-arcsec resolution maps of the central bright
region of M87 (core+jet and inner lobes), where the signal-to-noise was
sufficient for the MS-MFS algorithm to detect spectral curvature.
%The quantities displayed are the intensity at 1.5 GHz (left),
%the spectral index (middle) and the spectral curvature (right).
The quantities displayed are the intensity at 1.5 GHz (top left),
the residual image (top right),
%(peak on-source residual is 80 mJy/beam), 
the spectral index (bottom left) and the spectral curvature (bottom right).
The spectral index is near zero at the core, varies between $-$0.36 and
$-$0.6 along the jet and out into the lobes. The spectral curvature is
on average 0.5 which translates to $\triangle\alpha=0.2$ across L-band.
The peak of the source is 4.6 Jy, the on-source RMS is 40 mJy/beam 
and this gives an on-source signal-to-error ratio of about 100.
Note that the flux scale on the residual image (top right) is about 2 orders of magnitude
lower than the total-intensity image (top left).
}
\label{Fig:m87_lobes}
\end{figure}

\begin{figure}[t!]
\begin{center}
\epsfig{figure=pics/m87_specplot.eps,scale=0.3}
\end{center}
\caption[M87 core/jet/lobe : L-band spectrum]
{\small M87 core/jet/lobe : L-band spectrum : 
This plot shows the spectrum formed from the integrated flux within the central 
bright region between 1.1 and 1.8 GHz. The points are the integrated flux measured from single-spectral-window
model images, the curved line is the average spectrum that the MS-MFS algorithm
automatically fit to these data in this region. This spectrum corresponds to an average
$\alpha_{LL}=-0.52$ and a change of $\triangle\alpha\approx 0.2$ across the band (1.1 to 1.8 GHz).
The straight dashed lines represent pure power-law spectra with indices $-$0.42 and $-$0.62
and are another way of showing that the change in $\alpha$ across the band is about 0.2.
These numbers are consistent with two-point spectral indices computed between
327 MHz (P-band), 1.4 GHz (L-band), and 4.8 GHz (C-band) 
($-0.36 > \alpha_{PL} > -0.45$ and $-0.5 > \alpha_{LC} > -0.7$)
from existing images \cite{Owen_2000},[Owen, F. (private communication)]. 
}
\label{Fig:m87_spectrum}
\end{figure}

\section{MS-MFS error estimation and feasibility}\label{Sec:ErrorFeas}

\subsection{Dynamic-range vs $N_t$}

If continuum imaging is done with only MFS gridding and source spectra are ignored,
spectral structure will masquerade as spurious spatial structure. 
These errors will affect regions of the image both on-source and off-source
and their magnitudes depend on the available $uv$-coverage, the frequency range being covered,
the choice of reference frequency, and the intensity and spectral index of the source.
%
A rough rule of thumb for an EVLA-type $uv$-coverages (see section \ref{Sec:Err_order})
is that for a point source
of with spectral index $\alpha=-1.0$ measured between 1 and 2 GHz,
the peak error obtained if the spectrum is ignored is at a dynamic range of $<10^3$. 
Note that when all sources in the observed region of the sky have similar spectral indices, 
these errors can be reduced by dividing out an average spectral index
(one single number over the entire sky) from the visibilities before imaging them\footnote
{Note that such a division will reduce the signal-to-noise ratio of the higher-order
terms of the series (for the remaining spectral structure).
Therefore, although the removal of an average spectral index
could reduce the level of imaging artifacts obtained when source spectra are
ignored, the lower signal-to-noise ratio of the spectral signature 
could increase the error on the derived spectral index when MS-MFS is used. 
Note also that this point is not specific to the MS-MFS algorithm,
but is a general statement about how the accuracy of a fit depends on the SNR of the signal
being fitted.}.

%If this spectrum is fit by a straight line ($N_t=2$), then there still is an 
%approximation, and in addition to incurring an error on the value of $\alpha$, 
%errors will appear in the intensity image beyond a dynamic range of about $10^4$.

\subsubsection{3C286 field (EVLA)}

\begin{figure}[t!]
\begin{center}
\epsfig{figure=pics/pic1_3c286_nt1.eps,scale=0.2}
\epsfig{figure=pics/pic1_3c286_nt2.eps,scale=0.2}
\epsfig{figure=pics/pic1_3c286_nt3.eps,scale=0.2}
\epsfig{figure=pics/pic1_3c286_nt4.eps,scale=0.2}
\end{center}
\end{figure}

- frequency dependence of the primary-beam of the EVLA.....
- -1.4 at the HPBW (from SW paper and from EVLA sims).

- spectral indices of surrounding sources : xx yy zz..... corrected to xxxxx. Confirmed
by a second test observation pointed at those background sources.

\subsubsection{Error as a function of bandwidth ratio}

ADD FIGURE

ALSO REFER TO CCW formula.
Magnitude of diff Taylor terms :
Are they visible in image as errors (vs) SNR needed to solve.

This section shows an example of the errors obtained when the order of the polynomial
chosen for imaging is not sufficient to model the power-law spectrum of the
source. 
EVLA datasets (8 hour synthesis) were simulated for 5 different frequency ranges
around 2.0 GHz. The sky brightness distribution used for the simulation
was one point source whose flux is 1.0 Jy and spectral index is -1.0 with
no spectral curvature.
The bandwidth ratios\footnote
{There are two definitions of bandwidth ratio that are used in radio 
interferometry. One is the ratio of the highest to the lowest frequency
in the band, and is denoted as $\nu_{high}:\nu_{low}$. Another definition
is the ratio of the total bandwidth to the central frequency 
$(\nu_{high}-\nu_{low})/\nu_{mid}$ espressed as a percentage.
For example, the bandwidth ratio for
$\nu_{low}=1.0$ GHz, $\nu_{high}=2.0$ GHz is $2:1$ and $66\%$.
}for these 5 datasets were
100\%(3:1), 66\%(2:1), 50\%(1.67:1), 25\%(1.28:1), 10\%(1.1:1).

Figure.\ref{Fig:taylor_error} shows the measured peak residuals and absolute
measured errors on $I_{\nu_0},\alpha,\beta$ when these datasets were
imaged using multi-frequency deconvolution with 
$N_t=1$ to $N_t=7$ and a linear spectral basis (Eqn.~\ref{Eq:tfunc_1}). 
%
All these datasets were imaged using a maximum of 10 iterations, a 
loop-gain of 1.0, natural weighting and a flux threshold of $1.0 \mu Jy$. No noise
was added to these simulations (in order to isolate and measure 
numerical errors due to the spectral fits).
Peak residuals were measured over the entire $0^{th}$ order
residual image, and errors on $I_{\nu_0},\alpha,\beta$ were computed
at the location of the point source by taking differences with
the ideal values of $I_{\nu_0}=1.0,\alpha=-1.0,\beta=0.0$.

\noindent Noticeable trends from these plots are listed below.
\begin{enumerate}
\item All errors appear to decrease exponentially (linearly in log-space)
as a function of increasing order of the polynomial, and as a function
of decreasing total bandwidth. For very narrow bandwidths, the use of
high-order polynomials increases the error.
\item The peak residuals are much smaller than the error incurred on the 
peak source flux at the reference frequency $I_{\nu_0}$ and the
errors on $\alpha$ and $\beta$. 
\item As an example, for a 2:1 bandwidth ratio, 
a source with spectral index = -1.0, and $N_t=4$,
the achievable dynamic range (measured as the ratio of the peak flux to the 
off-source peak residual) is about $10^5$, the error on the peak flux
at the reference frequency is 1 part in $10^3$, and the absolute
errors on $\alpha$ anre $\beta$ are $~10^{-2}$ and $~10^{-1}$ respectively.
\end{enumerate}
\noindent Note that these trends are based on one simple example, 
and further
analysis is required to understand the source of these errors and
assess how they vary as a function 
of $\alpha$ and $\beta$.
\citet{MFCLEAN_CCW} suggest that for an $(N_t-1)$-order
polynomial, the peak residuals proportional to the product of
$\alpha$ and the peak sidelobe level of the next higher order
$N_t^{th}$ spectral PSF. However, the results of the above tests 
do not follow this rule for all bandwidth ratios.  
Further work is required to (a) understand these errors in terms of
signal-to-noise and in the presence of deconvolution errors and
(b) be able to predict limiting dynamic ranges and error-bars on
$\alpha$ and $\beta$.

Note that all the code implementations for this dissertation use the 
linear expansion given by Eqn.~\ref{Eq:tfunc_1} 
(a polynomial in $I ~ vs ~ (\nu-\nu_0)/\nu_0$ space) to model an 
arbitrary spectrum.
However, in the case of a power-law, a logarithmic expansion
given by Eqn.~\ref{Eq:tfunc_2} 
(a polynomial in $I ~ vs ~ \log(\nu/\nu_0)$ space)
might need fewer terms than the linear expansion to model a power-law spectrum
and yield better results.
\citet{MFCLEAN_CCW} state that the logarithmic expansion has better convergence
properties than the linear expansion when $\alpha << 1$, but
this is yet to be tested for arbitrary values of $\alpha$.
Further, for given values of $\alpha$ and $\beta$, the radius of 
convergence of each series expansion defines a maximum bandwidth 
that it can be used with.
Further work is required to
do a formal comparison between these two sets of spectral basis functions
and their convergence properties when applied to arbitrary spectral shapes.

\begin{figure}
\begin{center}
\epsfig{figure=pics/plot_point_errors.eps,scale=0.3}
\end{center}
\caption[Peak Residuals and Errors for MFS with different values of $N_t$]
{\small Peak Residuals and Errors for MFS with different values of $N_t$ :
These plots show the measured peak residuals (top left) and the errors
on $I_{\nu_0}$(top right), $\alpha$ (bottom left), and $\beta$ (bottom right)
when a point-source
of flux 1.0 Jy and $\alpha$=-1.0 was imaged using Taylor polynomials of different
orders ($N_t=1-7$) and a linear spectral basis (Eqn.~\ref{Eq:tfunc_2}). 
This simulation was done with EVLA $uv$-coverages
(for an 8 hour synthesis run) and 100\%,66\%,50\%,25\% and 10\% fractional 
bandwidths, with a reference frequency of 2.0 GHz. No noise was added to 
these simulations. All runs used a loop-gain of 1.0, used natural weighting,
and were terminated after either 10 iterations or a flux threshold of $1\mu Jy$.
The x-axis of all these plots show the value of $N_t$ used for the simulation.
Plots for $\alpha$ and $\beta$ begin from $N_t=2$ and $N_t=3$ respectively
because at least that many terms are required to calculate these derived
quantities.
Noticeable trends from these plots are
(a) The peak residuals decrease by about a factor of 8 with each increase
of 1 more polynomial coefficient. (b) The peak residuals are larger for 
larger fractional bandwidths. (c) The errors on $I_{\nu_0},\alpha,\beta$
are larger than the peak residuals, but they too decrease with increasing
$N_t$. For very narrow bandwidths, the use of a very high-order polynomial
increases the error.
For a 2:1 bandwidth ratio, a spectral index of -1.0 and very 
high signal-to-noise, a $5^{th}$ or $6^{th}$ order Taylor expansion
is most appropriate (when a linear spectral basis is used).
}
\label{Fig:taylor_error}
\end{figure}

\subsection{Error on Spectral Index}

\subsubsection{Propagation of deconvolution errors}
%Error bars on $\alpha$ can be estimated by 
%error propagation for the expression $\alpha = I_1/I_0$ using measured or estimated
%values for the error on $I_0$ and $I_1$ (see section \ref{Sec:error_prop}).
%For an isolated point source, ${\delta\I^{\alpha}}/{\I^{\alpha}} = \sqrt{\left( {\delta\I_1}/{\I_1} \right)^2 + \left( {\delta\I_0}/{\I_0} \right)^2}$.
Deconvolution errors contribute to the on-source error in the Taylor 
coefficient images, and these errors propagate to the spectral index map which  
is computed as a ratio of two coefficient images.
%and it is not just the stddev of the pixel values as measured from the image. 
Table~\ref{Tab:evlasim_err} lists the estimated and observed errors in spectral index
and curvature for a simulated example and shows that the deconvolution errors that
result when a point-source flux model is used to deconvolve extended emission, 
can increase the error bars on the spectral index and curvature by an order of magnitude.
%

\subsubsection{Overlapping sources}

\subsubsection{Errors as a function of SNR}
The accuracy to which $\alpha$ and $\beta$ can be determined also
depends on the noise per spectral data point, the number of sampled frequencies,
the total frequency range of the samples, and the number of 
spectral parameters $N_t$ in the fit. Section \ref{Sec:ErrPowerLaw} discusses empirically 
derived error bars for the spectral index based on these factors.
These results show that to measure $\alpha=-1.0$ with less than 100\% error bars, we
need a single-channel signal-to-noise ratio greater than about 6, and to measure
$\beta$ with less than 100\% error, we need a single-channel signal-to-noise of 
greater than about 100.  
%
Note that these numbers dictate the required accuracy of wide-band flux calibration.
%\end{enumerate}

REFER TO CCW BIAS : Error due to using too few Taylor terms to fit the exp with a poly.
CCW comment on a bias that occurs with a 2-term T-exp. This is just the use of insufficient terms of a polynomial to model an exponential.

%%%%%%%%%%%%%%%%%%%%%%
The errors on the polynomial coefficients and quantities derived from them
will depend on the number of measurements of the spectrum, 
the signal-to-noise ratio
of the measurements, and their distribution across a frequency range.
%\item 
They will also depend on the order of the polynomial used in the approximation.
Although the physical parameters $I_{\nu_0},\alpha$ and $\beta$ can be obtained from 
the first three coefficients of a Taylor expansion of a power-law with varying index
(Eqns.\ref{Eq:taylorexpansion} and \ref{Eq:series_ab}),
a higher order polynomial may be required during the fitting process to 
improve the accuracy of the first three coefficients\footnote
{\citet{MFCLEAN_CCW} comment on a bias that occurs with a 2-term Taylor expansion, due to the 
use of a polynomial of insufficient order to model an exponential.}.
%
%Empirical estimates for some of these errors are shown below.
In the case of very noisy spectra, 
errors can also arise from attempting to use too many terms in the polynomial fit.

\enlargethispage{\baselineskip}
Figure \ref{Fig:SpecError} illustrates the above trends for the value of $\alpha$
derived from a polynomial fit to a spectrum
constructed {\it via} Eqn.~\ref{EQN_POWERLAW1} 
($I^{true}_0=10.0,\alpha^{true}=-1.5,\beta^{true}=-0.5,\nu_0 = 2.4$GHz)
and evaluated between 1-4 GHz. 
Gaussian random noise was added to give measurement signal-to-noise
ratios of 100, 10 and 1 for three such spectra. 
These spectra were fitted using a linear least-squares method
on two series expansions (Eqns.\ref{Eq:taylorexpansion},\ref{Eq:series_ab})
for different numbers of
terms in the series $N=2,3,4,5$, and also by a non-linear least-squares method to
fit $\alpha$ and $\beta$ directly.
The plots show the error on the derived spectral index
$\delta\alpha=\alpha^{fitted}-\alpha^{true}$ for each case.
%
%%The top,middle and bottom rows correspond to data  with signal-to-noise
%%ratios of 100, 10, 1 respectively. 
%The LEFT column shows the error between the true and fitted values
%of $\alpha$, computed as $|\alpha_{fitted} - \alpha_{true}|$
%and averaged over an ensemble of 50 repetitions with different instances of noise.
%The four blocks represent polynomial approximations with $N_{t}=2,3,4,5$ terms
%in the series. The three bars in each block represent different spectral models 
%(Red : Fit for coefficients of a polynomial in $I$ vs $\dnuno$ space, 
%Blue : Fit for coefficients of a polynomial in $I$ vs $log\nuno$ space,
%Green : Direct fits to the power-law parameters $I_0,\alpha,\beta$).
%%For each case, the error (Y-axis) is computed as 
%%$Err = \sqrt{(\delta I_0/I_0)^2 + (\delta\alpha/\alpha)^2 + (\delta\beta/\beta)^2}$
%%and averaged over an ensemble of 50 repetitions with different instances of noise.
%The RIGHT column shows the corresponding spectra (Noisy Data and Spectral Fits) 
%for one instance of the noise, and for $N_{taylor}=3$.

\noindent Noticeable trends (based on $\delta\alpha$) are listed below.
\begin{enumerate}
\item For high SNR, higher order fits give better results.
For low SNR, higher order fits give larger errors.
\item In most cases, a Taylor expansion of a power law about $\alpha=0,\beta=0$ is a better 
choice than a Taylor expansion about $\nu=\nu_0$.
\item For spectra between 1 and 4 GHz with $\alpha \approx -1.5 \pm 0.4$, a 3rd or
4th order Taylor expansion (either form) is most appropriate.
\end{enumerate}
These trends can be used to choose the spectral basis function and number of terms $N_t$ 
to be used in the multi-frequency deconvolution algorithm (Section \ref{Sec:MFS-deconv}),
based on {\it a priori} knowledge of the average spectral index
and the signal-to-noise ratio of the measurements. 
When there are both high and low signal-to-noise sources, a multi-stage approach 
using different values of $N_t$ might be required.
For example, deconvolution runs can begin with $N_t>3$ but once the peak residual
reaches $10\sigma$, a switch to $N_t=2$ might be beneficial 
(note that this situation has not yet been tested).

\begin{figure}
\begin{center}
\epsfig{figure=pics/plot_aberror_snr100_N100_alpha.eps,scale=0.2}
\epsfig{figure=pics/plot_aberror_snr10_N100_alpha.eps,scale=0.2}
\epsfig{figure=pics/plot_aberror_snr1_N100_alpha.eps,scale=0.2}
\caption[Error Estimates for Spectral Index]
{\small These plots show the average error on the fitted spectral index
	($\delta\alpha = \alpha^{fitted}-\alpha^{true}$) from 
100 noisy measurements of a power-law spectrum defined by
$I^{true}_0=10.0,\alpha^{true}=-1.5,\beta^{true}=-0.5$.
The rows represent different signal-to-noise ratios (Top : 100, Middle : 10, Bottom : 1). 
The left column shows the average $\delta\alpha$ with $N_{t}$ = 2,3,4,5 terms in 
the series, for three different functional forms (Red/Left : $T(\nu=\nu_0)$ : Taylor expansion
of $I_{\nu}$ about $\nu_0$, Blue/Middle : $T(\alpha=0,\beta=0)$ : Taylor expansion of $I_{\nu}$ about
$\alpha=0,\beta=0$, Green/Right : Power Law with varying index).  
The right column shows the corresponding spectra for $N_{t}=3$.
Noticeable trends are 
(a) For high SNR, higher order fits give better results.
(b) For low SNR, higher order fits give larger errors.
(c) In most cases, a Taylor expansion about $\alpha=0,\beta=0$ is a better choice
than an expansion about $\nu_0$.
(d) For spectra between 1 and 4 GHz with $\alpha \approx -1.5 \pm 0.4$, a $3^{rd}$ or $4^{th}$ order
Taylor expansion is most appropriate.}
\label{Fig:SpecError}
\end{center}
\end{figure}

\subsection{Moderately resolved sources}

Consider a source with broad-band continuum emission and  
spatial structure that is either unresolved at all sampled frequencies
or unresolved at the low-frequency end of the band and resolved
at the high-frequency end. The intensity distribution as well as
the spectral index of such emission can be imaged
at the angular resolution allowed by the highest frequency in the band.
%This is true for the intensity image at the reference frequency as well as
%for the spectral index and curvature maps. 
%For sources with smooth broad-band spectra, it is possible to reconstruct spectral
%structure at this highest angular resolution.
This is because compact emission has a signature all across the spatial frequency plane
and its spectrum is well sampled by the measurements. The highest frequencies constrain
the spatial structure and the flux model (in which a spectrum is associated with each
flux component) naturally fits a spectrum at the angular resolution at which the spatial 
structure is modeled. 
%This reconstruction will be accurate for sources with
%smooth broad-band emission associated with each location on the source (as opposed to
Note that such a reconstruction is model-dependent and may require extra information
in order to distinguish between sources whose observed spectra are due to genuine 
changes in the shape of the source with frequency and those with broad-band (power-law)
emission emanating from each location on the source.
%Note that such a reconstruction is model-dependent and may not be accurate for 
%sources whose observed spectra are due to genuine changes in the shape of the source
%with frequency (as opposed to broad-band emission emanating from
%each location on the source).

\begin{figure}[t!]
\begin{center}
%%\epsfig{figure=feasibilit/mod_chanimages.eps,scale=0.8}
\epsfig{figure=pics/pic2_modres_chan0.eps,scale=0.15}
\epsfig{figure=pics/pic2_modres_chan1.eps,scale=0.15}
\epsfig{figure=pics/pic2_modres_chan2.eps,scale=0.15}
\epsfig{figure=pics/pic2_modres_chan3.eps,scale=0.15}
\epsfig{figure=pics/pic2_modres_chan4.eps,scale=0.15}
\epsfig{figure=pics/pic2_modres_chan5.eps,scale=0.15}
\end{center}
\caption[Moderately Resolved Sources : Single-Channel Images]
{\small Moderately Resolved Sources -- Single-Channel Images : 
These figures show the 6 single-channel images generated from simulated
EVLA data between 1 and 4 GHz in the EVLA D-configuration. 
The angular resolution at 1 GHz is 60 arcsec, and at 4 GHz is 15 arcsec
and the white circles in the lower left corner shows the resolution
element decreaseing in size as frequency increases.
The sky brightness consists of two point sources, each of flux 1.0 Jy at
a reference frequency of 2.5 GHz and separated by 18 arcsec.
%making it a moderately resolved source. 
The pixel size used in these images is 4.0 arcsec.
From these single-channel images we can see that the sources begin to 
be resolved only at the higher end of this frequency range, and at the
lower end of the band is barely distinguishable from a single point source
centered on the bottom point source. 
The top point source has a 
spectral index of $+$1.0  %(0.4 Jy to 1.5 Jy across the band) 
and the bottom one has a spectral index of $-$1.0.
%(2.5 Jy to 0.5 Jy across the band) 
%as shown in Fig.\ref{Fig:modres_spectra}.
}
\label{Fig:modres_channel}
\end{figure}
\paragraph{EVLA Simulation : }
Wide-band EVLA data were simulated for the D-configuration
across a frequency range of 3.0 GHz 
with 6 frequency channels between 1 and 4 GHz (600 MHz apart).
This wide
frequency range was chosen to emphasize the difference in angular resolution
at the two ends of the band (60 arcsec at 1 GHz, and 15 arcsec at 4.0 GHz).
The sky brightness chosen for this test consists of a pair of
point sources separated by a distance of 18 arcsec (about one resolution
element at the highest frequency), making this a moderately resolved
source. These point sources were given different spectral indices
($+$1.0 for the top source and $-$1.0 for the bottom one).
Figure \ref{Fig:modres_channel} shows the 6 single-channel images of this
source.
%and Fig.\ref{Fig:modres_spectra} shows the spectra of the two point 
%sources. 
At the low frequency end, the source is almost indistinguishable 
from a single flux component centered at the location of the bottom source
whose flux peaks at the low-frequency end. The source structure becomes
apparant only in the higher frequencies where the top source (with a positive
spectral index) is brighter.
%

%\enlargethispage{\baselineskip}
\paragraph{MS-MFS Imaging Results : }
\begin{enumerate}
\item These data were imaged using the MS-MFS algorithm with $N_t=3$ and 
$N_s=1$ with only one spatial scale (a $\delta$-function).
Figure \ref{Fig:modres_msmfs} shows the results of this imaging run.
The intensity distribution, spectral index and curvature of this source were
recovered at the angular resolution allowed by the 3.6 GHz samples
(18 arcsec). 
These results show that for a source that can be modeled as a set
of flux components (in this case point-sources) with polynomial spectra,
even partial spectral measurements at the highest angular resolution
are sufficient to reconstruct the full spectral structure.
%(for example a synchrotron source where the spectrum at 
%each location can be described by a power law).
%Section \ref{Sec:overlap_modres} later shows an example of imaging
%a moderately resolved source that also contains extended emission.

%Iterations were terminated using a stopping threshold of 1.0 mJy
%which was much higher than the theoretical because the goal of this
%test was to 
\item A second imaging run was performed using only 
the first and last channels (1.0 GHz and 4.0 GHz).
The source is almost completely 
unresolved at 1 GHz (point sources separated by 18 arcsec 
within a 60 arcsec resolution element), and just resolved at 4 GHz
(with an 15 arcsec resolution element).
The goal of this exercise was to test the limits of this algorithm
and the ability of the flux model to constrain the solution when
the data provide insufficient constraints.
The MS-MFS algorithm was run with $N_t=2$ and $N_s=1$ and used the
same number of iterations as the previous example.
Fig.\ref{Fig:modres_msmfs_ends} contains the resulting 
intensity image and spectral index map and shows that it
is still possible to resolve the source and measure its spectral
index at the resolution of the highest frequency. However, the
deconvolution errors are considerably higher.
The obtained peak residual of 5 mJy is not much larger than the
3 mJy level obtained when all 6 channels were used while imaging, 
indicating that this reconstruction is not well constrained by
the data and the model plays a very significant role.
\end{enumerate}

%\noindent These results show that for wide-band imaging in which
%visibilities are measured at a series of approximately regularly-spaced
%frequencies within a large range, the flux model used in the
%MS-MFS algorithm is able to constrain the solution well enough to
%reconstruct the spatial distribution of the spectral structure at
%the angular resolution allowed by the high-frequency measurements.

%\paragraph{Results : }

\begin{figure}[t!]
%\epsfig{figure=pics/mod_msmfsresults.eps,scale=0.8}
\begin{center}
\epsfig{figure=pics/pic2_modres_restored.eps,scale=0.2}
\epsfig{figure=pics/pic2_modres_residual.eps,scale=0.2}
\end{center}
\begin{center}
\epsfig{figure=pics/pic2_modres_alpha.eps,scale=0.2}
\epsfig{figure=pics/pic2_modres_beta.eps,scale=0.2}
\end{center}
\caption[Moderately Resolved Sources : MSMFS Images]
{\small Moderately Resolved Sources -- MSMFS Images : 
These images show the results of running MS-MFS
on EVLA data that was simulated to test the 
algorithm on moderately resolved sources. 
The test sky brightness distribution consists of two point sources 
with spectral indices +1.0 (North) and $-$1.0 (South) separated by one
resolution element at the highest frequency.
The four images shown here are the intensity at 2.5 GHz (top left),
the residual image with a peak residual of 3 mJy (top right),    
the spectral index showing a gradient between $-$1 and +1 (bottom left)
and the spectral curvature which peaks between the
two sources and falls off on either side (bottom right). These results
demonstrate that an appropriate flux model will constrain the solution to a physically
realistic one even when the spectral measurements are incomplete
at the highest resolution. 
}
\label{Fig:modres_msmfs}
\end{figure}

\subsection{Emission at large spatial scales}

%At the lower end of the sampled spatial-frequency range, the size of the central $uv$-hole
%increases with observing frequency. For very large spatial scales whose visibility
%functions are adequately sampled (more than 80\% of the integrated flux) only at the lower
%end of the frequency range, an ambiguity between spatial scale and spectrum can arise
%during the reconstruction. This is because the spectrum of this source is not well-sampled
%by the measurements. 
%A flat-spectrum extended source can be mistaken for a 
%steep spectrum less extended source, and {\it vice-versa}.  This problem can be 
%avoided by providing short-spacing flux constraints (from single-dish observations) 
%to bias the solution, or by flagging all 
%spatial frequencies below $u_{min}$ at $\nu_{max}$ (the smallest spatial frequency
%sampled by the highest observed frequency) to filter out these large spatial scales.
%

%The reconstruction of the visibility function at spatial frequencies
%within the central hole in the $uv$-coverage at each frequency
%involves an extrapolation of the measurements down to the zero $uv$-spacing. 
Consider a very large (extended) flat-spectrum source whose  
visibility function falls mainly within the
central hole in the $uv$-coverage at the highest observing frequency.
With multi-frequency measurements, the size of the central hole in the
$uv$-coverage increases with observing frequency,
and for this source the minimum spatial frequency sampled per 
channel will measure a decreasing peak flux level as frequency increases.
%the fraction of the total flux measured decreases 
%
%When a very large extended source with a flat spectrum is measured with a wide-band 
%interferometer, the minimum spatial frequency per channel
%measures a decreasing peak flux level as frequency increases.  
Since the reconstruction below the minimum spatial frequency
involves an extrapolation of the measurements % down to the zero $uv$-spacing
and is un-constrained by
the data, these decreasing peak visibility levels can be mistakenly interpreted as
the result of a source whose amplitude itself is decreasing with frequency
(a less-extended source with a steep spectrum).
%It can therefore be mistaken for a less-extended source with a steep spectrum.
Usually, a physically realistic flux model is used to 
apply constraints in these unsampled regions of the $uv$-plane
%For example, multi-scale imaging effectively
%fits the Fourier template of a scale function to measurements at the edge of the
%$uv$-hole to reconstruct it.  
and MS-MFS models the sky brightness with polynomial spectra associated
with a set of extended 2D symmetric flux components. 
However, with this model
a large flat-spectrum source and a smaller steep-spectrum source are both
allowed and considered equally probable.
This creates an ambiguity between the reconstructed scale and spectrum that 
cannot always be resolved directly from the data, and 
requires additional information (perhaps 
a low-frequency narrow-band image to constrain the spatial structure, 
low-resolution spectral information, or total-flux constraints).

\paragraph{EVLA Simulation : }
Wide-band EVLA data were simulated for the D-configuration
across a frequency range of 3.0 GHz centred at 2.5 GHz.
(6 frequency channels located 600 MHz apart between 1.0 and 4.0 GHz).
The size of the central hole in the $uv$-coverage was increased
by flagging all baselines shorter than 100 m 
and the wide frequency range was chosen to emphasize the difference
between the largest spatial scale measured at each frequency.
%in the size of the central hole in the $uv$-coverage.
%corresponding largest spatial scale measured at each frequency
%($ 0.8~k\lambda$ or $5~arcmin$ at 1.3 GHz, 
% and $1.3~k\lambda$ or $2.6~arcmin$at 3.9 GHz).
(0.3~k$\lambda$ or 10.3 arcmin at 1.0 GHz, 
 and 1.3~k$\lambda$ or 2.5 arcmin at 4.0 GHz).

The sky brightness chosen for this test consists of one large
flat-spectrum 2D Gaussian whose FWHM is 2.0 arcmin (corresponding
to 1.6~k$\lambda$ at the reference frequency of 2.5 GHz), and one
steep spectrum point-source ($\alpha$=-1.0) located on top of
this extended source at 30 arcsec away from its peak.
%Although the single extended source would suffice for the purpose of
%this test, the point-source is included as a control source.

\paragraph{MS-MFS Imaging Results : }
These data were imaged using the MS-MFS algorithm with $N_t=3$ and
$N_s=3$ with scale sizes given by [0,10,30] pixels. 
Two imaging runs were performed with these parameters and both
were terminated after 100 iterations in order to be able to
compare their performance in terms of the peak residuals.

\noindent Fig.~\ref{Fig:zero_visplot} shows the visibility amplitudes 
present in the simulated data (left column) as well as in the reconstructed model 
(right column) at each of the 6 frequencies for these two imaging runs (top,bottom).
Fig.~\ref{Fig:zero_images} shows images of the intensity, spectral
index and residuals for these runs and compares them to the true sky brightness
reconstructed when all frequencies sample at least 95\% of the total
flux of the source.

\begin{enumerate}
\item The first imaging run applied the MS-MFS algorithm to the 
simulated data after flagging all baselines below 200m. 
No additional constraints were used on the reconstruction.
The visibility plots and imaging results show that 
%
%Figure~\ref{Fig:zero_visplot} shows plots of the visibility amplitudes 
%in the simulated data (left) as well as in the reconstructed model 
%(right) at each of the 6 frequencies. 
%Fig.~\ref{Fig:zero_images} shows images of the intensity, spectral
%index and residuals and show that 
%
from these data it is not possible to distinguish large flat-spectrum 
source from a slightly less-extended steep spectrum source.
This occurs because the visibility function is unconstrained by the data
within the central $uv$ hole and 
%the spectrum is un-constrained by the data and 
given the MS-MFS flux model,
both source structures are equally probable.
Note that the spectrum of the point-source was correctly estimated as $-$1.0.
%
This run was repeated a few times with slightly different input scale
sizes, and the results changed between a flat-spectrum source and
a source with a steep spectrum. If a scale size corresponding
to the exact size of the source was present in the set, the algorithm
was able to reconstruct the correct flux and spectrum.

\item A second imaging run was performed on the same dataset, but
this time with additional information in the form of total-flux 
constraints at each observing frequency. These constraints were
added in by retaining a small number of very short-baseline 
measurements at each frequency in order to approximate the
presence of total-flux (or integrated flux) estimates 
(only baselines between 25 m and 100 m were flagged from the original
 EVLA D-configuration simulated data).
In practice, these constraints could be provided by single-dish 
measurements or estimates from existing low-resolution information
about the structure and spectrum of the source.
%
%Fig.~\ref{Fig:zero_dc_visplot} shows plots of the visibility amplitude
%used in the imaging run (left) and the corresponding reconstructed
%visibility function (right). Fig.~\ref{Fig:zero_dc_images} show
%the intensity, spectral index and residual images. 
%
The visibility plots and imaging results with this dataset 
show that the short-spacing flux estimates were sufficient to
bias the solution towards the correct solution in which the large
extended source has a flat spectrum and the point source has a
spectral index of $-$1.0.  Note that the residuals are at the same
level as in the previous run. This demonstrates that without 
the additional information about total-flux per frequency, both flux
models are equally poorly constrained by the data themselves.

\end{enumerate}

\noindent These results show that in the central unsampled region
of the $uv$-plane where there are no constraints from the data,
the MS-MFS flux model can produce ambiguous results and additional
information about the flux at low spatial-frequencies is required
(perhaps in the form of total-flux constraints per frequency).
%are required.  
%Note that this has been tested only for a simple
%spatial structure that can be fit with a single symmetric 2D Gaussian
%that the flux model supports. 
For complex spatial structure on these very large scales, the additional 
constraints may need to
come from existing low-resolution images of this field and the
associated spectra.
One way to avoid this problem altogether (but lose some information)
is to flag all spatial-frequencies smaller than $u_{min}$ at $\nu_{max}$
and not attempt to reconstruct any spatial scales larger than what
$\nu_{max}$ allows.

\begin{figure}[h!]
\begin{center}
\epsfig{figure=pics/plot_uvplot_space_data.eps,scale=0.2}
\epsfig{figure=pics/plot_uvplot_space_data_model.eps,scale=0.2}
\epsfig{figure=pics/plot_uvplot_withdc_data.eps,scale=0.2}
\epsfig{figure=pics/plot_uvplot_withdc_data_model.eps,scale=0.2}
\end{center}
\caption[Very Large Spatial Scales : Visibility plots]% ($>$100m)]
{\small Very Large Spatial Scales - Visibility plots : 
These plots show the observed (left) and 
 reconstructed (right) visibility functions for a
simulation in which a large extended flat-spectrum source is observed
with an interferometer with a large central hole in its $uv$-coverage.
%
The different colours/shades in these plot represent
6 frequency channels spread between 1 and 4 GHz.
These data were imaged in two runs. 
%
The first imaging run (top row) used only
baselines $b>$100 m to emphasize the changing size of the central
hole in the $uv$-coverage across the broad frequency range. 
The plot on the top left shows 
%the visibility function  dominated
%by the large flat-spectrum source and 
how the different frequencies measure 
very different fractions of the integrated flux of the large flat-spectrum source.
The plot on the right shows
that these data can be mistakenly fit using a less-extended source with a steep
spectrum (instead of the large single source with a flat spectrum).
This is possible because within the central $uv$ hole
the spectrum is un-constrained by the data and given the MS-MFS flux model,
both source structures are equally probable.
%
The second imaging run (bottom row) used baselines $b<$25m in addition to $b>$100m to
approximate the addition of nearly total-flux measurements to the first
dataset to attempt to constrain the solution.
The plot on the bottom right shows that this additional information in the
form of short-spacing
constraints (or very low-spatial frequency measurements) is sufficient to
be able to reconstruct the correct sky brightness distribution.
Figure \ref{Fig:zero_images} shows the images that resulted from these tests.
}
\label{Fig:zero_visplot}
\end{figure}

\begin{figure}[h!]
\begin{center}
\epsfig{figure=pics/piczero_restored_mfs.eps,scale=0.15}
\epsfig{figure=pics/piczero_alpha_mfs.eps,scale=0.15}
\epsfig{figure=pics/piczero_residual_mfs.eps,scale=0.15}
\epsfig{figure=pics/piczero_restored_zero.eps,scale=0.15}
\epsfig{figure=pics/piczero_alpha_zero.eps,scale=0.15}
\epsfig{figure=pics/piczero_residual_zero.eps,scale=0.15}
\epsfig{figure=pics/piczero_restored_dc.eps,scale=0.15}
\epsfig{figure=pics/piczero_alpha_dc.eps,scale=0.15}
\epsfig{figure=pics/piczero_residual_dc.eps,scale=0.15}
\end{center}
\caption[Very Large Spatial Scales : Intensity, Spectral Index, Residuals ]
{\small Very Large Spatial Scales - Intensity, Spectral Index, Residuals : 
These images show the intensity distribution (left), spectral index
(middle) and the residuals (right) for three different imaging runs that 
	applied the MS-MFS 
algorithm to the simulated EVLA D-configuration data described in this section
(note that the flux scale used for the residual images in the 
 right column is 3 orders of
 magnitude smaller than the scale used for the intensity image in the left column).
The true sky flux consists of one large flat-spectrum symmetric flux 
component and one steep-spectrum ($\alpha=-1.0$) point source. \\
Top Row : When all baselines are used for imaging,
each frequency samples more than 95\% of the integrated flux. This is
sufficient to reconstruct the true brightness distribution and spectrum.\\
Middle Row : When the central $uv$-hole is increased in size by using
only baseline $b>100m$, the reconstructed model is a slightly smaller flux
component (compare the left column of images) 
with a steep spectrum (compare the middle column of images).\\
Bottom Row : When very short spacing (approximately total-flux) estimates are
included during imaging (using spacings $b<25m$ and $b>100m$), the true sky
brightness distribution is again recovered.
Note that the large-scale residuals in
all three runs are at the same level (2 mJy).
These results show that the spectra are unconstrained by the data 
for very large spatial scales whose visibility functions
fall within the central $uv$-hole at the highest frequency in the band,
and additional information is required. 
}
\label{Fig:zero_images}
\end{figure}

\subsection{Band-limited signals}
spectral lines, continuum subtraction....

\subsection{Frequency dependence of the Primary Beam}

When wide-band imaging is done across wide fields-of-view, sources away from the
pointing center will be attenuated by the value of the primary beam at each
frequency. Wide-band imaging results from such data ignoring the primary beam
will contain spurious spectral structure.
For the EVLA primary beams between 1 and 2 GHz, this extra spectral index at the
half-power point is about -1.4 and about -0.6 at the 70\% point
(see Figs.\ref{PLPB2} and \ref{PLPB3}).
Note that even if the source has a flat spectrum, this artificial spectral index
can cause errors at the levels described for ignoring source spectra in the
restored intensity image.

%The accuracy with which the primary beam and its spectral variations can be
%corrected depends on several factors, the most influential being the fact that
%all image-domain corrections involve an approximate time-averaged primary beam. 
%Tests on 
%simulated and real data show that up to the 70\% point of the primary beam (at the
%reference frequency), the spectral index can be corrected to within 0.05 for point sources
%with signal-to-noise ratios of greater than 100, and to within 0.1 for point sources
%with signal-to-noise ratios of about 10. 
%For extended emission, the errors are dominated by the
%effects of multi-scale deconvolution errors and not primary-beam correction.
%On high signal-to-noise simulations (SNR$>$100) with extended sources located at 
%the 60\% point of the primary beam at the reference frequency, the spectral index 
%was recovered to within an error of 0.2.
%

%__________________________________________________________________

%__________________________________________________________________

\section{Discussion}\label{Sec:CONCLUSIONS}
\begin{enumerate}
\item { Summary : Can do this'n'that (image and spectrum reconstruction + astrophysical parameters)}
...at the angular resolution defined at the highest measured frequency.
\item Single-Channel Imaging Hybrid vs MS-MFS
\item Astrophysics : A variety of astrophysical observations could gain from these new instruments due to the 
increased sensitivity and large instantaneous bandwidth. Synchrotron spectra can be measured
within the instantaneous bandwidth, to detect/measure turnovers/breaks.
Snapshot imaging of highly variable sources 
(sunspots, supernovae, AGN) can be improved with the increased uv-coverage of mfs while
deriving the instantaneous spectra at multiple spatial scales.
Weak supernova remnants could be detected against a continuum background by their spectral signature.
Spectral information can be obtained at the highest angular resolution. This is especially 
significant for moderately resolved sources. 
The atmospheres of stars have an angular size of xxx and are resolvable at xxx GHz in the middle
of the EVLA frequency coverage. MFS with data between xx and yy GHz could yield wideband spectra
at the angular resolution defined by yy GHz.

Sunspot flares - magnetic loops - frequency probes different depths - band-limited signals.

VLBI position offsets as a function of frequency - can be resolved (defined) via MFS.

Multiple arrays can be combined by matching spatial frequency coverage across wide frequency ranges.
For example, EVLA-D between 5 and 50 GHz would match with EVLA-C in L and S bands... (as has always
been done, but it can be imaged together - not sure what final advantage this will bring though..
Snapshot imaging with EVLA subarrays could also achieve this.

\item Software (CASA and ASKAPSOFT), Cost and Parallelization (Hybrid vs MSMFS)
\item { Limitations }
- moderately resolved sources will need simultaneous deconvolution of multipel components (ASP)
- unresolved at lower freq, resolved at higher - 1D simulations show that it is possible to reconstruct, if at least XXX fraction of the bw has resolved info.
\item { Future Directions }
Out into the null of the PB (alternate parameterisations).
Full-Stokes imaging (ref SW)... parameterize differently in lm and freq but same idea.
\end{enumerate}

%__________________________________________________________________

\begin{acknowledgements}
The authors would like to thank the NRAO and NMT for support during
the PhD thesis project that resulted in this algorithm and implementation
in a usable software package. Many thanks to XX, YY, ZZ,... for useful 
discussions.
\end{acknowledgements}

\bibliographystyle{stylefiles/aa} % style aa.bst
\bibliography{msmfs_aa} % your references Yourfile.bib

\begin{appendix}

\subsubsection{Difference with SW-MF-Clean}

%\noindent{\bf Differences :}
\paragraph{Differences}
In the SW-MFCLEAN algorithm, the Hessian block kernels and dirty
images are computed {\it via} FFT-based convolutions in which 
gridded Taylor-weights
are multiplied with gridded visibilities : $(w_1+w_2)(V_1+V_2)$.
\begin{eqnarray}
\label{Eq:sw_psf}
\vec{I}^{psf,sw}_{t,q} &=& \vec{I}^{psf}_{t} \star \vec{I}^{psf}_{q} ~~~\mathrm{where}~~~~\vec{I}^{psf}_x = [\Fd\Sd\W^{mfs}_x\Wim] \vec{1} ~~\mathrm{for}~~x=t,q\\
\label{Eq:sw_dirty}
\vec{I}^{dirty,sw}_t &=& \vec{I}^{psf}_t \star \vec{I}^{dirty} ~~~\mathrm{where}~~~
\vec{I}^{dirty}=[\Fd\Sd\Wim] \vec{V}^{corr} 
\end{eqnarray}
According to Eqns.\ref{Eq:mfs_neqn_1} to \ref{Eq:mf_neqn_3} (MFCLEAN algorithm), 
the Hessian block kernels and dirty images are to be computed by 
multiplying the visibility measurments with the Taylor-weights
before gridding the result : $(w_1 V_1) + (w_2 V_2)$.
\begin{eqnarray}
%%\vec{I}^{psf}_{tq} &=& [\Fd\Sd\W^{mfs}_{t+q}\Wim] \vec{\bf 1}\\ 
\label{Eq:mfs_psf}
\vec{I}^{psf}_{t,q} &=& [\Fd\Sd\Wnt\Wnq\Wim] \vec{1}\\ 
\label{Eq:mfs_dirty}
\vec{I}^{dirty}_t &=& [\Fd\Sd\Wnt\Wim] \vec{V}^{corr}
\end{eqnarray}

%\noindent{\bf Conditions for equality :}
\paragraph{Conditions for equality}
The two methods listed above (Eqns.\ref{Eq:sw_psf},\ref{Eq:sw_dirty} and
Eqns.\ref{Eq:mfs_psf},\ref{Eq:mfs_dirty})
are equivalent only under certain conditions.
Consider Eqn.~\ref{Eq:mfs_neqn_2} for $[H_{t,q}]$. Iff $[\Sd] = [\Sd\Sa\Sd]$ and $[\Sa] = [\Sa\Sd\Sa]$, 
%(i.e. $[\Sd]$ is the 1-inverse\footnote
% {Definition of 1-inverse...
% }of $[\Sa]$ and vice-versa), 
then $[\Sd]$ can be
replaced by $[\Sd\Sa\Sd]$. Further, $[\Sa\Sd]$ and $[\Wntd]$ are both diagonal
matrices of size $nN_c\times nN_c$ and therefore commute.
In this case, $[H_{t,q}]$ becomes
\begin{eqnarray}
[H_{t,q}] &=& [\Fd\Sd\Wntd\Wim\Wnp\Sa\F]\\
	  &=& [\Fd\Sd\Wntd\Sa\F][\Fd\Sd\Wim\Sa\F] [\Fd\Sd\Wnp\Sa\F]\nonumber\\
\Rightarrow~~~~ \vec{I}^{psf}_{tq} &=& I^{psf}_t \star I^{psf} \star I^{psf}_q\\ 
& & \mathrm{where}~~~ I^{psf} = [\Fd\Sd\Wim]\vec{1} ~~~\mathrm{and}~~~ 
I^{psf}_t = [\Fd\Sd\Wnt]\vec{1} \nonumber
\end{eqnarray}
This is still not the same as Eqn.~\ref{Eq:sw_psf} which has two instances of $[\Wim]$.
Therefore, only when $[\Wim]$ is an identity matrix (equally weighted visibilities)
will the kernel functions  from both methods be identical 
$\vec{I}^{psf}_{t,q} = \vec{I}^{psf,sw}_{t,q}$.
A similar argument holds for the dirty images.
%
The restriction of $[\Sd] = [\Sd\Sa\Sd]$ and $[\Sa] = [\Sa\Sd\Sa]$ implies that 
each row and column in $[\Sa]$ has only one 1, with the rest being 0. Since $[\Sa]$ has dimensions
$nN_c\times m$, the maximum number of non-zero elements must be $m$.  
Therefore, any of the $m$ discrete spatial frequencies cannot be measured at more than one
baseline or frequency channel.  
However, consider the $m\times m$ diagonal matrix of gridded imaging weights
$[W^G_{\nu}] = [\Snd\Wimn\Sna]$ per frequency channel.
A projection operator $[\Sna^G]$ of shape ($m\times m$) can be constructed for each frequency
channel, with each diagonal element corresponding to one spatial frequency grid cell.
Measurements from multiple baselines that map onto the same spatial frequency grid cell are
treated as a single measurement in $[\Sna^G]$, with an increased weight in $[W^G_{\nu}]$.  
The use of uniform weighting will flatten out $[W^G_{\nu}]$ as required for equality with
Eqn.~\ref{Eq:sw_psf}. 
Written this way, with multiple frequencies, $[\Sna^G]$ has dimensions $mN_c \times m$, 
and the restriction of $[\Sd] = [\Sd\Sa\Sd]$ and $[\Sa] = [\Sa\Sd\Sa]$ means that any spatial 
frequency must not be measured in more than one frequency channel.
Therefore, a pure matched-filtering approach is strictly valid only for uniform 
weighting and when all filled spatial frequency grid cells contain 
measurements from only one frequency channel.

\end{appendix}

\end{document}

### 2016-04-19 aa/stylefiles/_region_.tex
More files needed for urvashi's equations

\message{ !name(msmfs_aa.tex)}%                                                                 aa.dem
% AA vers. 6.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                 (c) Springer-Verlag HD
%                                                revised by EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%
\documentclass[structabstract]{aa}  
%\documentclass[traditabstract]{aa} % for the abstract without structuration 
                                   % (traditional abstract) 
%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{epsfig}
\newcommand{\nuno}{{\left(\frac{\nu}{\nu_0}\right)}}
%%\newcommand{\nuno}{{({\nu}/{\nu_0})}}
\newcommand{\dnuno}{{\left(\frac{\nu-\nu_0}{\nu_0}\right)}}
%\newcommand{\dnuno}{{({\nu}/{\nu_0} - 1)}}
\newcommand{\F}{{F}}
\newcommand{\Fd}{{F^\dag}}
\newcommand{\E}{{E}}
\newcommand{\Ed}{{E^\dag}}
\newcommand{\G}{{G}}
\newcommand{\Gnu}{{G_{\nu}}}
\newcommand{\Gddnu}{{G_{\nu}^{dd}}}
\newcommand{\Gdddnu}{{{G_{\nu}^{dd}}^{\dag}}}
\newcommand{\Gd}{{G^\dag}}
\newcommand{\Gdnu}{{G_{\nu}^{\dag}}}
\newcommand{\Sa}{{S}}
\newcommand{\Sd}{{S^\dag}}
\newcommand{\Sna}{{S_{\nu}}}
\newcommand{\Snd}{{S_{\nu}^\dag}}
\newcommand{\Wim}{{W^{im}}}
\newcommand{\Wimd}{{{W^{im}}^\dag}}
\newcommand{\Wimn}{{W^{im}_{\nu}}}
\newcommand{\Wimnd}{{{W^{im}_{\nu}}^\dag}}
\newcommand{\wnt}{{w_{\nu}^t}}
\newcommand{\wnp}{{w_{\nu}^p}}
\newcommand{\wntp}{{w_{\nu}^{t+p}}}
\newcommand{\Wnt}{{W^{mfs}_t}}
\newcommand{\Wntd}{{{W^{mfs}_t}^\dag}}
\newcommand{\Wntn}{{w^{mfs}_{t,\nu}}}
\newcommand{\Wntnd}{{{w^{mfs}_{t,\nu}}^\dag}}
\newcommand{\Wnp}{{W^{mfs}_p}}
\newcommand{\Wnpd}{{{W^{mfs}_p}^\dag}}
\newcommand{\Wnpn}{{W^{mfs}_{p,\nu}}}
\newcommand{\Wnpnd}{{{W^{mfs}_{p,\nu}}^\dag}}
\newcommand{\Pnu}{{P_{\nu}}}
%\usepackage{amsmath}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{document}

\message{ !name(msmfs_aa.tex) !offset(-3) }

%\tableofcontents
%\newpage
%
   \title{A multi-scale multi-frequency deconvolution algorithm
for synthesis imaging in radio interferometry}

%   \subtitle{I. Overviewing the $\kappa$-mechanism}

   \author{U. Rau
          \inst{1}
          \and
          T.J. Cornwell \inst{2} %\fnmsep \thanks{Just to show the usage of the elements in the author field}
          }

   \institute{New Mexico Institute of Mining and Technology,Socorro, NM,USA\\
%              National Radio Astronomy Observatory, Socorro, NM, USA\\
%              Australia Telescope National Facility, CSIRO, Sydney, AU\\
              \email{rurvashi@aoc.nrao.edu}
         \and
             Australia Telescope National Facility, CSIRO, Sydney, AU \\
             \email{tim.cornwell@atnf.csiro.au}
%             \thanks{blah.}
             }

   \date{Received XXX XX, 2009; accepted XXX XX, 2009}
   %%\date{Received December 15, 2008; accepted XXX XX, 2009}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {The use of broad-band receivers in radio interferometery impacts standard continuum
    synthesis imaging in three ways. 
    First, correlations measured at different frequencies increase the uv-coverage of
    the synthesis array by sampling different spatial frequencies.
    Second, the sky brightness can change significantly across the large frequency 
    range that these receivers are sensitive to, violating a basic assumption of the
    Van Cittert Zernike theorem.  
    Third, the field-of-view of each array element is frequency-dependent, and this
    affects the spectral characteristics of the measured sky brightness during wide-field
    continuum imaging.
    Existing algorithms designed to account for some of the above, can be shown to be
    insufficient to achieve target image fidelity and dynamic range over the wide bands that 
    most new instruments are being built with.}
  % aims heading (mandatory)
   {This paper describes a multi-scale multi-frequency deconvolution algorithm for
    synthesis imaging that combines the information measured about the source spectrum
    with the additional uv-coverage offered by multi-frequency-synthesis to 
    reconstruct source structure and spectra at multiple spatial scales, %.
    while accounting for a frequency-dependent field-of-view.}
  % methods heading (mandatory)
   {The MS-MFS algorithm discussed in this paper extends the Sault-Wieringa MF-Clean 
    algorithm to reconstruct spectral curvature in addition to spectral index,
    and combines it with a  multi-scale deconvolution approach similar to that of MS-Clean.
    We also discuss a simpler hybrid of spectral-line and continuum imaging methods
    that may suffice in certain situations, and a more generic parametric approach.
    Frequency-dependent fields-of-view are corrected via existing wide-field imaging
    algorithms.
   }
  % results heading (mandatory)
   {We show via simulations and application to wideband (E)VLA data (1 GHz to 2 GHz),
    that it is possible to reconstruct both spatial and spectral structure of 
    compact and extended emission across wide fields of view.
    We also present an error analysis to show
    under what conditions these techniques are feasible and give accurate results. %(give numbers)
   }
  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{multi-frequency-synthesis --
                multi-scale-deconvolution --
                spectral index --
		spectral curvature --
               }

   \authorrunning{U.Rau and T.J.Cornwell}
   \titlerunning{Multi-Scale Multi-Frequency Synthesis Imaging in Radio Interferometry}
   \maketitle
%
\section{Introduction}

Instruments like the EVLA and e-MERLIN, are among a new 
generation of broad band radio interferometers designed to provide continuum sensitivities and
high dynamic range imaging capabilities superior to that of existing radio interferometers.
The wide bandwidths will allow detailed measurements of the spectral structure of astrophysical
sources, but to achieve the desired continuum sensitivities, imaging algorithms need to be 
sensitive to the effects of combining measurements from a large range of frequencies 
(namely varying ranges of sampled spatial scales, varying antenna fields-of-view, and
 the spectral structure of the sky brightness distribution).

%New algorithms are required to take full advantage of new broad-band 
%receivers for telescopes like the EVLA where broad-band sensitivity is the primary gain,
%as well as for arrays with fewer antennas like the VLBA, ATCA or Merlin where wide
%bandwidths are the key to measuring enough Fourier samples to reconstruct extended
%emission accurately.

Multi-Frequency-Synthesis (MFS) is the technique of combining measurements at multiple
discrete receiver frequencies during synthesis imaging.
MFS was initially done (REF) to increase the aperture-plane coverage
of sparse arrays by using narrow-band receivers and switching frequencies during the observations.
Wide bandwidth systems ( ~10\% ) presented the problem of bandwidth smearing, which was
eliminated (REF) by splitting the wide band into narrow-band channels and mapping them onto
their correct spatial frequencies during imaging. It was still assumed that
at the receiver sensitivities of the time, 
the sky brightness was constant across the observed bandwidth.
The next step was to consider a frequency-dependent sky brightness distribution.
Conway,Cornwell and Wilkenson,1990 describe a Double Deconvolution algorithm based on
spectral psfs, the instruments response to a series of spectral basis function,
in this case a Taylor series.
Sault,Wieringa,1994 describe a CLEAN-based MFS algorithm (MF-CLEAN) which models an image 
as a collection of point sources with power law spectra, and uses a linear approximation
to these spectra along with the above mentioned spectral psfs. A spectral index map
was produced as a by-product of the deconvolution.

So far, these CLEAN-based MFS deconvolution algorithms used 
point-source (zero-scale) flux components to model the sky emission,
a choice not well suited for extended emission.
We show in this paper that for MFS, deconvolution errors that occur with a point-source
model are enhanced in the spectral index image because of error propagation effects, and
that the use of a multi-scale technique can minimize this.
%In (REF1,REF1), XXX show that for high dynamic range imaging, point-source models
%are not well suited for extended emission, and describe several multi-scale
%techniques that are more capable at reconstructing complex extended emission accurately.
Multiscale deconvolution techniques model images using flux
components of varying scale size, and are more accurate at deconvolving large scale emission
features. Bhatnagar,Cornwell,2004 describe the ASP-CLEAN algorithm that explicitly fits 
for the parameters of Gaussian flux components and uses scale size to aid the separation of
signal from noise. Cornwell,Holdaway,2004 describe the MS-CLEAN algorithm which performs
matched filtering using templates constructed from the instrument response to various large 
scale flux components.

For high dynamic range imaging across wide fields-of-view,
direction-dependent instrumental effects that need to be accounted for are 
the w-term and the antenna power pattern, both of which are frequency-dependent. 
The XXX algorithm described by Bhatnagar,Cornwell,Golap 2007 describes the correction of
direction-dependent effects during imaging for the narrow-band case.

%Some other work related to this problem include a Spatio-spectral MEM (xxx).

   The MS-MFS algorithm discussed in this paper extends the Sault-Wieringa MF-Clean 
    algorithm to reconstruct spectral curvature in addition to spectral index,
    and combines it with a  multi-scale deconvolution approach similar to that of MS-Clean.
    Frequency-dependent wide-field instrumental effects are accounted for
    along with MFS as an extension of the AProjection and WProjection methods.
%    We also discuss a simpler hybrid of spectral-line and continuum imaging methods
%    that may suffice in certain situations.
   We show via simulations and application to (E)VLA observations of Cygnus-A,
    that for wide bandwidths (50\% up to xxx\%)
    it is possible to reconstruct both spatial and spectral structure from the
    combined measurements.
    We use the example of the 3C286 field to demonstrate
    frequency-dependent correction of the primary beam for wide-field imaging.
    We also present an error analysis to show
    under what conditions these techniques are feasible. %(give numbers)

%%Section \ref{PRIOR WORK} describes existing methods and their limitations.
%Section \ref{HYBRID} describes a simple MF algorithm for arrays with dense single-frequency
%uv-coverage.
%Section \ref{MS-MFS} describes the details of the multi-scale multi-frequency deconvolution algorithm 
%for arrays that use MFS to increase their uv-coverage. 
%Section \ref{SIMULATIONS} compares images of continuum flux, spectral index and curvature, obtained
%by the above two algorithms.
%Section \ref{DATA} compares results from both algorithms on real data (and compares to
%		previously existing algorithms).

%__________________________________________________________________

\section{Wide-Band Imaging}\label{IMAGING}
%__________________________________________________________________

%\subsection{Discussion}
%The primary use of broadband receivers in radio interferometry is to enhance the instruments
%continuum sensitivity and increase the range of spatial frequencies measured instantaneuosly.
%At these wide bandwidths, the frequency-dependence of the sky brightness distribution becomes
%relevant and must be taken into account during image reconstruction.  
%%Multi-frequency-synthesis (REF CCW/SW) is the technique of combining visibility 
%%measurements made at multiple receiver/sky frequencies to construct an image of the
%%sky and its frequency dependence.

We begin with a discussion of how well we can reconstruct both spatial and 
spectral information about the source
from incomplete measurements at multiple frequencies. 
The spatial frequency range per channel is given by 
$u_{\{min,max\}}(\nu) = \vec{b_{\{min,max\}}} \nu/c$, where
$\vec{b_{\{min,max\}}}$ represents the shortest and longest projected baseline vector on
the plane perpendicular to the direction of the source. 
The maximum spatial frequency measured at each frequency defines the angular resolution of
the instrument at that frequency as $\theta_{\nu} = 1/u_{max}(\nu)$.
The range of spatial frequencies between $u_{min}$ at $\nu_{max}$ and  $u_{max}$ at $\nu_{min}$
represents the region that is sampled at all frequencies in the band. In this region, 
both spatial and spectral information is measured in comparable detail. 
The spatial frequencies outside this region are sampled only by a fraction of the band, and 
the accuracy of the reconstruction depends on how well the spectral and spatial structure are
constrained by the choice of a flux model and the rest of the measurements.

%Multi-frequency-synthesis (REF CCW/SW) is the technique of combining visibility 
%measurements made at multiple receiver/sky frequencies to construct a single continuum image.
%The main advantages of MFS are the additional sensitivity and fidelity that comes with filling in 
%the spatial frequency plane and increasing the number of independant constraints
%on the spatial and spectral structure of the source, the increased angular resolution for
%imaging, and the measurement of the source spectrum across a wide bandwidth.

For a flat-spectrum source, measurements at multiple frequencies sample the same sky brightness
distribution at different ranges of spatial scales, increasing the signal-to-noise of the
measurements in regions of overlap, and providing better uv-coverage and angular resolution.
Therefore standard deconvolution algorithms applied to measurements combined via MFS will suffice
to reconstruct source structure across the full range of spatial scales measured across the band.

For sources with spectral structure, different frequencies measure different sky brightness 
distributions, and a direct combination of frequencies via MFS would generate spurious spatial
structure. The traditional method of imaging has been to make separate images at each frequency and
and derive spectral structure after tapering the measurements from higher frequencies to match
the angular resolution of the lowest frequency. 
The imaging sensitivity and fidelity is however restricted to be that of a single channel and its 
uv-coverage, and in most cases, it is possible to do better.

A compact, unresolved source with spectral structure is measured as a point
source at all frequencies, and $u_{max}$ at $\nu_{max}$ gives the maximum angular resolution at
which this source can be imaged. Since the visibility function of a point source is flat across
the entire spatial frequency plane, its spectrum is adequately sampled by the multi-frequency measurements,
and it is possible to completely reconstruct both spatial and spectral structure for such a source.

%%%???
For resolved sources with spectral structure, the accuracy of the reconstruction
across all spatial scales between $u_{min}$ at $\nu_{min}$ and  $u_{max}$ at $\nu_{max}$
depends on an appropriate the choice of flux model, and the constraints that it provides.

For example, a source with a synchrotron spectrum can be described by a brightness 
distribution at a particular frequency, and a power-law spectrum for each location. 
The angular resolution of the images are given by $u_{max}$ at $\nu_{max}$, with the 
assumption that different frequencies probe the same structure but different amplitudes. 
Measurements at the highest frequencies constrain the angular size of the flux 
components, and the combined spectra are constrained by the rest of the measurements.
This constraint is strong enough to correctly reconstruct even moderately resoloved sources
that are completely unresolved at the low end of the band, but resolved at the higher end.

On the other hand, a source whose structure itself changes across the band would break the
assumption that structure seen at the highest resolution is what is present all across the band.
One example is with multi-frequency observations of solar magnetic loops
where the different frequencies probe different layers in the xxx-sphere and can have
very different structures. In this case, a complete reconstruction would be possible only
between $u_{min}$ at $\nu_{min}$ and  $u_{max}$ at $\nu_{max}$, unless the flux model
includes constraints that bias the solution towards one appropriate for solar magnetic loops.
The imaging would still however operate at the continuum sensitivity level, and is therefore
better than single-channel imaging.

The lower end of the frequency range presents a different problem. 
The size of the central hole in the uv-coverage
increases with frequency, and source spectra are not measured adequately for
extended emission whose visibility function is non-zero only in
the range below $u_{min}$ at $\nu_{max}$. A flat-spectrum large-scale source is 
indistinguishable from a relatively smaller source with a steep spectrum, and vice versa.
For scale sizes in this range, additional constraints in the form 
of total-flux values may be required for each frequency.
% or spectrum is simple enough that half the band is enough to get it.

In addition to the sky flux being frequency dependent, the instrumental effects are also 
frequency dependent.
The HPBW of the primary beam of each array element scales as $HPBW = \frac{c}{\nu D}$. 
The reconstructed image will therefore have artificial spatial and spectral structure that can be
measured and removed to undo these effects.

\section{MS-MFS Algorithm}\label{ALGORITHMS}
The design of an image reconstruction algorithm begins with an appropriate choice of
parameterization of the sky brightness distribution and various instrumental effects.
This parameterization is then folded into the transfer function of the instrument
whose measurement process we are trying to invert. An optimization process is then
followed to obtain best-fit estimates for the parameters which can then be 
interpreted physically. This section describes each
of these stages in detail for one approach to broad-band synthesis imaging 
for radio interferometry.
The following is an extension of MF-CLEAN, combined with MS-CLEAN and AProjection.

\subsection{Parameterization of spatial structure}
For a multi-scale model, an image can be written as a linear combination of images
at different angular resolutions (Ref MS-CLEAN).
A multiscale representation of an image is given by 
\begin{equation}
I^m = \left\{ \sum_{s=0}^{N_s} \left[ I^m_{s,\delta}\star I^{blob}_s \right] \right\}
~~~ where ~~~ I^{blob}_s = XXX
\label{EQN_MS}
\end{equation}
and $N_s$ is the number of discrete spatial scales used to represent the image, and
$I^m_{s,\delta}$ represents a collection of $\delta$-functions that describe the locations
and integrated amplitudes of flux components of scale $s$ in the image.
$I^{blob}_s$ is a tapered truncated parabola of width proportional to $s$.
The symbol $\star$ denotes convolution.

\subsection{Parameterization of spectral structure}
The spectrum of each flux component is modeled by a polynomial in frequency
( a Taylor series expansion about $\nu_0$ ).
\begin{equation}
I^{obs}_{\nu} = \sum_{t=0}^{N_t} \dnuno^t I^m_t 
\label{EQN_TS}
\end{equation}
where $N_t$ is the order of the Taylor series expansion, and 
the $I^m_t$ represent multi-scale Taylor coefficient images (moment maps ??? ).
This decomposition is linear in the coefficients as well as in the basis functions.

These Taylor coefficients can be interpreted by choosing an astro-physically appropriate
spectral model and performing a Taylor expansion to derive expressions that each coefficient
maps to.
One choice for a spectral model is a power law with a varying index, represented by a 
second-order polynomial in $log(I)~vs~log\nuno$ space.
The variation of the spectral index with frequency is described by an average spectral
index $\alpha$ and a curvature term $\beta$.
\begin{equation}
I_{\nu}^{obs} = I_{\nu_0}^{obs} \nuno^{\alpha + \beta log \nuno}
\label{EQN_POWERLAW1}
\end{equation}
The main reason behind this choice is the fact that continuum synchrotron emission is 
usually modeled (and observed) as a power law distribution with frequency. Across the wide
frequency ranges that new receivers are now sensitive to, 
spectral breaks, steepening and turnovers need to be factored into
models, and the simplest way to include them and ensure smoothness, is spectral curvature.
(FIGURE).
%% Figure Caption
%% (1) Synchrotron spectrum with steepening, turnover, break.
(Wideband imaging algorithms describes in CCW/SW use a fixed spectral index across the band,
and handle slight curvature by performing multiple rounds of imaging after removing the 
dominant/average $\alpha$ at each stage. 
They also suggest using higher order polynomials to handle spectral curvature.)

A Taylor expansion of Eqn \ref{EQN_POWERLAW1} yields the following expressions for the first
three coefficients ($t=0,1,2$ in Eqn \ref{EQN_TS}) from which the spectral index $\alpha$ and curvature $\beta$ images 
can be computed algebraically.
\begin{equation}
I^m_0 = I^{obs}_{\nu_0} ~~;~~~ I^m_1 = \alpha I^{obs}_{\nu_0} ~~;~~~ I^m_2 = \left(\frac{\alpha(\alpha-1)}{2} + \beta\right) I^{obs}_{\nu_0}
\label{EQN_COEFFS}
\end{equation}
Note that with this choice of parameterization, 
we are using a polynomial to model an exponential, and this has caveats, etc which
are discussed in the section on errors.
Also, note that there can be other interpretations of the Taylor coefficients %poln ?
and other expansions. 
A power-series expansion about $\alpha$ and $\beta$ will yield a logarithmic
expansion i.e. I vs log nu. (CHECK if/when this is better)
It is however impractical to work directly in Log I and Log nu space because that
involves taking logs and exp of image pixel amplitudes and this is highly unstable numerically.

\subsubsection{Frequency-dependant instrumental effects}
Several instrumental effects are also frequency dependent, and need to be modeled/measured and
corrected for.
The field-of-view of the telescope depends on the physical size of the collector
in units of signal wavelength
(given by $\lambda/D$) and is therefore frequency dependent. Figure (FIGURE) shows the
the shape of the element response function (primary beam) of an EVLA antenna at multiple
frequencies. 
%% Figure Caption 
%% (1) HPBW at 1 GHz is near the 5% point at 2 GHz.
%% (2) Continuum sensitivity has no nulls until quite far out.
%% (3) HPBW of ref freq has a spectral index of about 1.4 (also derived in SW)
For a source away from the pointing-centre of these beams, the coefficient images now 
represent the sky and its spectrum, multiplied by the frequency dependent primary beam.  
The artificial spectral structure introduced by the antenna response pattern can also be
modeled by a polynomial in frequency, and divided out of coefficient images to correct 
them. A polynomial model for the primary beam can be either measured, or pre-computed.

%The continuum, spectral index and curvature images computed from Eqn \ref{EQN_COEFFS}
%can then be interpreted as 
%\begin{equation}
%I^{obs}_{\nu_0} = I^{sky}_{\nu_0}P_{\nu_0}~~; ~~~~ \alpha = \alpha_{sky} + \alpha_{pb}~~; ~~~~ 
%\beta = \beta_{sky} + \beta_{pb}
%\end{equation}
%where $P_{\nu_0}, \alpha_{pb}, \beta_{pb}$ describe the 
%effective spectrum of each direction on the sky due to the primary beam and
%can be measured and/or pre-computed. 
%The only reason for this choice is
%that since it is the same as the sky flux, the correction of spectral effects becomes linear
%and this controls error propagation.
%and the results can be corrected after the Minor cycle of deconvolution.
Other frequency-dependent instrumental effects include the w-term, and beam squint for 
off-axis feed systems, all of which are phase ramps in the aperture illumination pattern
and therefore correctable during the imaging process.

%%%%% Implementation

\subsection{Imaging Equations}

The {\it Measurement Equation\footnote
{
The measurement equation of an imaging instrument describes its transfer function (the effect of
the measurement process on the input signal).
For an ideal interferometer (a perfect spatial frequency filter, with no instrumental gains),
it can be written in matrix notation as follows. 
Let $I^{sky}_{m\times 1}$ be a pixelated image of the sky and let $V^{obs}_{n\times 1}$
be a vector of $n$ visibilities. 
Let $S_{n\times m}$ be a projection operator that describes
the uv-coverage as a mapping of $m$ discrete spatial 
frequencies (pixels on a grid) to $n$ visibility samples (usually $n>m$).
Let $F_{m\times m}$ be the Fourier transform operator.
Then, $[{\Sa}_{n\times m}] [F_{m\times m}] \vec{I}^{sky}_{m\times 1} = \vec{V}^{obs}_{n\times 1} $
}
}for a sky brightness distribution parameterized by Eqns.
\ref{EQN_MS}\&\ref{EQN_TS} is given by 
\begin{eqnarray}
V^{obs}_{\nu} &=& \sum_{t=0}^{N_t}\wnt [\Sna][\F] \left\{\sum_{s=0}^{N_s} I^{blob}_s \star I^{m}_{t,s,\delta} \right\} \\
		&=& \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} \wnt [\Sna][T_s] [\F] I^{m}_{t,s,\delta}
%%V^{corr}_{\nu} &=& \sum_{t=0}^{N_t}\wnt [\Sna][\F] [\Pnu] \left\{\sum_{s=0}^{N_s} I^{blob}_s \star I^{m}_{t,s,\delta} \right\} \\
%%		&=& \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} \wnt [\Sna][\Gnu][T_s] [\F] I^{m}_{t,s,\delta}
%%V^{corr}_{nc \times 1} = \sum_{t=0}^{N_t} \sum_{s=0}^{N_s} [\Wnt][\Sa][\F][\Fd T_s \F]I^{sky}_{t,s,\delta}
\end{eqnarray}
where $w_{\nu} = \dnuno$ and $\Sna$ represents the uv-coverage of the synthesis array at frequency $\nu$.
The image-domain convolution with $I^{blob}_s$ is written as a spatial-frequency
Taper function $[T_s]_{m\times m} = \F I^{blob}_s $. 
The image parameters to be solved for are the Taylor coefficients at each spatial scale
($I^{model}_{t,s,\delta}$ for all $t,s$), and data from all frequencies are used simultaneously
in the solution process.

The {\it Normal Equations\footnote
{
The Normal Equations are the linear system of equations whose solution 
gives a weighted least-squares estimate of a set of parameters in a model  
($\chi^2$ minimization). 
For an ideal interferometer, it is given by 
$[\Fd \Sd W \Sa \F ] I^{sky}_{m\times 1} = [\Fd \Sd W] V^{obs}_{n\times 1} = I^{dirty}_{m\times 1}$
where $W_{n\times n}$ is a diagonal matrix of signal-to-noise based measurement weights
and $\Sd$ denotes the mapping of measured visibilities onto a spatial frequency grid.
The Hessian (matrix on the LHS) is a {\it convolution} operator with 
$I^{psf}_{m\times 1} = diag[\Fd \Sd W \Sa]$ in each row. 
The {\it dirty} image on the RHS is therefore a result of a convolution of $I^{sky}_{m\times 1}$ with $I^{psf}$
and these equations can be solved by a {\it deconvolution}. 
}
}for each scale size $s$, and Taylor term $t$ are given by
\begin{equation}
\sum_{p=0}^{N_t}\sum_{q=0}^{N_s} H_{t,s,p,q} I^{m}_{p,q,\delta} = I^{dirty}_{s,t}
\label{HESS}
\end{equation}
\begin{eqnarray}
\label{HESS1}
%%%\mathrm{where} ~~~~ H_{t,s,p,q} &=& [\Fd T_s] \left\{\sum_{\nu} \wntp [\Snd\Wimn\Sna] \right\} [T_q \F] \\
\mathrm{where} ~~~~ H_{t,s,p,q} &=& \sum_{\nu} \wntp [\Fd T_s \Snd\Wimn\Sna T_q \F] \\
%%where ~~~~ H_{t,s,p,q} &=& \sum_{\nu} \wntp [\Fd T_s \Gdnu\Snd\Wimn\Gnu\Sna T_q \F] \\
%	&=& I^{blob}_s \star I^{psf}_{im,t,p} \star I^{blob}_q \\
\label{HESS2}
%%%\mathrm{and} ~~~~ I^{dirty}_{s,t} &=& [\Fd T_s] \left\{ \sum_{\nu}\wnt[\Snd\Wimn] V^{corr}_{\nu} \right\}
I^{dirty}_{s,t} &=&  \sum_{\nu}\wnt[\Fd T_s \Snd\Wimn] V^{corr}_{\nu} 
%%I^{dirty}_{s,t} &=&  \sum_{\nu}\wnt[\Fd T_s \Gdnu \Snd\Wimn] V^{corr}_{\nu} 
%	&=& I^{blob}_s \star I^{dirty}_t
\end{eqnarray}
%where $\Wimn_{n\times n}$ is a diagonal matrix of signal-to-noise based measurement weights for channel $\nu$
%and $\Snd$ denotes the mapping of measured visibilities onto a spatial frequency grid.
%and $I^{dirty}_{s,t} = [\Fd T_s \F] [\Fd\Sd\Wntd\Wim] V^{corr}$

%\begin{figure}[!t]\label{NEqn With PB}
%\epsfig{figure=pics/NEqn.cropped.pb.yes.eps,scale=0.4}
%\caption{Normal Equations for MFS Imaging with a Frequency Dependent Field of View}
%\label{FIG_NEQN_2}
%\end{figure}

Fig.\ref{FIG_NEQN_1} represents the Normal Equations for a three-term Taylor series
$N_t = 3$ and one spatial scale $N_s = 1$. 
	$I^{sky}_{3m\times 1}$ is a 1-D model of two point sources on an empty sky. 
	The three segments correspond to Taylor coefficients $I^{m}_p$ for $p=0,1,2$.
	The Hessian/Beam matrix on the LHS is comprised of $3 \times 3$ blocks each
	of size $m\times m$. Each block is a Toeplitz operator (constructed from spectral PSFs)
	that implements the shift/multiply/add sequence of convolution
	when applied to the corresponding segment of $I^{m}_t$.
	%The RHS vector represents the dirty image formed from direct Fourier inversion
	%of Taylor-weighted measured visibilities. 
	These equations show how the {\it dirty} image on the RHS can be
	written as a linear combination of convolutions of Taylor coefficient images with
	spectral PSFs. Each segment on the RHS is given by
	$I^{dirty}_t = \sum_{p=0}^{3} I^{psf}_{t+p} \star I^{m}_p$.
	Note : The equations corresponding to $t=0,p=0$ represent the standard convolution
	equation in synthesis imaging.

When all scales and Taylor terms are combined, 
the LHS of the Normal equations is made up of the full Hessian matrix with
$N_t N_s \times N_t N_s$ blocks each of size $m\times m$ and 
containing information from all frequency channels, 
and $N_t$ Taylor coefficient images each of size $m\times 1$.
Each block described by Eqn.\ref{HESS1} is a convolution operator with a {\it spectral} PSF of order $t+p$ 
in each row. 
The Spectral PSF $ I^{psf}_{t} = \sum_{\nu} \wnt I^{psf}_{im,\nu}$
describes the instrument response functions to a point source with Taylor-polynomial spectra of order $t$.
Similarly, the spatial PSF $ I^{psf}_s = I^{psf}_{im} \star I^{blob}_s $ is the instruments 
response function to a spatial scale $s$ and unit integrated flux.
%This taper gives lower spatial frequencies a relatively high weight compared to higher spatial frequencies,
%to tune the instrument's response towards a spatial scale larger than the
%angular resolution of the telescope.
Off-diagonal blocks represent the degree of coupling (non-orthogonality)
between the various spatial and spectral basis functions.
The RHS of this system of equations is a collection of {\it dirty} images obtained by
direct Fourier inversion of Taylor-weighted visibilities.

\begin{figure}[]
\epsfig{figure=pics/NEqn.cropped.pb.no.eps,scale=0.4}
\caption{A 1-D representation of the Normal Equations for MFS imaging with a $2^{nd}$ order
	Taylor expansion in frequency($N_t=3$) and a point-source flux model (no multi-scale).
%	$I^{sky}_{3m\times 1}$ is a 1-D model of two point sources on an empty sky. 
%	The three segments correspond to Taylor coefficients $I^{m}_p$ for $p=0,1,2$.
%	The Hessian/Beam matrix on the LHS is comprised of $3 \times 3$ blocks each
%	of size $m\times m$. Each block is a Toeplitz operator (constructed from spectral PSFs)
%	that implements the shift/multiply/add sequence of convolution
%	when applied to the corresponding segment of $I^{m}_t$.
%	%The RHS vector represents the dirty image formed from direct Fourier inversion
%	%of Taylor-weighted measured visibilities. 
%	These equations show how the {\it dirty} image on the RHS can be
%	written as a linear combination of convolutions of Taylor coefficient images with
%	spectral PSFs. Each segment on the RHS is given by
%	$I^{dirty}_t = \sum_{p=0}^{3} I^{psf}_{t+p} \star I^{m}_p$.
%	Note : The equations corresponding to $t=0,p=0$ represent the standard convolution
%	equation in synthesis imaging.
}
\label{FIG_NEQN_1}
\end{figure}

The Normal Equations are solved in two stages, following a CLEAN-type approach.
The parameters of the flux components are first solved for one component at a time.
Then, the Normal equations are used to calculate the contribution of these flux components
to the dirty image, which is then subtracted out.

For the purpose of finding flux components, several assumptions can be made about the Hessian.
First, each scale can be treated separately as in MS-CLEAN, and Hessian blocks that couple
different spatial scales are ignored. Each scale then has $[N_t \times N_t]$ blocks of size
$m\times m$ as shows in Fig.\ref{FIG_NEQN_1}.
Second, pixels are assumed to be independant of each other and a 
diagonal approximation to the Hessian blocks yields a single $N_t \times N_t$ matrix 
that applies to all pixels. The $t,s$ elements of this matrix come from the values of
$I^{psf}_{t+p}$, read off at the location of the peak of $I^{psf}_0$. These numbers
correspond to the sum of weights $\sum_{\nu} \wntp$.
These N.E. are the result of autocorrelations of Spectral PSFs and are 
(usually) guaranteed to be non-singular, and exactly solvable via an LU decomposition.
Taylor coefficients for each location in the image are obtained by
solving this $N_t \times N_t$ matrix for every component location.

The Normal equations are then evaluated for the components with the dominant Taylor0
coefficient (or picked out some other way), and the effect of that component is
subtracted out of the RHS vector. This is the update step in a formal iterative
optimization.

%\item When direction-dependant effects are included, the blocks in $H$ are no longer 
%convolution operators, and each row is the PSF scaled and multiplied by the PB.\\ 
%$H_{i,j} = P(i-i_{mid}) P(j-j_{mid}) I^{psf}(j-j_{mid}-i)$. 
%Therefore, can still compute only one PSF per block, and multiply and scale appropriately
%when updating the Residuals in the Minor Cycle.

The relation between this implementation and that of the SW MF-CLEAN is described in Appendix\ref{APP_A}.

\subsection{Algorithm Steps}
\subsubsection{Pre-compute Hessian}  
All $N_s N_t \times N_s N_t$ terms in the Hessian ($t,p \in [0,N_t]; s,q \in [0,N_s]$) are 
evaluated as in Eqn.\ref{HESS}.
%\begin{equation}
%[\Fd T_s T_q \F]I^{psf}_{im,t,p} = I^{blob}_{s} \star I^{psf}_{im,t,p} \star I^{blob}_{q} 
%\end{equation}
%where 
\begin{eqnarray}
H_{t,s,p,q} &=& I^{blob}_s \star \left\{ \sum_{\nu} \wntp I^{psf}_{im,\nu} \right\} \star I^{blob}_q \\
 I^{psf}_{im,\nu} &=& [I^{wt}_{pc}]^{-1} [ \Fd \G^{pc} \Snd \Wimn]  V^{1}
\end{eqnarray}
where $I^{wt}_{pc} = \Fd P_s$, where $P_s$ is an anti-aliasing function,
and $\G^{pc} = [\F I^{wt}_{pc}\Fd]$ is the corresponding gridding convolution operator.

Since each block is a Toeplitz matrix constructed from spectral and spatial PSFs, only one
row per block need be computed.

The values at the pixel at which the
{\it peak} of $I^{psf}_{im}$ occurs,
are separately recorded as $N_t\times N_t$ matrices ($H^{peak}_{t,s,p,s}$) for each scale $s$.

Note : Normalize all by the $w_{sum}$ of Taylor 0.

\subsubsection{Pre-compute Primary Beam} 
Taylor coefficients for a polynomial expansion of the PB across frequency are obtained
by applying the inverse of $H^{peak}$ to RHS functions computed as follows.
\begin{eqnarray}
\sum_{p=0}^{N_t} H_{t,p} P^{dd}_{p} &=& I^{wt}_{dd,t} \\
I^{wt}_{dd,t} = \sum_{\nu} \wnt I^{wt}_{dd,\nu} &=& \sum_{\nu} \wnt{ [\Fd \G_{\nu}^{{dd}^{\dag}} \Wimn \G_{\nu}^{dd} F] }
\end{eqnarray}
where $G_{\nu} = [\F P_{\nu} \Fd]$.
The effective $\alpha$ and $\beta$ in each direction on the sky, are computed
from Taylor coefficients. 
\begin{enumerate}
\item Note that $P^{dd}_{\nu_0},\alpha_{pb},\beta_{pb}$ as computed
directly from the coefficients have $PB^2$ in it, so take the sqrt of the PB and
divide the alpha and beta by two.
\item The PB and its frequency dependence are computed via weighted averages, using
the same imaging weights as used in computing the dirty/residual images. Also, baseline-dependent 
PBs at different parallactic angles can also be accounted for in this weighted average.
Therefore, $P^{dd}_{\nu_0},\alpha_{pb},\beta_{pb}$ represent the collective effect on the
final image.
\end{enumerate}

\subsubsection{Major and Minor Cycles}
The Normal Equations are solved iteratively by repeating steps \ref{stepRHS} to \ref{stepPredict} until 
some termination criterion is reached.
Steps \ref{stepRHS} and \ref{stepPredict} form one Major Cycle, and repetitions of Steps \ref{stepFind} and \ref{stepUpdate} form the Minor Cycle.

\begin{enumerate}
\item\label{stepRHS} {\bf Compute RHS} : Residual images for each 
$s\in[0,N_s]$ and $t\in[0,N_t]$ are computed as in Eqn.\ref{HESS2}.
\begin{eqnarray}
I^{dirty}_{s,t} &=& I^{blob}_s \star \left\{ \sum_{\nu} \wnt I^{dirty}_{\nu} \right\} \\
I^{dirty}_{\nu} &=& [\Pnu]^{-1} [I^{wt}_{pc}]^{-1} [ \Fd \G_{\nu}^{{dd}^{\dag}} \G^{pc} \Snd \Wimn ]  V^{residual}_{\nu}
\end{eqnarray}
where $V^{residual}_{\nu} = V^{corr}_{\nu}$ for the first Major cycle iteration.

%The normalization by $\Pnu$ takes out one instance of the PB
%and is equivalent to grid-correction.

$\G^{pc}$ and $\Gddnu$ are convolution operators used during gridding.
$\G^{pc}$ applies an anti-aliasing function, and $\Gddnu$ corrects for direction-dependent
phase effects (w-projection, pointing offsets and squint).
%Different versions are used in different
%variants of this process as discussed in section \ref{VARIANTS}.

%%\item\label{stepGridCorrect} {\bf Grid correction} : to take out one PB.

\item\label{stepFind} {\bf Find a Flux Component} :
For each scale $s$, the $N_t\times N_t$ Hessian matrix given by $H_{t,p,s,s}$
is inverted and applied to all pixels of the corresponding RHS vector 
($I^{dirty}_{t,s};s\in[0,N_s]$), to generate
solution sets $\{I^{model}_{p,\delta}\};p\in[0,N_t]$ for each $s\in[0,N_s]$. 
The $N_t$ element solution set of Taylor coefficients
with the dominant $p=0$ component across all scales and pixel locations,
is chosen the current flux component. 
Let the scale size for this set be $q$. 
The chosen solution set is given by $\{I^{model}_{p,q,\delta}\};p\in[0,N_t]$.

\item\label{stepUpdate} {\bf Updates }: 
Multiscale model images for each Taylor coefficient, and dirty images (RHS,residuals) are updated as 
\begin{equation}
I^{model}_p += I^{model}_{p,q,\delta} \star I^{blob}_q ~~~~~~~~ \forall p\in[0,N_t]
\end{equation}
\begin{equation}
I^{dirty}_{t,s} -= \sum_{p=0}^{N_t} \left[ I^{blob}_{s} \star I^{psf}_{im,t,p} \star I^{blob}_{q} \star I^{model}_{p,q,\delta} \right] ~~~~~~~~ \forall t,s
%= \sum_{p=0}^{N_t} \left[[\Fd T_s T_q \F]I^{psf}_{im,t,p}\right]\star I^{model}_{q,p,\delta}
\end{equation}

\item\label{stepCorrectPB} {\bf Correct for PB}: 
The model returned from the Minor cycle contains one instance of the PB along with its
frequency dependence, and this needs to be corrected before prediction. 
The model image is divided by the P0, the alpha and beta are corrected by subtracting
the alpha and beta of the PB.  The taylor coefficients are then recomputed using the
corrected PB0, alpha and beta.
\begin{eqnarray}
I^{skymodel}_0 &=& I^{model}_0 / P_{\nu_0} \\
I^{skymodel}_1 &=& I^{sky}_0 \left(\frac{I^{model}_1}{I^{model}_0} - \alpha_{PB}\right)\\
\cdots & & \forall p\in[0,N_t]
\end{eqnarray}

\item\label{stepPredict} {\bf Predict }: Model visibilities are computed as 
\begin{equation}
V^{model}_{\nu} = \sum_{p=0}^{N_t} \wnt [\Sna] [\G_{\nu}^{{dd}}] [\F]I^{skymodel}_{p}
%V^{model}_{nc\times 1} = \sum_{t=0}^{N_t}[\Wnt][\Sa \G^{pc}\F][I^{pc}]^{-1} I^{model}_{t}
\end{equation}
The PB-corrected models get back a PB in the predicted visibilities, 
    making them read to compare with the data.
Residual visibilities are computed as $V_{\nu}^{residual} = V_{\nu}^{corr} - V_{\nu}^{model}$ and
then processed as in Step \ref{stepRHS} to construct the residual image for the next Major cycle.

\subsubsection{Restoration / Trivial Solution}
Model images are usually smoothed by the {\it Clean Beam} and the residuals are added back in, to
create the final restored image. With MFS however, this step is different. 
Taylor coefficients need to be constructed from the residual images before they are added back in.

This is also relevant for the trivial solution - 
the case where deconvolution is not done, and where the dirty images
computed in step \ref{stepRHS} can be used as is, to compute Taylor coefficients. 
The $N_t \times N_t$ matrix formed from the Hessian peaks for the zero-scale $H^{peak}_{t,0,p,0}$,
is inverted and applied to all pixels of the images computed in step \ref{stepRHS}.
This step needs to be followed by a further correction by PB and its alpha and beta.

When there is no deconvolution, the Hessian blocks are diagonal, and the full hessian with
the PB is equivalent to multiplying the image with PB square..

\end{enumerate}

\subsection{Error Analysis}
\begin{enumerate}
\item Magnitude of diff Taylor terms :
Are they visible in image as errors (vs) SNR needed to solve.
\item Error bars on alpha and beta due to SNR
\item Error due to using too few Taylor terms to fit the exp with a poly.
CCW comment on a bias that occurs with a 2-term T-exp. This is just the use of insufficient terms of a polynomial to model an exponential.
\item Error due to too many Taylor terms for low SNR.
\item Expansion about $\nuno$ (logarithmic PSFs)
vs expand about $\alpha$ and $\beta$ (Taylor PSFs).
% NOTE : Can model it as a polynomial in I vs log(nu) space too.
% Cannot model it as a poly in log(I) vs log(nu) space because computing log(I) is a pain.
\item Spectrum of extended emission... artificially steepened due to short-spacing problem.
-- error is a certain amount, based on the percentage of flux measured at each freq.... 
\item Sources that are unresolved at one end of the band, but resolved at the other.
\item Wide-Band flux calibration : required accuracy.
\item Direction dependent sensitivity (Weight Image) (sources can pass through a Null at some freq and have less continuum sensitivity.)
\item Error in alpha due to uncorrected freq-dep primary beam.
\item Freq-dep PB correction with non-homogenous arrays : the Weight image is a weighted average
across all baselines,times, frequencies and represents a collective effect.
\end{enumerate}

\subsection{Variants}\label{VARIANTS}
The discussions in the preceding sections were based on a single pointing observation,
and a data-access pattern that treats frequencies independently (suitable for
data-partitioning by frequency in a parallel implementation). Also, the Minor cycle
operates on residual images that satisfy the convolution equation $PSF \star (P I^{sky})$, 
so that the update step in the Minor cycle is accurate. This is a flat-noise model, in that
the residual image is a signal-to-noise map and the model image represents $(P I^{sky}$.

The following are variants of this approach.
\begin{enumerate}
\item Using a Multi-Freq UV grid (P and 2 alpha) - This is an approximation in that the minor cycle update cannot be accurate, but the major cycle will be accurate. Perhaps useful when you need to touch all frequencies at once (mem-access-patterns).
\item Flat Sky vs Flat Noise (Mosaicing, Non-homogeneous arrays (Different PB for each antenna))
\end{enumerate}

\section{Hybrid Algorithm}\label{HYBRID}
This section describes a simple procedure for wide-band imaging as an extension of the traditional
single-channel imaging, and discusses situations in which this may be appropriate.

The simplest way to deal with a frequency-dependent sky is to image each frequency channel
separately, and then compute spectral characteristics. The advantage is that is does not depend
on any spectral flux model, can handle arbitrary spectral shapes, and is straightforward to
implement in code and parallelize. 
However, the accuracy of single-channel imaging depends on the uv-coverage obtained at one
single frequency (or a range of frequencies over which the sky is spectrally flat), so this method
is appropriate only for interferometers where the narrow-band coverage is sufficient to sample
all relevant spatial scales of a source.  
Further, the narrow-band imaging sensitivity can be 
significantly less than what the broadband instrument is capable of, but 
in some cases, a hybrid approach between single-channel and continuum imaging can be used to
overcome this problem.

\subsection{Procedure and Error Analysis}
\begin{enumerate}
\item Deconvolve each channel separately, and Clean down only to the single-channel sensitivity limit
$\sigma_{chan}$. 
\item Compute the continuum residual image, 
assume a flat spectrum, and clean down to the continuum sensitivity limit 
$\sigma_{cont} = {\sigma_{chan}}/{\sqrt{N_{chan}}}$.  
\end{enumerate}
When does the second stage work ? 
For bright sources visible in the single-channel maps, the spectral structure is detected and
removed by the end of Step 1, leaving flat-spectrum residuals that justify the assumption in
Step 2. 
Weak sources whose flux values lie below $\sigma_{chan}$ but above $\sigma_{cont}$ 
are detected only in Step 2, and the flat-spectrum assumption may be a problem for their
deconvolution. However, if these spectral errors are less than $\sigma_{cont}$, it will not
matter. CCW have shown that for a spectral index of -1.0 across a 2:1 band, spectral errors
are visible beyond a dynamic range of 1000. Therefore, as long as 
$\sigma_{chan}/\sigma_{cont} < 1000$,
the errors should not be visible. This implies $\sqrt{N_{chan}} < 1000$ which will almost
always be satisfied.

This method is straightforward for simple fields of compact sources where the single-channel
uv-coverage is sufficient to fully reconstruct the source structure. 
For complicated fields, or insufficient uv-coverage, this method is highly susceptible to
deconvolution errors that leave single-channel residuals that neither add up coherently in
Step 2, nor are random enough to average down by $\sqrt{N_{chan}}$. CCW also comment on this.

If there is sufficient uv-coverage per channel, then low-level undeconvolved flux can add up
coherently, and continuum residuals will satisfy a convolution equation that Step 2 can solve.
This idea has been tested on EVLA simulations, but it yet to be verified on real wide-band data
where all frequencies are measured simultaneously.

Since this is an inherently single-channel approach, the spectral structure can be obtained only
at the angular resolution allowed by the lowest frequency. The field of view is not a restriction,
but the continuum sensitivity degrades beyond the field of view given by the highest frequency.

%__________________________________________________________________

\section{Wide-band ASP-Clean}
\begin{enumerate}
\item Add a frequency dependence to each ASP. Simplest is a polynomial amplitude variation of each Gaussian in the image fit.
\item Can analyse N-unknowns and N-equations.
\item Need at least 2 measurements per Gaussian, and 2 freqs - to fit its scale and spectral params. This implies that it can work for very sparse sampling, eMerlin, VLBI, etc..
%For example, consider a flux parameterization of Gaussians with amplitudes given 
%by a polynomial spectrum. For extended emission, the spatial and spectral parameters 
%of a Gaussian can be fit as long as an adequate number of measurements ($>$ N parameters)
%are made in the spatial frequency range where the visibility function of the Gaussian is non-zero
%at at least two frequencies.
\item For moderately resolved sources, it gets multiple high-res Gaussians that satisfy the measured SUM at the lower ends.
\item Any limiting case for this ?? 
\item Some cases can get ambigious.
\item Very high dynamic range is difficult - since the fitting is done in the visibility domain. Good for extended emission. Difficult to get weak point source on top of extended emission.
\end{enumerate}

%__________________________________________________________________

\section{Simulations}\label{SIMULATIONS}
\begin{enumerate}
\item True Image : Range of spatial scales and 2nd order log spectra 
\item Data : Full range of scales is sampled only by multi-freq uv-coverage. One source is moderately resolved.
\item Without primary-beam : Comparison between MF-Clean (with 3terms), MS-MFS, Hybrid  in terms of 
image fidelity, angular resolution of spectral images and error-bars on alpha/beta.
\item With primary beam : Comparison between MS-MFS without and with primary-beam correction,
	sensitivity across field-of-view.  Use Data simulated with primary beam where
	there are sources outside the hpbw of the highest frequency, 
	but within the hpbw of the lowest frequency.  
	Also have one source near the null at some frequency.
\end{enumerate}

%__________________________________________________________________

\section{Data}\label{DATA}
\subsection{Narrow-field CygA, no Primary Beam Correction}
\begin{enumerate}
\item Data : Sparse single-freq uv-coverage at 9 frequencies.
\item Calibration (wideband)
\item Imaging Results : Hybrid vs MS-MFS (Continuum Image, Spectral Index Image,
Curvature ?) %for high SNR parts)
\end{enumerate}
\subsection{Wide-field 3C286, Primary Beam Correction}
\begin{enumerate}
\item Data : wide field of unresolved sources.
\item Calibration : PBMos + WProj + wideband model of 3C286
\item Imaging results : Continuum and spectral index images, without and with Primary Beam Correction
\end{enumerate}

%__________________________________________________________________

\section{Discussion}
\begin{enumerate}
\item { Summary : Can do this'n'that (image and spectrum reconstruction + astrophysical parameters)}
...at the angular resolution defined at the highest measured frequency.
\item Single-Channel Imaging Hybrid vs MS-MFS
\item Astrophysics : A variety of astrophysical observations could gain from these new instruments due to the 
increased sensitivity and large instantaneous bandwidth. Synchrotron spectra can be measured
within the instantaneous bandwidth, to detect/measure turnovers/breaks.
Snapshot imaging of highly variable sources 
(sunspots, supernovae, AGN) can be improved with the increased uv-coverage of mfs while
deriving the instantaneous spectra at multiple spatial scales.
Weak supernova remnants could be detected against a continuum background by their spectral signature.
Spectral information can be obtained at the highest angular resolution. This is especially 
significant for moderately resolved sources. 
The atmospheres of stars have an angular size of xxx and are resolvable at xxx GHz in the middle
of the EVLA frequency coverage. MFS with data between xx and yy GHz could yield wideband spectra
at the angular resolution defined by yy GHz.

Sunspot flares - magnetic loops - frequency probes different depths - band-limited signals.

VLBI position offsets as a function of frequency - can be resolved (defined) via MFS.

Multiple arrays can be combined by matching spatial frequency coverage across wide frequency ranges.
For example, EVLA-D between 5 and 50 GHz would match with EVLA-C in L and S bands... (as has always
been done, but it can be imaged together - not sure what final advantage this will bring though..
Snapshot imaging with EVLA subarrays could also achieve this.

\item Software (CASA and ASKAPSOFT), Cost and Parallelization (Hybrid vs MSMFS)
\item { Limitations }
- moderately resolved sources will need simultaneous deconvolution of multipel components (ASP)
- unresolved at lower freq, resolved at higher - 1D simulations show that it is possible to reconstruct, if at least XXX fraction of the bw has resolved info.
\item { Future Directions }
Out into the null of the PB (alternate parameterisations).
Full-Stokes imaging (ref SW)... parameterize differently in lm and freq but same idea.
\end{enumerate}

%__________________________________________________________________

\begin{acknowledgements}
      Thank you ! RJS, SB, RN, DW, FO, BC, 
\end{acknowledgements}

\begin{thebibliography}{}

  \bibitem[1994]{MFCLEAN} R.J. Sault and M.H. Wieringa... "
\end{thebibliography}

\begin{appendix}
\section{Relation to MF-CLEAN}\label{APP_A}

MF-CLEAN is formulated as a least-square method, but its implementation follows a matched-filtering method
where the Hessian and RHS are computed as post-gridding convolutions of images with point-spread functions.
(CHECK) . The two methods are equivalent only when mf-gridding is ignored.
%, and there too, the convo is with sqrt(imwt) in each PSF.
By computing the H and RHS as the equations dictate, 
there is a notable improvement in stability of the algorithm, and normalizations work out exactly.
%Also - this is reflected in a different way of computing Hessian elements as conv of PSFs vs a single PSF with the summed taylor-indices.

MF-Clean does the following : $I^{res}_t = I^{res}_0 \star B_t$ where $B_t = [\Fd]\left\{ \sum_{\nu} \wnt [\Snd][\Wimnd]V^{1} \right\} $. Also, the Hessian elements are computed as $H_{t,p} = B_t \star B_p$. 
Computed this way, $B_t \star B_p = B_{t+p}$ {\bf only} when $\Wimn$ is the 
Identity matrix (natural weighting) and 
$\Sna$ is a pseudo-unitary Boolean operator, or the UV-coverage of the array
is mutually independent across frequencies. 
This can be understood by thinking about when the following could be true 
$\left\{ \sum_{\nu} \wnt [\Snd\Wimnd] \right\}\left\{ \sum_{\nu} \wnp [\Snd\Wimnd]\right\} = \left\{ \sum_{\nu} \wntp [\Snd\Wimnd]\right\}$.

\end{appendix}

\end{document}

\message{ !name(msmfs_aa.tex) !offset(-962) }

### 2016-04-20 msmfs.tex
Algorithm added, figures updated.

@@
 % Symbol definitions
-\newcommand{\bytes}{b} % Length of a complex number in bytes
-\newcommand{\cbytes}{b} % Length of a real number in bytes
-\newcommand{\rbytes}{b} % Length of a complex number in bytes
-\newcommand{\bl}{B} % Baseline length
-\newcommand{\maxbl}{B_{\mathrm{max}}} % Maximum baseline length
-\newcommand{\membw}{B_\mathrm{mem}} % Memory bandwidth
-\newcommand{\pciebw}{B_\mathrm{PCIe}} % PCI-e bandwidth
-\newcommand{\viscompr}{C} % Compression factor of visibilities for gridding
-                          % for facets
-\newcommand{\vism}{\mathbb{D}} % Matrix of observed visibilities
-\newcommand{\da}{D_{\mathrm A}} % Extent of aperture
-\newcommand{\ds}{D_{\mathrm s}} % Diameter of antennas/stations
-\newcommand{\reff}{f_\mathrm{ref}} % Reference frequency
-\newcommand{\fcc}{F_{\mathrm{cc}}} % Convolution kernel computational cost
-\newcommand{\fci}{F_{\mathrm{ci}}} % Factor between bandwidth and FLOPS (bytes/ops)
-\newcommand{\fgrid}{F_{\mathrm{grid}}} % Gridding cost
-\newcommand{\fpatch}{f_\mathrm{patch}} % Fraction of the maximum baseline below which the uv coverage is almost filled
-\newcommand{\fpr}{F_\mathrm{pr}} % Cost for phase rotation
-\newcommand{\gainm}{\mathbb{G}} % 2x2 block diagonal matrix of complex gains
-% To change to N_something
-\newcommand{\modskyvism}{\mathbb{M}} % Matrix of model sky visibilities
-\newcommand{\skymodbufsize}{M_\mathrm{buf,skymod}} % Size of the sky model buffer
-\newcommand{\visbufsize}{M_\mathrm{buf,vis}} % Size of the visibility buffer
-\newcommand{\visbufsizespec}{M^\mathrm{spec}_\mathrm{buf,vis}} % Size of the visibility buffer needed for spectral line processing
-\newcommand{\visbufsizecont}{M^\mathrm{cont}_\mathrm{buf,vis}} % Size of the visibility buffer needed for continuum imaging
-\newcommand{\bufsize}{M_\mathrm{cu,buf}} % Size of buffer
-\newcommand{\slowmemsize}{M_\mathrm{cu,pool}} %	Size of slow working memory
-\newcommand{\cumemsize}{M_\mathrm{cu,work}} % Size of working memory of compute unit
-\newcommand{\imgsize}{M_\mathrm{image}} % Size of image (bytes)
-\newcommand{\imgsizeoneax}{M_\mathrm{image,ax}}	% Image size along one axis
-\newcommand{\nbyteperpix}{N_\mathrm{byte,pix}} % Number of bytes per pixel
-%\newcommand{\nbytepervis}{N_\mathrm{byte,vis}} % Number of bytes per pixel
-\newcommand{\targgridsize}{M_\mathrm{target}} %	Target grid size (per facet)
-\newcommand{\uvgridsize}{M_{uv,\mathrm{grid}}} % Size of the uv-grid (bytes)
-\newcommand{\vissize}{M_\mathrm{vis}} % Size of visibility data (bytes)
-\newcommand{\wgridsize}{M_\mathrm{weight\_grid}} % Size of weight grid (bytes)
-\newcommand{\gridcachesize}{M_{w,\mathrm{cache}}} % Cache size of
-                                % gridding kernel
-\newcommand{\ntel}{N_\mathrm{a}} % Number of antennas or stations in the array
-\newcommand{\nateam}{N_\mathrm{A}} % Number of A-team sources
-\newcommand{\ngridpakern}{N_\mathrm{AA}} % Number of grid points for A-kernel support (along one axis)
-\newcommand{\naccvis}{N_\mathrm{apv}} % Number of accesses per visibility
-\newcommand{\navg}{N_\mathrm{avg}} % Number of averages
-\newcommand{\nbins}{N_\mathrm{bin}} % Number of bins
-\newcommand{\nbeam}{N_\mathrm{beam}} % Number of beams simultaneously observed by the array
-\newcommand{\rma}{N_{rma}} % Number of flops in real multiply and add
-\newcommand{\cma}{N_{cma}} % Number of flops in complex multiply and add
-\newcommand{\nbits}{N_\mathrm{bit}} %	Number of bits
-\newcommand{\nbl}{N_\mathrm{bl}} % Number of baselines
-\newcommand{\nbyteacc}{N_\mathrm{bpa}} % Number of bytes per access
-\newcommand{\nbyte}{N_\mathrm{byte}} % Number of bytes
-\newcommand{\nch}{N_\mathrm{ch}XXXWANT TO USE NF HERE} % Number of channels
-\newcommand{\ncompu}{N_\mathrm{cu}} % Number of compute units
-\newcommand{\ndatait}{N_\mathrm{datait}} % Number of data items required per source (e.g. postion)
-\newcommand{\ngridp}{N_\mathrm{cvff}} % Number of grid points along one axis for the gridding convolution function
-\newcommand{\nfch}{N_\mathrm{f}} % Number of frequency channels
-\newcommand{\nfchcorr}{N_\mathrm{f,corr}} % Number of frequency channels output by the correlator
-\newcommand{\nfnosmear}{N_\mathrm{f,smear}} % Number of frequency channels set by smearing limit
-\newcommand{\nfchgrid}{N_\mathrm{f,grid}} % Number of frequency channels to be gridded
-\newcommand{\nfchdegrid}{N_\mathrm{f,de-grid}} % Number of frequency channels to be de-gridded
-\newcommand{\nfchout}{N_\mathrm{f,out}} % Number of output frequency channels
-                                % (to be FFT-ed and reprojected or output from
-                                % the CSP pulsar timing backend
-\newcommand{\nfchdist}{N_\mathrm{f,distribute}} % Minimum number of frequencies needed to maintain distributability
-\newcommand{\nfkernel}{N_\mathrm{f,kernel}} %Number of convolution kernels needed to cover frequency axis for a particular baseline
-\newcommand{\nfacet}{N_\mathrm{facet}} % Number of facets along one axis
-\newcommand{\facetoverlap}{r_\mathrm{facet}} %Facet overlap overhead
-\newcommand{\nflop}{N_\mathrm{FLOP}} % Number of floating point operations
-\newcommand{\nflopsamp}{N_\mathrm{FLOPsamp}} % Number of floating point
-                                % operations per sample
-\newcommand{\nflopvis}{N_\mathrm{FLOPvis}} % Number of floating point
-                                % operations per visibility 
-\newcommand{\npixgrid}{N_\mathrm{grid}} % Number of pixels in the grid (i.e. the square of the linear dimension)
-\newcommand{\ngridwkern}{N_\mathrm{GW}} % Number of grid points for w-kernel
-                                % support (along one axis)
-\newcommand{\nit}{N_\mathrm{it}} % Number of iterations
-\newcommand{\nawkern}{N_\mathrm{kernel}} % Linear size of combined A and w-kernels
-\newcommand{\nmaj}{N_\mathrm{major}} % Number of major CLEAN cycles
-\newcommand{\nmin}{N_\mathrm{minor}} % Number of minor CLEAN cycles
-\newcommand{\nselfcal}{N_\mathrm{SelfCal}} %Number of self cal loops
-% TODO: Factors should have another letter than N
-\newcommand{\facgridoffdiagmuller}{N_\mathrm{mm}} % Factor to account for
-                                % gridding of off-diagonal Mueller terms
-% TODO: Factors should have another letter than N
-\newcommand{\nops}{N_\mathrm{ops}} % Number of operations
-\newcommand{\npatchpix}{N_\mathrm{patch,pix}} % Number of pixels of PSF to be used in minor cycle decorrelation
-\newcommand{\npix}{N_\mathrm{pix}} % Number of pixels along one axis of a grid
-\newcommand{\npixfacet}{N_\mathrm{pix,facet}} % Number of pixels along one axis of a single image facet
-\newcommand{\npol}{N_\mathrm{pol}} % Number of polarisation products
-\newcommand{\nsubint}{N_\mathrm{subint}} % Number of sub-integrations
-\newcommand{\ntslot}{N_t} % Number of time slots
-\newcommand{\ntaylor}{N_\mathrm{Tt}} % Number of Taylor terms
-\newcommand{\ntslotave}{N_{t,\mathrm{av}}} % Number of time slots that are averaged (in de-mixing)
-\newcommand{\nvis}{N_\mathrm{vis}} % Number of visibilities
-\newcommand{\nvisrate}{R_{vis}} % Number of visibilities per second\\
-% TODO: what kind of qfac?
-\newcommand{\qfacbw}{Q_\mathrm{bw}} % Quality factor
-\newcommand{\qfacreuse}{Q_\mathrm{fcv}} % Factor to allow for the reuse of functions between neighbouring frequency channels
-\newcommand{\qfacfov}{Q_\mathrm{FoV}} % Quality factor defining how much bigger the diameter of the field of view is than the first zero of the Airy function
-\newcommand{\facgridovers}{Q_\mathrm{GCF}} % Gridding oversampling factor
-\newcommand{\facmsmf}{Q_\mathrm{MSMF}} % Factor to account for multiple subtractions (multi-frequency multi-scale CLEAN)
-\newcommand{\qfacbeamovers}{Q_\mathrm{pix}} % Quality factor defining the
-                                % oversampling of the synthesised beam
-% TODO: what kind of qfac?
-\newcommand{\qfac}{Q_t} % Quality factor
-\newcommand{\corrfacw}{Q_w} % Correction factor for w values (0 to 1)
-\newcommand{\earthrad}{R_\mathrm{Earth}} % Radius of the Earth ($6400\times 10^3$\,m)
-\newcommand{\comprate}{R} % Compute rate (FLOPS)
-\newcommand{\compratespec}{R^\mathrm{spec}} % Compute rate for spectral line processing
-\newcommand{\compratecont}{R^\mathrm{cont}} % Compute rate for continuum imaging
-\newcommand{\compratefast}{R^\mathrm{fast}} % Compute rate for fast imaging
-\newcommand{\covm}{\mathbb{R}} % Covariance matrix
-\newcommand{\covmfilt}{\mathbb{R}_\mathrm{filt}} % Filtered covariance matrix
-\newcommand{\bwtoworkmem}{R_\mathrm{bw,mem}} % Bandwidth to working memory
-\newcommand{\maxmembw}{R_\mathrm{bw,max}} % Maximum memory bandwidth of
-                                % compute unit
-\newcommand{\convkernflops}{R_\mathrm{CCF}} % Convolution kernel compute rate (FLOPS)
-\newcommand{\fftflops}{R_\mathrm{FFT}} % FFT compute rate (FLOPS)
-\newcommand{\flagflops}{R_\mathrm{flag}} % FLOP rate to do flagging on input visibilities
-\newcommand{\griddingflops}{R_\mathrm{grid}} % Gridding compute rate (FLOPS)
-\newcommand{\phaserotflops}{R_\mathrm{PR}} % Phase rotation compute rate (FLOPS)
-\newcommand{\reprojflops}{R_\mathrm{RP}} % Re-projection compute rate (FLOPS)
-\newcommand{\rcorr}{R_\mathrm{s}} % Correlator output rate (samples/sec)
-\newcommand{\peakflop}{R_\mathrm{peak}} % Peak FLOPS of compute unit
-% TODO - verschil met max mem bw of compute unit?
-% TODO - verschil memory bw and bw to buffer?
-\newcommand{\IO}{{R_\mathrm{IO}}} %IO rate
-\newcommand{\maxiobwbuff}{R_\mathrm{bw,I/O,max}} % Maximum I/O bandwidth of compute unit to buffer
-\newcommand{\bwbuff}{R_\mathrm{bw,I/O}} % Bandwidth to buffer
-\newcommand{\bwbuffcont}{R^\mathrm{cont}_\mathrm{bw,I/O}} % Bandwidth to
-                                % buffer for cotinuum imaging
-\newcommand{\bwbuffspec}{R^\mathrm{spec}_\mathrm{bw,I/O}} % Bandwidth to
-                                % buffer for spectral line processing
-\newcommand{\nsource}{N_\mathrm{source}} % Number of sources
-\newcommand{\telsensit}{S} % Telescope sensitivity
-\newcommand{\corrdumpt}{t_\mathrm{dump}} % Correlator dump time (sec)
-\newcommand{\compt}{t_{\mathrm{comp}}} % Computation time
-\newcommand{\gridkernupdate}{t_\mathrm{kernel,update}} % Update timescale for gridding kernel
-\newcommand{\lsmupdate}{t_\mathrm{LSM}} % Update timescale for the LSM
-\newcommand{\obst}{t_{\mathrm{obs}}} % Observation time
-\newcommand{\snapshot}{t_\mathrm{snap}} % Snapshot duration
-\newcommand{\minsnapshot}{t_\mathrm{snap,min}} % Minimum duration of a snapshot in which w-correction is not dominated by Earth's curvature
-\newcommand{\optsnapshot}{t_\mathrm{snap,opt}} % Optimum snapshot duration (sec)
-\newcommand{\subint}{t_\mathrm{subint}} % Subintegration time
-\newcommand{\uv}{uv} % Visibility coordinates
-\newcommand{\uvw}{u,v,w} % Visibility coordinates
-\newcommand{\w}{w} % Visibility w coordinate
-\newcommand{\nbytepervis}{N_\mathrm{byte,vis}} %Memory per individual visibility 
-\newcommand{\maxwdev}{\Delta w_\mathrm{max}} % Maximum deviation of the w
-                                % coordinate due to the Earth's curvature
-\newcommand{\minwdev}{\Delta w_\mathrm{min}} % Minimum deviation of the w
-                                % coordinate due to the Earth's curvature
-\newcommand{\spfiltm}{\mathbb{W}_\mathrm{filt}} % Spatial filter matrix
-\newcommand{\fres}{\Delta f} % Frequency resolution
-\newcommand{\tressmear}{t_\mathrm{smear}} % Smearing Time resolution (sec)
-\newcommand{\tres}{t_{res}} % Time resolution (sec)
-\newcommand{\fracconvfuncamp}{\epsilon_\omega} % Fraction of maximal amplitude of envelope of convolution function
-\newcommand{\compeff}{\eta_\mathrm{comp}} % Computational efficiency
-\newcommand{\fracfband}{\eta_f} % Fraction of frequency band
-\newcommand{\beamsq}{\theta_\mathrm{bs}} % Beam squint
-\newcommand{\fovdiam}{\theta_\mathrm{FoV}} % Angular diameter of the field of view (rad)
-\newcommand{\pixres}{\theta_\mathrm{pix}} % Angular resolution of pixel size (rad)
-\newcommand{\beamdiam}{\theta_\mathrm{PSF}} % Angular diameter of the synthesised beam (rad)
-\newcommand{\wavel}{\lambda} % Wavelength
-\newcommand{\maxwavel}{\lambda_\mathrm{max}} % Maximum wavelength
-\newcommand{\minwavel}{\lambda_\mathrm{min}} % Minimum wavelength
-\newcommand{\maxwavelsubb}{\lambda_{\mathrm{sub,max}}} % Maximum wavelength of an imaging sub-band (channels over which image and pixel size are matched)
-\newcommand{\minwavelsubb}{\lambda_\mathrm{sub,min}} % Minimum wavelength of an imaging sub-band (channels over which image and pixel size are matched)
-\newcommand{\earthrot}{\Omega_\mathrm{E}} % Angular velocity of the Earth (7.27 10-5 rad/sec)
-\newcommand{\epsilonf}{\epsilon_f}  %Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging
-% Other command definitions
-\newcommand{\Qkernel}{Q_\mathrm{kernel}} %Convolution kernel quality factor, for kernel re-use in frequency
-\newcommand{\bs}{\texttt{\char`\\}}
 
@@
 \newcommand{\He}{{\tens{H}}}
-\newcommand{\Sa}{{\tens{S}}}
+\newcommand{\Sa}{{\tens{S}{N}}}
 \newcommand{\Sd}{{\tens{S^\dag}}}
-\newcommand{\Sna}{\tens{{S_{\nu}}}}
+\newcommand{\Sna}{\tens{{S_{\nu}}{N}}}
 \newcommand{\Snd}{\tens{{S_{\nu}^\dag}}}
@@
 % Add here the symbols used in your document
-\clearpage \sdplistofsymbols
-\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
-\item[$\bytes$] Length of a complex number in bytes \vspace{-0.2cm}
-\item[$\cbytes$] Length of a complex number in bytes \vspace{-0.2cm}
-\item[$\rbytes$] Length of a real number in bytes \vspace{-0.2cm}
-\item[$\bl$] Baseline length / GPU memory bandwidth \vspace{-0.2cm}
-\item[$\maxbl$] Maximum baseline length \vspace{-0.2cm}
-\item[$\membw$] Memory bandwidth \vspace{-0.2cm}
-\item [$\pciebw$] PCI-e bandwidth \vspace{-0.2cm}
-\item[$\viscompr$] Compression factor of visibilities for grdding for facets
-  \textcolor{red}{This symbol is not used in the document, consider removing.}
-  \vspace{-0.2cm}
-\item[$\vism$] Matrix of observed visibilities \vspace{-0.2cm}
-\item[$\da$] Extent of aperture \vspace{-0.2cm}
-\item[$\ds$] Diameter of antennas/stations \vspace{-0.2cm}
-\item[$\reff$] Reference frequency\vspace{-0.2cm}
-\item[$\fcc$] Convolution kernel computational cost\vspace{-0.2cm}
-\item[$\fci$] Factor between bandwidth and FLOPS (bytes/operations)\vspace{-0.2cm}
-\item[$\fgrid$] Gridding cost\vspace{-0.2cm}
-\item[$\fpatch$] Fraction of the maximum baseline below which the $\uv$
-  coverage is almost filled\vspace{-0.2cm}
-\item[$\fpr$] Cost for phase rotation\vspace{-0.2cm}
-\item[$\gainm$] 2$\times$2 block diagonal matrix of complex
-  gains\vspace{-0.2cm}
-\item[$\modskyvism$] Matrix of model sky visibilities\vspace{-0.2cm}
-\item[$\skymodbufsize$] Size of the sky model buffer\vspace{-0.2cm}
-\item[$\visbufsize$] Size of the visibility buffer \vspace{-0.2cm}
-\item[$\visbufsizespec$] Size of the visibility buffer needed for spectral line processing\vspace{-0.2cm}
-\item[$\visbufsizecont$] Size of the visibility buffer needed for continuum imaging\vspace{-0.2cm}
-\item[$\bufsize$] Size of buffer \vspace{-0.2cm}
-\item[$\slowmemsize$] Size of slow working memory \vspace{-0.2cm}
-\item[$\cumemsize$] Size of working memory of compute unit \vspace{-0.2cm}
-\item[$\imgsize$] Size of image (bytes) \vspace{-0.2cm}
-\item[$\imgsizeoneax$] Image size along one axis \vspace{-0.2cm}
-\item[$\targgridsize$] Target grid size (per facet) \vspace{-0.2cm}
-\item[$\uvgridsize$] Size of the $\uv$ grid (bytes) \vspace{-0.2cm}
-\item[$\vissize$] Size of visibility data (bytes) \vspace{-0.2cm}
-\item[$\wgridsize$] Size of weight grid (bytes)\vspace{-0.2cm}
-\item[$\gridcachesize$] Cache size of gridding kernel \vspace{-0.2cm}
-\item[$\nbins$] Number of bins \vspace{-0.2cm}
-\item[$\ntel$] Number of antennas or stations in the array
-  \vspace{-0.2cm}
-\item[$\nateam$] Number of A-team sources \vspace{-0.2cm}
-\item[$\ngridpakern$] Number of grid points for A-kernel support (along one
-  axis) \vspace{-0.2cm}
-\item[$\naccvis$] Number of accesses per visibility \vspace{-0.2cm}
-\item[$\navg$] Number of averages\vspace{-0.2cm}
-\item[$\nbeam$] Number of beams simultaneously observed by the array
-  \vspace{-0.2cm}
-\item[$\nbits$] Number of bits\vspace{-0.2cm}
-\item[$\rma$] Number of flops in real multiply and add\vspace{-0.2cm}
-\item[$\cma$] Number of flops in complex multiply and add\vspace{-0.2cm}
-\item[$\nbl$] Number of baselines \vspace{-0.2cm}
-\item[$\nbyteacc$] Number of bytes per access \vspace{-0.2cm}
-\item[$\nbyte$] Number of bytes \vspace{-0.2cm}
-\item[$\nbyteperpix$] Number of bytes per pixel \vspace{-0.2cm}
-\item[$\nbytepervis$] Number of bytes per individual visibility
-
-\item[$\nch$] Number of channels \vspace{-0.2cm}
-\item[$\ncompu$] Number of compute units \vspace{-0.2cm}
-\item[$\ndatait$] Number of data items required per source (e.g. position)\vspace{-0.2cm}
-\item[$\ngridp$] Number of grid points along one axis for the gridding
-  convolution function \vspace{-0.2cm}
-\item[$\nfch$] Number of frequency channels/sub-bands \vspace{-0.2cm}
-\item[$\nfchcorr$] Number of frequency channels output by the correlator
-  \vspace{-0.2cm}
-\item[$\nfchgrid$] Number of frequency channels to be gridded
-  \vspace{-0.2cm}
-\item[$\nfchdegrid$] Number of frequency channels to be de-gridded
-  \vspace{-0.2cm}
-\item[$\nfchout$] Number of frequency channels to be FFT-ed and reprojected
-  (output) or output from the CSP pulsar timing backend\vspace{-0.2cm}
-\item[$\nfchdist$] Minimum number of frequency channels to maintain distributability in the SDP
-\item[$\nfkernel$] Number of convolution kernels needed to cover frequency axis for a given baseline
-\item[$\nfacet$] Number of facets along one axis \vspace{-0.2cm}
-\item[$\nflop$] Number of floating point operations
-  \vspace{-0.2cm}
-\item[$\nflopsamp$] Number of floating point operations per sample
-  \vspace{-0.2cm}
-\item[$\nflopvis$] Number of floating point operations per visibility
-\item[$\npixgrid$] Number of pixels in the grid (i.e. the square of the linear
-  dimension) \vspace{-0.2cm}
-\item[$\ngridwkern$] Number of grid points for w-kernel support (along one
-  axis) \vspace{-0.2cm}
-\item[$\nit$] Number of iterations \vspace{-0.2cm}
-\item[$\nawkern$] Linear size of combined A and $\w$-kernels \vspace{-0.2cm}
-\item[$\nmaj$] Number of major CLEAN cycles (within each self-cal loop if appropriate) \vspace{-0.2cm}
-\item[$\nselfcal$] Number of self-calibration cycles \vspace{-0.2cm}
-\item[$\nmin$] Number of minor CLEAN cycles (within each major cycle loop if appropriate)\vspace{-0.2cm}
-\item[$\facgridoffdiagmuller$] Factor to account for gridding of off-diagonal
-  M\"uller terms \vspace{-0.2cm}
-\item[$\nops$] Number of operations \vspace{-0.2cm}
-\item[$\npatchpix$] Number of pixels of PSF to be used in minor cycle
-  decorrelation \vspace{-0.2cm}
-\item[$\npix$] Number of pixels along one axis of a grid\vspace{-0.2cm}
-\item[$\npixfacet$] Number of pixels on one side of a facet \vspace{-0.2cm}
-\item[$\npol$] Number of polarisation products \vspace{-0.2cm}
-\item[$\nsource$] Number of sources \vspace{-0.2cm}
-\item[$\ntslot$] Number of time slots \vspace{-0.2cm}
-\item[$\ntaylor$] Number of Taylor terms (in MS-MFS)\vspace{-0.2cm}
-\item[$\ntslotave$] Number of time slots that are averaged (in de-mixing)
-  \vspace{-0.2cm}
-\item[$\nvis$] Number of visibilities in an observation \vspace{-0.2cm}
-\item[$\qfacbw$] Quality factor \vspace{-0.2cm}
-\item[$\qfacreuse$] Factor to allow for the reuse of functions between
-  neighbouring frequency channels \vspace{-0.2cm}
-\item[$\qfacfov$] Quality factor defining how much bigger the diameter of the
-  field of view is than the first zero of the Airy function \vspace{-0.2cm}
-\item[$\facgridovers$] Gridding oversampling factor \vspace{-0.2cm}
-\item[$\facmsmf$] Factor to account for multiple subtractions (multi-frequency
-  multi-scale CLEAN\vspace{-0.2cm}
-\item[$\qfacbeamovers$] Quality factor defining the oversampling of the
-  synthesised beam \vspace{-0.2cm}
-\item[$\qfac$] Quality factor \vspace{-0.2cm}
-\item[$\Qkernel$] Convolution kernel quality factor, for kernel re-use in frequency
-\item[$\corrfacw$] Correction factor for w values (0 to 1) \vspace{-0.2cm}
-\item[$\earthrad$] Radius of the Earth ($6400\times 10^3$\,m) \vspace{-0.2cm}
-\item[$\covm$] Covariance matrix \vspace{-0.2cm}
-\item[$\covmfilt$] Filtered covariance matrix \vspace{-0.2cm}
-\item[$\comprate$] Compute rate (FLOPS) \vspace{-0.2cm}
-\item[$\compratespec$] Compute rate (FLOPS) for spectral line processing \vspace{-0.2cm}
-\item[$\compratecont$] Compute rate (FLOPS) for continuum imaging \vspace{-0.2cm}
-\item[$\compratefast$] Compute rate (FLOPS) for fast imaging \vspace{-0.2cm}
-\item[$\bwtoworkmem$] Bandwidth to working memory \vspace{-0.2cm}
-\item[$\maxmembw$] Maximum memory bandwidth of compute unit \vspace{-0.2cm}
-\item[$\peakflop$] Peak FLOPS of compute unit \vspace{-0.2cm}
-\item[$\maxiobwbuff$] Maximum I/O bandwidth of compute unit to buffer
-  \vspace{-0.2cm}
-\item[$\bwbuff$] Bandwidth to buffer \vspace{-0.2cm}
-\item[$\bwbuffcont$] Bandwidth to buffer for continuum imaging\vspace{-0.2cm}
-\item[$\bwbuffspec$] Bandwidth to buffer for spectral line processing\vspace{-0.2cm}
-\item[$\convkernflops$] Convolution kernel compute rate (FLOPS)\vspace{-0.2cm}
-\item[$\fftflops$] FFT compute rate (FLOPS)\vspace{-0.2cm}
-\item[$\griddingflops$] Gridding compute rate (FLOPS)\vspace{-0.2cm}
-\item[$\phaserotflops$] Phase rotation compute rate (FLOPS)\vspace{-0.2cm}
-\item[$\reprojflops$] Re-projection compute rate (FLOPS)\vspace{-0.2cm}
-\item[$\rcorr$] Correlator output rate\vspace{-0.2cm} TRY TO REMOVE
-\item[$\facetoverlap$]  ratio for linear size increase of facets to account for overlap, typically 1.2-1.5 \vspace{-0.2cm}
-\item[$\telsensit$] Telescope sensitivity \vspace{-0.2cm} TRY TO REMOVE
-\item[$\corrdumpt$] Correlator dump time (sec) \vspace{-0.2cm}
-\item[$\compt$] Computation time \vspace{-0.2cm}
-\item[$\gridkernupdate$] Update timescale for gridding kernel \vspace{-0.2cm}
-\item[$\lsmupdate$] Update timescale for the LSM \vspace{-0.2cm}
-\item[$\obst$] Observation time \vspace{-0.2cm}
-\item[$\snapshot$] Snapshot duration \vspace{-0.2cm}
-\item[$\minsnapshot$] Minimum duration of a snapshot in which w-correction is
-  not dominated by the Earth's curvature \vspace{-0.2cm}
-\item[$\optsnapshot$] Optimum snapshot duration (sec) \vspace{-0.2cm}
-\item[$\subint$] Sub-integration time\vspace{-0.2cm}
-\item[$\uvw$] Visibility coordinates \vspace{-0.2cm}
-\item [$\w$] Visibility $w$ coordinate \vspace{-0.2cm}
-\item[$\maxwdev$] Maximum deviation of the $w$ coordinate due to the Earth's
-  curvature \vspace{-0.2cm}
-\item[$\minwdev$] Minimum deviation of the $w$ coordinate due to the Earth's
-  curvature \vspace{-0.2cm}
-\item[$\spfiltm$] Spatial filter matrix \vspace{-0.2cm}
-\item[$\fres$] Frequency resolution \vspace{-0.2cm}
-\item[$\tres$] Time resolution (sec) \vspace{-0.2cm}
-\item[$\tressmear$]  Smearing Time resolution (sec)  \vspace{-0.2cm}
-\item[$\fracconvfuncamp$] Fraction of maximal amplitude of envelope of
-  convolution function \vspace{-0.2cm}
-\item[$\compeff$] Computational efficiency \vspace{-0.2cm}
-\item[$\fracfband$] Fraction of frequency band \vspace{-0.2cm}
-\item[$\beamsq$] Beam squint \vspace{-0.2cm}
-\item[$\fovdiam$] Angular diameter of the field of view (rad) \vspace{-0.2cm}
-\item[$\pixres$] Angular resolution of pixel size (rad) \vspace{-0.2cm}
-\item[$\beamdiam$] Angular diameter of the synthesised beam (rad)
-  \vspace{-0.2cm}
-\item[$\wavel$] Wavelength \vspace{-0.2cm}
-\item[$\maxwavel$] Maximum wavelength \vspace{-0.2cm}
-\item[$\minwavel$] Minimum wavelength \vspace{-0.2cm}
-\item[$\maxwavelsubb$] Maximum wavelength of an imaging sub-band (channels
-  over which image and pixel size are matched) \vspace{-0.2cm}
-\item[$\minwavelsubb$] Minimum wavelength of an imaging sub-band (channels
-  over which image and pixel size are matched) \vspace{-0.2cm}
-\item[$\earthrot$] Angular velocity of the Earth
-  ($7.27\times10^{-5}$\,rad/sec) \vspace{-0.2cm}
-\item[$\epsilonf$] Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging \vspace{-0.2cm}
-\end{basedescript}
+%\clearpage \sdplistofsymbols
+%\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
+%\item[$\bytes$] Length of a complex number in bytes \vspace{-0.2cm}
+%\item[$\cbytes$] Length of a complex number in bytes \vspace{-0.2cm}
+%\item[$\rbytes$] Length of a real number in bytes \vspace{-0.2cm}
+%\item[$\bl$] Baseline length / GPU memory bandwidth \vspace{-0.2cm}
+%\item[$\maxbl$] Maximum baseline length \vspace{-0.2cm}
+%\item[$\membw$] Memory bandwidth \vspace{-0.2cm}
+%\item [$\pciebw$] PCI-e bandwidth \vspace{-0.2cm}
+%\item[$\viscompr$] Compression factor of visibilities for grdding for facets
+%  \textcolor{red}{This symbol is not used in the document, consider removing.}
+%  \vspace{-0.2cm}
+%\item[$\vism$] Matrix of observed visibilities \vspace{-0.2cm}
+%\item[$\da$] Extent of aperture \vspace{-0.2cm}
+%\item[$\ds$] Diameter of antennas/stations \vspace{-0.2cm}
+%\item[$\reff$] Reference frequency\vspace{-0.2cm}
+%\item[$\fcc$] Convolution kernel computational cost\vspace{-0.2cm}
+%\item[$\fci$] Factor between bandwidth and FLOPS (bytes/operations)\vspace{-0.2cm}
+%\item[$\fgrid$] Gridding cost\vspace{-0.2cm}
+%\item[$\fpatch$] Fraction of the maximum baseline below which the $\uv$
+%  coverage is almost filled\vspace{-0.2cm}
+%\item[$\fpr$] Cost for phase rotation\vspace{-0.2cm}
+%\item[$\gainm$] 2$\times$2 block diagonal matrix of complex
+%  gains\vspace{-0.2cm}
+%\item[$\modskyvism$] Matrix of model sky visibilities\vspace{-0.2cm}
+%\item[$\skymodbufsize$] Size of the sky model buffer\vspace{-0.2cm}
+%\item[$\visbufsize$] Size of the visibility buffer \vspace{-0.2cm}
+%\item[$\visbufsizespec$] Size of the visibility buffer needed for spectral line processing\vspace{-0.2cm}
+%\item[$\visbufsizecont$] Size of the visibility buffer needed for continuum imaging\vspace{-0.2cm}
+%\item[$\bufsize$] Size of buffer \vspace{-0.2cm}
+%\item[$\slowmemsize$] Size of slow working memory \vspace{-0.2cm}
+%\item[$\cumemsize$] Size of working memory of compute unit \vspace{-0.2cm}
+%\item[$\imgsize$] Size of image (bytes) \vspace{-0.2cm}
+%\item[$\imgsizeoneax$] Image size along one axis \vspace{-0.2cm}
+%\item[$\targgridsize$] Target grid size (per facet) \vspace{-0.2cm}
+%\item[$\uvgridsize$] Size of the $\uv$ grid (bytes) \vspace{-0.2cm}
+%\item[$\vissize$] Size of visibility data (bytes) \vspace{-0.2cm}
+%\item[$\wgridsize$] Size of weight grid (bytes)\vspace{-0.2cm}
+%\item[$\gridcachesize$] Cache size of gridding kernel \vspace{-0.2cm}
+%\item[$\nbins$] Number of bins \vspace{-0.2cm}
+%\item[$\ntel$] Number of antennas or stations in the array
+%  \vspace{-0.2cm}
+%\item[$\nateam$] Number of A-team sources \vspace{-0.2cm}
+%\item[$\ngridpakern$] Number of grid points for A-kernel support (along one
+%  axis) \vspace{-0.2cm}
+%\item[$\naccvis$] Number of accesses per visibility \vspace{-0.2cm}
+%\item[$\navg$] Number of averages\vspace{-0.2cm}
+%\item[$\nbeam$] Number of beams simultaneously observed by the array
+%  \vspace{-0.2cm}
+%\item[$\nbits$] Number of bits\vspace{-0.2cm}
+%\item[$\rma$] Number of flops in real multiply and add\vspace{-0.2cm}
+%\item[$\cma$] Number of flops in complex multiply and add\vspace{-0.2cm}
+%\item[$\nbl$] Number of baselines \vspace{-0.2cm}
+%\item[$\nbyteacc$] Number of bytes per access \vspace{-0.2cm}
+%\item[$\nbyte$] Number of bytes \vspace{-0.2cm}
+%\item[$\nbyteperpix$] Number of bytes per pixel \vspace{-0.2cm}
+%\item[$\nbytepervis$] Number of bytes per individual visibility 
+%
+%\item[$\nch$] Number of channels \vspace{-0.2cm}
+%\item[$\ncompu$] Number of compute units \vspace{-0.2cm}
+%\item[$\ndatait$] Number of data items required per source (e.g. position)\vspace{-0.2cm}
+%\item[$\ngridp$] Number of grid points along one axis for the gridding
+%  convolution function \vspace{-0.2cm}
+%\item[$\nfch$] Number of frequency channels/sub-bands \vspace{-0.2cm}
+%\item[$\nfchcorr$] Number of frequency channels output by the correlator
+%  \vspace{-0.2cm}
+%\item[$\nfchgrid$] Number of frequency channels to be gridded
+%  \vspace{-0.2cm}
+%\item[$\nfchdegrid$] Number of frequency channels to be de-gridded
+%  \vspace{-0.2cm}
+%\item[$\nfchout$] Number of frequency channels to be FFT-ed and reprojected
+%  (output) or output from the CSP pulsar timing backend\vspace{-0.2cm}
+%\item[$\nfchdist$] Minimum number of frequency channels to maintain distributability in the SDP
+%\item[$\nfkernel$] Number of convolution kernels needed to cover frequency axis for a given baseline
+%\item[$\nfacet$] Number of facets along one axis \vspace{-0.2cm}
+%\item[$\nflop$] Number of floating point operations
+%  \vspace{-0.2cm}
+%\item[$\nflopsamp$] Number of floating point operations per sample
+%  \vspace{-0.2cm}
+%\item[$\nflopvis$] Number of floating point operations per visibility
+%\item[$\npixgrid$] Number of pixels in the grid (i.e. the square of the linear
+%  dimension) \vspace{-0.2cm}
+%\item[$\ngridwkern$] Number of grid points for w-kernel support (along one
+%  axis) \vspace{-0.2cm}
+%\item[$\nit$] Number of iterations \vspace{-0.2cm}
+%\item[$\nawkern$] Linear size of combined A and $\w$-kernels \vspace{-0.2cm}
+%\item[$\nmaj$] Number of major CLEAN cycles (within each self-cal loop if appropriate) \vspace{-0.2cm}
+%\item[$\nselfcal$] Number of self-calibration cycles \vspace{-0.2cm}
+%\item[$\nmin$] Number of minor CLEAN cycles (within each major cycle loop if appropriate)\vspace{-0.2cm}
+%\item[$\facgridoffdiagmuller$] Factor to account for gridding of off-diagonal
+%  M\"uller terms \vspace{-0.2cm}
+%\item[$\nops$] Number of operations \vspace{-0.2cm}
+%\item[$\npatchpix$] Number of pixels of PSF to be used in minor cycle
+%  decorrelation \vspace{-0.2cm}
+%\item[$\npix$] Number of pixels along one axis of a grid\vspace{-0.2cm}
+%\item[$\npixfacet$] Number of pixels on one side of a facet \vspace{-0.2cm}
+%\item[$\npol$] Number of polarisation products \vspace{-0.2cm}
+%\item[$\nsource$] Number of sources \vspace{-0.2cm}
+%\item[$\ntslot$] Number of time slots \vspace{-0.2cm}
+%\item[$\ntaylor$] Number of Taylor terms (in MS-MFS)\vspace{-0.2cm}
+%\item[$\ntslotave$] Number of time slots that are averaged (in de-mixing)
+%  \vspace{-0.2cm}
+%\item[$\nvis$] Number of visibilities in an observation \vspace{-0.2cm}
+%\item[$\qfacbw$] Quality factor \vspace{-0.2cm}
+%\item[$\qfacreuse$] Factor to allow for the reuse of functions between
+%  neighbouring frequency channels \vspace{-0.2cm}
+%\item[$\qfacfov$] Quality factor defining how much bigger the diameter of the
+%  field of view is than the first zero of the Airy function \vspace{-0.2cm}
+%\item[$\facgridovers$] Gridding oversampling factor \vspace{-0.2cm}
+%\item[$\facmsmf$] Factor to account for multiple subtractions (multi-frequency
+%  multi-scale CLEAN\vspace{-0.2cm}
+%\item[$\qfacbeamovers$] Quality factor defining the oversampling of the
+%  synthesised beam \vspace{-0.2cm}
+%\item[$\qfac$] Quality factor \vspace{-0.2cm}
+%\item[$\Qkernel$] Convolution kernel quality factor, for kernel re-use in frequency
+%\item[$\corrfacw$] Correction factor for w values (0 to 1) \vspace{-0.2cm}
+%\item[$\earthrad$] Radius of the Earth ($6400\times 10^3$\,m) \vspace{-0.2cm}
+%\item[$\covm$] Covariance matrix \vspace{-0.2cm}
+%\item[$\covmfilt$] Filtered covariance matrix \vspace{-0.2cm}
+%\item[$\comprate$] Compute rate (FLOPS) \vspace{-0.2cm}
+%\item[$\compratespec$] Compute rate (FLOPS) for spectral line processing \vspace{-0.2cm}
+%\item[$\compratecont$] Compute rate (FLOPS) for continuum imaging \vspace{-0.2cm}
+%\item[$\compratefast$] Compute rate (FLOPS) for fast imaging \vspace{-0.2cm}
+%\item[$\bwtoworkmem$] Bandwidth to working memory \vspace{-0.2cm}
+%\item[$\maxmembw$] Maximum memory bandwidth of compute unit \vspace{-0.2cm}
+%\item[$\peakflop$] Peak FLOPS of compute unit \vspace{-0.2cm}
+%\item[$\maxiobwbuff$] Maximum I/O bandwidth of compute unit to buffer
+%  \vspace{-0.2cm}
+%\item[$\bwbuff$] Bandwidth to buffer \vspace{-0.2cm}
+%\item[$\bwbuffcont$] Bandwidth to buffer for continuum imaging\vspace{-0.2cm}
+%\item[$\bwbuffspec$] Bandwidth to buffer for spectral line processing\vspace{-0.2cm}
+%\item[$\convkernflops$] Convolution kernel compute rate (FLOPS)\vspace{-0.2cm}
+%\item[$\fftflops$] FFT compute rate (FLOPS)\vspace{-0.2cm}
+%\item[$\griddingflops$] Gridding compute rate (FLOPS)\vspace{-0.2cm}
+%\item[$\phaserotflops$] Phase rotation compute rate (FLOPS)\vspace{-0.2cm}
+%\item[$\reprojflops$] Re-projection compute rate (FLOPS)\vspace{-0.2cm}
+%\item[$\rcorr$] Correlator output rate\vspace{-0.2cm} TRY TO REMOVE
+%\item[$\facetoverlap$]  ratio for linear size increase of facets to account for overlap, typically 1.2-1.5 \vspace{-0.2cm}
+%\item[$\telsensit$] Telescope sensitivity \vspace{-0.2cm} TRY TO REMOVE
+%\item[$\corrdumpt$] Correlator dump time (sec) \vspace{-0.2cm}
+%\item[$\compt$] Computation time \vspace{-0.2cm}
+%\item[$\gridkernupdate$] Update timescale for gridding kernel \vspace{-0.2cm}
+%\item[$\lsmupdate$] Update timescale for the LSM \vspace{-0.2cm}
+%\item[$\obst$] Observation time \vspace{-0.2cm}
+%\item[$\snapshot$] Snapshot duration \vspace{-0.2cm}
+%\item[$\minsnapshot$] Minimum duration of a snapshot in which w-correction is
+%  not dominated by the Earth's curvature \vspace{-0.2cm}
+%\item[$\optsnapshot$] Optimum snapshot duration (sec) \vspace{-0.2cm}
+%\item[$\subint$] Sub-integration time\vspace{-0.2cm}
+%\item[$\uvw$] Visibility coordinates \vspace{-0.2cm}
+%\item [$\w$] Visibility $w$ coordinate \vspace{-0.2cm}
+%\item[$\maxwdev$] Maximum deviation of the $w$ coordinate due to the Earth's
+%  curvature \vspace{-0.2cm}
+%\item[$\minwdev$] Minimum deviation of the $w$ coordinate due to the Earth's
+%  curvature \vspace{-0.2cm}
+%\item[$\spfiltm$] Spatial filter matrix \vspace{-0.2cm}
+%\item[$\fres$] Frequency resolution \vspace{-0.2cm}
+%\item[$\tres$] Time resolution (sec) \vspace{-0.2cm}
+%\item[$\tressmear$]  Smearing Time resolution (sec)  \vspace{-0.2cm}
+%\item[$\fracconvfuncamp$] Fraction of maximal amplitude of envelope of
+%  convolution function \vspace{-0.2cm}
+%\item[$\compeff$] Computational efficiency \vspace{-0.2cm}
+%\item[$\fracfband$] Fraction of frequency band \vspace{-0.2cm}
+%\item[$\beamsq$] Beam squint \vspace{-0.2cm}
+%\item[$\fovdiam$] Angular diameter of the field of view (rad) \vspace{-0.2cm}
+%\item[$\pixres$] Angular resolution of pixel size (rad) \vspace{-0.2cm}
+%\item[$\beamdiam$] Angular diameter of the synthesised beam (rad)
+%  \vspace{-0.2cm}
+%\item[$\wavel$] Wavelength \vspace{-0.2cm}
+%\item[$\maxwavel$] Maximum wavelength \vspace{-0.2cm}
+%\item[$\minwavel$] Minimum wavelength \vspace{-0.2cm}
+%\item[$\maxwavelsubb$] Maximum wavelength of an imaging sub-band (channels
+%  over which image and pixel size are matched) \vspace{-0.2cm}
+%\item[$\minwavelsubb$] Minimum wavelength of an imaging sub-band (channels
+%  over which image and pixel size are matched) \vspace{-0.2cm}
+%\item[$\earthrot$] Angular velocity of the Earth
+%  ($7.27\times10^{-5}$\,rad/sec) \vspace{-0.2cm}
+%\item[$\epsilonf$] Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging \vspace{-0.2cm}
+%\end{basedescript}
 
@@
 
- \begin{center}{
+ \begin{center}
  \begin{tabularx}{\textwidth}{|l|X|}
@@
      \hline
-   \end{tabularx}}
+   \end{tabularx}
 \end{center}
 
+\clearpage
 \subsection*{Reference Documents}
@@
    {\bf [RD07]} & SKA-TEL-SDP-0000031, M. Johnston-Hollitt: Science Analysis Pipeline\\
-    {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent
-    Averaging\\
+   {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent Averaging\\
    {\bf [RD09]} & SKA-TEL-SDP-0000058, S. Salvini: Fast Fourier Transforms\\
    {\bf [RD10]} & SKA-TEL-SDP-0000057, C. Skipper: Time and Channel Averaging\\
-    {\bf [RD11]} & Alexander et al., Software and Computing CoDR, Analysis of requirements derived from the DRM, WP2-050.020.010-RR-001\\
-    {\bf [RD12]} & SKA-TEL-SDP-0000016, R. Bolton: SDP Archive Size
-    Estimates\\
    {\bf [RD13]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
-    {\bf [RD14]} & \url{https://confluence.ska-sdp.org/pages/viewpage.action?title=MS+in+UML&spaceKey=ARCH}\\
-    {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821,
-    \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
-    {\bf [RD16]} & \url{http://www.lofar.org/operations/lib/exe/fetch.php?media=:public:documents:casa_image_for_lofar_0.03.00.pdf}\\
-    {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis
-    imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
+   {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
+   {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
    {\bf [RD18]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
-    {\bf [RD19]} & Cornwell, T. J., Voronkov, M. A., Humphreys, B. (2012). Wide field imaging for the Square Kilometre Array.\\
-    {\bf [RD20]} & An Efficient Work-Distribution Strategy for Gridding Radio-Telescope Data on GPUs, John W.
-Romein, ICS’12, June 25–29, 2012, San Servolo Island, Venice, Italy.
-Source code available at: \url{ftp://ftp.astron.nl/outgoing/romein/Gridding-0.2.tar.bz}\\
-   {\bf [RD21]} & Mitchell, D. a. (2014). Analysis of w-projection kernel size. SKA SDP Consortium\\
    {\bf [RD22]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
-   {\bf [RD23]} & A parametric model for the calibration and imaging costs of SKA1, T.J. Cornwell, 12 Feb 2013, SKA Draft Memo\\
    {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
-     \hline
-   \end{tabularx}}
-\end{center}
-
- \begin{center}{
- \begin{tabularx}{\textwidth}{|l|X|}
-     \hline
-     \bf{Reference} & \bf{Reference}\\
-     \bf{Number} & \\
-     \hline
-   {\bf [RD25]} & Tasse, C. a. (2013). Applying full polarization A-Projection to very wide field of view instruments: An imager for LOFAR. A\&A.\\
-   {\bf [RD26]} & Faceted imaging in AIPS, Kogan and Greisen, AIPS Memo 113, 2009\\
-   {\bf [RD27]} & Radio-interferometric imaging of very large fields - The
-   problem of non-coplanar arrays, Cornwell, T. J.; Perley, R. A., Astronomy
-   and Astrophysics (ISSN 0004-6361), vol. 261, no. 1, p. 353-364\\
-    {\bf [RD28]} & Wide Field Imaging at Low Frequencies, Sault, R.
-J.,  The Universe at Low Radio Frequencies, Proceedings of IAU Symposium 199, held 30 Nov - 4 Dec 1999, Pune, India.
-Edited by A.
-Pramesh Rao, G.
-Swarup, and Gopal-Krishna, 2002., p.508\\
-    {\bf [RD29]} & Roofline: An Insightful Visual Performance Model for
-    Multicore Architectures, Williams, Waterman and Patterson,
-    DOI:10.1145/1498765.1498785, April 2009, Vol. 52, No. 4, Communications of
-    the ACM\\
-    {\bf [RD30]} & Whiting, M., 2012, Duchamp: a 3D source finder for spectral-line data, Mon. Not. R. Astron. Soc., Volume 421, Issue 4,  pages 3242–3256, April 2012 \\
-    {\bf [RD31]} & Hales, C.
-A., Murphy, T., Curran, J.
-R., Middelberg, E., Gaensler, B.
-M., Norris, R.
-P., blobcat: software to catalogue flood-filled blobs in radio images of total intensity and linear polarization, Mon.
-Not.
-R.
-Astron.
-Soc., Volume 425,  Issue 2, pages 1365-2966, September 2012\\
-    {\bf [RD32]} & Hancock, P.J., Murphy, T., Gaensler, B.M., Hopkins, A.,
-    Curran, J.R., (2012), Compact continnum source finding for next generation
-    radio surveys, Mon. Not. R. Astron. Soc. 422, 1812–1824\\
-   {\bf [RD33]} & Whiting, M.,  Humphreys, B., (2012) Source-Finding for the
-   Australian Square Kilometre Array Pathfinder, Publications of the
-   Astronomical Society of Australia, Volume 29, Special Issue 03, 2012,
-   pp 371-381.\\
-   {\bf [RD34]} Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
-   {\bf [RD35]} Conway, J.
+   {\bf [RD34]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
+   {\bf [RD35]} & Conway, J.
 E., Cornwell, T.
-J., & Wilkinson, P.
+J., and Wilkinson, P.
 N.
@@
 Monthly Notices of the Royal Astronomical Society, 246, 490.\\
-
    \hline
- \end{tabularx}}
+ \end{tabularx}
 \end{center}
@@
 
-\subsection{Purpose of the document}
+\section{Purpose of the document}
 
 The purpose of this document is to describe the Multi-Scale MultiFrequency Synthesis algorithm, the variant implementations in CASA and ASKAPsoft, and the scaling for insertion in the SDP Performance Model.
+We conclude with some recommendations.
 
-\subsection{Scope of the document}
+\section{Scope of the document}
 
-\clearpage
-\section{Overview of MSMFS}
-\label{sec:overview}
+\section{Background}
 
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\textwidth]{./MSMFS_Predict_UV.pdf}
-  \caption{Structure of UV-weighting Predict, as used in ASKAPsoft. }
-  \label{fig:predictuv}
-\end{figure}
+\clearpage
+\section{Overview of MSMFS}
+\label{sec:overview}
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_UV.pdf}
-  \caption{Structure of UV-weighting MakeImage, as used in ASKAPsoft. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
-  \label{fig:predictimage}
-\end{figure}
+\subsection{Mathematical description}
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\textwidth]{./MSMFS_Predict_Image.pdf}
-  \caption{Structure of Image plane-weighting Predict, as used in CASA. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
-  \label{fig:makeimageuv}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_Image.pdf}
-  \caption{Structure of Image plane-weighting MakeImage, as used in CASA. Note that the three loops (denoted by the rectangle with curved ends) can be permuted at will.}
-  \label{fig:makeimageimage}
-\end{figure}
-
-\clearpage
-\section{Serial}
-\label{sec:serial}
-
-The serial algorithm is as shown in Algorithm \ref{alg:MSMFS_1}.
+The MSMFS algorithm decomposes the image into a set of blobs, using a greedy algorithm in which the peak (in scale space) is found by searching for the peak in the reference frequency residual images each convolved with scale.
+For bright pixels, the principal solution is calculated by applying the inverse Hessian of the PSF convolved with scale.
+For faint pixels, just the peak in the reference frequency image is used.
 
@@
 
-In words, the MSMFS algorithm decomposes the image into a set of blobs, using a greedy algorithm in which the peak (in scale space) is found by searching for the peak in the reference frequency residual images each convolved with scale.
-For bright pixels, the principal solution is calculated by applying the inverse Hessian of the PSF convolved with scale.
-For faint pixels, just the peak in the reference frequency image is used.
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_Predict_UV.pdf}
+  \caption{Structure of UV-weighting Predict, as used in ASKAPsoft. }
+  \label{fig:predictuv}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_UV.pdf}
+  \caption{Structure of UV-weighting MakeImage, as used in ASKAPsoft. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
+  \label{fig:predictimage}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_Predict_Image.pdf}
+  \caption{Structure of Image plane-weighting Predict, as used in CASA. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
+  \label{fig:makeimageuv}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_Image.pdf}
+  \caption{Structure of Image plane-weighting MakeImage, as used in CASA. Note that the three loops (denoted by the rectangle with curved ends) can be permuted at will.}
+  \label{fig:makeimageimage}
+\end{figure}
+
+\clearpage
+\section{Serial and Parallel implementations}
+
+\subsection{Serial}
+\label{sec:serial}
+
+The serial algorithm is as shown in Figure \ref{fig:algoCASA}.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./algorithm.pdf}
+  \caption{The MSMFS algorithm as implemented in CASA [RD??]}
+  \label{fig:algoCASA}
+\end{figure}
 
@@
 
-\begin{algorithm}[t!]
-\scriptsize
-\label{alg:MSMFS_1}
-%  \SetKwRepeat{Repeat}{Repeat}{Until}
-%  \SetKwFor{ForEach}{ForEach}{Do}{End}
-%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{shp}_s~\forall~s\in\{0,N_s$-$1\}, ~ [\Sna]~ \forall \nu$ }
-%\KwResult{ $\vec{I}^{psf}_{{sp}\atop{tq}},[{H^{peak}_s}], \vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ }
-  \KwData{calibrated visibilities : $\vec{V}^{corr}_{\nu}~~\forall \nu$}
-  \KwData{$uv$-sampling function : $\Sna$}
-  \KwData{image noise threshold and loop gain $\sigma_{thr}, g_s$}
-  \KwData{scale basis functions : $\I^{shp}_s ~ \forall s\in\{0,N_s-1\}$}
-  \KwResult{model coefficient images : $\I^{m}_{q}~ \forall q\in\{0,N_t-1\}$}
-  \KwResult{spectral index and curvature : $\I^{m}_{\alpha},\I^{m}_{\beta}$}
+%\begin{algorithm}[htp!]
+%\scriptsize
+%\label{alg:MSMFS_1}
+%  \KwData{calibrated visibilities : $\vec{V}^{corr}_{\nu}~~\forall \nu$}
+%  \KwData{Fourier plane sampling function : {\tens{S_\nu}{4} }}
+%  \KwData{image noise threshold and loop gain $\sigma_{thr}, g_s$}
+%  \KwData{scale basis functions : $\I^{shp}_s ~ \forall s\in\{0,N_s-1\}$}
+%  \KwResult{model coefficient images : $\I^{m}_{q}~ \forall q\in\{0,N_t-1\}$}
+%  \KwResult{spectral index and curvature : $\I^{m}_{\alpha},\I^{m}_{\beta}$}
+%
+%  \vspace{0.5cm} 
+%  \For{ $t \in \{0,N_t-1\}, q \in \{t,N_t-1\}$}
+%  {
+%        { Compute the spectral PSF } $\vec{I}^{psf}_{tq}$\;
+%        \For{ $s \in \{0,N_s-1\}, p \in \{s,N_s-1\}$}
+%	{
+%		{Compute the scale-spectral PSF} $\vec{I}^{psf}_{{sp}\atop{tq}} = \vec{I}^{shp}_s \star \vec{I}^{shp}_p \star \vec{I}^{psf}_{tq} $\;
+%	}
+%  }
+%  \For { $s \in \{0,N_s-1\}$}
+%  {
+%     Construct $[{H^{peak}_s}]$ from $mid(I^{psf}_{{s,s}\atop{t,q}})$ and compute $[{H^{peak}_s}^{-1}]$\;
+%  }
+%  Initialize the model $\vec{I}^{m}_t$ for all $t \in \{0,N_t-1\}$ and compute $f_{sidelobe}$ \;
+%  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{res}_0 < \sigma_{thr}$ }
+%  {
+%    \For{$t \in \{0,N_t$-$1\}$}
+%    {
+%      Compute the residual image $\vec{I}^{res}_t$ \tcc*[r]{Makeimage}
+%      \For{$s \in \{0,N_s$-$1\}$}
+%      {
+%	      Compute $\vec{I}^{res}_{{s},{t}} = \vec{I}^{shp}_s \star \vec{I}^{res}_t$
+%      }
+%    }
+%    Calculate $f_{limit}$ from $\vec{I}^{res}_{0,0}$\;
+%    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{res}_{0,0} < f_{limit} $ } 
+%    {
+%     \For{$s \in \{0,N_s$-$1\}$} 
+%     {
+%      \uIf{Peak of $\vec{I}^{res}_{s,0} > 10~\sigma_{thr} $}
+%      {
+%       \ForEach{pixel}
+%       {
+%          Construct $I^{rhs}_s$, an $N_t\times 1$ vector from $I^{res}_{s,t} ~~\forall ~ t \in \{0,N_t$-$1\}$\;
+%          Compute principal solution $I^{sol}_s = [{H^{peak}_s}^{-1}] I^{rhs}_s$\;
+%       }
+%       Choose $I^{sol} = max\{I^{sol}_{t=0},~\forall~s\in\{0,N_s$-$1\}\}$ \;
+%      }
+%      \Else
+%      {
+%       Find the location of the peak in $\vec{I}^{res}_{s,0},~\forall~s\in\{0,N_s$-$1\}$\;
+%       Construct $I^{rhs}_s$, from $I^{res}_{s,t}$ for the chosen $s$, at this location\;
+%       Compute $I^{sol} = [{H^{peak}_s}^{-1}] I^{rhs}_s$ at this location\;
+%      }
+%     }
+%       \For{$t \in \{0,N_t-1\}$}
+%       {
+%        Update the model image : $I^{m}_t = I^{m}_t + g_s ~ I^{shp}_{s_i} \star I^{sol}_t$ \;
+%        \For{$s \in \{0,N_s$-$1\}$}
+%	{
+%          Update the residual image : $I^{res}_{s,t} = I^{res}_{s,t} - g ~\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1}[I^{psf}_{{sp}\atop{tq}} \star I^{sol}_q]$\;
+%	}
+%       }
+%    }
+%   Compute model visibilities $V^{m}_{\nu}$ from  $I^{m}_t~\forall t\in \{0.N_t-1\}$\tcc*[r]{Predict} 
+%   Compute a new residual image $I^{res}$ from residual visibilities $V^{corr}_{\nu}-V^{m}_{\nu}$ \tcc*[r]{Makeimage}
+%}
+%Calculate $\vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ from $I^{m}_t~\forall t\in \{0.N_t-1\}$ and restore the results \;
+%If required, remove average primary beam : $\vec{I}^{new}_{\nu_0}=\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}, \vec{I}^{new}_{\alpha}=\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}, \vec{I}^{new}_{\beta}=\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}$\;
+%\vspace{0.5cm}
+%
+%  \caption[MS-MFS Algorithm]
+%          {MS-MFS Algorithm : }
+%\end{algorithm}
 
-  \vspace{0.5cm} 
-%{Compute the dirty image $\vec{I}^{dirty}$ and psf $\vec{I}^{psf}$}\;
-  \For{ $t \in \{0,N_t-1\}, q \in \{t,N_t-1\}$}
-  {
-        { Compute the spectral PSF } $\vec{I}^{psf}_{tq}$\;
-        \For{ $s \in \{0,N_s-1\}, p \in \{s,N_s-1\}$}
-	{
-		{Compute the scale-spectral PSF} $\vec{I}^{psf}_{{sp}\atop{tq}} = \vec{I}^{shp}_s \star \vec{I}^{shp}_p \star \vec{I}^{psf}_{tq} $\;
-	}
-  }
-  \For { $s \in \{0,N_s-1\}$}
-  {
-     Construct $[{H^{peak}_s}]$ from $mid(I^{psf}_{{s,s}\atop{t,q}})$ and compute $[{H^{peak}_s}^{-1}]$\;
-  }
-%%%%%%%%%%%%%%%%%%%%%%
-%%%\vspace{0.5cm}
-%%  \caption[MS-MFS CLEAN : Pre-Deconvolution Setup]
-%%          {MS-MFS CLEAN : Pre-Deconvolution Setup}
-%%\end{algorithm}
-%%%\newpage
-%%
-%%\begin{algorithm}[H]
-%%  \SetLine
-%%  \linesnumbered
-%%  \dontprintsemicolon
-%%%%%%%%%%%%%%%%%%%%%%%  
-%%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{shp}_s, \vec{I}^{psf}_{{sp}\atop{tq}},[{H^{peak}_s}]~~~\forall~~s \in \{0,N_s-1\}, p \in \{s,N_s-1\}, \sigma_{thr}$ }
-%%\KwResult{$I^{m}_{q}~ \forall q \in \{0,N_t-1\}$}
-%%%%%%%%%%%%%%%%%%%%%  
-  Initialize the model $\vec{I}^{m}_t$ for all $t \in \{0,N_t-1\}$ and compute $f_{sidelobe}$ \;
-  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{res}_0 < \sigma_{thr}$ }
-  {
-    \For{$t \in \{0,N_t$-$1\}$}
-    {
-      Compute the residual image $\vec{I}^{res}_t$ \tcc*[r]{Makeimage}
-      \For{$s \in \{0,N_s$-$1\}$}
-      {
-	      Compute $\vec{I}^{res}_{{s},{t}} = \vec{I}^{shp}_s \star \vec{I}^{res}_t$
-      }
-    }
-    Calculate $f_{limit}$ from $\vec{I}^{res}_{0,0}$\;
-    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{res}_{0,0} < f_{limit} $ } 
-    {
-%     Compute $I^{m}_q~\forall q\in \{0.N_t-1\}$ and update $\vec{I}^{res}_{{s},{t}}~\forall s,t$ (Algorithm \vref{ALGO:MSMFS_2})\;
-     \For{$s \in \{0,N_s$-$1\}$} 
-     {
-      \uIf{Peak of $\vec{I}^{res}_{s,0} > 10~\sigma_{thr} $}
-      {
-       \ForEach{pixel}
-       {
-          Construct $I^{rhs}_s$, an $N_t\times 1$ vector from $I^{res}_{s,t} ~~\forall ~ t \in \{0,N_t$-$1\}$\;
-          Compute principal solution $I^{sol}_s = [{H^{peak}_s}^{-1}] I^{rhs}_s$\;
-       }
-       Choose $I^{sol} = max\{I^{sol}_{t=0},~\forall~s\in\{0,N_s$-$1\}\}$ \;
-      }
-      \Else
-      {
-       Find the location of the peak in $\vec{I}^{res}_{s,0},~\forall~s\in\{0,N_s$-$1\}$\;
-       Construct $I^{rhs}_s$, from $I^{res}_{s,t}$ for the chosen $s$, at this location\;
-       Compute $I^{sol} = [{H^{peak}_s}^{-1}] I^{rhs}_s$ at this location\;
-      }
-     }
-       \For{$t \in \{0,N_t-1\}$}
-       {
-        Update the model image : $I^{m}_t = I^{m}_t + g_s ~ I^{shp}_{s_i} \star I^{sol}_t$ \;
-        \For{$s \in \{0,N_s$-$1\}$}
-	{
-          Update the residual image : $I^{res}_{s,t} = I^{res}_{s,t} - g ~\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1}[I^{psf}_{{sp}\atop{tq}} \star I^{sol}_q]$\;
-	}
-       }
-    }
-   Compute model visibilities $V^{m}_{\nu}$ from  $I^{m}_t~\forall t\in \{0.N_t-1\}$\tcc*[r]{Predict} 
-   Compute a new residual image $I^{res}$ from residual visibilities $V^{corr}_{\nu}-V^{m}_{\nu}$ \tcc*[r]{Makeimage}
-}
-Calculate $\vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ from $I^{m}_t~\forall t\in \{0.N_t-1\}$ and restore the results \;
-If required, remove average primary beam : $\vec{I}^{new}_{\nu_0}=\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}, \vec{I}^{new}_{\alpha}=\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}, \vec{I}^{new}_{\beta}=\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}$\;
-\vspace{0.5cm}
+\clearpage
+\subsection{Parallel processing}
+\label{sec:parallel}
 
-  \caption[MS-MFS Algorithm]
-          {MS-MFS Algorithm : }
-\end{algorithm}
-
-Notes:
+When moving the MSMFS algorithm to a distributed (slow interconnect) and parallel (fast interconnect) architecture such as the SDP architecture, there are a large number of important factors that must be considered.
 \begin{description}
-\item[Line 13]	
+\item[Order of iterations]
+\item[Coupling across partitions] e.g. is the same pixel addressed in two different partitions? If so, is there a satisfactory approach to reconciling the two values?
+\item[Synchronisation points] e.g. is a global synchronisation point required so that the deconvolutions can be made consistent? Is there an acceptable algorithm for reconciling separate facets?
+\item[Amount of CU memory]
+\item[Loading of CU memory backplane]
+\item[Access to visibility store]
 \end{description}
 
-\clearpage
-\section{Partitioning}
-\label{sec:partitioning}
+\subsubsection{Minor cycle}
+\label{subsec:minorcycle}
 
-The SDP pipeline framework [RD:??], allows partitioning across multiple axes. In this case, the natural partitions are:
+\subsubsection{Partitioning}
+\label{subsec:partitoning}
+
+The SDP pipeline framework [RD:??], allows partitioning across multiple axes. In this case, the natural partitions and typical values are:
 
 \begin{description}
-\item[Frequency]
-\item[Time]
-\item[Facet]
-\item[Scales]	
+\item[Polarisation] Only Stokes I can be modelled using MSMFS
+\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing
+\item[Sub-bands] O(1-3) partitions of the coarse channels are required to ensure that the power law approximation is adequate.
+\item[Time] O(43200) ~ O(1s) correlator samples are required to avoid smearing on the longest baselines
+\item[Facet] O(100) facets to limit the size of the AW projection kernel
+%\item[Snapshot time] O(100s) long gridding/degridding sequences are required to limit smearing due to w-plane rotation
+\item[Scales] O(5) scales are typically required to model extended structure
+\item[Taylor terms] O(5) terms are required to represent the spectral behaviour
 \end{description}
+
+According to the logic of the MSMFS algorithm, Sub-bands can be defined as the limit of the power law approximation for brightness so consequently can always be used as the coarsest partition.
+For Predict and MakeImage the remaining choices are between (most rapid first):
+
+\begin{table}[htp]
+ \caption{Possible ordering of partitions $UV$ weighting version of Predict and MakeImage}
+ \begin{center}
+ \begin{tabular}{|c|c|c|c|c|}
+\hline
+Frequency & Time & Facet & Sub-Band \\
+Frequency & Facet & Time & Sub-Band \\
+Time & Facet & Frequency & Sub-Band \\
+Time & Frequency & Facet & Sub-Band \\
+Facet & Time & Frequency & Sub-Band \\
+Facet & Frequency & Time & Sub-Band \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Frequency and Taylor terms are tight coupled and so to avoid excess inter-CU communication, the order should be [Taylor, Frequency].
+The remaining choices are then between the two orderings of Time and Facet.
+Since deconvolution is non-linear, we prefer that all times be represented in a single facet rather than the other way around.
+
+Hence according to this level of analysis, the best ordering is [Frequency, Time, Facet, Sub-Band].
+Parallelisation over Compute Nodes starts from the left to the right, Distribution over Compute Islands works from the right to the left (see Figure \ref{fig:order}):
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_Order.pdf}
+  \caption{Preferred order of parallelisation and distribution for MSMFS algorithm}
+  \label{fig:order}
+\end{figure}
+
+In this approach, the facets will almost certainly be processed on different compute islands. Note we have a dilemma on how to deal with the deconvolution of the facets.
+
+\begin{itemize}	
+\item Construct the facets with some padding, deconvolve each separately, and reconcile after deconvolution by facet to facet broadcast.
+\item Construct the facets without padding, send all facets to one compute island, perform deconvolution, distribute models to facets. 
+\end{itemize}
+
+The first approach will inevitably introduce edge effects in the final image revealing the facet grid, while the second will block processing while deconvolution proceeds on the different sub-band compute nodes.
+We would then require two different types of nodes:
+\begin{description}
+\item[Predict and MakeImage] Requires the visibility data but not many scratch images. Stalled during MinorCycle.
+\item[MinorCycle] Requires a substantial number of scratch images, but not the visibility data. Stalled during Predict and MakeImage.	
+\end{description}
+
+\subsection{Minor Cycles}
+
+\subsubsection{CASA}
+
+\subsubsection{ASKAPsoft}
+The core algorithm is in C++ Template DeconvolverMultiTermBasisFunction.tcc.
+The basis function is abstracted and can be any class having the interface DeconvolverMultiTermBasisFunction.
+There are two forms present: one for point sources, and one for the same blobs used in CASA - a truncated upside down parabola.
+
+The expansion in Taylor terms is generalised to any form obeying an equation like: $D = B(0)*I(0) + B(1)*I(1) + B(2)*I(2)$, where D is the dirty image, B(?) are the spectral dirty beams, and I(?) are the images for each term.
+The coupling matrices between the terms are calculated for each basis function.
+Once a suitable peak component is found, the inverse of the coupling matrix is used to decouple the different terms for the optimum scale.
+The optimum blob location and scale is found by using one of a number of criteria:
+\begin{description}
+\item[MAXBASE0] The peak of term 0 across after decoupling in term.
+\item[MAXCHISQ] The peak of chisquared across scale after decoupling in term.
+\end{description}
+
+The second criterion is expensive to compute and usually MAXBASE0 is used.
+
+Once the optimum blob scale and location is found, the vector in term-space is found by decoupling.
+The model images (one for each term) are then updated, and the effects of this blob removed from all the cached images.
 

### 2016-04-22 msmfs.tex
Mostly complete to and including section 7

@@
 \usepackage{hhtensor}
+\usepackage{color}
+\usepackage{txfonts}
+
+%\usepackage{stylefiles/fmtcount}
+\usepackage{natbib}
+
 \graphicspath{ {images/} }
@@
     \item[CDR] Critical Design Review \vspace{-0.2cm}
-    \item[CPU] Central Processing Unit\vspace{-0.2cm}
-    \item[CSP] Central Signal Processor\vspace{-0.2cm}
-    \item[CUFFT] \textcolor{red}{Define it.}\vspace{-0.2cm}
-    \item[DFT] \textcolor{red}{Only used once, in a title. Remove/expand?}\vspace{-0.2cm}
-    \item[DM] \textcolor{red}{Define it.}\vspace{-0.2cm}
-    \item[EoR] Epoch of Reionisation\vspace{-0.2cm}
     \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
@@
     \item[FoV] Field of View\vspace{-0.2cm}
-    \item[GPU] Graphics Processing Unit \vspace{-0.2cm}
-    \item[ICD] Interface Control Document\vspace{-0.2cm}
-    \item[LMA] Levenberg-Marquardt Algorithm \vspace{-0.2cm}
-    \item[LOFAR] LOw Frequency ARray\vspace{-0.2cm}
-    \item[LSM] Local Sky Model\vspace{-0.2cm}
     \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
-    \item[NVRAM] \textcolor{red}{Define it.} \vspace{-0.2cm}
-    \item[OCLD] Optimised Candidate List and Data \vspace{-0.2cm}
-    \item[PDR] Preliminary Design Review\vspace{-0.2cm}
-    \item[RFI] Radio Frequency Interference\vspace{-0.2cm}
-    \item[PSF] \textcolor{red}{Define it.} \vspace{-0.2cm}
-    \item[PSS] Pulsar Search\vspace{-0.2cm}
-    \item[PST] Pulsar Timing\vspace{-0.2cm}
+    \item[PSF] Point Spread Function \vspace{-0.2cm}
     \item[SDP] Scientific Data Processor\vspace{-0.2cm}
@@
     \item[SKAO] SKA Organisation\vspace{-0.2cm}
-    \item[WCS] \textcolor{red}{Define it.} \vspace{-0.2cm}
 %    \item[] \vspace{-0.2cm}
@@
 \newcommand{\Bd}{\vec{b^\dag}}
-\newcommand{\A}{{\tens{A}}}
-\newcommand{\Ad}{{\tens{A^\dag}}}
-\newcommand{\F}{{\tens{F}}}
-\newcommand{\Fd}{{\tens{F^\dag}}}
-\newcommand{\He}{{\tens{H}}}
-\newcommand{\Sa}{{\tens{S}{N}}}
-\newcommand{\Sd}{{\tens{S^\dag}}}
-\newcommand{\Sna}{\tens{{S_{\nu}}{N}}}
-\newcommand{\Snd}{\tens{{S_{\nu}^\dag}}}
-\newcommand{\T}{{\tens{T}}}
-\newcommand{\W}{{\tens{W}}}
-\newcommand{\Wd}{{\tens{W^\dag}}}
-\newcommand{\Pb}{{P_b}}
+\newcommand{\V}{\vec{V}}
+\newcommand{\Vd}{\vec{V^\dag}}
+\newcommand{\A}{{\tens{A}{s}}}
+\newcommand{\Ad}{{\tens{A^\dag}{s}}}
+\newcommand{\F}{{\tens{F}{s}}}
+\newcommand{\Fd}{{\tens{F^\dag}{s}}}
+\newcommand{\He}{{\tens{H}{s}}}
+\newcommand{\Sa}{{\tens{S}{s}}}
+\newcommand{\Sd}{{\tens{S^\dag}{s}}}
+\newcommand{\Sna}{\tens{{S_{\nu}}{s}}}
+\newcommand{\Snd}{\tens{{S_{\nu}^\dag}{s}}}
+\newcommand{\T}{{\tens{T}{s}}}
+\newcommand{\W}{{\tens{W}{s}}}
+\newcommand{\Wd}{{\tens{W^\dag}{s}}}
+\newcommand{\Pb}{{\vec{P}}}
 
-\newcommand{\Wim}{{\tens{W^{im}}}}
-\newcommand{\Wimd}{{\tens{{W^{im}}^\dag}}}
-\newcommand{\Wnt}{{\tens{W^{mfs}_t}}}
-\newcommand{\Wntd}{{\tens{{W^{mfs}_t}^\dag}}}
-\newcommand{\Wnp}{{\tens{W^{mfs}_p}}}
-\newcommand{\Wnpd}{{\tens{{W^{mfs}_p}^\dag}}}
-\newcommand{\Wnq}{{\tens{W^{mfs}_q}}}
-\newcommand{\Wnqd}{{\tens{{W^{mfs}_q}^\dag}}}
-\newcommand{\Wimn}{{\tens{W^{im}_{\nu}}}}
-\newcommand{\Wimnd}{{\tens{{W^{im}_{\nu}}^\dag}}}
+%\newcommand{\Wim}{{\tens{W^{im}}}}
+%\newcommand{\Wimd}{{\tens{{W^{im}}^\dag}}}
+%\newcommand{\Wnt}{{\tens{W^{\rm {mfs}}_t}}}
+%\newcommand{\Wntd}{{\tens{{W^{\rm {mfs}}_t}^\dag}}}
+%\newcommand{\Wnp}{{\tens{W^{\rm mfs}_p}}}
+%\newcommand{\Wnpd}{{\tens{{W^{\rm mfs}_p}^\dag}}}
+%\newcommand{\Wnq}{{\tens{W^{\rm mfs}_q}}}
+%\newcommand{\Wnqd}{{\tens{{W^{\rm {mfs}}_q}^\dag}}}
+%\newcommand{\Wimn}{{\tens{W^{im}_{\nu}}}}
+%\newcommand{\Wimnd}{{\tens{{W^{im}_{\nu}}^\dag}}}
+
+\newcommand{\Wim}{{{\W^{\rm im}}}}
+\newcommand{\Wimd}{{{{\W^{\rm im}}^\dag}}}
+\newcommand{\Wnt}{{{\W^{\rm {mfs}}_t}}}
+\newcommand{\Wntd}{{{{\W^{\rm {mfs}}_t}^\dag}}}
+\newcommand{\Wnp}{{{\W^{\rm mfs}_p}}}
+\newcommand{\Wnpd}{{{{\W^{\rm mfs}_p}^\dag}}}
+\newcommand{\Wnq}{{{\W^{\rm mfs}_q}}}
+\newcommand{\Wnqd}{{{{\W^{\rm {mfs}}_q}^\dag}}}
+\newcommand{\Wimn}{{{\W^{\rm im}_{\nu}}}}
+\newcommand{\Wimnd}{{{{\W^{\rm im}_{\nu}}^\dag}}}
 
@@
 \newcommand{\wntq}{{w_{\nu}^{t+q}}}
-\newcommand{\Wntn}{{\tens{w^{mfs}_{t,\nu}}}}
-\newcommand{\Wntnd}{{\tens{{w^{mfs}_{t,\nu}}^\dag}}}
-\newcommand{\Wnpn}{{\tens{W^{mfs}_{p,\nu}}}}
-\newcommand{\Wnpnd}{{\tens{{W^{mfs}_{p,\nu}}^\dag}}}
+%\newcommand{\Wntn}{{\tens{w^{\rm mfs}_{t,\nu}}}}
+%\newcommand{\Wntnd}{{\tens{{w^{\rm mfs}_{t,\nu}}^\dag}}}
+%\newcommand{\Wnpn}{{\tens{W^{\rm mfs}_{p,\nu}}}}
+%\newcommand{\Wnpnd}{{\tens{{W^{\rm mfs}_{p,\nu}}^\dag}}}
 
@@
 \newcommand{\I}{{\vec{I}}}
+
+\newcommand{\Nt}{N_{\rm t}}
+\newcommand{\Ns}{N_{\rm s}}
+\newcommand{\Nc}{N_{\rm c}}
 
@@
 
+\clearpage
 \section{Scope of the document}
 
+\clearpage
 \section{Background}
@@
 
-A canonical major/minor cycle algorithm is shown in Figure \ref{fig:major/minor}.
+A canonical major/minor cycle algorithm is shown in Figure \ref{fig:majorminor}.
 The two functions Predict and MakeImage are sub-pipelines.
-ASKAPSoft and CASA both follow this form but with different ways of performing Predict (\ref{fig:predictuv} and \ref{fig:predictimage}) and MakeImage (\ref{fig:makeimageuv} and \ref{fig:makeimageimage})
+ASKAPSoft and CASA both follow this form but with different ways of performing Predict (Figures \ref{fig:predictuv} and \ref{fig:predictimage}) and MakeImage (Figures \ref{fig:makeimageuv} and \ref{fig:makeimageimage})
 
@@
 
-The MSMFS algorithm decomposes the image into a set of blobs, using a greedy algorithm in which the peak (in scale space) is found by searching for the peak in the reference frequency residual images each convolved with scale.
-For bright pixels, the principal solution is calculated by applying the inverse Hessian of the PSF convolved with scale.
-For faint pixels, just the peak in the reference frequency image is used.
+In the Multi-Scale CLEAN [RD??], the sky is modelled as a set of blobs.
+Performing this decomposition using a gradient search algorithm is very demanding of resources.
+Instead, in MSCLEAN, a greedy algorithm obtained by generalising the CLEAN algorithm is used.
+The greedy algorithm locates the amplitude, scale and location of the blob by a straightforward search in the set of residual images, each smoothed with the appropriate blob.
+For each minor cycle this produces an amplitude, location, and scale.
+The smoothed residual images and model images are updated, and iteration proceeds.
 
@@
 
+The MSMFS algorithm decomposes the image into a set of $N_s$ blobs each with $N_t$ Taylor terms.
+
 \begin{eqnarray}
@@
 the $I^m_t$ represent multi-scale Taylor coefficient images.
+
+Before performing any search, it is necessary to decouple the pixels in Taylor term space as well as possible.
+In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used.
+The greedy algorithm then locates the scale and location in the reference-frequency image (as convolved with the set of scales).
+For each minor cycle this produces a location, scale, and values for all Taylor terms.
+The residual images and model images are updated, and iteration proceeds.
+The actual algorithm is shown in Figure \ref{fig:CASA}.
 
@@
 coefficient.
+
+\clearpage
+\section{CASA and ASKAPsoft implementations}
+
+The MS-MFS algorithm as implemented in CASA shown in Algorithm \ref{CASA}. The ASKAPsoft version is shown in Algorithm \ref{ASKAP} with the variant lines in red.
+
+%\begin{figure}[htb]
+%  \centering
+%  \includegraphics[width=\textwidth]{./algorithm.pdf}
+%  \caption{The MSMFS algorithm as implemented in CASA [RD??]}
+%  \label{fig:algoCASA}
+%\end{figure}
+
+\clearpage
+
+\begin{algorithm}
+\scriptsize
+  \SetLine
+  \linesnumbered
+  \dontprintsemicolon
+%  \SetKwRepeat{Repeat}{Repeat}{Until}
+%  \SetKwFor{ForEach}{ForEach}{Do}{End}
+%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{\rm shp}_s~\forall~s\in\{0,\Ns $-$1\}, ~ [\Sna]~ \forall \nu$ }
+%\KwResult{ $\vec{I}^{\rm psf}_{{sp}\atop{tq}},[{H^{\rm peak}_s}], \vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ }
+  \KwData{calibrated visibilities : $\vec{V}^{\rm obs}_{\nu}~~\forall \nu$}
+  \KwData{$uv$-sampling function and weights : $[{\Sa}_{\nu}], [\Wimn]$}
+  \KwData{input : number of Taylor-terms $\Nt $, number of scales $\Ns $}
+  \KwData{input : image noise threshold, $\sigma_{\rm thr}$, loop gain $g$}
+  \KwData{input : scale basis functions : $\I^{\rm shp}_s ~ \forall s\in\{0,\Ns -1\}$}
+  \KwData{input : reference frequency $\nu_0$ to compute $w_{\nu}=\dnuno$}
+  \KwResult{model coefficient images : $\I^{\rm m}_{t}~ \forall t\in\{0,\Nt -1\}$}
+  \KwResult{intensity, spectral index and curvature : $\I^{\rm m}_{\nuup_0},\I^{\rm m}_{\alphaup},\I^{\rm m}_{\betaup}$}
+
+%{Compute the dirty image $\vec{I}^{\rm dirty}$ and psf $\vec{I}^{\rm psf}$}\;
+  \For{ $t \in \{0,\Nt -1\}, q \in \{t,\Nt -1\}$}
+  {
+        { Compute the spectral Hessian kernel } $\vec{I}^{\rm psf}_{tq} = \sum_{\nu} \wntq \vec{I}^{\rm psf}_{\nu}$\;
+        \For{ $s \in \{0,\Ns -1\}, p \in \{s,\Ns -1\}$}
+	{
+		{Compute scale-spectral kernels} $\vec{I}^{\rm psf}_{{sp}\atop{tq}} = \vec{I}^{\rm shp}_s \star \vec{I}^{\rm shp}_p \star \vec{I}^{\rm psf}_{tq} $\;
+	}
+  }
+  \For { $s \in \{0,\Ns -1\}$}
+  {
+     Construct scale-dependent Taylor space Hessian $[{\He^{\rm peak}_s}]$ from $mid(\I^{\rm psf}_{{s,s}\atop{t,q}})$ and compute $[{\He^{\rm peak}_s}^{-1}]$\;
+  }
+%%%%%%%%%%%%%%%%%%%%%%
+%%%\vspace{0.5cm}
+%%  \caption[MS-MFS CLEAN : Pre-Deconvolution Setup]
+%%          {MS-MFS CLEAN : Pre-Deconvolution Setup}
+%%\end{algorithm}
+%%%\newpage
+%%
+%%\begin{algorithm}[\He]
+%%  \SetLine
+%%  \linesnumbered
+%%  \dontprintsemicolon
+%%%%%%%%%%%%%%%%%%%%%%%  
+%%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{\rm shp}_s, \vec{I}^{\rm psf}_{{sp}\atop{tq}},[{H^{\rm peak}_s}]~~~\forall~~s \in \{0,\Ns -1\}, p \in \{s,\Ns -1\}, \sigma_{thr}$ }
+%%\KwResult{$I^{\rm m}_{q}~ \forall q \in \{0,\Nt -1\}$}
+%%%%%%%%%%%%%%%%%%%%%  
+  Initialize the model $\vec{I}^{\rm m}_t$ for all $t \in \{0,\Nt -1\}$\; % and compute $f_{sidelobe}$ \;
+%  \vspace{0.5cm} 
+  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{\rm res}_0 < \sigma_{\rm thr}$ }
+  {
+    \For{$t \in \{0,\Nt $-$1\}$}
+    {
+      Compute $\vec{I}^{\rm res}_t = \sum_{\nu} \wnt \vec{I}^{\rm res}_{\nu}$ from frequency-weighted residual images $\I^{\rm res}_{\nu}$\;
+      \For{$s \in \{0,\Ns $-$1\}$}
+      {
+	    Convolve with $s^{th}$ scale-function $\vec{I}^{\rm res}_{{s}\atop{t}} = \vec{I}^{\rm shp}_s \star \vec{I}^{\rm res}_t$
+      }
+    }
+    Calculate minor-cycle threshold $f_{\rm limit}$ from $\vec{I}^{\rm res}_{{0}\atop{0}}$\;
+%  \vspace{0.5cm} 
+    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{\rm res}_{{0}\atop{0}} < f_{\rm limit} $ } 
+    {
+%     Compute $I^{\rm m}_q~\forall q\in \{0.\Nt -1\}$ and update $\vec{I}^{\rm res}_{{s},{t}}~\forall s,t$ (Algorithm \vref{MSMFS_2})\;
+     \For{$s \in \{0,\Ns $-$1\}$}
+     {
+%      \uIf{Peak of $\vec{I}^{\rm res}_{s,0} > 10~\sigma_{thr} $}
+%      {
+       \ForEach{pixel}
+       {
+          Construct $\I^{\rm rhs}_s$, an $\Nt \times 1$ vector from $\I^{\rm res}_{{s}\atop{t}} ~~\forall ~ t \in \{0,\Nt $-$1\}$\;
+          Compute principal solution $\I^{\rm sol}_s = [{\He^{\rm peak}_s}^{-1}] \I^{\rm rhs}_s$\;
+          Fill solution $\I^{\rm sol}_s$ into model images $\forall t$ : $\I^{\rm m,sol}_{{s}\atop{t}}$
+       }
+ %     }
+  %    \Else
+%      {
+ %      Find the location of the peak in $\vec{I}^{\rm res}_{s,0},~\forall~s\in\{0,\Ns $-$1\}$\;
+ %      Construct $I^{\rm rhs}_s$, from $I^{\rm res}_{s,t}$ for the chosen $s$, at this location\;
+ %      Compute $I^{sol} = [{H^{\rm peak}_s}^{-1}] I^{\rm rhs}_s$ at this location\;
+ %     }
+    }
+       Choose $\I^{\rm m}_{{p}\atop{t}} = max\{\I^{\rm m,sol}_{{s}\atop{t=0}},~\forall~s\in\{0,\Ns $-$1\}\}$ \;
+       \For{$t \in \{0,\Nt -1\}$}
+       {
+        Update the model image : $\I^{\rm m}_t = I^{\rm m}_t + g ~[ \I^{\rm shp}_{p} \star \I^{\rm m}_{{p}\atop{t}}]$ \;
+        \For{$s \in \{0,\Ns $-$1\}$}
+	{
+          Update the residual image : $\I^{\rm res}_{{s}\atop{t}} = \I^{\rm res}_{{s}\atop{t}} - g ~\sum_{q=0}^{\Nt -1}[\I^{\rm psf}_{{sp}\atop{tq}} \star \I^{\rm m}_{{p}\atop{t}}]$\;
+	}
+       }
+    }
+%  \vspace{0.5cm} 
+   Compute model visibilities $\V^{\rm m}_{\nu}$ from  $\I^{\rm m}_t~\forall t\in \{0.\Nt -1\}$\;
+   Compute residual visibilities $\V^{\rm res}_{\nu} = \V^{\rm obs}_{\nu}-\V^{\rm m}_{\nu}$\;
+  }
+  \vspace{0.5cm} 
+Restore the model Taylor-coefficients $\I^{\rm m}_t~\forall t\in \{0.\Nt -1\}$ \;
+Calculate $\vec{I}^{\rm m}_{\nuup_0}, \vec{I}^{\rm m}_{\alphaup}, \vec{I}^{\rm m}_{\betaup}$ from $\I^{\rm m}_t~\forall t\in \{0.\Nt -1\}$\;
+If required, remove average primary beam : $\vec{I}^{\rm new}_{\nuup_0}=\vec{I}^{\rm m}_{\nuup_0}/\vec{P}_{\nuup_0};  \vec{I}^{\rm new}_{\alphaup}=\vec{I}^{\rm m}_{\alphaup}-\vec{P}_{\alphaup};  \vec{I}^{\rm new}_{\betaup}=\vec{I}^{\rm m}_{\betaup}-\vec{P}_{\betaup}$\;
+%\vspace{0.5cm}
+
+  \caption[MS-MFS Algorithm]
+         {MS-MFS, as implemented in CASA}\label{CASA}
+
+\end{algorithm}
+
+\begin{algorithm}
+\scriptsize
+  \SetLine
+  \linesnumbered
+  \dontprintsemicolon
+%  \SetKwRepeat{Repeat}{Repeat}{Until}
+%  \SetKwFor{ForEach}{ForEach}{Do}{End}
+%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{\rm shp}_s~\forall~s\in\{0,\Ns $-$1\}, ~ [\Sna]~ \forall \nu$ }
+%\KwResult{ $\vec{I}^{\rm psf}_{{sp}\atop{tq}},[{H^{\rm peak}_s}], \vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ }
+  \KwData{calibrated visibilities : $\vec{V}^{\rm obs}_{\nu}~~\forall \nu$}
+  \KwData{$uv$-sampling function and weights : $[{\Sa}_{\nu}], [\Wimn]$}
+  \KwData{input : number of Taylor-terms $\Nt $, number of scales $\Ns $}
+  \KwData{input : image noise threshold, $\sigma_{\rm thr}$, loop gain $g$}
+  \KwData{input : scale basis functions : $\I^{\rm shp}_s ~ \forall s\in\{0,\Ns -1\}$}
+  \KwData{input : reference frequency $\nu_0$ to compute $w_{\nu}=\dnuno$}
+  \KwResult{model coefficient images : $\I^{\rm m}_{t}~ \forall t\in\{0,\Nt -1\}$}
+  \KwResult{intensity, spectral index and curvature : $\I^{\rm m}_{\nuup_0},\I^{\rm m}_{\alphaup},\I^{\rm m}_{\betaup}$}
+
+%{Compute the dirty image $\vec{I}^{\rm dirty}$ and psf $\vec{I}^{\rm psf}$}\;
+  \For{ $t \in \{0,\Nt -1\}, q \in \{t,\Nt -1\}$}
+  {
+{\color{red}        { Compute the spectral Hessian kernel $\vec{I}^{\rm psf}_{tq}$ from transform of frequency weight sampling } $\sum_{\nu} \wntq \Sa_{\nu}$\;}
+        \For{ $s \in \{0,\Ns -1\}, p \in \{s,\Ns -1\}$}
+	{
+		{Compute scale-spectral kernels} $\vec{I}^{\rm psf}_{{sp}\atop{tq}} = \vec{I}^{\rm shp}_s \star \vec{I}^{\rm shp}_p \star \vec{I}^{\rm psf}_{tq} $\;
+	}
+  }
+  \For { $s \in \{0,\Ns -1\}$}
+  {
+     Construct scale-dependent Taylor space Hessian $[{\He^{\rm peak}_s}]$ from $mid(\I^{\rm psf}_{{s,s}\atop{t,q}})$ and compute $[{\He^{\rm peak}_s}^{-1}]$\;
+  }
+%%%%%%%%%%%%%%%%%%%%%%
+%%%\vspace{0.5cm}
+%%  \caption[MS-MFS CLEAN : Pre-Deconvolution Setup]
+%%          {MS-MFS CLEAN : Pre-Deconvolution Setup}
+%%\end{algorithm}
+%%%\newpage
+%%
+%%\begin{algorithm}[\He]
+%%  \SetLine
+%%  \linesnumbered
+%%  \dontprintsemicolon
+%%%%%%%%%%%%%%%%%%%%%%%  
+%%\KwData{ $\vec{V}^{corr}_{\nu}, \vec{I}^{\rm shp}_s, \vec{I}^{\rm psf}_{{sp}\atop{tq}},[{H^{\rm peak}_s}]~~~\forall~~s \in \{0,\Ns -1\}, p \in \{s,\Ns -1\}, \sigma_{thr}$ }
+%%\KwResult{$I^{\rm m}_{q}~ \forall q \in \{0,\Nt -1\}$}
+%%%%%%%%%%%%%%%%%%%%%  
+  Initialize the model $\vec{I}^{\rm m}_t$ for all $t \in \{0,\Nt -1\}$\; % and compute $f_{sidelobe}$ \; 
+  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{\rm res}_0 < \sigma_{\rm thr}$ }
+  {
+    \For{$t \in \{0,\Nt $-$1\}$}
+    {
+{\color{red}      Compute $\vec{I}^{\rm res}_t = \sum_{\nu} \wnt \vec{I}^{\rm res}_{\nu}$ from frequency-weighted residual visibilities $\V^{\rm res}_{\nu}$\;}
+      \For{$s \in \{0,\Ns $-$1\}$}
+      {
+	    Convolve with $s^{th}$ scale-function $\vec{I}^{\rm res}_{{s}\atop{t}} = \vec{I}^{\rm shp}_s \star \vec{I}^{\rm res}_t$
+      }
+    }
+    Calculate minor-cycle threshold $f_{\rm limit}$ from $\vec{I}^{\rm res}_{{0}\atop{0}}$\;
+%  \vspace{0.5cm} 
+    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{\rm res}_{{0}\atop{0}} < f_{\rm limit} $ } 
+    {
+%     Compute $I^{\rm m}_q~\forall q\in \{0.\Nt -1\}$ and update $\vec{I}^{\rm res}_{{s},{t}}~\forall s,t$ (Algorithm \vref{MSMFS_2})\;
+     \For{$s \in \{0,\Ns $-$1\}$}
+     {
+%      \uIf{Peak of $\vec{I}^{\rm res}_{s,0} > 10~\sigma_{thr} $}
+%      {
+       \ForEach{pixel}
+       {
+          Construct $\I^{\rm rhs}_s$, an $\Nt \times 1$ vector from $\I^{\rm res}_{{s}\atop{t}} ~~\forall ~ t \in \{0,\Nt $-$1\}$\;
+          Compute principal solution $\I^{\rm sol}_s = [{\He^{\rm peak}_s}^{-1}] \I^{\rm rhs}_s$\;
+          Fill solution $\I^{\rm sol}_s$ into model images $\forall t$ : $\I^{\rm m,sol}_{{s}\atop{t}}$
+       }
+ %     }
+  %    \Else
+%      {
+ %      Find the location of the peak in $\vec{I}^{\rm res}_{s,0},~\forall~s\in\{0,\Ns $-$1\}$\;
+ %      Construct $I^{\rm rhs}_s$, from $I^{\rm res}_{s,t}$ for the chosen $s$, at this location\;
+ %      Compute $I^{sol} = [{H^{\rm peak}_s}^{-1}] I^{\rm rhs}_s$ at this location\;
+ %     }
+     }
+     Choose $\I^{\rm m}_{{p}\atop{t}} = max\{\I^{\rm m,sol}_{{s}\atop{t=0}},~\forall~s\in\{0,\Ns $-$1\}\}$ \;
+       \For{$t \in \{0,\Nt -1\}$}
+       {
+        Update the model image : $\I^{\rm m}_t = I^{\rm m}_t + g ~[ \I^{\rm shp}_{p} \star \I^{\rm m}_{{p}\atop{t}}]$ \;
+        \For{$s \in \{0,\Ns $-$1\}$}
+	{
+          Update the residual image : $\I^{\rm res}_{{s}\atop{t}} = \I^{\rm res}_{{s}\atop{t}} - g ~\sum_{q=0}^{\Nt -1}[\I^{\rm psf}_{{sp}\atop{tq}} \star \I^{\rm m}_{{p}\atop{t}}]$\;
+	}
+       }
+    }
+%  \vspace{0.5cm} 
+   {\color{red} Compute model visibilities $\V^{\rm m}_{\nu}$ from  $\V^{\rm m}_t~\forall t\in \{0.\Nt -1\}$}\;
+   Compute residual visibilities $\V^{\rm res}_{\nu} = \V^{\rm obs}_{\nu}-\V^{\rm m}_{\nu}$\;
+  } 
+Restore the model Taylor-coefficients $\I^{\rm m}_t~\forall t\in \{0.\Nt -1\}$ \;
+Calculate $\vec{I}^{\rm m}_{\nuup_0}, \vec{I}^{\rm m}_{\alphaup}, \vec{I}^{\rm m}_{\betaup}$ from $\I^{\rm m}_t~\forall t\in \{0.\Nt -1\}$\;
+If required, remove average primary beam : $\vec{I}^{\rm new}_{\nuup_0}=\vec{I}^{\rm m}_{\nuup_0}/\vec{P}_{\nuup_0};  \vec{I}^{\rm new}_{\alphaup}=\vec{I}^{\rm m}_{\alphaup}-\vec{P}_{\alphaup};  \vec{I}^{\rm new}_{\betaup}=\vec{I}^{\rm m}_{\betaup}-\vec{P}_{\betaup}$\;
+%\vspace{0.5cm}
+
+  \caption[MS-MFS Algorithm]
+         {MS-MFS, as implemented in ASKAPsoft: {\color{red} red denotes difference from CASA}}\label{ASKAP}
+
+\end{algorithm}
+
+We can relate these back to the graphical form Figure \ref{fig:majorminor} by introducing variant sub pipelines for Predict and MakeImage.
+These are graphically shown in Figures \ref{fig:predictuv} to \ref{fig:makeimageimage}:
 
@@
 
-\clearpage
-\section{Serial and Parallel implementations}
+The CASA and ASKAPsoft approaches should be equivalent apart from numerical differences such as due to the convolution function.
+Hence either can be used according to the most favourable scaling.
+Compared to CASA, ASKAPsoft does approximately two orders of magnitude less FFTs and 5 times as much gridding.
+Both forms should therefore be available in the SDP Performance Model and used as the context dictates.
 
-\subsection{Serial}
-\label{sec:serial}
-
-The serial algorithm is as shown in Figure \ref{fig:algoCASA}.
+The Minor Cycle is identical in CASA and ASKAPsoft. The Minor cycle is depicted graphically in Figure \ref{fig:minor}.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{./algorithm.pdf}
-  \caption{The MSMFS algorithm as implemented in CASA [RD??]}
-  \label{fig:algoCASA}
+  \includegraphics[width=\textwidth]{./MSMFS_Minor.pdf}
+  \caption{Structure of Minor cycle for MSMFS as used in both CASA and ASKAPsoft.}
+  \label{fig:minor}
 \end{figure}
@@
 \clearpage
-
-%\begin{algorithm}[htp!]
-%\scriptsize
-%\label{alg:MSMFS_1}
-%  \KwData{calibrated visibilities : $\vec{V}^{corr}_{\nu}~~\forall \nu$}
-%  \KwData{Fourier plane sampling function : {\tens{S_\nu}{4} }}
-%  \KwData{image noise threshold and loop gain $\sigma_{thr}, g_s$}
-%  \KwData{scale basis functions : $\I^{shp}_s ~ \forall s\in\{0,N_s-1\}$}
-%  \KwResult{model coefficient images : $\I^{m}_{q}~ \forall q\in\{0,N_t-1\}$}
-%  \KwResult{spectral index and curvature : $\I^{m}_{\alpha},\I^{m}_{\beta}$}
-%
-%  \vspace{0.5cm} 
-%  \For{ $t \in \{0,N_t-1\}, q \in \{t,N_t-1\}$}
-%  {
-%        { Compute the spectral PSF } $\vec{I}^{psf}_{tq}$\;
-%        \For{ $s \in \{0,N_s-1\}, p \in \{s,N_s-1\}$}
-%	{
-%		{Compute the scale-spectral PSF} $\vec{I}^{psf}_{{sp}\atop{tq}} = \vec{I}^{shp}_s \star \vec{I}^{shp}_p \star \vec{I}^{psf}_{tq} $\;
-%	}
-%  }
-%  \For { $s \in \{0,N_s-1\}$}
-%  {
-%     Construct $[{H^{peak}_s}]$ from $mid(I^{psf}_{{s,s}\atop{t,q}})$ and compute $[{H^{peak}_s}^{-1}]$\;
-%  }
-%  Initialize the model $\vec{I}^{m}_t$ for all $t \in \{0,N_t-1\}$ and compute $f_{sidelobe}$ \;
-%  \Repeat (\tcc*[f]{Major Cycle}) { Peak residual in $\vec{I}^{res}_0 < \sigma_{thr}$ }
-%  {
-%    \For{$t \in \{0,N_t$-$1\}$}
-%    {
-%      Compute the residual image $\vec{I}^{res}_t$ \tcc*[r]{Makeimage}
-%      \For{$s \in \{0,N_s$-$1\}$}
-%      {
-%	      Compute $\vec{I}^{res}_{{s},{t}} = \vec{I}^{shp}_s \star \vec{I}^{res}_t$
-%      }
-%    }
-%    Calculate $f_{limit}$ from $\vec{I}^{res}_{0,0}$\;
-%    \Repeat (\tcc*[f]{Minor Cycle}){ Peak residual in $\vec{I}^{res}_{0,0} < f_{limit} $ } 
-%    {
-%     \For{$s \in \{0,N_s$-$1\}$} 
-%     {
-%      \uIf{Peak of $\vec{I}^{res}_{s,0} > 10~\sigma_{thr} $}
-%      {
-%       \ForEach{pixel}
-%       {
-%          Construct $I^{rhs}_s$, an $N_t\times 1$ vector from $I^{res}_{s,t} ~~\forall ~ t \in \{0,N_t$-$1\}$\;
-%          Compute principal solution $I^{sol}_s = [{H^{peak}_s}^{-1}] I^{rhs}_s$\;
-%       }
-%       Choose $I^{sol} = max\{I^{sol}_{t=0},~\forall~s\in\{0,N_s$-$1\}\}$ \;
-%      }
-%      \Else
-%      {
-%       Find the location of the peak in $\vec{I}^{res}_{s,0},~\forall~s\in\{0,N_s$-$1\}$\;
-%       Construct $I^{rhs}_s$, from $I^{res}_{s,t}$ for the chosen $s$, at this location\;
-%       Compute $I^{sol} = [{H^{peak}_s}^{-1}] I^{rhs}_s$ at this location\;
-%      }
-%     }
-%       \For{$t \in \{0,N_t-1\}$}
-%       {
-%        Update the model image : $I^{m}_t = I^{m}_t + g_s ~ I^{shp}_{s_i} \star I^{sol}_t$ \;
-%        \For{$s \in \{0,N_s$-$1\}$}
-%	{
-%          Update the residual image : $I^{res}_{s,t} = I^{res}_{s,t} - g ~\sum_{p=0}^{N_s-1}\sum_{q=0}^{N_t-1}[I^{psf}_{{sp}\atop{tq}} \star I^{sol}_q]$\;
-%	}
-%       }
-%    }
-%   Compute model visibilities $V^{m}_{\nu}$ from  $I^{m}_t~\forall t\in \{0.N_t-1\}$\tcc*[r]{Predict} 
-%   Compute a new residual image $I^{res}$ from residual visibilities $V^{corr}_{\nu}-V^{m}_{\nu}$ \tcc*[r]{Makeimage}
-%}
-%Calculate $\vec{I}^m_{\nu_0}, \vec{I}^{\alpha}, \vec{I}^{\beta}$ from $I^{m}_t~\forall t\in \{0.N_t-1\}$ and restore the results \;
-%If required, remove average primary beam : $\vec{I}^{new}_{\nu_0}=\vec{I}^m_{\nu_0}/\vec{\Pb}_{\nu_0}, \vec{I}^{new}_{\alpha}=\vec{I}^m_{\alpha}-\vec{\Pb}_{\alpha}, \vec{I}^{new}_{\beta}=\vec{I}^m_{\beta}-\vec{\Pb}_{\beta}$\;
-%\vspace{0.5cm}
-%
-%  \caption[MS-MFS Algorithm]
-%          {MS-MFS Algorithm : }
-%\end{algorithm}
-
-\clearpage
-\subsection{Parallel processing}
+\section{Parallel processing}
 \label{sec:parallel}
 
-When moving the MSMFS algorithm to a distributed (slow interconnect) and parallel (fast interconnect) architecture such as the SDP architecture, there are a large number of important factors that must be considered.
+When moving the MSMFS algorithm to a distributed (loosely coupled) and parallel (tightly coupled) architecture such as the SDP architecture, there are a large number of important factors that must be considered.
 \begin{description}
-\item[Order of iterations]
+\item[Optimum partitioning] 
 \item[Coupling across partitions] e.g. is the same pixel addressed in two different partitions? If so, is there a satisfactory approach to reconciling the two values?
@@
 
-\subsubsection{Minor cycle}
-\label{subsec:minorcycle}
-
-\subsubsection{Partitioning}
-\label{subsec:partitoning}
+\subsection{Partitioning for Predict and MakeImage}
+\label{subsec:partitoningpredictmakeimage}
 
@@
 According to the logic of the MSMFS algorithm, Sub-bands can be defined as the limit of the power law approximation for brightness so consequently can always be used as the coarsest partition.
-For Predict and MakeImage the remaining choices are between (most rapid first):
+For Predict and MakeImage the remaining choices are as shown in Figure \ref{tab:partitions} (most rapid first):
 
@@
  \caption{Possible ordering of partitions $UV$ weighting version of Predict and MakeImage}
+ \label{tab:partitions}
  \begin{center}
- \begin{tabular}{|c|c|c|c|c|}
+ \begin{tabular}{|c|c|c|c|}
 \hline
@@
 
-Frequency and Taylor terms are tight coupled and so to avoid excess inter-CU communication, the order should be [Taylor, Frequency].
-The remaining choices are then between the two orderings of Time and Facet.
-Since deconvolution is non-linear, we prefer that all times be represented in a single facet rather than the other way around.
+The only choice is between the two orderings of Time and Facet. Since deconvolution is non-linear, we prefer that all times be represented in a single facet rather than the other way around.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{./MSMFS_Order.pdf}
-  \caption{Preferred order of parallelisation and distribution for MSMFS algorithm}
+  \includegraphics[width=\textwidth]{./Order.pdf}
+  \caption{Preferred order of parallelisation and distribution for MakeImage and Predict sub-pipelines.}
   \label{fig:order}
@@
 The first approach will inevitably introduce edge effects in the final image revealing the facet grid, while the second will block processing while deconvolution proceeds on the different sub-band compute nodes.
-We would then require two different types of nodes:
+We would then benefit from two different types of nodes:
 \begin{description}
@@
 
-\subsection{Minor Cycles}
+\subsection{Partitioning for Minor Cycle}
 
-\subsubsection{CASA}
+In the minor cycle, the first calculation is [Pixel, Taylor Term], and then the Scale convolutions are calculated.
+Hence the natural order overall is [Pixel, Taylor Term, Scale].
+This is used in both CASA and ASKAPsoft.
+Both use a single threaded Minor Cycle.
+For SDP, a multi-threaded but not-distributed Minor Cycle should be prototyped.
 
-\subsubsection{ASKAPsoft}
-The core algorithm is in C++ Template DeconvolverMultiTermBasisFunction.tcc.
+We have a modelling problem that is not going to be wished away.
+The science demands seamless imaging in multiple axes, whereas computing would like to partition the image cube in those same axes.
+There is a tendency to assert that one axis or another does not have to be seamless but that has not be conclusively argued.
+Distributing the minor cycle across Compute Islands may be possible in some cases (e.g.
+spectral line after continuum estimation and subtraction), but we should also be considering how best to parallelise the Minor Cycle on a single, appropriately specified Compute Island.
+
+\subsection{Implementation in CASA}
+
+The CASA tclean program supports parallel execution using python to run copies on different nodes.
+The visibility data are partitioned in time across the available processors.
+Gridding and transforms (i.e.
+Predict and MakeImage) are done separately and in parallel per partition.
+Then the Images (and weights) are gathered and then normalised.
+A serial minor cycle is run, blocking all other processing, and then the model image is scattered to all processes.
+The Major cycle is done in parallel per partition.
+
+Thus in summary, the Major Cycles are parallelised while the deconvolution runs serially.
+
+\subsection{Implementation in ASKAPsoft}
+
+The ASKAPsoft code is mirrored in the SDP Github repository.
+The ASKAPsoft cimager is an MPI-enabled program.
+Since it was written as parallel program from the beginning, the structure is easier to follow than the CASA version.
+Distribution over MPI is via frequency, rather than time as in CASA.
+Predict and MakeImage are distributed using MPI (via a thin bookkeeping layer).
+The MinorCycle is performed on the master, and thus blocks until the deconvolution steps have completed.
+
+The minor cycle  algorithm is in C++ Template DeconvolverMultiTermBasisFunction.tcc.
 The basis function is abstracted and can be any class having the interface DeconvolverMultiTermBasisFunction.
 There are two forms present: one for point sources, and one for the same blobs used in CASA - a truncated upside down parabola.
+The basis functions may optionally be orthogonalised using the Gramm-Schmidt algorithm.
 
-The expansion in Taylor terms is generalised to any form obeying an equation like: $D = B(0)*I(0) + B(1)*I(1) + B(2)*I(2)$, where D is the dirty image, B(?) are the spectral dirty beams, and I(?) are the images for each term.
-The coupling matrices between the terms are calculated for each basis function.
-Once a suitable peak component is found, the inverse of the coupling matrix is used to decouple the different terms for the optimum scale.
-The optimum blob location and scale is found by using one of a number of criteria:
+The optimum blob location and scale is found by using one of two criteria:
+
 \begin{description}
@@
 The second criterion is expensive to compute and usually MAXBASE0 is used.
+Thus for each scale and for each pixel the pixel value at the reference frequency has to be determined by pre-multipling by the inverse Hessian (in term-term space).
+This means that the cost of the search is:
 
-Once the optimum blob scale and location is found, the vector in term-space is found by decoupling.
-The model images (one for each term) are then updated, and the effects of this blob removed from all the cached images.
+\begin{equation}
+Rflop_{MSMFSminor} = N_{minor} N_{pix}^2 N_s N_T	
+\end{equation}
+
+To save on calculation, a number of cached images are used. For ASKAPsoft, these are shown in Table \ref{tab:askapmemory}.
+
+\begin{table}[tbh]
+\caption{ASKAPsoft cimager use of memory during minor cycles}\label{tab:askapmemory}
+
+\begin{tabular}{|l|c|c|c|}
+\hline
+Item & Axes & Size (words) & Needed? \\
+\hline
+Basis function FFT & scales & $N_{pix} N_S $ & During initialisation \\
+PSF FFT & 2 * Taylor terms & $2 N_{pix} N_T $ & During initialisation \\
+Spectral Dirty image & Taylor terms, scales & $N_{pix} N_T N_s$ & During iteration \\
+Spectral Dirty PSF & Taylor terms, scales & $N_{patch} N_T N_s$ & During iteration \\
+Spectral Model image & Taylor terms & $N_{pix} N_T $ & During iteration \\	
+Basis functions & scales & $N_{pix} N_s$ & During iteration \\
+Inverse of Hessian & Taylor terms, Taylor terms, scales & $N^2_T N_s$ & During Iteration \\
+\hline
+\end{tabular}	
+\end{table}
+
+where $N_{pix}$ is the number of pixels in an image, $N_{patch}$ is the number of pixels in the patch of the Dirty beam used, $N_T$ is the number of Taylor terms, and $N_S$ is the number of scales/basis functions.
+
+Thus, keeping only the dominant terms, the total memory used during iteration is:
+
+\begin{equation}
+M_{MSMFS} = (N_T (N_s + 1) + N_s) N_{pix} 
+\end{equation}
+
+Using typical values of $N_t=5$ and $N_s=5$, we find that the memory use is about $35 N_{pix}$.
+For a 200,000 by 200,000 pixel image, this requires 5.6 TB total.
+Spread over O(100) Compute Islands (partitioning by facet and sub-band), this requires 56 GB per CI.
 
@@
 \clearpage
-\section{Modelling MSMFS in the SDP Performance Manager}
-\label{sec:modelling}
+\section{Recommendations}
+\label{sec:recommendations}
 
-\clearpage
-\section{Resource usage}
-\label{sec:resource}
+\begin{enumerate}
+\item Implement both Image and Visibility based calculation of the Taylor term images in the Performance Model.
+\item Update the Performance Model to reflect the memory usage:
+\begin{equation}
+M_{MSMFS} = (N_T (N_s + 1) + N_s) N_{pix} 
+\end{equation}
+\item A multi-threaded Minor Cycle on one Compute Island should be prototyped.
+\end{enumerate}
 

### 2016-04-26 msmfs.tex
Added much material
Is MSMFS needed?
Equations about the Hessian etc
More recommendations

@@
 
-\documentclass[11pt,a4paper]{article}
+\documentclass[11pt,a4paper,variablewidth]{article}
 
@@
 \usepackage{fmtcount}
+
+\usepackage{array}
+\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}
 
@@
 \newcommand{\Vd}{\vec{V^\dag}}
-\newcommand{\A}{{\tens{A}{s}}}
-\newcommand{\Ad}{{\tens{A^\dag}{s}}}
+\newcommand{\A}{{\tens{A}{}}}
+\newcommand{\Ad}{{\tens{A^\dag}{}}}
 \newcommand{\F}{{\tens{F}{s}}}
-\newcommand{\Fd}{{\tens{F^\dag}{s}}}
-\newcommand{\He}{{\tens{H}{s}}}
-\newcommand{\Sa}{{\tens{S}{s}}}
-\newcommand{\Sd}{{\tens{S^\dag}{s}}}
-\newcommand{\Sna}{\tens{{S_{\nu}}{s}}}
-\newcommand{\Snd}{\tens{{S_{\nu}^\dag}{s}}}
-\newcommand{\T}{{\tens{T}{s}}}
-\newcommand{\W}{{\tens{W}{s}}}
-\newcommand{\Wd}{{\tens{W^\dag}{s}}}
+\newcommand{\Fd}{{\tens{F^\dag}{}}}
+\newcommand{\He}{{\tens{H}{}}}
+\newcommand{\Sa}{{\tens{S}{}}}
+\newcommand{\Sd}{{\tens{S^\dag}{}}}
+\newcommand{\Sna}{\tens{{S_{\nu}}{}}}
+\newcommand{\Snd}{\tens{{S_{\nu}^\dag}{}}}
+\newcommand{\T}{{\tens{T}{}}}
+\newcommand{\W}{{\tens{W}{}}}
+\newcommand{\Wd}{{\tens{W^\dag}{}}}
 \newcommand{\Pb}{{\vec{P}}}
@@
 \newcommand{\Nc}{N_{\rm c}}
-
-% Add here the symbols used in your document
-%\clearpage \sdplistofsymbols
-%\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
-%\item[$\bytes$] Length of a complex number in bytes \vspace{-0.2cm}
-%\item[$\cbytes$] Length of a complex number in bytes \vspace{-0.2cm}
-%\item[$\rbytes$] Length of a real number in bytes \vspace{-0.2cm}
-%\item[$\bl$] Baseline length / GPU memory bandwidth \vspace{-0.2cm}
-%\item[$\maxbl$] Maximum baseline length \vspace{-0.2cm}
-%\item[$\membw$] Memory bandwidth \vspace{-0.2cm}
-%\item [$\pciebw$] PCI-e bandwidth \vspace{-0.2cm}
-%\item[$\viscompr$] Compression factor of visibilities for grdding for facets
-%  \textcolor{red}{This symbol is not used in the document, consider removing.}
-%  \vspace{-0.2cm}
-%\item[$\vism$] Matrix of observed visibilities \vspace{-0.2cm}
-%\item[$\da$] Extent of aperture \vspace{-0.2cm}
-%\item[$\ds$] Diameter of antennas/stations \vspace{-0.2cm}
-%\item[$\reff$] Reference frequency\vspace{-0.2cm}
-%\item[$\fcc$] Convolution kernel computational cost\vspace{-0.2cm}
-%\item[$\fci$] Factor between bandwidth and FLOPS (bytes/operations)\vspace{-0.2cm}
-%\item[$\fgrid$] Gridding cost\vspace{-0.2cm}
-%\item[$\fpatch$] Fraction of the maximum baseline below which the $\uv$
-%  coverage is almost filled\vspace{-0.2cm}
-%\item[$\fpr$] Cost for phase rotation\vspace{-0.2cm}
-%\item[$\gainm$] 2$\times$2 block diagonal matrix of complex
-%  gains\vspace{-0.2cm}
-%\item[$\modskyvism$] Matrix of model sky visibilities\vspace{-0.2cm}
-%\item[$\skymodbufsize$] Size of the sky model buffer\vspace{-0.2cm}
-%\item[$\visbufsize$] Size of the visibility buffer \vspace{-0.2cm}
-%\item[$\visbufsizespec$] Size of the visibility buffer needed for spectral line processing\vspace{-0.2cm}
-%\item[$\visbufsizecont$] Size of the visibility buffer needed for continuum imaging\vspace{-0.2cm}
-%\item[$\bufsize$] Size of buffer \vspace{-0.2cm}
-%\item[$\slowmemsize$] Size of slow working memory \vspace{-0.2cm}
-%\item[$\cumemsize$] Size of working memory of compute unit \vspace{-0.2cm}
-%\item[$\imgsize$] Size of image (bytes) \vspace{-0.2cm}
-%\item[$\imgsizeoneax$] Image size along one axis \vspace{-0.2cm}
-%\item[$\targgridsize$] Target grid size (per facet) \vspace{-0.2cm}
-%\item[$\uvgridsize$] Size of the $\uv$ grid (bytes) \vspace{-0.2cm}
-%\item[$\vissize$] Size of visibility data (bytes) \vspace{-0.2cm}
-%\item[$\wgridsize$] Size of weight grid (bytes)\vspace{-0.2cm}
-%\item[$\gridcachesize$] Cache size of gridding kernel \vspace{-0.2cm}
-%\item[$\nbins$] Number of bins \vspace{-0.2cm}
-%\item[$\ntel$] Number of antennas or stations in the array
-%  \vspace{-0.2cm}
-%\item[$\nateam$] Number of A-team sources \vspace{-0.2cm}
-%\item[$\ngridpakern$] Number of grid points for A-kernel support (along one
-%  axis) \vspace{-0.2cm}
-%\item[$\naccvis$] Number of accesses per visibility \vspace{-0.2cm}
-%\item[$\navg$] Number of averages\vspace{-0.2cm}
-%\item[$\nbeam$] Number of beams simultaneously observed by the array
-%  \vspace{-0.2cm}
-%\item[$\nbits$] Number of bits\vspace{-0.2cm}
-%\item[$\rma$] Number of flops in real multiply and add\vspace{-0.2cm}
-%\item[$\cma$] Number of flops in complex multiply and add\vspace{-0.2cm}
-%\item[$\nbl$] Number of baselines \vspace{-0.2cm}
-%\item[$\nbyteacc$] Number of bytes per access \vspace{-0.2cm}
-%\item[$\nbyte$] Number of bytes \vspace{-0.2cm}
-%\item[$\nbyteperpix$] Number of bytes per pixel \vspace{-0.2cm}
-%\item[$\nbytepervis$] Number of bytes per individual visibility 
-%
-%\item[$\nch$] Number of channels \vspace{-0.2cm}
-%\item[$\ncompu$] Number of compute units \vspace{-0.2cm}
-%\item[$\ndatait$] Number of data items required per source (e.g. position)\vspace{-0.2cm}
-%\item[$\ngridp$] Number of grid points along one axis for the gridding
-%  convolution function \vspace{-0.2cm}
-%\item[$\nfch$] Number of frequency channels/sub-bands \vspace{-0.2cm}
-%\item[$\nfchcorr$] Number of frequency channels output by the correlator
-%  \vspace{-0.2cm}
-%\item[$\nfchgrid$] Number of frequency channels to be gridded
-%  \vspace{-0.2cm}
-%\item[$\nfchdegrid$] Number of frequency channels to be de-gridded
-%  \vspace{-0.2cm}
-%\item[$\nfchout$] Number of frequency channels to be FFT-ed and reprojected
-%  (output) or output from the CSP pulsar timing backend\vspace{-0.2cm}
-%\item[$\nfchdist$] Minimum number of frequency channels to maintain distributability in the SDP
-%\item[$\nfkernel$] Number of convolution kernels needed to cover frequency axis for a given baseline
-%\item[$\nfacet$] Number of facets along one axis \vspace{-0.2cm}
-%\item[$\nflop$] Number of floating point operations
-%  \vspace{-0.2cm}
-%\item[$\nflopsamp$] Number of floating point operations per sample
-%  \vspace{-0.2cm}
-%\item[$\nflopvis$] Number of floating point operations per visibility
-%\item[$\npixgrid$] Number of pixels in the grid (i.e. the square of the linear
-%  dimension) \vspace{-0.2cm}
-%\item[$\ngridwkern$] Number of grid points for w-kernel support (along one
-%  axis) \vspace{-0.2cm}
-%\item[$\nit$] Number of iterations \vspace{-0.2cm}
-%\item[$\nawkern$] Linear size of combined A and $\w$-kernels \vspace{-0.2cm}
-%\item[$\nmaj$] Number of major CLEAN cycles (within each self-cal loop if appropriate) \vspace{-0.2cm}
-%\item[$\nselfcal$] Number of self-calibration cycles \vspace{-0.2cm}
-%\item[$\nmin$] Number of minor CLEAN cycles (within each major cycle loop if appropriate)\vspace{-0.2cm}
-%\item[$\facgridoffdiagmuller$] Factor to account for gridding of off-diagonal
-%  M\"uller terms \vspace{-0.2cm}
-%\item[$\nops$] Number of operations \vspace{-0.2cm}
-%\item[$\npatchpix$] Number of pixels of PSF to be used in minor cycle
-%  decorrelation \vspace{-0.2cm}
-%\item[$\npix$] Number of pixels along one axis of a grid\vspace{-0.2cm}
-%\item[$\npixfacet$] Number of pixels on one side of a facet \vspace{-0.2cm}
-%\item[$\npol$] Number of polarisation products \vspace{-0.2cm}
-%\item[$\nsource$] Number of sources \vspace{-0.2cm}
-%\item[$\ntslot$] Number of time slots \vspace{-0.2cm}
-%\item[$\ntaylor$] Number of Taylor terms (in MS-MFS)\vspace{-0.2cm}
-%\item[$\ntslotave$] Number of time slots that are averaged (in de-mixing)
-%  \vspace{-0.2cm}
-%\item[$\nvis$] Number of visibilities in an observation \vspace{-0.2cm}
-%\item[$\qfacbw$] Quality factor \vspace{-0.2cm}
-%\item[$\qfacreuse$] Factor to allow for the reuse of functions between
-%  neighbouring frequency channels \vspace{-0.2cm}
-%\item[$\qfacfov$] Quality factor defining how much bigger the diameter of the
-%  field of view is than the first zero of the Airy function \vspace{-0.2cm}
-%\item[$\facgridovers$] Gridding oversampling factor \vspace{-0.2cm}
-%\item[$\facmsmf$] Factor to account for multiple subtractions (multi-frequency
-%  multi-scale CLEAN\vspace{-0.2cm}
-%\item[$\qfacbeamovers$] Quality factor defining the oversampling of the
-%  synthesised beam \vspace{-0.2cm}
-%\item[$\qfac$] Quality factor \vspace{-0.2cm}
-%\item[$\Qkernel$] Convolution kernel quality factor, for kernel re-use in frequency
-%\item[$\corrfacw$] Correction factor for w values (0 to 1) \vspace{-0.2cm}
-%\item[$\earthrad$] Radius of the Earth ($6400\times 10^3$\,m) \vspace{-0.2cm}
-%\item[$\covm$] Covariance matrix \vspace{-0.2cm}
-%\item[$\covmfilt$] Filtered covariance matrix \vspace{-0.2cm}
-%\item[$\comprate$] Compute rate (FLOPS) \vspace{-0.2cm}
-%\item[$\compratespec$] Compute rate (FLOPS) for spectral line processing \vspace{-0.2cm}
-%\item[$\compratecont$] Compute rate (FLOPS) for continuum imaging \vspace{-0.2cm}
-%\item[$\compratefast$] Compute rate (FLOPS) for fast imaging \vspace{-0.2cm}
-%\item[$\bwtoworkmem$] Bandwidth to working memory \vspace{-0.2cm}
-%\item[$\maxmembw$] Maximum memory bandwidth of compute unit \vspace{-0.2cm}
-%\item[$\peakflop$] Peak FLOPS of compute unit \vspace{-0.2cm}
-%\item[$\maxiobwbuff$] Maximum I/O bandwidth of compute unit to buffer
-%  \vspace{-0.2cm}
-%\item[$\bwbuff$] Bandwidth to buffer \vspace{-0.2cm}
-%\item[$\bwbuffcont$] Bandwidth to buffer for continuum imaging\vspace{-0.2cm}
-%\item[$\bwbuffspec$] Bandwidth to buffer for spectral line processing\vspace{-0.2cm}
-%\item[$\convkernflops$] Convolution kernel compute rate (FLOPS)\vspace{-0.2cm}
-%\item[$\fftflops$] FFT compute rate (FLOPS)\vspace{-0.2cm}
-%\item[$\griddingflops$] Gridding compute rate (FLOPS)\vspace{-0.2cm}
-%\item[$\phaserotflops$] Phase rotation compute rate (FLOPS)\vspace{-0.2cm}
-%\item[$\reprojflops$] Re-projection compute rate (FLOPS)\vspace{-0.2cm}
-%\item[$\rcorr$] Correlator output rate\vspace{-0.2cm} TRY TO REMOVE
-%\item[$\facetoverlap$]  ratio for linear size increase of facets to account for overlap, typically 1.2-1.5 \vspace{-0.2cm}
-%\item[$\telsensit$] Telescope sensitivity \vspace{-0.2cm} TRY TO REMOVE
-%\item[$\corrdumpt$] Correlator dump time (sec) \vspace{-0.2cm}
-%\item[$\compt$] Computation time \vspace{-0.2cm}
-%\item[$\gridkernupdate$] Update timescale for gridding kernel \vspace{-0.2cm}
-%\item[$\lsmupdate$] Update timescale for the LSM \vspace{-0.2cm}
-%\item[$\obst$] Observation time \vspace{-0.2cm}
-%\item[$\snapshot$] Snapshot duration \vspace{-0.2cm}
-%\item[$\minsnapshot$] Minimum duration of a snapshot in which w-correction is
-%  not dominated by the Earth's curvature \vspace{-0.2cm}
-%\item[$\optsnapshot$] Optimum snapshot duration (sec) \vspace{-0.2cm}
-%\item[$\subint$] Sub-integration time\vspace{-0.2cm}
-%\item[$\uvw$] Visibility coordinates \vspace{-0.2cm}
-%\item [$\w$] Visibility $w$ coordinate \vspace{-0.2cm}
-%\item[$\maxwdev$] Maximum deviation of the $w$ coordinate due to the Earth's
-%  curvature \vspace{-0.2cm}
-%\item[$\minwdev$] Minimum deviation of the $w$ coordinate due to the Earth's
-%  curvature \vspace{-0.2cm}
-%\item[$\spfiltm$] Spatial filter matrix \vspace{-0.2cm}
-%\item[$\fres$] Frequency resolution \vspace{-0.2cm}
-%\item[$\tres$] Time resolution (sec) \vspace{-0.2cm}
-%\item[$\tressmear$]  Smearing Time resolution (sec)  \vspace{-0.2cm}
-%\item[$\fracconvfuncamp$] Fraction of maximal amplitude of envelope of
-%  convolution function \vspace{-0.2cm}
-%\item[$\compeff$] Computational efficiency \vspace{-0.2cm}
-%\item[$\fracfband$] Fraction of frequency band \vspace{-0.2cm}
-%\item[$\beamsq$] Beam squint \vspace{-0.2cm}
-%\item[$\fovdiam$] Angular diameter of the field of view (rad) \vspace{-0.2cm}
-%\item[$\pixres$] Angular resolution of pixel size (rad) \vspace{-0.2cm}
-%\item[$\beamdiam$] Angular diameter of the synthesised beam (rad)
-%  \vspace{-0.2cm}
-%\item[$\wavel$] Wavelength \vspace{-0.2cm}
-%\item[$\maxwavel$] Maximum wavelength \vspace{-0.2cm}
-%\item[$\minwavel$] Minimum wavelength \vspace{-0.2cm}
-%\item[$\maxwavelsubb$] Maximum wavelength of an imaging sub-band (channels
-%  over which image and pixel size are matched) \vspace{-0.2cm}
-%\item[$\minwavelsubb$] Minimum wavelength of an imaging sub-band (channels
-%  over which image and pixel size are matched) \vspace{-0.2cm}
-%\item[$\earthrot$] Angular velocity of the Earth
-%  ($7.27\times10^{-5}$\,rad/sec) \vspace{-0.2cm}
-%\item[$\epsilonf$] Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging \vspace{-0.2cm}
-%\end{basedescript}
 
@@
 Monthly Notices of the Royal Astronomical Society, 246, 490.\\
+   {\bf [RD36]} & http://www.aanda.org/component/citedby/?task=crossref\&doi=10.1051/0004-6361/201117104 \\  
+   {\bf [RD37]} & Junklewitz, H., Bell, M.
+R., and Ensslin, T.
+(2015).
+A new approach to multifrequency synthesis in radio interferometry.
+Astronomy \& Astrophysics, 581, A59.
+http://doi.org/10.1051/0004-6361/201423465 \\
    \hline
@@
 
+\clearpage
 \section{Purpose of the document}
@@
 
-MS-MFS was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g.
+The imaging performance of a radio synthesis telescope is related to the uv coverage.
+The uv coverage is dependent on the observing wavelength so diversity in uv can be obtained by observing over a range of frequencies [RD35].
+This Multi-Frequency Synthesis (MFS) only works for radio sources in which broadband emission is present, giving a way to relate measurements at different frequencies.
+There have been multiple attempts at an MFS algorithm (e.g.
+[RD35]).
+MFS is usually most important for sources having extended emission.
+Hence multi-scale algorithms are favoured.
+
+The Multi-Scale, Multi-Frequency Synthesis (MS-MFS) algorithm was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g.
 [RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data.
 The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law expanded into a Taylor series.
+It has been used in JVLA, ALMA, and ATCA observations [RD36].
 
@@
 \clearpage
+
+\section{Is MFS needed?}
+\label{sec:mfsneeded}
+
+There is an L1 requirement, $SKA1-SYS\_REQ-2324$ specifying that MSMFS must be used (see Table\ref{tab:L1}).
+
+\begin{table}[htp]
+\begin{tabular}{|c|L|c|c|c|}
+\hline
+ID & Requirement & Status & Parent Requirement & Verification \\
+\hline
+$SKA1-SYS\_REQ-2324$ &Multi-frequency synthesis imaging.
+All imaging shall construct and make use of frequency dependent image models over the entire observed bandwidth.
+& Accepted & Reference [2] & Demonstration \\
+\hline 
+\end{tabular}	
+\caption{Relevant L1 requirement}\label{tab:L1}	
+\end{table}
+
+This is reflected in the L2 requirements shown in Table\ref{tab:L2}.
+\begin{table}[htp]
+\begin{tabular}{|L|L|L|L|L|}
+\hline
+\multicolumn{2}{|c|}{Function}&\multicolumn{3}{c|}{Satisfies requirement}\\
+\hline
+Number & 	Name & 	Number	& Name	& Description \\
+\hline
+F1.9 & Calibrate and Image & $SDP\_REQ-390$	&Multi-frequency synthesis
+  imaging	& The SDP shall construct and make
+  use of frequency-dependent image models over the entire observed bandwidth. \\
+\hline
+F1.9 & Calibrate and Image & $SDP\_REQ-676$	& Continuum imaging deconvolution
+  scales &	While performing continuum
+  imaging the SDP shall deconvolve using a maximum number of 10 (TBC) scales.\\
+\hline
+F1.9 & Calibrate and Image & $SDP\_REQ-677$	& Continuum imaging deconvolution
+  Taylor terms &	While performing continuum
+  imaging the SDP shall deconvolve using a maximum number of 5 (TBC) Taylor
+  terms.\\
+\hline
+F1.9 & Calibrate and Image & $SDP\_REQ-678$	& Continuum imaging deconvolution
+  by sub-band	& While performing continuum
+  imaging the SDP shall deconvolve by sub-band (TBC).\\
+\hline 
+\end{tabular}
+\caption{Relevant L2 requirements}	\label{tab:L2}
+\end{table}
+
+The last three L2 requirements, $SDP\_REQ-676$, $SDP\_REQ-677$, $SDP\_REQ-678$, are an attempt to give some constraints to $SKA1-SYS\_REQ-2324$.
+However, they beg the same question as the original L1 requirement: is MSMFS needed to meet the scientific goals?
+If so, how many scales, Taylor terms, and sub-bands are required?
+In addition, there is also a potentially missing requirement concerning the connection between the models for different sub-bands - is there allowed to be a seam in the continuum-subtracted spectral line cube at the transitions between sub-bands?
+
+Finally we could ask is there a superior algorithm already published? (see e.g. [RD37]).
+
+Some caution is advised in advocating that any form of MFS is not needed.
+Clearly ALMA and JVLA, which also have relatively good uv coverage, have benefited from the use of MSMFS.
+Furthermore, although the uv-coverage of SKA1-LOW and SKA1-MID is excellent after a long observation, there will inevitably be pressure to observe with shorter observation times.
+We should remember than the VLA was supposed to take 8 hours to image every field.
+With the advent of CLEAN and self-calibration, it was known even in the early eighties that 400 sources could be imaged in a day.
+
+There is no mechanism to address these questions beyond simulations. Arguing from uv coverage metrics such as the UVGAP only helps in a relative sense.
+
+\clearpage
+
 \section{Overview of MSMFS}
@@
 
-Before performing any search, it is necessary to decouple the pixels in Taylor term space as well as possible.
-In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used.
-The greedy algorithm then locates the scale and location in the reference-frequency image (as convolved with the set of scales).
-For each minor cycle this produces a location, scale, and values for all Taylor terms.
-The residual images and model images are updated, and iteration proceeds.
-The actual algorithm is shown in Figure \ref{fig:CASA}.
-
 The image flux model at each frequency can be written as a linear sum of  
@@
 
+The visibility function corresponding to \ref{eqn:msmf_model} is:
+\begin{eqnarray}
+\vec{V}^{\rm obs}_{\nu} &=& [\Sna][\F]\vec{I}^{\rm m}_{\nu}
+= \sum_{t=0}^{\Nt } \sum_{s=0}^{\Ns } \wnt [\Sna][\T_s][\F] \vec{I}^{\rm sky}_{s\atop t}
+\label{eqn:msmfs_meqn}
+\end{eqnarray}
+
+$\vec{V}^{\rm obs}_{\nu}$ is a vector of $n\times 1$ visibilities
+measured at frequency $\nu$.
+$w_{\nu}$ are Taylor-weights (shown in Eqn.\ref{Eq:msmf_model}).
+$[\Sna]$ is an $n\times m$ projection operator that represents the 
+spatial-frequency sampling function for frequency $\nu$. 
+The image-domain convolution of model $\vec{I}^{\rm sky}_{s\atop t}$ with
+ $\I^{\rm shp}_s$ is written as a spatial-frequency-domain multiplication.
+$ \vec{I}^{\rm shp}_s \star \vec{I}^{\rm sky}_{s\atop t} = [\Fd][\T_s][\F] \vec{I}^{\rm sky}_{s\atop t}$
+where $[\T_s]_{m\times m} = diag([\F] \vec{I}^{\rm shp}_s)$ is a spatial-frequency
+taper function.  All images are vectors of shape $m\times 1$.
+
+The normal equations 
+for the system described in Eqn. \ref{Eq:msmfs_meqn}
+can be written in block matrix form, with each block-row 
+(for scale size $s$, and Taylor term $t$) given by
+\begin{eqnarray}
+\label{Eq:msmfs_neqn_1}
+\sum_{p=0}^{\Ns -1}\sum_{q=0}^{\Nt -1} \left[\He_{{s,p}\atop{t,q}}\right] \vec{I}^{\rm sky}_{{p}\atop{q}} &=& \vec{I}^{\rm dirty}_{{s}\atop{t}}~~~  \forall~ s \in [0,\Ns -1], t\in[0,\Nt -1]
+\end{eqnarray}
+Here, each $\left[\He_{{s,p}\atop{t,q}} \right]$ is an $m\times m$ block of the 
+Hessian matrix, and $\vec{I}^{\rm dirty}_{{s}\atop{t}}$ is one of $\Ns  \Nt $ dirty images.
+\begin{eqnarray}
+\left[\He_{{s,p}\atop{t,q}} \right] &=& \left[\A_{{s}\atop{t}}^{\dag}\right][\Wim] \left[\A_{{p}\atop{q}}\right] \\
+ &=&    [\Fd \T_s  \Sd \Wntd ]  [\Wim]  [\W^{\rm mfs}_q \Sa \T_p \F] \\
+ &=&    [\Fd \T_s \F] [\Fd  \Sd \Wntd  \Wim  \W^{\rm mfs}_q \Sa  \F] [\Fd \T_p \F]\\
+% &=&    [\Fd \T_s \F] [\He^{\rm mfs}_{t,q}] [\Fd \T_p \F]
+ &=& [\Fd \T_s \F] \left\{  \sum_{\nu} \wntq [\Fd\Snd\Wimn\Sna\F] \right\} [\Fd \T_p \F] \\
+ &=& [\Fd \T_s \F] \left\{  \sum_{\nu} \wntq [\He_{\nu}\} \right\} [\Fd \T_p \F] 
+\end{eqnarray}
+
+$[\Wim]$ is a diagonal matrix of data-weights (and imaging-weights) and  
+$[\W^{\rm mfs}_t]$ is a diagonal matrix containing Taylor-weights $\wnt$.
+$[\He_{\nu}] =  [\Fd\Snd\Wimn\Sna\F]$ is the Hessian matrix formed using only one
+ frequency channel, and
+is a convolution operator containing a shifted version of the single-frequency 
+point-spread-function $\I^{\rm psf}_{\nu} = diag[\Fd \Sd \Wimn \Sa]$ in each row 
+(see appendix \ref{App:normal} for details).
+$[\Fd \T_s\F]$ and $[\Fd\T_p\F]$ are also convolution operators 
+with $\I^{\rm shp}_s$ and $\I^{\rm shp}_p$ as their kernels. 
+The process of convolution is associative and commutative, and 
+therefore, $\left[\He_{{s,p}\atop{t,q}}\right]$  is also a convolution operator 
+whose kernel is given by 
+\begin{equation}
+\label{Eq:msmfs_neqn_2.5}
+\vec{I}^{\rm psf}_{{s,p}\atop{t,q}} = \I^{\rm shp}_s  \star \left\{ \sum_{\nu} \wntq \vec{I}^{\rm psf}_{\nu} \right\} \star \I^{\rm shp}_p 
+%&=& \I^{\rm shp}_s \star I^{\rm psf}_{t,q} \star \I^{\rm shp}_p
+\end{equation}
+
+\noindent The dirty images on the RHS of Eqn.\ref{Eq:msmfs_neqn_1} can be written as follows.
+\begin{eqnarray}
+\label{Eq:msmfs_neqn_3}
+\vec{I}^{\rm dirty}_{{s}\atop{t}}  &=&  \left[\A_{{s}\atop{t}}^{\dag}\right][\Wim]  \vec{V}^{\rm obs}\\
+&=& [\Fd \T_s \Sd\Wntd\Wim] \vec{V}^{\rm obs} \\
+&=& [\Fd \T_s \F][\Fd\Sd\Wntd\Wim] \vec{V}^{\rm obs} \\
+&=& [\Fd \T_s \F] \left\{  \sum_{\nu} \wnt [\Fd\Snd\Wimn] \vec{V}_{\nu}^{\rm obs}  \right\} \\
+&=& \I^{\rm shp}_s \star \left\{\sum_{\nu} \wnt \vec{I}^{\rm dirty}_{\nu}  \right\} 
+%&=& [\Fd \T_s \F] \vec{I}^{\rm dirty}_t = \I^{\rm shp}_s \star \vec{I}^{\rm dirty}_t\\
+\label{Eq:msmfs_neqn_3a}
+%\vec{I}^{\rm dirty}_{t}  &=& [\Fd\Sd\Wntd\Wim] \vec{V}^{corr}\\
+% &=& \sum_{\nu} \wnt [\Fd\Snd\Wimn] \vec{V}_{\nu}^{corr} = \sum_{\nu} \wnt \vec{I}^{\rm dirty}_{\nu}\\
+%\label{Eq:specPSF}
+%\vec{I}^{\rm psf}_{t,q}&=& \sum_{\nu} \wntq [\Fd\Snd\Wimn] \vec{1} = \sum_{\nu} \wntq \vec{I}^{\rm psf}_{\nu}
+\end{eqnarray}
+where ${\I}_{\nu}^{\rm dirty} = [\Fd \Snd \Wimn] \vec{V}^{\rm obs}_{\nu} $ is the dirty
+image formed by direct Fourier inversion of weighted visibilities from one frequency channel.
+
+When all scales and Taylor terms are combined, 
+the full Hessian matrix contains
+$\Nt  \Ns  \times \Nt  \Ns $ blocks each of size $m\times m$, 
+and $\Nt $ Taylor coefficient images each of size $m\times 1$, 
+for all $\Ns $ spatial scales.
+
+The normal equations in block matrix form for the example 
+in Eqn.\ref{meqn_msmfs_math} for $\Nt =3, \Ns =2$ is shown
+in Eqn.\ref{Eq:msmfs_neqn_matrix}.
+The Hessian matrix consists of $\Ns \times \Ns =2\times 2$ blocks 
+(the four quandrants of the matrix), each for one pair of spatial scale $s,p$ (the upper indices).
+Within each quadrant, the $\Nt \times \Nt  = 3\times 3$
+matrices correspond to various pairs of $t,q$ (Taylor coefficient indices; the lower indices).
+This layout shows how the multi-scale and multi-frequency aspects of
+this imaging problem are combined and illustrates the 
+dependencies between the spatial and spectral basis functions.
+
+\begin{equation}\small
+\left[\begin{array}{llllll} 
+\noalign{\medskip}
+   \left[\He_{{ 0, 0}\atop{ 0, 0}}\right] & \left[\He_{{ 0, 0}\atop{ 0, 1}}\right] & \left[\He_{{ 0, 0}\atop{ 0, 2}}\right] & \left[\He_{{ 0, 1}\atop{ 0, 0}}\right] & \left[\He_{{ 0, 1}\atop{ 0, 1}}\right] & \left[\He_{{ 0, 1}\atop{ 0, 2}}\right] \\
+\noalign{\medskip}
+   \left[\He_{{ 0, 0}\atop{ 1, 0}} \right] & \left[\He_{{ 0, 0}\atop{ 1, 1}}\right] & \left[\He_{{ 0, 0}\atop{ 1, 2}}\right] & \left[\He_{{ 0, 1}\atop{ 1, 0}}\right] & \left[\He_{{ 0, 1}\atop{ 1, 1}}\right] & \left[\He_{{ 0, 1}\atop{ 1, 2}}\right] \\
+\noalign{\medskip}
+   \left[\He_{{ 0, 0}\atop{ 2, 0}} \right] & \left[\He_{{ 0, 0}\atop{ 2, 1}}\right] & \left[\He_{{ 0, 0}\atop{ 2, 2}}\right] & \left[\He_{{ 0, 1}\atop{ 2, 0}}\right] & \left[\He_{{ 0, 1}\atop{ 2, 1}}\right] & \left[\He_{{ 0, 1}\atop{ 2, 2}}\right] \\
+\noalign{\medskip}
+%   \hline
+\noalign{\medskip}
+   \left[\He_{{ 1, 0}\atop{ 0, 0}} \right] & \left[\He_{{ 1, 0}\atop{ 0, 1}}\right] & \left[\He_{{ 1, 0}\atop{ 0, 2}}\right] & \left[\He_{{ 1, 1}\atop{ 0, 0}}\right] & \left[\He_{{ 1, 1}\atop{ 0, 1}}\right] & \left[\He_{{ 1, 1}\atop{ 0, 2}}\right] \\
+\noalign{\medskip}
+   \left[\He_{{ 1, 0}\atop{ 1, 0}} \right] & \left[\He_{{ 1, 0}\atop{ 1, 1}}\right] & \left[\He_{{ 1, 0}\atop{ 1, 2}}\right] & \left[\He_{{ 1, 1}\atop{ 1, 0}}\right] & \left[\He_{{ 1, 1}\atop{ 1, 1}}\right] & \left[\He_{{ 1, 1}\atop{ 1, 2}}\right] \\
+\noalign{\medskip}
+   \left[\He_{{ 1, 0}\atop{ 2, 0}} \right] & \left[\He_{{ 1, 0}\atop{ 2, 1}}\right] & \left[\He_{{ 1, 0}\atop{ 2, 2}}\right] & \left[\He_{{ 1, 1}\atop{ 2, 0}}\right] & \left[\He_{{ 1, 1}\atop{ 2, 1}}\right] & \left[\He_{{ 1, 1}\atop{ 2, 2}}\right] \\
+\noalign{\medskip}
+   \end{array} \right]
+\left[\begin{array}{l} 
+\noalign{\medskip}
+		       \vec{I}^{\rm sky}_{{ 0}\atop{ 0}} \\ 
+\noalign{\medskip}
+                       \vec{I}^{\rm sky}_{{ 0}\atop{ 1}}\\ 
+\noalign{\medskip}
+		       \vec{I}^{\rm sky}_{{ 0}\atop{ 2}}\\ 
+\noalign{\medskip}
+%\hline
+\noalign{\medskip}
+		       \vec{I}^{\rm sky}_{{ 1}\atop{ 0}} \\ 
+\noalign{\medskip}
+		       \vec{I}^{\rm sky}_{{ 1}\atop{ 1}}\\ 
+\noalign{\medskip}
+		       \vec{I}^{\rm sky}_{{ 1}\atop{ 2}}\\
+\noalign{\medskip}
+			       \end{array}\right] =
+\left[\begin{array}{l} 
+\noalign{\medskip}
+                       \vec{I}^{\rm dirty}_{{ 0}\atop{ 0}}  \\ 
+\noalign{\medskip}
+                       \vec{I}^{\rm dirty}_{{ 0}\atop{ 1}} \\ 
+\noalign{\medskip}
+		       \vec{I}^{\rm dirty}_{{ 0}\atop{ 2}}\\ 
+\noalign{\medskip}
+%\hline
+\noalign{\medskip}
+		       \vec{I}^{\rm dirty}_{{ 1}\atop{ 0}} \\ 
+\noalign{\medskip}
+		       \vec{I}^{\rm dirty}_{{ 1}\atop{ 1}} \\ 
+\noalign{\medskip}
+		       \vec{I}^{\rm dirty}_{{ 1}\atop{ 2}} \\
+\noalign{\medskip}
+			       \end{array}\right] 
+\label{Eq:msmfs_neqn_matrix}
+\end{equation}
+
+This is the system of equations to be solved.
+% to obtain estimates of the model parameters $I^{\rm sky}_{p\atop q}$.
+The spatial-frequency sampling  of a real interferometer is always incomplete 
+($[\Sa]$ is rank-deficient).
+Therefore, each Hessian block, and the entire Hessian matrix is singular, and an exact inverse
+does not exist\footnote
+{Even if $[\He]$ were invertible, it is impractical to evaluate and invert the full Hessian
+(each row of each Hessian block represents an image).
+}. An accurate reconstruction can be obtained only via successive approximation
+(iterative numerical optimization).
+
+However, the Hessian is still useful since some of the diagonal terms can used to decouple the pixels in Taylor term space as well as possible.
+In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used.
+The greedy algorithm then locates the scale and location in the reference-frequency image (as convolved with the set of scales).
+For each minor cycle this produces a location, scale, and values for all Taylor terms.
+The residual images and model images are updated, and iteration proceeds.
+The actual algorithm is shown in Algorithm \ref{algo:CASA} (CASA form) and Algorithm \ref{algo:ASKAP} (ASKAP form).
+
+In summary, MSMFS decouples Taylor terms using the Hessian in Taylor-space evaluated at the peak of the PSF, and uses a greedy CLEAN-like algorithm to decouple in scale and sky position.
+The scales can be decoupled by an orthogonalisation process prior to deconvolution but the resulting images tend to have non-physical regions of negative brightness.
+
 \clearpage
@@
 
-The MS-MFS algorithm as implemented in CASA shown in Algorithm \ref{CASA}. The ASKAPsoft version is shown in Algorithm \ref{ASKAP} with the variant lines in red.
+The MS-MFS algorithm as implemented in CASA shown in Algorithm \ref{algo:CASA}. The ASKAPsoft version is shown in Algorithm \ref{algo:ASKAP} with the variant lines in red.
+
+We can relate these back to the graphical form Figure \ref{fig:majorminor} by introducing variant sub pipelines for Predict and MakeImage.
+These are graphically shown in Figures \ref{fig:predictuv} to \ref{fig:makeimageimage}:
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_Predict_UV.pdf}
+  \caption{Structure of UV-weighting Predict, as used in ASKAPsoft. }
+  \label{fig:predictuv}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_UV.pdf}
+  \caption{Structure of UV-weighting MakeImage, as used in ASKAPsoft. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
+  \label{fig:predictimage}
+\end{figure}
 
@@
   \caption[MS-MFS Algorithm]
-         {MS-MFS, as implemented in CASA}\label{CASA}
+         {MS-MFS, as implemented in CASA}\label{algo:CASA}
 
@@
   \caption[MS-MFS Algorithm]
-         {MS-MFS, as implemented in ASKAPsoft: {\color{red} red denotes difference from CASA}}\label{ASKAP}
+         {MS-MFS, as implemented in ASKAPsoft: {\color{red} red denotes difference from CASA}}\label{algo:ASKAP}
 
 \end{algorithm}
-
-We can relate these back to the graphical form Figure \ref{fig:majorminor} by introducing variant sub pipelines for Predict and MakeImage.
-These are graphically shown in Figures \ref{fig:predictuv} to \ref{fig:makeimageimage}:
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\textwidth]{./MSMFS_Predict_UV.pdf}
-  \caption{Structure of UV-weighting Predict, as used in ASKAPsoft. }
-  \label{fig:predictuv}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\textwidth]{./MSMFS_MakeImage_UV.pdf}
-  \caption{Structure of UV-weighting MakeImage, as used in ASKAPsoft. Note that the three loops (denoted by the rectangle with curved ends) can be permutated at will.}
-  \label{fig:predictimage}
-\end{figure}
 
@@
 \item Construct the facets without padding, send all facets to one compute island, perform deconvolution, distribute models to facets. 
+\item Each CI holds a distinct facet, which do not overlap.
+Synchronise after peak finding and residual updates.
+After a peak is found, all islands receive the global peak and update accordingly.
+Since the limit of coupling is assumed to be the CLEAN patch size, all peaks sufficiently far apart can be processed simultaneously.
 \end{itemize}
@@
 \end{description}
+
+The third approach to deconvolution requires low latency communications. It should be investigated.
 
@@
 There are two forms present: one for point sources, and one for the same blobs used in CASA - a truncated upside down parabola.
-The basis functions may optionally be orthogonalised using the Gramm-Schmidt algorithm.
+The basis functions may optionally be orthogonalised using the Gramm-Schmidt algorithm though there is little relevant experience.
 
@@
 
-Using typical values of $N_t=5$ and $N_s=5$, we find that the memory use is about $35 N_{pix}$.
-For a 200,000 by 200,000 pixel image, this requires 5.6 TB total.
-Spread over O(100) Compute Islands (partitioning by facet and sub-band), this requires 56 GB per CI.
-
-\clearpage
-\section{A projection}
-\label{sec:minor}
-
-\clearpage
-\section{Wide-band behaviour}
-\label{sec:wideband}
-
-\begin{itemize}
-\item Does naive broadband work?
-\item Bhatnagar et al WB algorithm	
-\end{itemize}
+Using L2 assumptions values of $N_t=5$ and $N_s=10$, we find that the memory use is about $65 N_{pix}$.
+For a 200,000 by 200,000 pixel image, this requires 10.4 TB total.
+Spread over O(100) Compute Islands (partitioning by facet and sub-band), this requires 104 GB per CI.
+The overhead for performing the MSMFS minor cycle is a factor of fifty, and for the major cycle a factor of 5 (if ASKAP) in gridding, and a factor of O(500) in FFT (if CASA).
+
+%
+%\clearpage
+%\section{A projection}
+%\label{sec:minor}
+%
+%\clearpage
+%\section{Wide-band behaviour}
+%\label{sec:wideband}
+%
+%\begin{itemize}
+%\item Does naive broadband work?
+%\item Bhatnagar et al WB algorithm	
+%\end{itemize}
+%
 
@@
 
+Parallel versions of both variants exist.
+These work by distributing in time (CASA) and frequency (ASKAPsoft).
+Parallelisation of the Minor Cycle is not straightforward and consequently both CASA and ASKAPsoft use single-threaded inner loops.
+This is unlikely to be successful for SKA, and instead the Minor Cycle has to be parallelised inside one CI and perhaps additionally distributed over multiple CI's.
+
+In light of this memo, the following recommendations are made:
+
 \begin{enumerate}
-\item Implement both Image and Visibility based calculation of the Taylor term images in the Performance Model.
-\item Update the Performance Model to reflect the memory usage:
+\item Perform simulations to explore the need for MSMFS in SKA.
+Currently an L1 requirement directs SDP to implement MSMFS.
+However there is no justification given for this very computationally demanding processing.
+Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable.
+Adding to the desire to see some more justification of the requirement, we note that the L1 requirements do not address the number of scales required, the number of Taylor terms, or the possibility of sub-banding (all of which appear in the SDP L2 requirements.
+We note three possible outcomes of the simulations: MFS can be ignored, MFS is required and MSMFS is sufficient, or MFS is needed and MFMFS is not sufficient.
+In the case of the latter outcome, investigation of other deconvolution techniques (e.g.
+[RD37]) would be required.
+\item Implement both Image (CASA) and Visibility based (ASKAPsoft) calculation of the Taylor term images in the Performance Model.
+The only reasons to choose one over the other are to do with the relative speed of gridding and FFTs.
+When it comes to implementation during construction, the costs of implementing both approaches should be minimal, and unimportant compared to the risk of implementing the wrong choice.
+\item Update the SDP Performance Model to reflect the MSMFS memory usage:
 \begin{equation}
@@
 \end{equation}
-\item A multi-threaded Minor Cycle on one Compute Island should be prototyped.
+\item Update the SDP Performance Model to reflect the MSMFS Flops:
+\begin{equation}
+Rflop_{MSMFSminor} = N_{minor} N_{pix}^2 N_s N_T	
+\end{equation}
+\item Prototype a multi-threaded Minor Cycle on one Compute Island. GPUMultiscaleCLEAN might be a good starting point. Do MS and then MSMFS. This can be done without implementing a major cycle.
+\item Model and then, if promising, prototype a distributed Minor Cycle on multiple CI's where synchronisation occurs every iteration.
+Do MS and then MSMFS.
+This can be done without implementing a major cycle.
+\item The SKAO Science Team should be consulted about the possibility of seams in the images or spectral cubes.
+These may occur both spatially (if the facets are deconvolved separately, or in frequency, because of the disjoint use of MSMFS in the different sub-bands.
 \end{enumerate}

### 2016-04-27 msmfs.tex
Version 0.3, ready for review by Steve and Anna.

@@
 \usepackage{txfonts}
+\usepackage{enumitem}
 
@@
 % Revision
-\newcommand{\revision}{2\xspace}
+\newcommand{\revision}{3\xspace}
 % Author(s)
@@
 % the document
-\newcommand{\firstsigname}{Name 1 - to be filled in}
-\newcommand{\firstsigdesignation}{Designation 1 - to be filled in}
-\newcommand{\firstsigaffiliation}{Affiliation 1 - to be filled in}
-
-\newcommand{\secondsigname}{Name 2 - to be filled in}
-\newcommand{\secondsigdesignation}{Designation 2 - to be filled in}
-\newcommand{\secondsigaffiliation}{Affiliation 2 - to be filled in}
+\newcommand{\firstsigname}{T.J. Cornwell}
+\newcommand{\firstsigdesignation}{Author}
+\newcommand{\firstsigaffiliation}{TCC}
+
+\newcommand{\secondsigname}{T.J. Cornwell}
+\newcommand{\secondsigdesignation}{Lead Consultant}
+\newcommand{\secondsigaffiliation}{TCC}
 
@@
     \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
+    \hline
+    0.3 & 27/04/2016 & T.J. Cornwell & Ready for review by Anna and Steve \\
     \hline
@@
       \hline
-      Name & Science Data Processor Consortium\\
+      Name & Tim Cornwell Consulting\\
       \hline
@@
 
+Parallel versions of both variants of MSMFS exist.
+These work by distributing in time (CASA) and frequency (ASKAPsoft).
+Parallelisation of the Minor Cycle is not straightforward and consequently both CASA and ASKAPsoft currently use single-threaded inner loops.
+This is unlikely to be successful for SKA, and instead the Minor Cycle has to be parallelised inside one CI and perhaps additionally distributed over multiple CI's.
+
+In light of this memo, the following recommendations are made concerning the need for MFS:
+
+\begin{enumerate}
+\item Perform simulations to explore the need for MSMFS in SKA.
+Currently an L1 requirement directs SDP to implement MSMFS.
+However there is no justification given for this very computationally demanding processing.
+Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable, although arguing from the work of Urvashi [RD??] does indicate that with MFS, the dynamic range would be limited to $10^4$ for wide frequency span.
+Either CASA or ASKAPsoft could be used for the reconstructions.
+\item Adding to the desire to see some more justification of the requirement, we note that the L1 requirements do not address the number of scales required, the number of Taylor terms, or the possibility of sub-banding (all of which appear in the SDP L2 requirements.
+We note three possible outcomes of the simulations: MFS can be ignored, MFS is required and MSMFS is sufficient, or MFS is needed and MFMFS is not sufficient.
+In the case of the latter outcome, investigation of other deconvolution techniques (e.g.
+[RD14]) would be required.
+\item Assess the importance of the considerations in Section \ref{sec:other}.
+\end{enumerate}
+
+The following recommendations address the resource usage:
+\begin{enumerate}[resume]
+\item Implement both Image (CASA) and Visibility based (ASKAPsoft) calculation of the Taylor term images in the SDP Performance Model.
+The main reasons to choose one over the other are to do with the relative speed of gridding and FFTs.
+When it comes to implementation during construction, the costs of implementing both approaches should be minimal, and unimportant compared to the risk of implementing the wrong choice.
+\item Update the SDP Performance Model to reflect the MSMFS memory usage:
+\begin{equation}
+M_{MSMFS} = (N_T (N_s + 1) + N_s) N_{pix}^2 
+\end{equation}
+\item Update the SDP Performance Model to reflect the MSMFS Flops:
+\begin{equation}
+Rflop_{MSMFSminor} = N_{minor} N_{pix}^2 N_s N_T	
+\end{equation}
+\item Prototype a multi-threaded Minor Cycle on one Compute Island. GPUMultiscaleCLEAN might be a good starting point. Do MS and then MSMFS. This can be done without implementing a major cycle.
+\item Model and then, if promising, prototype a distributed Minor Cycle on multiple CI's where synchronisation occurs every iteration.
+Do MS and then MSMFS.
+This can be done without implementing a major cycle.\end{enumerate}
+
+The following recommendations address wider issues:
+\begin{enumerate}[resume]
+\item The SKAO Science Team should be consulted about the impact of seams in the images or spectral cubes.
+These may occur both spatially (if the facets are deconvolved separately, or in frequency, because of the disjoint use of MSMFS in the different sub-bands.
+It seems questionable that these can be tolerated.
+
+\item We should note that MSMFS was developed in a serial platform where computation was expensive and memory access cheap.
+In the SDP context, the processing is distributed and parallel, and computation is cheap and memory access expensive.
+Although SDP has no remit to develop new algorithms, it could encourage researchers to address this new context.
+\end{enumerate}
+
 \newpage
@@
      \hline
-   {\bf [RD01]} & SKA-TEL-SDP-0000041, F. Malan: iPython Performance Model Description\\
-   {\bf [RD02]} & SKA-TEL-SDP-0000027, R. Nijboer: Pipelines Element Subsystem Design\\
-   {\bf [RD03]} & SKA-TEL-SDP-0000038, R. Bolton: High Priority Science Objectives: Performance Analysis\\
-   {\bf [RD04]} & SKA-TEL-SDP-0000028, A.-J. Boonstra: Ingest Pipeline\\
-   {\bf [RD05]} & SKA-TEL-SDP-0000029, S. Salvini: Pipelines: Calibration\\
-   {\bf [RD06]} & SKA-TEL-SDP-0000030, A. Scaife: Imaging Pipeline\\
-   {\bf [RD07]} & SKA-TEL-SDP-0000031, M. Johnston-Hollitt: Science Analysis Pipeline\\
-   {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent Averaging\\
-   {\bf [RD09]} & SKA-TEL-SDP-0000058, S. Salvini: Fast Fourier Transforms\\
-   {\bf [RD10]} & SKA-TEL-SDP-0000057, C. Skipper: Time and Channel Averaging\\
-   {\bf [RD13]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
-   {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
-   {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
-   {\bf [RD18]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
-   {\bf [RD22]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
-   {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
-   {\bf [RD34]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
-   {\bf [RD35]} & Conway, J.
+   {\bf [RD01]} & SKA-TEL-SDP-0000027, R. Nijboer: Pipelines Element Subsystem Design\\
+   {\bf [RD02]} & SKA-TEL-SDP-0000038, R. Bolton: High Priority Science Objectives: Performance Analysis\\
+   {\bf [RD03]} & SKA-TEL-SDP-0000029, S. Salvini: Pipelines: Calibration\\
+   {\bf [RD04]} & SKA-TEL-SDP-0000030, A. Scaife: Imaging Pipeline\\
+   {\bf [RD05]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
+   {\bf [RD06]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
+   {\bf [RD07]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
+   {\bf [RD08]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
+   {\bf [RD09]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
+   {\bf [RD10]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
+   {\bf [RD11]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
+   {\bf [RD12]} & Conway, J.
 E., Cornwell, T.
@@
 Monthly Notices of the Royal Astronomical Society, 246, 490.\\
-   {\bf [RD36]} & http://www.aanda.org/component/citedby/?task=crossref\&doi=10.1051/0004-6361/201117104 \\  
-   {\bf [RD37]} & Junklewitz, H., Bell, M.
+   {\bf [RD13]} & http://www.aanda.org/component/citedby/?task=crossref\&doi=10.1051/0004-6361/201117104 \\  
+   {\bf [RD14]} & Junklewitz, H., Bell, M.
 R., and Ensslin, T.
@@
 http://doi.org/10.1051/0004-6361/201423465 \\
+   {\bf [RD15]} & Cornwell, T.J. (2016) Extension of the SDP performance model for cross data-island connectivity, TCC-SDP-151123-1-1 \\
+	{\bf [RD16]} & Rau, U., \& Cornwell, T.
+J.
+(2011, June 14).
+A multi-scale multi-frequency deconvolution algorithm for synthesis imaging in radio interferometry.
+arXiv.org.
+EDP Sciences.
+http://doi.org/10.1051/0004-6361/201117104   \\
+	{\bf [RD17]} & Bhatnagar, S., Rau, U., \& Golap, K.
+(2013).
+Wide-field wide-band Interferometric Imaging: The WB A-Projection and Hybrid Algorithms.
+The Astrophysical Journal, 770(2), 91.
+http://doi.org/10.1088/0004-637X/770/2/91 \\
 	\hline
@@
 
-% The actual content goes here
-\newpage
-\section{Introduction}
-
 \clearpage
@@
 
-\clearpage
 \section{Scope of the document}
 
-\clearpage
+The scope includes only the published Multi-Scale, Multi-Frequency algorithm, existing implementations, and inclusion in the SDP processing pipelines. No other MFS algorithms are explored.
+
 \section{Background}
@@
 The imaging performance of a radio synthesis telescope is related to the uv coverage.
-The uv coverage is dependent on the observing wavelength so diversity in uv can be obtained by observing over a range of frequencies [RD35].
+The uv coverage is dependent on the observing wavelength so diversity in uv can be obtained by observing over a range of frequencies [RD12].
 This Multi-Frequency Synthesis (MFS) only works for radio sources in which broadband emission is present, giving a way to relate measurements at different frequencies.
 There have been multiple attempts at an MFS algorithm (e.g.
-[RD35]).
+[RD12]).
 MFS is usually most important for sources having extended emission.
@@
 
-The Multi-Scale, Multi-Frequency Synthesis (MS-MFS) algorithm was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g.
-[RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data.
+The Multi-Scale, Multi-Frequency Synthesis (MS-MFS) algorithm was developed by Urvashi [RD07], melding together the concepts from multi-scale clean [RD11] and multi-frequency synthesis e.g.
+[RD12], with the goal of improving the reconstruction of sky brightness from radio interferometric data.
 The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law expanded into a Taylor series.
-It has been used in JVLA, ALMA, and ATCA observations [RD36].
-
-The algorithm is an elaboration of the Multiscale CLEAN. It consists of two parts:
+It has been used in JVLA, ALMA, and ATCA observations [RD13].
+
+The algorithm is an elaboration of the Multiscale CLEAN [RD11]. It consists of two parts:
 \begin{description}
@@
 
-There is an L1 requirement, $SKA1-SYS\_REQ-2324$ specifying that MSMFS must be used (see Table\ref{tab:L1}).
+There is an L1 requirement [AD03], $SKA1-SYS\_REQ-2324$ specifying that MSMFS must be used for continuum imaging (see Table\ref{tab:L1}).
 
@@
 All imaging shall construct and make use of frequency dependent image models over the entire observed bandwidth.
-& Accepted & Reference [2] & Demonstration \\
+& Accepted & Reference & Demonstration \\
 \hline 
@@
 
-This is reflected in the L2 requirements shown in Table\ref{tab:L2}.
+This is reflected in the L2 requirements shown in Table \ref{tab:L2}.
 \begin{table}[htp]
+\begin{center}
 \begin{tabular}{|L|L|L|L|L|}
@@
 \caption{Relevant L2 requirements}	\label{tab:L2}
+\end{center}
 \end{table}
@@
 
-Finally we could ask is there a superior algorithm already published? (see e.g. [RD37]).
+Finally we could ask is there a superior algorithm already published?
+(see e.g.
+[RD14]).
+The degree of tuning required for MSMFS (number of scales and number of Taylor terms) is acceptable in a hands-on data processing scenario but may be not acceptable for a general purpose pipeline.
 
@@
 
+Urvashi has discussed the use of MSMFS for the JVLA (section 8.4.2.1 of [RD??]. With regard to just ignoring the effect of spectral index variations, she says:
+
+\begin{quotation}
+
+Dynamic-range limits when source spectra are ignored : If continuum imaging is done with only MFS gridding and source spectra are ignored, spectral structure will masquerade as spurious spatial structure.
+These errors will affect regions of the image both on-source and off-source and their magnitudes depend on the available uv-coverage, the frequency range being covered, the choice of reference frequency, and the intensity and spectral index of the source.
+A rough rule of thumb for an EVLA-type uv-coverages (see section 6.2.4.2) is that for a point source of with spectral index α = −1.0 measured between 1 and 2 GHz, the peak error obtained if the spectrum is ignored is at a dynamic range of $ < 10^3$.
+Note that when all sources in the observed region of the sky have similar spectral indices, these errors can be reduced by dividing out an average spectral index (one single number over the entire sky) from the visibilities before imaging the\end{quotation}
+
+If we suppose that the SKA PSF is ten times better than that of the JVLA (an over estimate since the RMS sidelobe level goes only as the inverse of the number of antennas), the dynamic range limitation for SKA1-LOW and SKA1-MID still appears at about $10^4$, which is clearly insufficient.
+Furthermore, the primary beam will vary across the frequency span, and this will lead to spurious structure if spectral effects are ignored (cf Urvashi).
+The introduction of sub-bands does ameliorate these problems to some degree but then the continuum structure model is only defined in pieces that are not necessarily consistent at the edges.
+
 There is no mechanism to address these questions beyond simulations. Arguing from uv coverage metrics such as the UVGAP only helps in a relative sense.
@@
 
-In the Multi-Scale CLEAN [RD??], the sky is modelled as a set of blobs.
+In the Multi-Scale CLEAN [RD11], the sky is modelled as a set of blobs.
 Performing this decomposition using a gradient search algorithm is very demanding of resources.
@@
 
-The MSMFS algorithm decomposes the image into a set of $N_s$ blobs each with $N_t$ Taylor terms.
+The MSMFS [RD07] algorithm decomposes the image into a set of $N_s$ blobs each with $N_t$ Taylor terms. The following description is drawn from [RD07]
 
@@
 
-%\begin{figure}[htb]
-%  \centering
-%  \includegraphics[width=\textwidth]{./algorithm.pdf}
-%  \caption{The MSMFS algorithm as implemented in CASA [RD??]}
-%  \label{fig:algoCASA}
-%\end{figure}
-
 \clearpage
@@
 \item[Optimum partitioning] 
-\item[Coupling across partitions] e.g. is the same pixel addressed in two different partitions? If so, is there a satisfactory approach to reconciling the two values?
+\item[Coupling across partitions] e.g.
+is the same pixel addressed in two different partitions?
+If so, is there a satisfactory approach to reconciling the two values?
+This is a generic question that should be considered for all distributed and parallel processing.
 \item[Synchronisation points] e.g. is a global synchronisation point required so that the deconvolutions can be made consistent? Is there an acceptable algorithm for reconciling separate facets?
 \item[Amount of CU memory]
-\item[Loading of CU memory backplane]
-\item[Access to visibility store]
+\item[Loading of CU memory backplane] The intrinsic operations in MSMFS are mostly simple SAXBY type operations. Hence the memory backplane will be overloaded, and the CPU data-starved.
+\item[Access to visibility store] What access to the visibility store is required?
+A straightforward implementation of the Major/Minor cycle algorithm loads the observed data every major cycle.
+However, as discussed in [RD15], this is actually only true of the initialisation and any self-calibration cycles.
+For other cycles, it is necessary only to load the uvw coordinates and flagging information.
 \end{description}
@@
 
-The SDP pipeline framework [RD:??], allows partitioning across multiple axes. In this case, the natural partitions and typical values are:
+The SDP pipeline framework [RD04], allows partitioning across multiple axes. In this case, the natural partitions and typical values are:
 
 \begin{description}
-\item[Polarisation] Only Stokes I can be modelled using MSMFS
-\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing
+\item[Polarisation] Only Stokes I can be modelled using MSMFS. Modelling the Faraday Rotation is possible but there is no existing algorithm.
+\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing/ There is no advantage to finer channels (see e.g. section 8.4 in [RD??]).
 \item[Sub-bands] O(1-3) partitions of the coarse channels are required to ensure that the power law approximation is adequate.
@@
 %\item[Snapshot time] O(100s) long gridding/degridding sequences are required to limit smearing due to w-plane rotation
-\item[Scales] O(5) scales are typically required to model extended structure
-\item[Taylor terms] O(5) terms are required to represent the spectral behaviour
+\item[Scales] O(5) scales are typically required to model extended structure (see section 8.4 in [RD??])
+\item[Taylor terms] O(5) terms are required to represent the spectral behaviour for moderately complex spectrum sources (see section 8.4 in [RD??])
 \end{description}
 
-According to the logic of the MSMFS algorithm, Sub-bands can be defined as the limit of the power law approximation for brightness so consequently can always be used as the coarsest partition.
+According to the logic of the MSMFS algorithm, sub-bands can be defined as the limit of the power law approximation for brightness so consequently can always be used as the coarsest partition.
 For Predict and MakeImage the remaining choices are as shown in Figure \ref{tab:partitions} (most rapid first):
@@
 
-The only choice is between the two orderings of Time and Facet. Since deconvolution is non-linear, we prefer that all times be represented in a single facet rather than the other way around.
-
+The choice is between the two orderings of Time and Facet.
+Since deconvolution is non-linear, we prefer that all times be represented in a single facet rather than the other way around.
 Hence according to this level of analysis, the best ordering is [Frequency, Time, Facet, Sub-Band].
@@
 \item Construct the facets without padding, send all facets to one compute island, perform deconvolution, distribute models to facets. 
-\item Each CI holds a distinct facet, which do not overlap.
-Synchronise after peak finding and residual updates.
-After a peak is found, all islands receive the global peak and update accordingly.
+\item Each CI holds a distinct facet, which do not need to overlap.
+Synchronise in each minor cycle after peak finding.
+After a peak is found, all islands receive all peaks and choose the same global peak and update accordingly.
 Since the limit of coupling is assumed to be the CLEAN patch size, all peaks sufficiently far apart can be processed simultaneously.
@@
 \begin{description}
-\item[Predict and MakeImage] Requires the visibility data but not many scratch images. Stalled during MinorCycle.
-\item[MinorCycle] Requires a substantial number of scratch images, but not the visibility data. Stalled during Predict and MakeImage.	
+\item[Predict and MakeImage] Requires the visibility data but not many scratch images. Stalls during MinorCycle.
+\item[MinorCycle] Requires a substantial number of scratch images, but not the visibility data. Stalls during Predict and MakeImage.	
 \end{description}
@@
 
-We have a modelling problem that is not going to be wished away.
+However, in the minor cycle, we have a modelling problem that is not going to be wished away.
 The science demands seamless imaging in multiple axes, whereas computing would like to partition the image cube in those same axes.
@@
 The CASA tclean program supports parallel execution using python to run copies on different nodes.
+This mode of parallel execution fits the CASA execution model.
 The visibility data are partitioned in time across the available processors.
@@
 The ASKAPsoft code is mirrored in the SDP Github repository.
-The ASKAPsoft cimager is an MPI-enabled program.
-Since it was written as parallel program from the beginning, the structure is easier to follow than the CASA version.
-Distribution over MPI is via frequency, rather than time as in CASA.
+The ASKAPsoft cimager is MPI-enabled, written as parallel program from the beginning/ Distribution over MPI is via frequency, rather than time as in CASA.
 Predict and MakeImage are distributed using MPI (via a thin bookkeeping layer).
@@
 The second criterion is expensive to compute and usually MAXBASE0 is used.
-Thus for each scale and for each pixel the pixel value at the reference frequency has to be determined by pre-multipling by the inverse Hessian (in term-term space).
+Thus for each scale and for each pixel the pixel value at the reference frequency has to be determined by pre-multiplying by the inverse Hessian (in term-term space).
 This means that the cost of the search is:
@@
 \hline
-Basis function FFT & scales & $N_{pix} N_S $ & During initialisation \\
-PSF FFT & 2 * Taylor terms & $2 N_{pix} N_T $ & During initialisation \\
-Spectral Dirty image & Taylor terms, scales & $N_{pix} N_T N_s$ & During iteration \\
-Spectral Dirty PSF & Taylor terms, scales & $N_{patch} N_T N_s$ & During iteration \\
-Spectral Model image & Taylor terms & $N_{pix} N_T $ & During iteration \\	
-Basis functions & scales & $N_{pix} N_s$ & During iteration \\
+Basis function FFT & scales & $N_{pix}^2 N_S $ & During initialisation \\
+PSF FFT & 2 * Taylor terms & $2 N_{pix}^2 N_T $ & During initialisation \\
+Spectral Dirty image & Taylor terms, scales & $N_{pix}^2 N_T N_s$ & During iteration \\
+Spectral Dirty PSF & Taylor terms, scales & $N_{patch}^2 N_T N_s$ & During iteration \\
+Spectral Model image & Taylor terms & $N_{pix}^2 N_T $ & During iteration \\	
+Basis functions & scales & $N_{pix}^2 N_s$ & During iteration \\
 Inverse of Hessian & Taylor terms, Taylor terms, scales & $N^2_T N_s$ & During Iteration \\
@@
 
-where $N_{pix}$ is the number of pixels in an image, $N_{patch}$ is the number of pixels in the patch of the Dirty beam used, $N_T$ is the number of Taylor terms, and $N_S$ is the number of scales/basis functions.
+where $N_{pix}$ is the number of pixels on an axis of the image, $N_{patch}$ is the number of pixels in the patch of the Dirty beam used, $N_T$ is the number of Taylor terms, and $N_S$ is the number of scales/basis functions.
 
@@
 \begin{equation}
-M_{MSMFS} = (N_T (N_s + 1) + N_s) N_{pix} 
+M_{MSMFS} = (N_T (N_s + 1) + N_s) N_{pix}^2 
 \end{equation}
@@
 \clearpage
+\section{Other considerations}
+\label{sec:other}
+
+Ignoring the behaviour of the MSMFS algorithm, there are other considerations arising from the use of any MFS technique (this discussion drawn from section 8.4 of [RD07]).
+
+\begin{description}
+\item[Calibration]	
+The use of MSMFS or any technique taking advantage of simple models of broadband behaviour requires that the calibration across the observing band be well understood.
+Otherwise errors in bandpass calibration will couple into derived spatial structure.
+Hence there is a derived system requirement on bandpass accuracy, even just for continuum imaging.
+This should be evaluated via simulations and analysis.
+\item[Anomalous spectral index objects] such as some pulsars, GHz-peaked sources, very extended radio galaxies, may limit dynamic range.
+\item[Primary beam model] If the true spectrum is to be recovered, the primary beam must be modelled accurately.
+\item[UV coverage effects] The fine structure is constrained by the high frequency measurements and the extended emission is constrained by the low frequency measurements.
+In principle, these biases can be avoided by limiting the range of uv coordinates to a common set at all frequencies, but this does somewhat defeat the purpose of MFS.
+It is not clear how serious this effect will be for SKA.
+\end{description}
+
+In addition, Bhatnagar et al.
+[RD18] have drawn attention to the difficulties of performing primary beam correction over a wide bandwidth.
+They note some difficulties with the straightforward application of what they call the Narrow Band MSMFS algorithm (which we call simply MSMFS) to large bandwidths.
+Specifically the apparent Noise to Signal Ratio becomes very large outside the primary beam.
+They argue that analysis of the images scientifically relies upon understanding the noise level outside the primary beam sampled area.
+This is undoubtedly correct, and MIRIAD and ASKAPsoft address this by designating a primary beam cutoff beyond which the noise alone is used without normalisation.
+In their approach, Bhatnagar et al.
+advocate correcting the primary beam to a desired effective primary beam.
+This is quite sensible.
+Some evaluation of this approach should be performed using simulations, and the scientific analysis implications should be considered.
+
+\clearpage
 \section{Recommendations}
@@
 
-Parallel versions of both variants exist.
+Parallel versions of both variants of MSMFS exist.
 These work by distributing in time (CASA) and frequency (ASKAPsoft).
-Parallelisation of the Minor Cycle is not straightforward and consequently both CASA and ASKAPsoft use single-threaded inner loops.
+Parallelisation of the Minor Cycle is not straightforward and consequently both CASA and ASKAPsoft currently use single-threaded inner loops.
 This is unlikely to be successful for SKA, and instead the Minor Cycle has to be parallelised inside one CI and perhaps additionally distributed over multiple CI's.
 
-In light of this memo, the following recommendations are made:
+In light of this memo, the following recommendations are made concerning the need for MFS:
 
@@
 However there is no justification given for this very computationally demanding processing.
-Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable.
-Adding to the desire to see some more justification of the requirement, we note that the L1 requirements do not address the number of scales required, the number of Taylor terms, or the possibility of sub-banding (all of which appear in the SDP L2 requirements.
+Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable, although arguing from the work of Urvashi [RD??] does indicate that with MFS, the dynamic range would be limited to $10^4$ for wide frequency span.
+Either CASA or ASKAPsoft could be used for the reconstructions.
+\item Adding to the desire to see some more justification of the requirement, we note that the L1 requirements do not address the number of scales required, the number of Taylor terms, or the possibility of sub-banding (all of which appear in the SDP L2 requirements.
 We note three possible outcomes of the simulations: MFS can be ignored, MFS is required and MSMFS is sufficient, or MFS is needed and MFMFS is not sufficient.
 In the case of the latter outcome, investigation of other deconvolution techniques (e.g.
-[RD37]) would be required.
-\item Implement both Image (CASA) and Visibility based (ASKAPsoft) calculation of the Taylor term images in the Performance Model.
-The only reasons to choose one over the other are to do with the relative speed of gridding and FFTs.
+[RD14]) would be required.
+\item Assess the importance of the considerations in Section \ref{sec:other}.
+\end{enumerate}
+
+The following recommendations address the resource usage:
+\begin{enumerate}[resume]
+\item Implement both Image (CASA) and Visibility based (ASKAPsoft) calculation of the Taylor term images in the SDP Performance Model.
+The main reasons to choose one over the other are to do with the relative speed of gridding and FFTs.
 When it comes to implementation during construction, the costs of implementing both approaches should be minimal, and unimportant compared to the risk of implementing the wrong choice.
@@
 Do MS and then MSMFS.
-This can be done without implementing a major cycle.
-\item The SKAO Science Team should be consulted about the possibility of seams in the images or spectral cubes.
+This can be done without implementing a major cycle.\end{enumerate}
+
+The following recommendations address wider issues:
+\begin{enumerate}[resume]
+\item The SKAO Science Team should be consulted about the impact of seams in the images or spectral cubes.
 These may occur both spatially (if the facets are deconvolved separately, or in frequency, because of the disjoint use of MSMFS in the different sub-bands.
+It seems questionable that these can be tolerated.
+
+\item We should note that MSMFS was developed in a serial platform where computation was expensive and memory access cheap.
+In the SDP context, the processing is distributed and parallel, and computation is cheap and memory access expensive.
+Although SDP has no remit to develop new algorithms, it could encourage researchers to address this new context.
 \end{enumerate}

### 2016-05-12 msmfs.tex
Final version submitted
Minor changes per Anna.

@@
 % Revision
-\newcommand{\revision}{3\xspace}
+\newcommand{\revision}{1.0\xspace}
 % Author(s)
@@
 % Status of the document (draft/final/etc.)
-\newcommand{\docstatus}{Draft\xspace}
+\newcommand{\docstatus}{Released\xspace}
 
@@
     0.3 & 27/04/2016 & T.J. Cornwell & Ready for review by Anna and Steve \\
+    1.0 & 12/05/2016 & T.J. Cornwell & To be signed \\
     \hline
@@
 However there is no justification given for this very computationally demanding processing.
-Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable, although arguing from the work of Urvashi [RD??] does indicate that with MFS, the dynamic range would be limited to $10^4$ for wide frequency span.
+Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable, although arguing from the work of Urvashi [RD07] does indicate that with MFS, the dynamic range would be limited to $10^4$ for wide frequency span.
 Either CASA or ASKAPsoft could be used for the reconstructions.
@@
 
-Urvashi has discussed the use of MSMFS for the JVLA (section 8.4.2.1 of [RD??]. With regard to just ignoring the effect of spectral index variations, she says:
+Urvashi has discussed the use of MSMFS for the JVLA (section 8.4.2.1 of [RD07]. With regard to just ignoring the effect of spectral index variations, she says:
 
@@
 For each minor cycle this produces an amplitude, location, and scale.
-The smoothed residual images and model images are updated, and iteration proceeds.
+The smoothed residual images and model images are updated, and iteration proceeds\footnote{To exclude any errors in transcription, the LaTeX code for the equations in this section is drawn directly from [RD16] via arxiv.org}.
 
@@
 
-However, the Hessian is still useful since some of the diagonal terms can used to decouple the pixels in Taylor term space as well as possible.
+However, the Hessian is still useful since some of the terms can used to decouple the pixels in Taylor term space as well as possible.
 In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used.
@@
 
-In summary, MSMFS decouples Taylor terms using the Hessian in Taylor-space evaluated at the peak of the PSF, and uses a greedy CLEAN-like algorithm to decouple in scale and sky position.
+In summary, MSMFS decouples Taylor terms using the (non-diagonal) Hessian in Taylor-space evaluated at the peak of the PSF, and uses a greedy CLEAN-like algorithm to decouple in scale and sky position.
 The scales can be decoupled by an orthogonalisation process prior to deconvolution but the resulting images tend to have non-physical regions of negative brightness.
@@
 \item[Polarisation] Only Stokes I can be modelled using MSMFS. Modelling the Faraday Rotation is possible but there is no existing algorithm.
-\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing/ There is no advantage to finer channels (see e.g. section 8.4 in [RD??]).
+\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing/ There is no advantage to finer channels (see e.g. section 8.4 in [RD07]).
 \item[Sub-bands] O(1-3) partitions of the coarse channels are required to ensure that the power law approximation is adequate.
@@
 %\item[Snapshot time] O(100s) long gridding/degridding sequences are required to limit smearing due to w-plane rotation
-\item[Scales] O(5) scales are typically required to model extended structure (see section 8.4 in [RD??])
-\item[Taylor terms] O(5) terms are required to represent the spectral behaviour for moderately complex spectrum sources (see section 8.4 in [RD??])
+\item[Scales] O(5) scales are typically required to model extended structure (see section 8.4 in [RD07])
+\item[Taylor terms] O(5) terms are required to represent the spectral behaviour for moderately complex spectrum sources (see section 8.4 in [RD07])
 \end{description}

### 2018-05-10 mmm-minorcycle.tex
Draft with most elements filled in

%rubber: module pdflatex

\documentclass[11pt,a4paper,variablewidth]{article}

\usepackage[sans]{tcc_doc} % SDP style file
\usepackage[english]{babel}
\usepackage{amssymb,amsmath} % for the math symbols
\usepackage{mdwlist}
\usepackage{geometry}
\usepackage{marginnote}
\usepackage{datetime}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{mathtools}
\usepackage{hhtensor}
\usepackage{color}
\usepackage{txfonts}
\usepackage{enumitem}
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage{float}

%\usepackage{stylefiles/fmtcount}
\usepackage{natbib}

\graphicspath{ {images/} }

\usepackage[boxruled,linesnumbered]{algorithm2e}
\usepackage{fmtcount}

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}

% TODO:
% 1. Check if all new symbols have been explained when used for the first
% time.
% 2. Check if all symbols are actually being used.
% 3. Check the symbol definitions: rates, numbers, etc. to use the same main letter
% 4. Is there a way to use proper referencing of applicable and reference
% documents?

%%%%%%%%% START OF USER SETTINGS %%%%%%%%%%%%%%%%%%

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-CASS-001\xspace}
% Context
\newcommand{\context}{(ASKAPsoft)}
% Revision
\newcommand{\revision}{0.1\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Released\xspace}

% The definitions below are used for the signature tables in the preamble of
% the document
\newcommand{\firstsigname}{T.J. Cornwell}
\newcommand{\firstsigdesignation}{Author}
\newcommand{\firstsigaffiliation}{TCC}

\newcommand{\secondsigname}{T.J. Cornwell}
\newcommand{\secondsigdesignation}{Lead Consultant}
\newcommand{\secondsigaffiliation}{TCC}

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.1 & 03/04/2018 & T.J. Cornwell & Initial version \\
    \hline
  \end{tabularx}
}

%%%%%%%%%%%%% END OF USER SETTINGS %%%%%%%%%%%%%%%%%%%

% Table with signatures
\newcommand{\signaturetable}[3]{
  \begin{tabularx}{\textwidth}{|X|X|X|}
    \hline
    Name & Designation & Affilitation\\
    \hline
    {#1} & {#2} & {#3}\\
    \hline
    Signature \& Date: & & \\
    & & \\
    & & \\
    \hline
  \end{tabularx}
}

% Table with affiliations
\newcommand{\organisationtable}{
  \begin{center}
    \sffamily{\bf ORGANISATION DETAILS}\end{center}
  \begin{table}[htb]
    \centering
    \begin{tabular}[htb]{|l|l|}
      \hline
      Name & Tim Cornwell Consulting\\
      \hline
    \end{tabular}
  \end{table}
}

%%%%%%%%%%%%%%%%%%%% START YOUR DOCUMENT HERE %%%%%%%%%%%%%%%%%

\newcommand{\subsubsubsection}[1]{\noindent{\bf{#1}}}

%%%% Please DO USE the symbol definitions listed below to preserve
%%%% consistency.

%%%% If you need a new symbol, please add it here. If really want to modify
%%%% one, do so, just take care to do a search-and-replace for the entire document
% Symbol definitions

\begin{document}

% load automatic pages
\tccfrontpage

\sdptableofcontents

% Add here the abbreviations used in your document
\sdplistofabbreviations
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
    \item[ASKAP] Australian Square Kilometer Array Pathfinder\vspace{-0.2cm}
    \item[ARL] Algorithm Reference Library\vspace{-0.2cm}
    \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
    \item[FLOP] Floating Point Operations \vspace{-0.2cm}
    \item[FLOPS] Floating Point Operations per Second \vspace{-0.2cm}
    \item[FoV] Field of View\vspace{-0.2cm}
    \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[MAM] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[PSF] Point Spread Function \vspace{-0.2cm}
    \item[SDP] Scientific Data Processor\vspace{-0.2cm}
    \item[SKA] Square Kilometre Array\vspace{-0.2cm}
%    \item[] \vspace{-0.2cm}
\end{basedescript}

\newcommand{\nuno}{{\left(\frac{\nu}{\nu_0}\right)}}
\newcommand{\dnuno}{{\left(\frac{\nu-\nu_0}{\nu_0}\right)}}

\newcommand{\dg}{^\dag}
\newcommand{\X}{\vec{x}}
\newcommand{\Xd}{\vec{{x}^\dag}}
\newcommand{\B}{\vec{b}}
\newcommand{\Bd}{\vec{b^\dag}}
\newcommand{\V}{\vec{V}}
\newcommand{\Vd}{\vec{V^\dag}}
\newcommand{\A}{{\tens{A}{}}}
\newcommand{\Ad}{{\tens{A^\dag}{}}}
\newcommand{\F}{{\tens{F}{s}}}
\newcommand{\Fd}{{\tens{F^\dag}{}}}
\newcommand{\He}{{\tens{H}{}}}
\newcommand{\Sa}{{\tens{S}{}}}
\newcommand{\Sd}{{\tens{S^\dag}{}}}
\newcommand{\Sna}{\tens{{S_{\nu}}{}}}
\newcommand{\Snd}{\tens{{S_{\nu}^\dag}{}}}
\newcommand{\T}{{\tens{T}{}}}
\newcommand{\W}{{\tens{W}{}}}
\newcommand{\Wd}{{\tens{W^\dag}{}}}
\newcommand{\Pb}{{\vec{P}}}

%\newcommand{\Wim}{{\tens{W^{im}}}}
%\newcommand{\Wimd}{{\tens{{W^{im}}^\dag}}}
%\newcommand{\Wnt}{{\tens{W^{\rm {mfs}}_t}}}
%\newcommand{\Wntd}{{\tens{{W^{\rm {mfs}}_t}^\dag}}}
%\newcommand{\Wnp}{{\tens{W^{\rm mfs}_p}}}
%\newcommand{\Wnpd}{{\tens{{W^{\rm mfs}_p}^\dag}}}
%\newcommand{\Wnq}{{\tens{W^{\rm mfs}_q}}}
%\newcommand{\Wnqd}{{\tens{{W^{\rm {mfs}}_q}^\dag}}}
%\newcommand{\Wimn}{{\tens{W^{im}_{\nu}}}}
%\newcommand{\Wimnd}{{\tens{{W^{im}_{\nu}}^\dag}}}

\newcommand{\Wim}{{{\W^{\rm im}}}}
\newcommand{\Wimd}{{{{\W^{\rm im}}^\dag}}}
\newcommand{\Wnt}{{{\W^{\rm {mfs}}_t}}}
\newcommand{\Wntd}{{{{\W^{\rm {mfs}}_t}^\dag}}}
\newcommand{\Wnp}{{{\W^{\rm mfs}_p}}}
\newcommand{\Wnpd}{{{{\W^{\rm mfs}_p}^\dag}}}
\newcommand{\Wnq}{{{\W^{\rm mfs}_q}}}
\newcommand{\Wnqd}{{{{\W^{\rm {mfs}}_q}^\dag}}}
\newcommand{\Wimn}{{{\W^{\rm im}_{\nu}}}}
\newcommand{\Wimnd}{{{{\W^{\rm im}_{\nu}}^\dag}}}

\newcommand{\wnt}{{w_{\nu}^t}}
\newcommand{\wnq}{{w_{\nu}^q}}
\newcommand{\wntq}{{w_{\nu}^{t+q}}}
%\newcommand{\Wntn}{{\tens{w^{\rm mfs}_{t,\nu}}}}
%\newcommand{\Wntnd}{{\tens{{w^{\rm mfs}_{t,\nu}}^\dag}}}
%\newcommand{\Wnpn}{{\tens{W^{\rm mfs}_{p,\nu}}}}
%\newcommand{\Wnpnd}{{\tens{{W^{\rm mfs}_{p,\nu}}^\dag}}}

\newcommand{\pd}{{\partial}}
\newcommand{\mi}{{m_{I}}}
\newcommand{\R}{{R}}
\newcommand{\Rd}{{R^\dag}}
\newcommand{\I}{{\vec{I}}}

\newcommand{\Nt}{N_{\rm Taylor}}
\newcommand{\Ns}{N_{\rm scales}}
\newcommand{\Nc}{N_{\rm chan}}
\newcommand{\Np}{N_{\rm pix}}
\newcommand{\Nf}{N_{\rm Facet}}

\newcommand{\MAM}{M\&M}

\newcommand{\ARL}{ARL}

\sdplistoffigures

\sdplistoftables

% Add here the executive summary
\sdpsummary

The MSMFS algorithm is integral to the data processing for ASKAP and SKA. The inner loops for ASKAP continuum processing use the Multi-Scale Multi-Frequency Synthesis (MSMFS, also \MAM\footnote{SDP adopted this name since it is easier to say.}) to model the changing structure of the source over frequency. By synthesising across frequency, the imaging performance of the telescope can be improved.

\MAM\ consists of a major cycle that compares the current model with the visibility data, and a minor cycle that deconvolves in (RA, Dec, scale, moment) space. The major cycle is composed to gridding, degridding, and FFT steps and has seen the most effort in optimisation. The minor cycle of \MAM\ consists of a modified and extended CLEAN in image space. This is currently executed in a single thread and can constitute a significant overhead in that processing is blocked until this CLEAN finishes.

The amount of processing in typical use of \MAM\ is comparable to that in the invert and predict steps. Hence since the \MAM\ step is currently serial, it will inevitably result in a bottleneck. The cure is to reduce the amount of work in \MAM\ and and/or distribute the work across multiple workers.

We investigate the various ways that the distribution can be performed.

\newpage

\subsection*{Reference Documents}

The following documents are referenced in this document. In the event of
conflict between the contents of the referenced documents and this document,
\emph{this document} shall take precedence.

 \begin{center}
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
   {\bf [RD01]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
   {\bf [RD02]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
   {\bf [RD03]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
   {\bf [RD04]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
   {\bf [RD05]} & Junklewitz, H., Bell, M. R., and Ensslin, T. (2015). A new approach to multifrequency synthesis in radio interferometry. Astronomy \& Astrophysics, 581, A59. http://doi.org/10.1051/0004-6361/201423465 \\
	{\bf [RD06]} & Rau, U., \& Cornwell, T. J. (2011, June 14). A multi-scale multi-frequency deconvolution algorithm for synthesis imaging in radio interferometry. arXiv.org. EDP Sciences. http://doi.org/10.1051/0004-6361/201117104   \\
	{\bf [RD07]} & Bhatnagar, S., Rau, U., \& Golap, K. (2013). Wide-field wide-band Interferometric Imaging: The WB A-Projection and Hybrid Algorithms. The Astrophysical Journal, 770(2), 91. http://doi.org/10.1088/0004-637X/770/2/91 \\
	{\bf [RD08]} & Cornwell, T.J. (2016), Implementation of the Rau-Cornwell MSMFS algorithm, TCC-SDP-151123-2. \\
	\hline
 \end{tabularx}
\end{center}

\pagebreak
\section{Purpose of the document}

The purpose of this document is to describe options for distributing the minor cycle of the Multi-Scale MultiFrequency Synthesis algorithm, as implemented in ASKAPsoft,.

\section{Scope of the document}

The scope includes only the published Multi-Scale, Multi-Frequency algorithms, existing implementations, and inclusion in the ASKAPsoft processing pipelines. Only the minor cycle of that algorithm is addressed. In addition, the many possible threading possibilities are not explored.

\clearpage

\section{\MAM\ algorithm in ASKAPsoft}

The \MAM\ algorithm implementation has been discussed extensively in RD08. We use the same terminology and symbols.

The (serial) minor cycle from the Urvashi-Cornwell algorithm [RD03, RDO6] is shown below in both pseudo-code (Algorithm \ref{alg:minor_cycle_UC2011}) and python (Figure \ref{fig:minor_cycle_python}).

\begin{algorithm}
%\scriptsize
  \SetLine
  \linesnumbered
  \dontprintsemicolon
  \KwData{input : number of Taylor-terms $\Nt $, number of scales $\Ns $}
  \KwData{input : image noise threshold, $\sigma_{\rm thr}$, loop gain $g$}
  \KwData{input : scale basis functions : $\I^{\rm shp}_s ~ \forall s\in\{0,\Ns -1\}$}
  \KwResult{model coefficient images : $\I^{\rm m}_{t}~ \forall t\in\{0,\Nt -1\}$}
  \Repeat { Peak residual in $\vec{I}^{\rm res}_{{0}\atop{0}} < f_{\rm limit} $ } 
    {
     \For{$s \in \{0,\Ns $-$1\}$}
     {
       {
          Construct $\I^{\rm rhs}_s$, an $\Nt \times 1$ vector from $\I^{\rm res}_{{s}\atop{t}} ~~\forall ~ t \in \{0,\Nt $-$1\}$\;
          Compute principal solution $\I^{\rm sol}_s = [{\He^{\rm peak}_s}^{-1}] \I^{\rm rhs}_s$\;
          Fill solution $\I^{\rm sol}_s$ into model images $\forall t$ : $\I^{\rm m,sol}_{{s}\atop{t}}$
       }
     }
     Choose $\I^{\rm m}_{{p}\atop{t}} = max\{\I^{\rm m,sol}_{{s}\atop{t=0}},~\forall~s\in\{0,\Ns $-$1\}\}$ \;
       \For{$t \in \{0,\Nt -1\}$}
       {
        Update the model image : $\I^{\rm m}_t = I^{\rm m}_t + g ~[ \I^{\rm shp}_{p} \star \I^{\rm m}_{{p}\atop{t}}]$ \;
        }
       \For{$t \in \{0,\Nt -1\}$}
       {
        \For{$s \in \{0,\Ns $-$1\}$}
	{
          Update the residual image : $\I^{\rm res}_{{s}\atop{t}} = \I^{\rm res}_{{s}\atop{t}} - g ~\sum_{q=0}^{\Nt -1}[\I^{\rm PSF}_{{sp}\atop{tq}} \star \I^{\rm m}_{{p}\atop{t}}]$\;
	}
       }
    }
\caption[Minor cycle of \MAM\ Algorithm]
         {Minor cycle of \MAM}\label{algo:ASKAP}
\label{alg:minor_cycle_UC2011}
\end{algorithm}

\begin{figure}
\begin{tcolorbox}
\inputminted{python}{minor_cycle.py}
\end{tcolorbox}
\caption{Minor cycle in python}
\label{fig:minor_cycle_python}
\end{figure}

\clearpage

\pagebreak
\section{Scale of problem}

To study the scale of the problem without running large numbers of simulations, we have adapted the SKA Performance Model to ASKAP processing. The PM is a python library and accompanying Jupyter notebooks. The various steps in SKA processing have been expressed using the symbolic math capabilities of sympy. This allows both evaluation of a particular configuration of processing as a function of a range of parameters. We have developed a branch of the code specially adapted to ASKAP processing.

The SKA ICAL pipeline performs continuum imaging and self-calibration. A typical case for ASKAP is shown in figure \ref{fig:processing_histogram}.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./processing_histogram.png}
  \caption{Estimation of the cost of processing steps from the SKA Performance Model as adapted to ASKAP. Left column is for large PSF patch size, right is for patch size of 512, which is the recommended value for ASKAP processing.}
  \label{fig:processing_histogram}
\end{figure}

Identify Component in the SKA model corresponds to line 2 - 6 in \ref{alg:minor_cycle_UC2011}, and Subtract Image Component to lines 7 - 15. The other parts of the processing in the histogram do distribute with reasonably high efficiency.

The minor cycle is currently purely serial and therefore constitutes a loss of processing efficiency in a parallel processing context. For N processors, and with a fraction p being amenable to parallelisation, Amdahl's law constrains the maximum speedup to be:

\begin{equation}
S = \frac{1}{(1-p) + \frac{p}{N}}
\end{equation}

The resulting curves are shown in figure \ref{fig:amdahl}.

The fraction of processing, $p$, for \MAM\ is about 4 \%. Thus the maximum possible speedup of the ICAL pipeline is about 25.
\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./AmdahlsLaw.png}
  \caption{Amdahl's law for various fractions of parallel portion.}
  \label{fig:amdahl}
\end{figure}

\clearpage

\pagebreak
\section{Distribution}
\label{sec:distribution}

In this section, we discuss the opportunities for distribution of the inner loop over multiple processing nodes.

The axes of the inner loop are (RA, Dec, Scale, Moment). The numerical ranges of these parameters are as follows:
\begin{description}
\item[RA, Dec] The image sizes for ASKAP are typically 4k x 4k or less. For SKA, numbers are much larger, up to 100k x 100k.
\item[Scale] Typically 3 - 5 scales are used, in a roughly geometric progression such as [0, 3, 10, 30, 100].
\item[Moment] Between 3 and 5 (frequency) moments are typical.
\end{description}

Distributing over scale will at most gain a factor of order 5. The network costs will be very significant.

More opportunity for distribution lies in the spatial axes. Spatially the coupling occurs for two reasons: the point spread function and the MS blobs. The point spread function can usually be assumed to be sufficiently well-behaved that updating of the residuals is only need in a limited neighbourhood. This statement is scale dependent and larger scales will inevitably show more coupling. It will be necessary to deal with components close to the partition boundary. Thus we will require enforcing some continuity constraint at the partition boundaries. There are a number of ways of doing so. We consider each of these in order of increasing complexity:

\begin{description}
\item[S1. Partition into sub-images]

Divide the image space into a number of disjoint partitions and clean each separately. The drawback to this approach is that large scales close to the boundaries will be penalised, leading to discontinuities in the final image.

\item[S2. Overlap the partitions]

To avoid the penalisation of large scales in S1, each partition could be extended with a suitably sized guard band. At the end of the minor cycle, neighbours could be smoothly interpolated.

\item[S3. Message passing inside the minor loop]

To avoid conflicts between components found in the overlap region, we could allow the neighbours to communicate every component found. We could partition as in S2 and communicate components found during the minor cycles, either peer-to-peer or to a central broker such as the master. Overlap by at least the largest scale size is required.

Each worker can check to see if a component overlaps with another worker or set of workers. If so, the component can be sent peer-to-peer or via the master. In theory MPI offers mechanisms for such communication but some experimentation might be necessary.

A message would be a flux value[scale, moment], a direction, and a shape, plus some ancillary information such as peak residual, taking O(100) bytes. For scaling, the communication of update events must occur in the same or less time as the worker execution times. The rate of overlap events is proportional to the fraction of the total image space that is subject to overlaps.

\end{description}

S1 and S2 are relatively minor changes and can be implemented and tested with a moderate amount of effort. S3 is considerably more complex to implement and might be subject of the vagaries of MPI.

\subsection{Data movement for S1, S2}
\label{sec:data_movement}

In ASKAPsoft, at the end of the major cycle the different frequency residual images are sent to the master and combined into the residual frequency cubes. In the current single-threaded approach in ASKAPsoft, the residual images (per Taylor term) and PSF are then deconvolved on the master. Thus there are two synchronisation points: one to assemble the Taylor term cubes and one to await the end of the deconvolution.

In options S1 and S2, the simplest change would be to still combine the image planes on the master but to extract the patch residual images and PSF and distribute those to the workers. Each worker would then apply \MAM\ to it's patch and return the resulting deconvolved image. The entire data flow to the master would be:

\begin{equation}
	R_{IO,{to\ master}} = \Np * \Nc
\end{equation}

Alternatively, instead of the data flow being centralised on the master, the workers could either broadcast the entire image of their frequency term to all workers or each worker could send just the appropriate patch of frequency to each worker.

\begin{equation}
	R_{IO,{pair-wise}} = {\Np/\Nf} * \Nc * \Nf
\end{equation}

Thus the data flow is the same in both cases but is bottlenecked to the master in the first approach.

The other important factor to consider is the working set size in both cases. In the 
${to\ master}$ approach the memory required in the master is significantly larger because the
entire image cube has to be assembled before distribution to the workers

\subsection{Overlaps and edge effects}

S1 and S2 inevitably create seams in the resulting images where the deconvolution has been performed from different pixels either side of the seam. Making the seam narrow is unlikely to be satisfactory since even for the Hogbom clean the discontinuity will lead to different deconvolution either side of the seam. For the MultiScale clean, the same effect will occur at least on all scales in the deconvolution. Thus we can expect that the taper should be extend over more pixels than the largest scale.

Some possible tapering functions are shown in Figure \ref{fig:tapers}. The simplest is simple addition or averaging, here shown as 'flat'. We improve by using a linear interpolation but this shifts discontinuities into the first derivative. The well-known Tukey filter removes these discontinuities by apply a cosine-taper, thus ensuring that only the second derivative is substantial. The computational cost in using the Tukey filter is negligible so we recommend always using it in preference to the flat of linear tapers.

The loss of efficiency due to the overlaps is non-negligible:

\begin{equation}
\epsilon_{overlap} = \left(\left(N_{pixel} - N_{facets} N_{overlap}\right)/N_{pixel}\right)^2
\end{equation}

For a typical ASKAP case, we would need about 2048 pixels on a side, 8 facets on a side and an overlap of 32, say. For this example, the efficiency due to overlap is $(2048 - 256)/2048) = 0.875$. Increasing the overlap by a factor of two to 64, the efficiency is 0.75. However, we can expect to gain much more than this factor by distribution so we regard this level of inefficiency as acceptable.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./tapers.png}
  \caption{Tapers for feathering facet images. The vertical pink lines show the where the overlaps start. This is for 256 pixels and overlap of 64. For this case, the efficiency is roughly $(64/256)^2 = 1/16$}
  \label{fig:tapers}
\end{figure}

\subsection{Load balancing}

The primary cause of load in-balance is that different sub-images will have different numbers of sources and hence different elapsed time for deconvolution. In the graph-based approach used in the \ARL\ this is somewhat mitigated by the intrinsic flexibility of the graph-based work allocation. In MPI, it is possible that some explicit load-balancing may be required. We therefore provide some analysis of the need for load balancing.

If the image is made with a guard band of one or two primary beams, the outermost sub-images will mostly have weak sources and require little deconvolution initially. For the example above of 8 by 8 sub-images, about 16 will have substantial flux to deconvolve. This means that at the beginning of a major/minor cycle algorithm, the maximum scaling is about 16, although we divided the image into 64 sub-images. Using smaller sub-images by a factor of two would bring this scaling limit up to about 64. Essentially the scaling is determined by the number of sub-images we can define across the primary beam.

We can improve on this rough argument by some straightforward modeling. To a reasonable approximation, the processing time for a given sub-image will be proportional to the fraction of a flux in that sub-image. For a given model of the sky, we can calculate the fraction of time taken for each sub-image\footnote{See section \ref{sec:simulations} for more details of the simulations}. We illustrate that point in figures \ref{fig:fractions_without_pb}, \ref{fig:fractions_with_pb}, and \ref{fig:fractions_clipped_image}. The first plot is a realistic astronomical field constructed from the S3-SEX catalog, selected for all sources brighter than 10uJy/beam. The curve shows the fraction of flux per sub-image against rank of that flux. The slope comes from the logN/logS statistics of the S3-SEX catalog. The peak value shows roughly the processing for the worst sub-image. This provides an upper bound on the scaling possible. The variable nproc is an estimate for the number of processors derived by dividing the total amount of processing by the peak value needed.

The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam. Most sub-images do not require cleaning and all the processing goes to the brightest fields. As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes. At this point, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy. The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.

So, to summarise, we expect the scaling to be poor for the initial major cycles (typically $\sim 5$ for 16 facets per axis) where the brightest sources in the main lobes are removed, but it will improve (typically to $\sim 50$) as the cleaning goes deeper. Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./fluxes_without_pb_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) without a primary beam applied. Scales are [0, 3, 10], and the overlap is 16 pixels. The slope for all curves is mainly due to the statistics of image which is derived from radio source LogN/LogS relationship.}
  \label{fig:fluxes_without_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./fluxes_with_pb_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) with a primary beam applied. Scales are [0, 3, 10], and the overlap is 16 pixels. Compared to figure \ref{fig:fluxes_without_pb}}, the primary beam application shifts the high rank sub-images lower in flux.
  \label{fig:fluxes_with_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./fluxes_clipped_image_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank for various numbers of sub-images, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels. There are now no bright sources left.}
  \label{fig:fractions_clipped_image}
\end{figure}

To get the overall scaling curve we must sum the work required for a sequence of major cycles going successively deeper. The thresholds start at 0.1 Jy and diminish by $\sqrt(10)$ as we proceed through 10 major cycles. We sum the results for each threshold weighting according to the number of sources at that level. We perform this analysis for both 2km and 6km configurations. The results are shown in figure \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}. The scaling is driven by the larger number of weak sources. The flatness of the curves as a function of rank means that strong scaling is possible for 200 - 300 processors. Adding more processors beyond this limit brings diminishing returns, and the overall efficiency drops.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./aggregate_scaling_ASKAP_2km.png}
  \caption{Aggregate scaling curve for 2km ASKAP configuration. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels.}
  \label{fig:aggregate_scaling_2km}
\end{figure}

We have repeated this calculation for the full 36 antenna 6km ASKAP (see figure \ref{fig:aggregate_scaling_6km}).

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./aggregate_scaling_ASKAP_6km.png}
  \caption{Aggregate predicted scaling curve for 6km configuration. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10, 30], and the overlap is 48 pixels.}
  \label{fig:aggregate_scaling_6km}
\end{figure}

\pagebreak
\section{Simulations}
\label{sec:simulations}

The Algorithm Reference Library\footnote{https://github.com/SKA-ScienceDataProcessor/algorithm-reference-library} was used for investigations. The ARL distributes imaging processing using the Dask library\footnote{http://dask.pydata.org}. Since the ARL is python-based, the absolute processing times are not to be taken as realistic for ASKAPsoft, but we use only the relative times in our analysis.

We made simulations as follows:
\begin{itemize}
	\item ASKAP antennas in 2km configuration, 
	\item 11 frequencies spread from 0.85GHz to 1.15GHz, enabling fitting of  Taylor series up to 5 terms (quintic). The bandwidth is sufficiently broad that over the frequency range far out sources can appear and disappear at different frequencies. Hence 5 terms are used for many investigations.
	\item The model was generated by querying the S3-SEX database\footnote{http://s-cubed.physics.ox.ac.uk/s3\_sex}, with a flux cutoff at 10uJy/beam. The spectral index was derived from the 1400MHz and 610MHz fluxes.
	\item Integrations spaced at 12 minutes, no noise added,
	\item Field center at ra=+30deg, dec=-60deg,
	\item Primary beam of 12m antenna with 2m blockage, no feed-legs.
	\item For some simulations the w term was set to zero to concentrate on the cleaning timing.
\end{itemize}

Figure \ref{fig:ASKAP_sim_restored} shows a typical restored image obtained by running 5 major cycles MSMFS with 3 Taylor terms. The residual image (\ref{fig:ASKAP_sim_residual}) shows emission out into the first lobe of the primary beam.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./ASKAP_sim_restored.png}
  \caption{Test field imaged with the MSMFS, major/minor cycle algorithm in the ARL. The fit is to the first three Taylor terms (i.e. quadratic). No faceting has been used for this image and the PSF support is to the entire field.}
  \label{fig:ASKAP_sim_restored}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./ASKAP_sim_residual.png}
  \caption{Residual for test field imaged with the MSMFS, major/minor cycle algorithm in the ARL. The fit is to the first three Taylor terms (i.e. quadratic). No faceting has been used for this image and the PSF support is to the entire field.}
  \label{fig:ASKAP_sim_residual}
\end{figure}

A typical result is shown in figures \ref{fig:ASKAP_sim_restored} (restored image) and \ref{fig:ASKAP_sim_residual} (residual image).

\subsection{Number of facets, and overlap}

\subsection{PSF support}

For a single iteration in the minor cycles, the processing breaks down into a search, scaling as the number of pixels, and a subtraction scaling as the product of the number of pixels in the image and the size of the PSF support. The value of the PSF support used thus directly affects the processing time. Furthermore, the processing via sub-image inevitably means that a limit PSF support will be used.

For the simulated case, the RMS side-lobes of the synthesized beam are about 0.3\% or less for pixels more than a few beams from the PSF peak. Hence we have reason to be optimistic that small values of PSF support will be acceptable given that we use a major/minor cycle. It is therefore important to determine the relationship is between final image quality and the PSF support.

We investigated the impact of limit PSF support using the simulations described above. In figure \ref{fig:error_vs_support}, we show the fractional error between the images obtained with the restricted PSF and the full PSF as a function of the support. The curve shows that the the effects of limited PSF support are much less than the criterion to switch to a major cycle (typical 10\% or less).

The corresponding run times (wall clock) are shown in figure \ref{fig:time_vs_support}. The scaling is as expected: for small supports, the time is roughly independent of the PSF support, whereas for larger values, the scaling steepens as the component subtraction comes to dominated.

Clearly, small PSF support should be used: the speedup is significant and there is no significant increase in error. As a corollary, the use of sub-images in the deconvolution should be acceptable within a major/minor cycle algorithm.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./error_versus_support.png}
  \caption{Fractional discrepancy as a function of PSF support. This shows a parallel 2048 by 2048 pixel, 3 term MSMFS clean in ARL. The computer was running 4 Dask workers with 32GB each. The discrepancy is the difference between the image processed with PSF support and the image processed with PSF support set to cover the entire image.}
  \label{fig:error_vs_support}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./time_versus_support}
  \caption{Processing time (seconds) as a function of PSF support. This shows  a parallel 1024 by 1024 pixel, 3 term MSMFS clean in ARL. The computer was running 4 Dask workers with 32GB each.}
  \label{fig:time_vs_support}
\end{figure}

\subsection{Distribution of residuals}

Do the residuals for no overlap and overlap have similar distributions?

\subsection{Spatial structure induced by partitioning}

Do the residuals show excess power on the scale of the overlap?

\pagebreak
\section{Porting to MPI}

For distributed processing, ASKAPsoft uses a master/worker pattern implemented using MPI. In the imager, the master apportions calculation of the normal equations (essentially the residual images) to the workers, gathers and sums the results, and performs a full image single threaded deconvolution using MSMFS. The updated model is then broadcast to all workers, and processing repeats with the new model. Different frequency channels are assigned to different workers. Forming and summing the normal equations is a reduction across the entire set of workers.

Since the scaling curves are relatively flat (see e.g.figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing. Load balancing would then only be needed for the high rank tail.

\pagebreak
\section{Work plan}

\pagebreak
\section{Conclusions}

\begin{enumerate}
	\item We have analysed the performance of the current single threaded MSMFS algorithm in ASKAPSoft. According to the SKA Parametric Model as adapted to ASKAP, the processing can consume anything in the range 30\% - close to 100\%
	\item The motivation for a distributed algorithm is therefore clear. We have developed and demonstrated a distributed MSMFS algorithm that tiles the image space into partially overlapping sub-images and runs the single-threaded MSMFS on each sub-image. The sub-images are then appropriately combined back into a single final image cube, using a Tukey tapering function to reconcile the overlap regions.
	\item This algorithm has been implemented in the SKA Algorithm Reference Library. The distribution of the processing is performed using a graph-based approach.
	\item We have performed an analysis of the large N scaling of this algorithm. The scaling is ultimately determined by the number of sub-images that can be used without distorting the multi-scale deconvolution. For the ASKAP 2km configuration using scales of 0, 3, 10 pixels, this translated to about 32 - 64 sub-images (per axis). For the initial deconvolution cycles, the processing is limited to a few bright sub-images and the strong scaling is limited to 3 - 10 processors. However, as the algorithm proceeds deeper, the sources to be deconvolved occupy more of the image and so more sub-images are active, allowing strong scaling up to 30 - 100 processors. On aggregate, the algorithm shows strong scaling up to 50 - 100 processors (32 - 64 sub-images per axis with 16 pixel overlap).
	\item For the 6km configuration strong scaling is predicted for up to 100 to 300 processors.
	\item This algorithm does not preclude the use of parallelization, such as OpenACC or CUDA to speed up the single sub-image MSMFS. As the speed of the MSMFS processing of a single sub-image improves, the overall speed should improve. 
	\item The flatness of the scaling up to a few hundred processors. There will still need to be some work distribution to deal with the tail in the scaling curves. However, this should fit naturally into the master/worker paradigm used in ASKAPsoft.
\end{enumerate}

% Add the bibliography
\clearpage

\addcontentsline{toc}{section}{References}
%\bibliography{your_bibfile}%

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-05-10 minor_cycle.py

@@ -1,0 +1,24 @@
+for i in range(niter):
+
+    # Lines 2 - 6
+    mscale, mx, my, mval = find_global_optimum(hsmmpsf,
+    	ihsmmpsf, smresidual, windowstack, findpeak)
+    scale_counts[mscale] += 1
+    scale_flux[mscale] += mval[0]
+
+    # Line 7
+    peak = numpy.max(numpy.fabs(mval))
+    lhs, rhs = overlapIndices(ldirty[0, ...],  psf[0, ...], mx, my)
+
+    # Lines 8 - 10
+    m_model = update_moment_model(m_model, 
+    	pscalestack, lhs, rhs, gain, mscale, mval)
+    
+    # Lines 11 - 15
+    smresidual = update_scale_moment_residual(smresidual, 
+        ssmmpsf, lhs, rhs, gain, mscale, mval)
+
+    if peak < absolutethresh:
+        break
+
+

### 2018-05-29 mmm-minorcycle.tex
Sub image MAM memo as submitted

@@
 % Revision
-\newcommand{\revision}{0.1\xspace}
+\newcommand{\revision}{0.9\xspace}
 % Author(s)
@@
 % Status of the document (draft/final/etc.)
-\newcommand{\docstatus}{Released\xspace}
+\newcommand{\docstatus}{Draft for review\xspace}
 
@@
     \hline
-    0.1 & 03/04/2018 & T.J. Cornwell & Initial version \\
+    0.9 & 30/04/2018 & T.J. Cornwell & Initial version \\
     \hline
@@
     \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
-    \item[FLOP] Floating Point Operations \vspace{-0.2cm}
-    \item[FLOPS] Floating Point Operations per Second \vspace{-0.2cm}
     \item[FoV] Field of View\vspace{-0.2cm}
@@
 
-We investigate the various ways that the distribution can be performed.
+We analyze the various ways that the distribution of processing can be performed and what performance we can expect.
+We develop and demonstrate an algorithm.
+Out analysis shows that strong scaling will possible up to hundreds of nodes.
+
+The algorithm has been coded in python using a graph-based distribution package, Dask. We describe the steps necessary to implement the algorithm in ASKAPsoft using MPI to achieve the distribution.
 
@@
      \hline
-   {\bf [RD01]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
-   {\bf [RD02]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
-   {\bf [RD03]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
-   {\bf [RD04]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
-   {\bf [RD05]} & Junklewitz, H., Bell, M.
-R., and Ensslin, T.
-(2015).
-A new approach to multifrequency synthesis in radio interferometry.
-Astronomy \& Astrophysics, 581, A59.
-http://doi.org/10.1051/0004-6361/201423465 \\
-	{\bf [RD06]} & Rau, U., \& Cornwell, T.
-J.
-(2011, June 14).
-A multi-scale multi-frequency deconvolution algorithm for synthesis imaging in radio interferometry.
-arXiv.org.
-EDP Sciences.
-http://doi.org/10.1051/0004-6361/201117104   \\
-	{\bf [RD07]} & Bhatnagar, S., Rau, U., \& Golap, K.
-(2013).
-Wide-field wide-band Interferometric Imaging: The WB A-Projection and Hybrid Algorithms.
-The Astrophysical Journal, 770(2), 91.
-http://doi.org/10.1088/0004-637X/770/2/91 \\
-	{\bf [RD08]} & Cornwell, T.J. (2016), Implementation of the Rau-Cornwell MSMFS algorithm, TCC-SDP-151123-2. \\
+     {\bf [RD01]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
+     {\bf [RD02]} & U.Rau, T.J.Cornwell, A multi-scale multi-frequency deconvolution algorithm for synthesis imaging in radio-interferometry, , Astronomy and Astrophysics, Volume 532, August 2011.\\
+	{\bf [RD03]} & Cornwell, T.J. (2016), Implementation of the Rau-Cornwell MSMFS algorithm, TCC-SDP-151123-2. \\
 	\hline
@@
 
-The purpose of this document is to describe options for distributing the minor cycle of the Multi-Scale MultiFrequency Synthesis algorithm, as implemented in ASKAPsoft,.
+The purpose of this document is to describe options for distributing the minor cycle of the Multi-Scale MultiFrequency Synthesis algorithm, as implemented in ASKAPsoft.
 
@@
 
-The scope includes only the published Multi-Scale, Multi-Frequency algorithms, existing implementations, and inclusion in the ASKAPsoft processing pipelines.
-Only the minor cycle of that algorithm is addressed.
+The scope includes only the published Multi-Scale, Multi-Frequency algorithm, existing implementations, and inclusion in the ASKAPsoft processing pipelines.
+Only distribution of the minor cycle of that algorithm is addressed.
 In addition, the many possible threading possibilities are not explored.
@@
 
-The \MAM\ algorithm implementation has been discussed extensively in RD08. We use the same terminology and symbols.
+The \MAM\ algorithm implementation has been discussed extensively in RD03. We use the same terminology and symbols.
 
-The (serial) minor cycle from the Urvashi-Cornwell algorithm [RD03, RDO6] is shown below in both pseudo-code (Algorithm \ref{alg:minor_cycle_UC2011}) and python (Figure \ref{fig:minor_cycle_python}).
+The (serial) minor cycle from the Urvashi-Cornwell algorithm [RD01, RD02, RDO3] is shown below in both pseudo-code (Algorithm \ref{alg:minor_cycle_UC2011}) and python (Figure \ref{fig:minor_cycle_python}).
 
@@
   \centering
-  \includegraphics[width=\textwidth]{./processing_histogram.png}
+  \includegraphics[width=\textwidth]{./pngs/processing_histogram.png}
   \caption{Estimation of the cost of processing steps from the SKA Performance Model as adapted to ASKAP.
@@
   \centering
-  \includegraphics[width=\textwidth]{./AmdahlsLaw.png}
+  \includegraphics[width=\textwidth]{./pngs/AmdahlsLaw.png}
   \caption{Amdahl's law for various fractions of parallel portion.}
@@
   \centering
-  \includegraphics[width=\textwidth]{./tapers.png}
+  \includegraphics[width=\textwidth]{./pngs/tapers.png}
   \caption{Tapers for feathering facet images.
@@
   \label{fig:tapers}
-\end{figure}
-
-\subsection{Load balancing}
-
-The primary cause of load in-balance is that different sub-images will have different numbers of sources and hence different elapsed time for deconvolution.
-In the graph-based approach used in the \ARL\ this is somewhat mitigated by the intrinsic flexibility of the graph-based work allocation.
-In MPI, it is possible that some explicit load-balancing may be required.
-We therefore provide some analysis of the need for load balancing.
-
-If the image is made with a guard band of one or two primary beams, the outermost sub-images will mostly have weak sources and require little deconvolution initially.
-For the example above of 8 by 8 sub-images, about 16 will have substantial flux to deconvolve.
-This means that at the beginning of a major/minor cycle algorithm, the maximum scaling is about 16, although we divided the image into 64 sub-images.
-Using smaller sub-images by a factor of two would bring this scaling limit up to about 64.
-Essentially the scaling is determined by the number of sub-images we can define across the primary beam.
-
-We can improve on this rough argument by some straightforward modeling.
-To a reasonable approximation, the processing time for a given sub-image will be proportional to the fraction of a flux in that sub-image.
-For a given model of the sky, we can calculate the fraction of time taken for each sub-image\footnote{See section \ref{sec:simulations} for more details of the simulations}.
-We illustrate that point in figures \ref{fig:fractions_without_pb}, \ref{fig:fractions_with_pb}, and \ref{fig:fractions_clipped_image}.
-The first plot is a realistic astronomical field constructed from the S3-SEX catalog, selected for all sources brighter than 10uJy/beam.
-The curve shows the fraction of flux per sub-image against rank of that flux.
-The slope comes from the logN/logS statistics of the S3-SEX catalog.
-The peak value shows roughly the processing for the worst sub-image.
-This provides an upper bound on the scaling possible.
-The variable nproc is an estimate for the number of processors derived by dividing the total amount of processing by the peak value needed.
-
-The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam.
-Most sub-images do not require cleaning and all the processing goes to the brightest fields.
-As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes.
-At this point, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy.
-The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.
-
-So, to summarise, we expect the scaling to be poor for the initial major cycles (typically $\sim 5$ for 16 facets per axis) where the brightest sources in the main lobes are removed, but it will improve (typically to $\sim 50$) as the cleaning goes deeper.
-Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{./fluxes_without_pb_ASKAP_2km.png}
-  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) without a primary beam applied.
-Scales are [0, 3, 10], and the overlap is 16 pixels.
-The slope for all curves is mainly due to the statistics of image which is derived from radio source LogN/LogS relationship.}
-  \label{fig:fluxes_without_pb}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{./fluxes_with_pb_ASKAP_2km.png}
-  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) with a primary beam applied.
-Scales are [0, 3, 10], and the overlap is 16 pixels.
-Compared to figure \ref{fig:fluxes_without_pb}}, the primary beam application shifts the high rank sub-images lower in flux.
-  \label{fig:fluxes_with_pb}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{./fluxes_clipped_image_ASKAP_2km.png}
-  \caption{Integrated sub-image flux versus rank for various numbers of sub-images, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam.
-Scales are [0, 3, 10], and the overlap is 16 pixels.
-There are now no bright sources left.}
-  \label{fig:fractions_clipped_image}
-\end{figure}
-
-To get the overall scaling curve we must sum the work required for a sequence of major cycles going successively deeper.
-The thresholds start at 0.1 Jy and diminish by $\sqrt(10)$ as we proceed through 10 major cycles.
-We sum the results for each threshold weighting according to the number of sources at that level.
-We perform this analysis for both 2km and 6km configurations.
-The results are shown in figure \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}.
-The scaling is driven by the larger number of weak sources.
-The flatness of the curves as a function of rank means that strong scaling is possible for 200 - 300 processors.
-Adding more processors beyond this limit brings diminishing returns, and the overall efficiency drops.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{./aggregate_scaling_ASKAP_2km.png}
-  \caption{Aggregate scaling curve for 2km ASKAP configuration.
-Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam.
-Scales are [0, 3, 10], and the overlap is 16 pixels.}
-  \label{fig:aggregate_scaling_2km}
-\end{figure}
-
-We have repeated this calculation for the full 36 antenna 6km ASKAP (see figure \ref{fig:aggregate_scaling_6km}).
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{./aggregate_scaling_ASKAP_6km.png}
-  \caption{Aggregate predicted scaling curve for 6km configuration.
-Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam.
-Scales are [0, 3, 10, 30], and the overlap is 48 pixels.}
-  \label{fig:aggregate_scaling_6km}
 \end{figure}
@@
 	\item 11 frequencies spread from 0.85GHz to 1.15GHz, enabling fitting of  Taylor series up to 5 terms (quintic).
-The bandwidth is sufficiently broad that over the frequency range far out sources can appear and disappear at different frequencies.
+The bandwidth is sufficiently broad that over the frequency range far out sources can appear and disappear and then appear again at different frequencies.
 Hence 5 terms are used for many investigations.
-	\item The model was generated by querying the S3-SEX database\footnote{http://s-cubed.physics.ox.ac.uk/s3\_sex}, with a flux cutoff at 10uJy/beam.
-The spectral index was derived from the 1400MHz and 610MHz fluxes.
+	\item The source catalogs generated by querying the S3-SEX database\footnote{http://s-cubed.physics.ox.ac.uk/s3\_sex}, with a flux cutoff at 10uJy/beam and 100uJy/beam.
+The spectral index was derived from the 1400MHz and 610MHz fluxes.\footnote{Caveat emptor: S3-SEX contains a 730 Jy source in the catalog.
+We have removed this prior to these simulations on the basis that it should be dealt with by peeling instead of cleaning.}
 	\item Integrations spaced at 12 minutes, no noise added,
@@
 	\item For some simulations the w term was set to zero to concentrate on the cleaning timing.
+	\item The subimages are calculated so that the size is constant.
+Thus if the overlap is non-zero, the actual number of sub-images on an axis is two less than the nominal {\em e.g.} If the nominal is 8 then the actual is 6 and then there will be 36 sub-images spread over the two dimensions.
+We chose this approach to avoid having to  analyze the behaviour for different size subimages.
 \end{itemize}
 
-Figure \ref{fig:ASKAP_sim_restored} shows a typical restored image obtained by running 5 major cycles MSMFS with 3 Taylor terms.
+Figure \ref{fig:ASKAP_sim_restored} shows a typical restored image obtained by running 5 major cycles \MAM\ with 5 Taylor terms.
 The residual image (\ref{fig:ASKAP_sim_residual}) shows emission out into the first lobe of the primary beam.
@@
   \centering
-  \includegraphics[width=\textwidth]{./ASKAP_sim_restored.png}
+  \includegraphics[width=\textwidth]{./pngs/ASKAP_sim_restored.png}
   \caption{Test field imaged with the MSMFS, major/minor cycle algorithm in the ARL.
-The fit is to the first three Taylor terms (i.e.
-quadratic).
-No faceting has been used for this image and the PSF support is to the entire field.}
+The fit is to the first five Taylor terms (i.e.
+quintic).
+16 facets on each axis were used, with overlap 16 pixels.}
   \label{fig:ASKAP_sim_restored}
@@
   \centering
-  \includegraphics[width=\textwidth]{./ASKAP_sim_residual.png}
+  \includegraphics[width=\textwidth]{./pngs/ASKAP_sim_residual.png}
   \caption{Residual for test field imaged with the MSMFS, major/minor cycle algorithm in the ARL.
-The fit is to the first three Taylor terms (i.e.
-quadratic).
-No faceting has been used for this image and the PSF support is to the entire field.}
+The fit is to the first five Taylor terms (i.e.
+quintic polynomial).
+16 facets on each axis were used, with overlap 16 pixels.}
   \label{fig:ASKAP_sim_residual}
@@
 
-\subsection{Number of facets, and overlap}
-
 \subsection{PSF support}
 
-For a single iteration in the minor cycles, the processing breaks down into a search, scaling as the number of pixels, and a subtraction scaling as the product of the number of pixels in the image and the size of the PSF support.
+The first possible stumbling block is that the cleaning requires a PSF that crosses facet boundaries. If this was actually correct then the sub-image \MAM\ would not be feasible.
+
+In the current single threaded \MAM\ the minor cycles, the processing consists of a search step, scaling as the number of pixels, and a subtraction step,  scaling as the product of the number of pixels in the image and the size of the PSF support.
 The value of the PSF support used thus directly affects the processing time.
@@
 
-For the simulated case, the RMS side-lobes of the synthesized beam are about 0.3\% or less for pixels more than a few beams from the PSF peak.
+For the simulated case, the RMS side-lobes of the (uniform weighted) synthesized beam are about 0.3\% or less for pixels more than a few beams from the PSF peak.
 Hence we have reason to be optimistic that small values of PSF support will be acceptable given that we use a major/minor cycle.
@@
 In figure \ref{fig:error_vs_support}, we show the fractional error between the images obtained with the restricted PSF and the full PSF as a function of the support.
-The curve shows that the the effects of limited PSF support are much less than the criterion to switch to a major cycle (typical 10\% or less).
+The curve shows that the the effects of limited PSF support are much less than the criterion to switch to a major cycle (typically 10\% or less).
 
@@
 
-Clearly, small PSF support should be used: the speedup is significant and there is no significant increase in error.
-As a corollary, the use of sub-images in the deconvolution should be acceptable within a major/minor cycle algorithm.
+Clearly, in the context of a major/minor cycle algorithm, small PSF support can be used: the speedup is significant and there is no significant increase in error compared to the cutoff for the transition from minor to major cycle.
+This is confirmed by evaluating the difference between images processed using \MAM\ and Sub-image \MAM\ after ten major cycles (see figure \ref{fig:error_vs_support_ci}.
+
+We should emphasize that this result also holds for the current use of the single threaded \MAM\: the psf support can be set to a value much smaller than 512.
+One possible limit would be where the subtraction step becomes comparable to the search cost.
+This is obviously context dependent but we would expect values in the range 32 - 128.
+However, figure \ref{fig:time_vs_support} shows that the gains are relatively small.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{./error_versus_support.png}
+  \includegraphics[width=\textwidth]{./pngs/error_versus_support.png}
   \caption{Fractional discrepancy as a function of PSF support.
-This shows a parallel 2048 by 2048 pixel, 3 term MSMFS clean in ARL.
+This shows a parallel 2048 by 2048 pixel, 3 term \MAM\ clean in ARL.
 The computer was running 4 Dask workers with 32GB each.
@@
   \centering
-  \includegraphics[width=\textwidth]{./time_versus_support}
-  \caption{Processing time (seconds) as a function of PSF support. This shows  a parallel 1024 by 1024 pixel, 3 term MSMFS clean in ARL. The computer was running 4 Dask workers with 32GB each.}
+  \includegraphics[width=\textwidth]{./pngs/error_versus_support_ci.png}
+  \caption{Fractional discrepancy as a function of PSF support after five major cycles.
+This shows a parallel 2048 by 2048 pixel, 3 term \MAM\ clean in ARL.
+The computer was running 4 Dask workers with 32GB each.
+The discrepancy is the difference between the image processed with PSF support and the image processed with PSF support set to cover the entire image.}
+  \label{fig:error_vs_support_ci}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/time_versus_support}
+  \caption{Processing time (seconds) as a function of PSF support. This shows  a parallel 1024 by 1024 pixel, 3 term \MAM\ clean in ARL. The computer was running 4 Dask workers with 32GB each.}
   \label{fig:time_vs_support}
@@
 
-\subsection{Distribution of residuals}
+\subsection{Number of facets, and overlap}
 
-Do the residuals for no overlap and overlap have similar distributions?
+We turn now to the connection between the partitioning of the image by the used of overlapped facets and the performance of the sub-imaging algorithm.
+Accepting the performance as given by the existing single-threaded code running in parallel on many sub-images, the key question is whether the processing load is well-balance and provides strong scaling.
 
-\subsection{Spatial structure induced by partitioning}
+The primary cause of load imbalance is that different sub-images will have different numbers of sources and hence different elapsed time for deconvolution.
+In the graph-based approach used in the \ARL\ this is observed directly.
+In ARL it is somewhat mitigated by the intrinsic flexibility of the graph-based work allocation.
 
-Do the residuals show excess power on the scale of the overlap?
+If the image is made with a guard band of one or two primary beams, the outermost sub-images will mostly have weak sources and require little deconvolution initially.
+For the example above of 8 by 8 sub-images, about 16 will have substantial flux to deconvolve.
+This means that at the beginning of a major/minor cycle algorithm, the maximum scaling is about 16, although we divided the image into 64 sub-images.
+Using smaller sub-images by a factor of two would bring this scaling limit up to about 64.
+Essentially the scaling is determined by the number of sub-images we can define across the primary beam.
+
+We can improve on this rough argument by some straightforward modeling.
+To a reasonable approximation, the processing time for a given sub-image will be proportional to the fraction of a flux in that sub-image.
+For a given model of the sky, we can calculate the fraction of time taken for each sub-image.
+We illustrate that point in figures \ref{fig:fluxes_without_pb}, \ref{fig:fluxes_with_pb}, and \ref{fig:fractions_clipped_image}.
+The first plot is a realistic astronomical field constructed from the S3-SEX catalog, selected for all sources brighter than 10uJy/beam.
+The curve shows the fraction of flux per sub-image against rank of that flux.
+The slope comes from the logN/logS statistics of the S3-SEX catalog.
+The peak value shows roughly the processing for the worst sub-image.
+This provides an upper bound on the scaling possible.
+The mean number of processors is an estimate derived by dividing the total amount of processing by the peak value needed.
+
+The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam.
+Most sub-images do not require cleaning and all the processing goes to the brightest fields.
+As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes.
+For the final major cycles, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy.
+The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.
+
+So, to summarise, we expect the scaling to be poor for the initial major cycles (typically $\sim 5$ for 16 facets per axis) where the brightest sources in the main lobes are removed, but it will improve (typically to $\sim 50$) as the cleaning goes deeper.
+Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/fluxes_No_primary_beam_ASKAP_2km_2048.png}
+  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) without a primary beam applied.
+Scales are [0, 3, 10], and the overlap is 16 pixels.
+The slope for all curves is mainly due to the statistics of image which is derived from radio source LogN/LogS relationship.
+This shows the complete absence of strong scaling i.e.
+adding more processors does not bring commensurate improvements in performance.}
+  \label{fig:fluxes_without_pb}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/fluxes_with_pb_ASKAP_2km.png}
+  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) with a primary beam applied.
+Scales are [0, 3, 10], and the overlap is 16 pixels.
+Compared to figure \ref{fig:fluxes_without_pb}, the primary beam application shifts the high rank sub-images lower in flux.
+The scaling has improved slightly.}
+  \label{fig:fluxes_with_pb}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/fluxes_clipped_image_ASKAP_2km.png}
+  \caption{Integrated sub-image flux versus rank for various numbers of sub-images, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 100uJy/beam.
+Scales are [0, 3, 10], and the overlap is 16 pixels.}
+  \label{fig:fractions_clipped_image}
+\end{figure}
+
+To get the overall scaling curve we must sum the work required for a sequence of major cycles going successively deeper.
+The thresholds start at $\sqrt{0.1}$ Jy and diminish by $\sqrt{10}$ as we proceed through 10 major cycles.
+We sum the results for each threshold.
+We perform this analysis for both 2km and 6km configurations.
+The results are shown in figure \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}.
+The scaling is driven by the larger number of weak sources.
+The flatness of the curves as a function of rank means that strong scaling is possible for 200 - 300 processors.
+Adding more processors beyond this limit brings diminishing returns, and the overall efficiency drops.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/aggregate_scaling_single_ASKAP_beam_ASKAP_2km_2048.png}
+  \caption{Aggregate scaling curve for 2km ASKAP configuration on multiple major cycles.
+Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam.
+Scales are [0, 3, 10], and the overlap is 16 pixels.
+The image size is 2048 by 2048 pixels.
+The scaling of work with numbers of sub-images (i.e.
+independent processors) is much closer to flat.}
+  \label{fig:aggregate_scaling_2km}
+\end{figure}
+
+We have repeated this calculation for the full 36 antenna, 6km ASKAP (see figure \ref{fig:aggregate_scaling_6km}).
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/aggregate_scaling_single_ASKAP_beam_ASKAP_6km_6144.png}
+  \caption{Aggregate predicted scaling curve for 6km configuration.
+Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam.
+Scales are [0, 3, 10, 30], and the overlap is 48 pixels.
+The image size is 6144 by 6144 pixels.}
+  \label{fig:aggregate_scaling_6km}
+\end{figure}
+
+If the deconvolution were to be done on the mosaiced data then the full mosaiced primary beam (see figure \ref{fig:full_askap_pb}) would be appropriate.
+This primary beam is larger and flatter so we would expect that the behaviour would be similar to the no-primary beam approach.
+The results for mosaiced primary beam are shown in figure \ref{fig:full_askap_aggregate_scaling_2km}.
+(Compare this to figure \ref{fig:aggregate_scaling_2km}).
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/full_askap_pb_4096.png}
+  \caption{Full ASKAP primary formed by summing PAF beams in quadrature. The image size is 4096 by 4096 pixels.}
+  \label{fig:full_askap_pb}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/full_askap_aggregate_scaling_ASKAP_2km_4096.png}
+  \caption{Aggregate scaling curve for 2km ASKAP configuration using the full mosaic primary beam.
+Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam.
+Scales are [0, 3, 10], and the overlap is 16 pixels.
+The image size is 4096 by 4096 pixels.}
+  \label{fig:full_askap_aggregate_scaling_2km}
+\end{figure}
+
+Figure \ref{fig:number_active_ASKAP_2km} shows the scaling of number of active sub-images with increasing cycle number in major/minor cycles. The processing scaling improves with cycle number.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/number_active_Single_ASKAP_beam_ASKAP_2km_2048.png}
+  \caption{Number of active ranks vs major/minor cycle, for 2km ASKAP configuration using the single primary beam.
+The threshold for each cycle starts at $\sqrt{0.1}$ Jy/beam and proceeds in geometric progression to 0.1 mJy/beam.}
+  \label{fig:number_active_ASKAP_2km}
+\end{figure}
+
+\subsection{Confirmation of scaling}
+
+Although the above behaviour does need to be verified on a large scale cluster, there is much that can be learned using small scale simulations on an 8 core Linux desktop.
+
+In figure \ref{fig:culm_time_vs_cycle} we show time taken per cycle for a simulation of the 2km ASKAP configuration using the single primary beam.
+In figure \ref{fig:time_vs_rank} we show the processing time as a function of sub-image for each cycle.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/culm_time_per_cycle.png}
+  \caption{Cumulative Measured time taken for the clean of each sub-image vs sub-image rank for each major/minor cycle, for 2km ASKAP configuration using the single primary beam.
+The threshold for each cycle starts at $\sqrt{0.1}$ Jy/beam and proceeds in geometric progression to 0.1 mJy/beam.
+The work increases steadily with cycle.
+Note that there is a stopping problem that can consume extra time, as well as the costs of a major cycle.}
+  \label{fig:culm_time_vs_cycle}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\textwidth]{./pngs/time_versus_rank.png}
+  \caption{Measured time taken for the clean of each sub-image vs sub-image rank for each major/minor cycle, for 2km ASKAP configuration using the single primary beam.
+The threshold for each cycle starts at $\sqrt{0.1}$ Jy/beam and proceeds in geometric progression to 0.1 mJy/beam.
+The work increases steadily with cycle.
+Note that there is a stopping problem that can consume extra time, as well as the costs of a major cycle.}
+  \label{fig:time_vs_rank}
+\end{figure}
 
 \pagebreak
-\section{Porting to MPI}
+\section{Porting to ASKAPsoft}
+
+The sub-image \MAM\ algorithm is now present in the ARL and can be run on both simulated and real data.
+Porting the algorithm to ASKAPsoft requires implementation in C++.
+ASKAPsoft has much of the necessary library capabilities, with the exception of sub-image scatter/gather.
+
+The core ASKAPsoft \MAM\ algorithm, DeconvolverMultiTermBasisFunction.tcc, can be retained as is.
 
@@
 
-Since the scaling curves are relatively flat (see e.g.figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing.
-Load balancing would then only be needed for the high rank tail.
+For the deconvolution step, it may be convenient to use a separate MPI work group. The ASKAPsoft master/worker framework nominally supports this.
+
+Since the scaling curves are relatively flat (see figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing.
+Load balancing would then only be needed for the high rank tail and should be performed automatically by the master/worker pattern.
+
+The major python-based ARL capabilities that would need to be reproduced in ASKAPsoft are for scatter and gather of an image into sub-images, allowing overlap and tapering.
+
+At the high numbers of processors, reduction in the scatter/gather steps may be needed.
+The total amount of network traffic is the full image size times the number of Taylor terms.
+Before beginning the sub-image cleans, the sub-images have to be scattered to the workers.
+
+To facilitate development, the sub-image \MAM\ should first be implemented in cdeconvolver to resolve the issues around use of the master-worker scheme and then ported to cimager.
 
@@
 \section{Work plan}
+
+The work plan to bring the Sub-image \MAM\ to ASKAPsoft. is straightforward. The steps required are:
+
+\begin{itemize}
+	\item Develop sub-image scatter/gather in ASKAPsoft.
+	\item Develop master/worker version of scatter/gather
+	\item Deploy in cdeconvolver as a test harness
+	\item Perform scaling tests and adjust as necessary
+	\item Deploy in cimager
+	\item Check for strong scaling
+	\item Run on simulated and real ASKAP and SKA data
+	\item Investigate scaling to SKA
+	\item Write report
+\end{itemize}
+
+As part of this work, the stopping problem should be pursued further since this is leading to substantial overhead in unnecessary major cycles.
+
+Access to ASKAP computing facilities will be necessary.
 
@@
 \begin{enumerate}
-	\item We have analysed the performance of the current single threaded MSMFS algorithm in ASKAPSoft.
-According to the SKA Parametric Model as adapted to ASKAP, the processing can consume anything in the range 30\% - close to 100\%
+	\item We have analysed the performance of the current single threaded \MAM\ algorithm in ASKAPSoft.
+As if there were any doubt, according to the SKA Parametric Model as adapted to the full scale ASKAP, the processing can consume anything in the range 30\% - close to 100\%.
 	\item The motivation for a distributed algorithm is therefore clear.
-We have developed and demonstrated a distributed MSMFS algorithm that tiles the image space into partially overlapping sub-images and runs the single-threaded MSMFS on each sub-image.
+We have developed and demonstrated a distributed \MAM\ algorithm that tiles the image space into partially overlapping sub-images and runs the single-threaded \MAM\ on each sub-image.
 The sub-images are then appropriately combined back into a single final image cube, using a Tukey tapering function to reconcile the overlap regions.
-	\item This algorithm has been implemented in the SKA Algorithm Reference Library. The distribution of the processing is performed using a graph-based approach.
-	\item We have performed an analysis of the large N scaling of this algorithm.
+We call this algorithm the Sub-image \MAM\ .
+	\item Sub-image \MAM\ has been implemented in the SKA Algorithm Reference Library. The distribution of the processing is performed using a graph-based approach.
+	\item Although we have not had access to a cluster with large numbers of processors, we have performed an analysis of the large N scaling of Sub-image \MAM\ .
 The scaling is ultimately determined by the number of sub-images that can be used without distorting the multi-scale deconvolution.
@@
 	\item For the 6km configuration strong scaling is predicted for up to 100 to 300 processors.
+	\item There will still need to be some work distribution to deal with the tail in the scaling curves. However, this should fit naturally into the master/worker paradigm used in ASKAPsoft.
 	\item This algorithm does not preclude the use of parallelization, such as OpenACC or CUDA to speed up the single sub-image MSMFS.
-As the speed of the MSMFS processing of a single sub-image improves, the overall speed should improve.
-	\item The flatness of the scaling up to a few hundred processors.
-There will still need to be some work distribution to deal with the tail in the scaling curves.
-However, this should fit naturally into the master/worker paradigm used in ASKAPsoft.
+As the speed of the \MAM\ processing of a single sub-image improves, the overall speed should improve.
+	\item Finally, it is worth emphasizing that this approach works because the PSF of ASKAP (for the simulated) data is really good with rms sidelobes about 0.3\%.
+Nearly all of the removal of the PSF sidelobes is actually performed via the major cycles.
+	\item The Sub-image \MAM\ is not well suited for the case where many scales are needed to represent the structure.
+In that case it may be necessary to perform successive processing with increasing maximum uv baseline.
+	\item The predicted scaling has yet to be confirmed on a configuration with many (hundreds) nodes.
 \end{enumerate}
-
-% Add the bibliography
-\clearpage
-
-\addcontentsline{toc}{section}{References}
-%\bibliography{your_bibfile}%
 

### 2018-06-18 mmm-minorcycle.tex
Rev 1.0

@@
 % Status of the document (draft/final/etc.)
-\newcommand{\docstatus}{Draft for review\xspace}
+\newcommand{\docstatus}{For release\xspace}
 
@@
 \newcommand{\firstsigaffiliation}{TCC}
-
-\newcommand{\secondsigname}{T.J. Cornwell}
-\newcommand{\secondsigdesignation}{Lead Consultant}
-\newcommand{\secondsigaffiliation}{TCC}
 
@@
     0.9 & 30/04/2018 & T.J. Cornwell & Initial version \\
+    \hline
+    1.0 & 18/06/2018 & T.J. Cornwell & Corrections from Mitch \\
     \hline
@@
 
-In the current single threaded \MAM\ the minor cycles, the processing consists of a search step, scaling as the number of pixels, and a subtraction step,  scaling as the product of the number of pixels in the image and the size of the PSF support.
+In the current single threaded \MAM\ the minor cycles, the processing consists of a search step, scaling as the number of pixels, and a subtraction step,  scaling as the square of the size of the PSF support.
 The value of the PSF support used thus directly affects the processing time.
@@
 Clearly, in the context of a major/minor cycle algorithm, small PSF support can be used: the speedup is significant and there is no significant increase in error compared to the cutoff for the transition from minor to major cycle.
-This is confirmed by evaluating the difference between images processed using \MAM\ and Sub-image \MAM\ after ten major cycles (see figure \ref{fig:error_vs_support_ci}.
+This is confirmed by evaluating the difference between images processed using \MAM\ and Sub-image \MAM\ after ten major cycles (see figure \ref{fig:error_vs_support_ci}).
 
-We should emphasize that this result also holds for the current use of the single threaded \MAM\: the psf support can be set to a value much smaller than 512.
+We should emphasize that this result also holds for the current use of the single threaded \MAM\ : the psf support can be set to a value much smaller than 512.
 One possible limit would be where the subtraction step becomes comparable to the search cost.
@@
 
-The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam.
+The next plot (figure \ref{fig:fluxes_with_pb}) shows the effect of the ASKAP (single) primary beam.
 Most sub-images do not require cleaning and all the processing goes to the brightest fields.
@@
 So, to summarise, we expect the scaling to be poor for the initial major cycles (typically $\sim 5$ for 16 facets per axis) where the brightest sources in the main lobes are removed, but it will improve (typically to $\sim 50$) as the cleaning goes deeper.
-Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}
+Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}.
 
@@
 
-Since the scaling curves are relatively flat (see figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing.
+Since the scaling curves are relatively flat (see figures \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing.
 Load balancing would then only be needed for the high rank tail and should be performed automatically by the master/worker pattern.
@@
 
+The gather/scatter step could be implemented either centrally or via neighbourhoods of workers. The latter avoids the use of a large amount of memory on the master to store the entire image.
+
 To facilitate development, the sub-image \MAM\ should first be implemented in cdeconvolver to resolve the issues around use of the master-worker scheme and then ported to cimager.
+The implementation should accommodate both central and localised gather/scatter as described in the previous paragraph.
 

### 2018-10-08 mmm-minorcycle-ska.tex
SKA version of Distribution memo

--- 2018-06-18 mmm-minorcycle.tex
@@
 
-\usepackage[sans]{tcc_doc} % SDP style file
+\usepackage[sans]{sdp_doc} % SDP style file
 \usepackage[english]{babel}
@@
 % on the front pages (will be filled in via the \sdpfrontpage command)
-\newcommand{\bigdoctitle}{Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
+\newcommand{\bigdoctitle}{SKA Memo ???: Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
 % Title to go in the "Document Status Sheet" Document number
-\newcommand{\docnr}{TCC-CASS-001\xspace}
+\newcommand{\doctype}{Memo}
+\newcommand{\docnr}{SKA Memo ???\xspace}
 % Context
-\newcommand{\context}{(ASKAPsoft)}
+\newcommand{\context}{(SDP.PIP)}
 % Revision
@@
 % Lead author (goes in the footer)
-\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
+\newcommand{\leadauthorname}{T.J.\ Cornwell\xspace}
+\newcommand{\leadauthordesignation}{Author}
+\newcommand{\leadauthoraffiliation}{TCC}
+\newcommand{\ownername}{T.J.\ Cornwell\xspace}
+\newcommand{\footerauthor}{T.J.\ Cornwell\xspace}
+\newcommand{\ownerdesignation}{Author}
+\newcommand{\owneraffiliation}{TCC}
+\newcommand{\approvername}{P.\ Alexander\xspace}
+\newcommand{\approverdesignation}{SDP Lead}
+\newcommand{\approveraffiliation}{U.Cam.}
+\newcommand{\releasername}{P.\ Alexander\xspace}
+\newcommand{\releaserdesignation}{SDP Lead}
+\newcommand{\releaseraffiliation}{U.Cam.}
 % Release date (YYYY-MM-DD)
@@
 % Status of the document (draft/final/etc.)
-\newcommand{\docstatus}{For release\xspace}
+\newcommand{\docstatus}{Draft\xspace}
 
@@
 
+\newcommand{\secondsigname}{T.J. Cornwell}
+\newcommand{\secondsigdesignation}{Lead Consultant}
+\newcommand{\secondsigaffiliation}{TCC}
+
 % Table with version numbers
@@
     \hline
-    0.9 & 30/04/2018 & T.J. Cornwell & Initial version \\
-    \hline
-    1.0 & 18/06/2018 & T.J. Cornwell & Corrections from Mitch \\
+    0.9 & 08/10/2018 & T.J. Cornwell & Ready for review \\
+    1.0 & ?? & T.J. Cornwell & To be signed \\
     \hline
@@
 }
-
 %%%%%%%%%%%%% END OF USER SETTINGS %%%%%%%%%%%%%%%%%%%
@@
 % Table with signatures
-\newcommand{\signaturetable}[3]{
+\newcommand{\signaturetable}[4]{
   \begin{tabularx}{\textwidth}{|X|X|X|}
     \hline
-    Name & Designation & Affilitation\\
+    {#1} & Designation & Affiliation\\
     \hline
-    {#1} & {#2} & {#3}\\
+    {#2} & {#3} & {#4}\\
     \hline
@@
 
-% Table with affiliations
+% Table with the SDP details
+\clearpage
 \newcommand{\organisationtable}{
@@
     \sffamily{\bf ORGANISATION DETAILS}\end{center}
-  \begin{table}[htb]
+  \begin{table}[htbp]
     \centering
-    \begin{tabular}[htb]{|l|l|}
+    \begin{tabular}[htbp]{|c|l|}
       \hline
-      Name & Tim Cornwell Consulting\\
+      Name & Science Data Processor Consortium\\
+      \hline
+      Address & Astrophysics\\
+      & Cavendish Laboratory\\
+      & JJ Thomson Avenue\\
+      & Cambridge CB3 0HE\\
+      \hline
+      Website & http://ska-sdp.org\\
+      \hline
+      Email & ska-sdp-pa@mrao.cam.ac.uk\\
       \hline
@@
 % load automatic pages
-\tccfrontpage
+\sdpfrontpage
 
@@
     \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
-    \item[MAM] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
+    \item[M\&M] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
     \item[PSF] Point Spread Function \vspace{-0.2cm}
@@
 
-\sdplistoftables
+%\sdplistoftables
 
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
   \centering
@@
 
-\pagebreak
-\section{Work plan}
-
-The work plan to bring the Sub-image \MAM\ to ASKAPsoft. is straightforward. The steps required are:
-
-\begin{itemize}
-	\item Develop sub-image scatter/gather in ASKAPsoft.
-	\item Develop master/worker version of scatter/gather
-	\item Deploy in cdeconvolver as a test harness
-	\item Perform scaling tests and adjust as necessary
-	\item Deploy in cimager
-	\item Check for strong scaling
-	\item Run on simulated and real ASKAP and SKA data
-	\item Investigate scaling to SKA
-	\item Write report
-\end{itemize}
-
-As part of this work, the stopping problem should be pursued further since this is leading to substantial overhead in unnecessary major cycles.
-
-Access to ASKAP computing facilities will be necessary.
-
+%\pagebreak
+%\section{Work plan}
+%
+%The work plan to bring the Sub-image \MAM\ to ASKAPsoft is straightforward. The steps required are:
+%
+%\begin{itemize}
+%	\item Develop sub-image scatter/gather in ASKAPsoft.
+%	\item Develop master/worker version of scatter/gather
+%	\item Deploy in cdeconvolver as a test harness
+%	\item Perform scaling tests and adjust as necessary
+%	\item Deploy in cimager
+%	\item Check for strong scaling
+%	\item Run on simulated and real ASKAP and SKA data
+%	\item Investigate scaling to SKA
+%	\item Write report
+%\end{itemize}
+%
+%As part of this work, the stopping problem should be pursued further since this is leading to substantial overhead in unnecessary major cycles.
+%
+%Access to ASKAP computing facilities will be necessary. 
+%
 \pagebreak
@@
 
+\pagebreak
+\section{Acknowledgements}
+
+I thank Daniell Mitchell for comments. This work ws supported by CSIRO CASS.
+
 \end{document}
